---
title: "Fort Collins Sonde Data"
author: "Sam Struthers- CSU ROSSyndicate"
date: "`r Sys.Date()`"
output: html_document
---

Installing Packages and library necessary ones

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


### ----- Load packages ----- ###
package_load <- function(package_names){
  for(i in 1:length(package_names)){
    if(!package_names[i] %in% installed.packages()){
      install.packages(package_names[i])
    }
    library(package_names[i],character.only = TRUE)
  }
}
#vector of packages
pack_req <- c( "tidyverse","lubridate","padr","janitor","padr", "broom","arrow","zoo",
  "rjson", "rvest", "dataRetrieval", "httr", "jsonlite", "RSelenium","devtools")
package_load(pack_req)

#Simple function to negate %in%
`%nin%` = Negate(`%in%`)

remove(pack_req, package_load)
```
This script will allow you to pull in sonde data from the City of Fort Collins Contrail site to access their sonde data from Manner's Bridge (site code: PMAN) and Indian Meadows (Site Code: PBR). This will grab all the available parameters from each sonde: Specific Conductivity, Turbidity, pH, Dissolved Oxygen, Temperature, and Stage. 

NOTE: This is RAW data and may include errors!!

# Set up user inputs

Enter below where you want the data to be saved, what to save it as and the start + end dates of the pull


```{r cars}
#start date time in format: YYYY-MM-DD HH:MM:SS in Mountain Time
start_DT <- "2024-08-01 12:00:00"
#end date time in format: YYYY-MM-DD HH:MM:SS in Mountain Time
#end_DT <- "2024-12-01 23:45:00"
### Or use the current time by uncommenting the line below ### 
end_DT <- as.character(round_date(Sys.time(), "15 minutes"))

#Check to make sure the end DT is greater than the start DT
if(as.POSIXct(end_DT) < as.POSIXct(start_DT)){
  stop("End date is before start date")
}


# Where do you want it to be saved?
folder_path <- "data/sensor_data/FC_sondes/"
# what do you want to save the final dataset as?
file_name <- "FC_sonde_data_2017_to_20240823"
# What file format do you want (csv, xslx or RDS)
format <- "RDS"

# What type of system are you using?
system <- "Mac" # Options are "Mac" or "PC" ... this is important!

```

Taking user input and converting it to the correct format and creating directory for data to be saved
 
```{r}


downloadPath <- file.path(getwd(), paste0(folder_path, "raw_data"))

if(system == "PC") {downloadPath <- downloadPath %>% stringr::str_replace_all("/", "\\\\\\\\")}

#Little code snippet to make the data folder if you don't have it. 
if(!file.exists(downloadPath)){
  dir.create(downloadPath)
}
```

If you want to remove old data, uncomment the following line:
```{r}
# delete old instances of the download: 
#do.call(file.remove, list(list.files(downloadPath, full.names = TRUE)))
```


# Setup R selenium and log into Contrail website

```{r}

creds = yaml::read_yaml("data/testing/credentials.yml") %>%
  unlist()
username = as.character(creds["username"])
password = as.character(creds["password"])


# make sure server downloads data into appropriate folder (i.e., our temporary folder)
fprof <- RSelenium::makeFirefoxProfile(list(browser.download.dir = downloadPath,
                                            browser.download.folderList = 2L,
                                            browser.download.manager.showWhenStarting = FALSE,
                                            browser.helperApps.neverAsk.openFile = "text/csv",
                                            browser.helperApps.neverAsk.saveToDisk = "text/csv"))

exCap <- list(firefox_profile = fprof$firefox_profile#,
              # REMOVE  "Pop-up" CONSOLE:
              #"moz:firefoxOptions" = list(args = list('--headless'))
)


# start up RSelenium session:

rD <- RSelenium::rsDriver(browser = "firefox", port = sample.int(1000, 1), extraCapabilities = exCap)

remDr <- rD$client

# log in to contrail system using username and password
remDr$navigate('https://contrail.fcgov.com/')
Sys.sleep(1) # provide time for site navigation

# Fill in username and password:
dropper <- remDr$findElement(using = 'xpath', value = '//*[@id="username"]')
Sys.sleep(1) # provide time for navigation and selection
try(dropper$sendKeysToElement(list(paste0(username))))
Sys.sleep(1)
dropper <- remDr$findElement(using = 'xpath', value = '//*[@id="password"]')
try(dropper$sendKeysToElement(list(paste0(password), key = "enter")))



```


# Pulling in the data

Depending on the time frame you have selected, this may take a fair amount of time (~ 5 min per year). If you select a time period where there is no data, you will recieve a message "Data not available for this date range" but it should not error out. 
Raw data will be saved to the folder you inputted above in a sub folder named `Raw Data`. If you selected multiple years, files will be saved with the '(i)'. I suggest runnning the following chunk before inspecting data. 

```{r}

FC_downloader <- function( start_DT, end_DT){
  
  # Convert the date string to URL-encoded format
start_DT_encoded <- gsub(" ", "%20", start_DT)%>%gsub(":", "%3A", .)
end_DT_encoded <- gsub(" ", "%20", end_DT)%>%gsub(":", "%3A", .)

metas <- read_csv( paste0(file.path(getwd(), paste0(folder_path),'device_urls.csv')), show_col_types = F)%>%
  mutate(encoded_url = paste0("https://contrail.fcgov.com/sensor/?time_zone=US%2FMountain",device_urls,"&bin=86400&range=Custom%20Range&markers=false&legend=true&thresholds=true&refresh=off&show_raw=true&show_quality=true&data_start=", start_DT_encoded, "&data_end=", end_DT_encoded))

download_data <- function(encoded_url){

remDr$navigate(encoded_url)
Sys.sleep(5)


apply_button <- tryCatch({
 remDr$findElement(using = "xpath", value = '//*[@id="dropdown-button-action-download"]')

}, error = function(e) {
  print("Data not available for this date range")
  return(NULL)
  stop()
})
  if(is.null(apply_button)){
   return(NULL)
    stop()
  }
  Sys.sleep(1)  # Pause to ensure the element is interactable
  apply_button$clickElement()
  
  # Attempt to find and click the "CSV" button
  csv_button <- remDr$findElement(using = "xpath", value = "//a[contains(text(), 'CSV')]")
  Sys.sleep(1)
  csv_button$clickElement()
  
  Sys.sleep(5)  # Pause to allow for download or other processes

}
walk(metas$encoded_url, download_data)


}

#determine if there is more than 1 year's worth of data to grab (365 days)
if(as.POSIXct(end_DT) - as.POSIXct(start_DT) > 365){
  multiple_years <- TRUE
  # split the start and end DT into 365 day chunks
  split_dates <- data.frame(start_DT_seq = seq.POSIXt(as.POSIXct(start_DT), as.POSIXct(end_DT), by = "365 DSTdays"))%>%
    mutate(end_DT_seq = lead(start_DT_seq - minutes(30), default = as.POSIXct(end_DT)))%>%
    mutate(start_DT = as.character(start_DT_seq),
           end_DT = as.character(end_DT_seq))%>%
    select(start_DT, end_DT)
 
}

if(multiple_years){
  for(i in 1:nrow(split_dates)){
    start_DT <- split_dates[i,1]
    end_DT <- split_dates[i,2]
    #print(paste0(start_DT," end: ", end_DT))
    FC_downloader(start_DT, end_DT)
  }
  #quit out of the RSelenium session
  remDr$close()
  rD[["server"]]$stop()

}else{
  FC_downloader(start_DT, end_DT)
  #quit out of the RSelenium session
  remDr$close()
rD[["server"]]$stop()

}



# end the RSelenium server session
#remDr$close()
rD[["server"]]$stop()

print('Download Complete')
```


# Tidying up data

This chunk will read in the data, clean it up, and save it as a  file in the folder you specified above.
This code will add the site codes/names to the file (Manners/PMAN, IndianMeadows/PBR), convert the reading time to MST and identify the parameter being measured.

```{r}

raw_files <- list.files(downloadPath, full.names = T)

collate_sensor_data <- function(file){
  
  site_name <- ifelse(grepl("Manners", file), "Manners", "Indian Meadows")
  site_cd <- ifelse(grepl("Manners", file), "PMAN", "PBR")
  
  data <- read_csv(file, show_col_types = F)%>%
    mutate(DT_MT = force_tz(Reading, tzone = "America/Denver"), 
           DT_mst = with_tz(DT_MT, tzone = "MST"),
           parameter = case_when(Unit == "uS/cm" ~ "Specific Conductivity",
                                  Unit == "mg/L" ~ "DO",
                                  Unit == "ntu" ~ "Turbidity",
                                  Unit == "pH" ~ "pH",
                                  Unit == "C" ~ "Temperature",
                                  Unit == "ft" ~ "Depth"), 
           site = site_name, 
           site_code = site_cd)%>%
    drop_na(parameter)%>%
    select(DT_mst, parameter, value = Value, `Data Quality`, site_code)
  
}

all_data <- map(raw_files, collate_sensor_data)%>%
  bind_rows()%>%
  distinct()

#save the consolidated file with the chosen name and format
file_save <- file.path(getwd(), paste0(folder_path, file_name))
if(system == "PC") {file_save <- file_save %>% stringr::str_replace_all("/", "\\\\\\\\")}

if(format == "RDS"){
  write_rds(all_data, paste0(file_save, ".RDS"))
}
if(format == "csv") {
  
  write_csv(all_data, paste0(file_save ,".csv"))
}
if(format == "xlsx") {
  
  write_xlsx(all_data, paste0(file_save, ".xlsx"))
}

```

# Checking out the data

```{r}
#read in data
if(format == "RDS"){
  all_data <- read_rds( paste0(file_save, ".RDS"))
}
if(format == "csv") {
  
 all_data<- read_csv(paste0(file_save ,".csv"), show_col_types = F)
}
if(format == "xlsx") {
  
  all_data <- readxl::read_xlsx(paste0(file_save, ".xlsx"))
}

ggplot(all_data, aes(DT_mst, value, color = site_code))+
  geom_point()+
  facet_wrap(~parameter, scales = "free_y")+
  theme_minimal()+
  labs(title = "Sensor Data", x = "Date/Time (MST)", y = "Value", color = "Site")

```
