---
title: "Figures"
author: "Sam Struthers"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("paper_analysis/00_analysis_setup.R")
library(here)

library(rstatix)
library(tidyverse)
library(car)
#basic function to save a plot in standard size, dpi and width into figures folder
save_plot_image<- function(plot,name){
  
  ggsave(paste0("docs/cpf_reservoir/paper_draft/",name,".jpg"), dpi = 300, width = 18, height = 10)
  
}

paper_draft_dir <- here("docs/cpf_reservoir/paper_draft/")
spatial_dir <- "data/raw/spatial/res_study"
external_spatial <- "data/raw/external_datasets"


#Seasonal Color values
season_clp_color_vals <- c("Winter" = "#03045e",  "Rising Limb" = '#047E82',  "Peak Flow" = '#59B851', "Falling Limb"='#FA850E', "Fall" = "#772f1a")

```

# Grab data

```{r}

#discrete sample data and location metadata from most recent pub
data_version = "v2025.11.14"


data_folder <- here("data/raw/chem/ross_clp_chem/v2025.11.14")

most_recent_chem <- read_rds(here(data_folder, "data/cleaned/CLP_chemistry_up_to_20251114.rds"))%>%
  mutate(flow_gauge_id = ifelse(flow_gauge_source == "USGS", paste0("0", flow_gauge_id), flow_gauge_id))
most_recent_meta <- read_csv_arrow(here(data_folder,"data/metadata/location_metadata.csv"))
units_meta_file <- here(data_folder, "data/metadata/water_chemistry_units.xlsx" )
chem_units <- readxl::read_xlsx(units_meta_file)
all_params <- readxl::read_xlsx(units_meta_file)%>%
  pull(simple)

rm(units_meta_file)


tidy_chem <- most_recent_chem%>%
  #remove random pre fire data 
  filter(Year > 2020 & Year < 2025)%>%
  filter(watershed %nin% c("CLP Mainstem - Lower","CLP Tributary - Lower"))%>%
  # pfal + slep represent the same site but were moved to match future sensor location
  mutate(distance_upstream_km = case_when(site_code == "PFAL" ~ 73.9, TRUE ~ distance_upstream_km))%>%
  mutate(location_type = if_else(site_code == "PFAL", "Mainstem", location_type))%>%
  # add order table which gives the sites an order to be plotted in
  left_join(order_table, by = "site_code")%>%
  #calculate nitrogen variables as mg/L of N instead of mg/L of NH4 or NO3
  mutate(n_nh4 = NH4 * 0.7765, 
         n_no3 = NO3 * 0.2259, 
         #calculate DIN and DON
         DIN = n_nh4 + n_no3,
         DON = DTN - DIN)%>%
  #filter site code BRNR with chla value of 0.02, seems to be sample error
  filter(!(site_code == "BRNR" & ChlA == 0.02))%>%
  #filter site code CBRR with TSS value > 1000, seems to be sample error
  mutate( TSS = case_when((site_code == "CBRR" & TSS > 1000) ~ NA_real_,
                          TRUE ~ TSS), 
          Year = as.character(Year), 
          month = as.character(month(Date)),
          season = case_when(
            month %in% c("3", "4") ~ "Rising Limb", 
            month %in% c( "5","6") ~ "Peak Flow",
            month %in% c("7", "8") ~ "Falling Limb",
            month %in% c("9", "10", "11") ~ "Fall Baseflow", 
            TRUE ~ "Winter"),
          burn_group = case_when(site_code %in% c("CBRR", "BRNR", "PTRR") ~ "burned", 
                                 TRUE ~ "unburned"), 
          burn_year = paste(burn_group, Year, sep = "_"), 
          site_year = paste(site_code, Year, sep = "_")
  )

#site order now added to tidy chem and thus not needed
rm( most_recent_chem, data_version)
```

## Run above to get data loaded

```{r}

```


# SI Tables

## PO4 

Looking at where PO4 is present (above 0.01 detection limit) and if that is changing over time

```{r}

upper_PO4 <- filter(tidy_chem, watershed %nin% c("CLP Mainstem - Lower","CLP Tributary - Lower"))%>%
  select(site_code,location_type, Date, Turbidity, PO4, Year)%>%
  mutate(location_type = case_when(
    site_code %in% c("JWC", "BEAV", "LBEA", "PENN", "SFM", "SAWM", "JOEI", "CBRI", "COMI", "HORI") ~ "Tributary",
    TRUE ~ location_type))%>%
  group_by(location_type, Year)%>%
  # fina what percent of each year had PO4 levels above 0.01 and get the mean level for each location type and year
  summarise(total_samples = n(),
            PO4_perc_dect = sum(PO4 >= 0.01, na.rm = TRUE)/n()* 100,
            #calculate mean of PO4 for only samples above 0.01, excluding samples lower than 0.01
            mean_PO4 = mean(PO4[PO4 >= 0.01], na.rm = TRUE),
            samples = paste0(sum(PO4 >= 0.01, na.rm = TRUE),"/",n() ))%>%
  # pivot wider by year
  pivot_wider(names_from = Year, values_from = c(total_samples, PO4_perc_dect, mean_PO4, samples)) %>%
  mutate(across(total_samples_2021:mean_PO4_2024, ~replace_na(.x, 0)))%>%
  mutate(across(samples_2021:samples_2024, ~replace_na(.x, "0/0")))%>%
  # if a column is numeric, round it to 2 decimal places
  mutate(across(where(is.numeric), ~round(.x, 2)))%>%
  arrange(factor(location_type, levels = c("Inflow", "Reservoir", "Outflow", "Mainstem", "Tributary")))%>%
  select(location_type, total_samples_2021:mean_PO4_2024)

#### Creating Table ####
# Add spacer columns
upper_PO4_with_spacers <- upper_PO4 %>%
  mutate(sp1 = NA, sp2 = NA, sp3 = NA) %>%
  select(1, 2:5, sp1, 6:9, sp2, 10:13, sp3)

# Rename columns with valid names
colnames(upper_PO4_with_spacers) <- c(
  "Site_Type",           # 1
  "n_2021", "n_2022", "n_2023", "n_2024", "sp1",  # 2-6
  "pct_2021", "pct_2022", "pct_2023", "pct_2024", "sp2",  # 7-11
  "mean_2021", "mean_2022", "mean_2023", "mean_2024", "sp3" # 12-16
)

# Create the flextable
ft1 <- flextable(upper_PO4_with_spacers) %>%
  set_header_labels(
    Site_Type = "",
    n_2021 = "2021", n_2022 = "2022", n_2023 = "2023", n_2024 = "2024",
    pct_2021 = "2021", pct_2022 = "2022", pct_2023 = "2023", pct_2024 = "2024",
    mean_2021 = "2021", mean_2022 = "2022", mean_2023 = "2023", mean_2024 = "2024",
    sp1 = "", sp2 = "", sp3 = ""
  ) %>%
  add_header_row(
    values = c("Site Type", "Samples per Year", "", "Percent above Detection Limit", "", 
               "Mean PO4 Above Detection Limit", ""),
    colwidths = c(1, 4, 1, 4, 1, 4, 1)
  ) %>%
  theme_vanilla() %>%
  align(align = "center", part = "all") %>%
  valign(valign = "center", part = "all") %>%
  fontsize(size = 11, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  set_table_properties(layout = "autofit", width = 1)%>%
  autofit()

ft1
save_as_docx(ft1,path = here(paper_draft_dir,"SI_1_Table.docx"))


```

## Key Constiuent Table

Calculate mean and max for key constituents for the 5 main study reservoirs

```{r}

key_table <- tidy_chem %>%
  filter(site_code %in% c("LNGR", "JOER","COMR", "HORR",  "CBRR", "PTRR", "BRNR")) %>%
  select(site_code, Year, pH,SC,TSS, DOC, DTN, K,NH4,NO3, SO4) %>%
  pivot_longer(cols = c(pH,SC,TSS, DOC, DTN, K,NH4,NO3, SO4),
               names_to = "parameter",
               values_to = "value") %>%
  #get the median and max for each site
  group_by(site_code, parameter) %>%
  summarise(median = median(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE),
            max = max(value, na.rm = TRUE)) %>%
  ungroup()%>%
  pivot_longer(cols = c(median, max, sd),
               names_to = "stat",
               values_to = "stat_value")%>%
  unite("param_stat", parameter, stat, sep = " ") %>%
  pivot_wider(names_from = param_stat, values_from = stat_value)%>%
  arrange(factor(site_code, levels = c("LNGR", "JOER", "HORR", "COMR", "CBRR", "PTRR", "BRNR")))%>%
  select(site_code:`NO3 sd`,`SO4 median`:`pH sd`, `SC median`,`SC max`,`SC sd`)%>%
  #round all numeric values to 2 digits
  mutate(across(where(is.numeric), ~ round(.x, 2)))


param_oi <- c("DOC", "DTN", "K", "NH4", "NO3", "SO4", "TSS", "pH", "SC")

params <- chem_units %>%
  filter(simple %in% param_oi) %>%
  arrange(match(simple, param_oi)) %>%
  pull(Parameters)


# Create a basic flextable
ft2 <- key_table %>%
  rename(Site = site_code) %>%
  flextable() %>%
  set_header_labels(
    .labels = setNames(rep(c("mean", "max", "sd"), times = length(params)),
                      nm = as.vector(t(outer("", c("median", "max","sd"), paste, sep = " "))))) %>%
  delete_part(part = "header")%>%
  add_header_row(
    values = c("", rep(c("Median", "Max", "SD"), times = length(params))),  
    colwidths = c(1, rep(1, length(params)*3))  # Adjust colwidths for Site and parameters
  ) %>%
  add_header_row(
    values = c("Site", params), 
    colwidths = c(1, rep(3, length(params)))  # Adjust colwidths for Site and parameters
  ) %>%
  
  add_header_row(
    values = c("", 
                "mg L⁻¹", "mg L⁻¹", "mg L⁻¹", 
               "mg L⁻¹", "mg L⁻¹", "mg L⁻¹", "mg L⁻¹",  " ", "µS cm⁻¹"), 
    colwidths = c(1, rep(3, length(params)))  # Adjust colwidths again to match
  ) %>%
  merge_v(j = ~Site) %>%
  merge_h(part = "header") %>%
  border_remove() %>%
  border_outer() %>%
  border_inner_v() %>%
  border_inner_h() %>%
  align(align = "center", part = "all") %>%
  valign(valign = "center", part = "all") %>%
  fontsize(size = 11, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  set_table_properties(layout = "autofit", width = 1)%>%
  autofit()

sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    width = 8.3, height = 11.7
  ),
  type = "continuous",
  page_margins = page_mar()
)

save_as_docx( ft2, path = here(paper_draft_dir,"SI_2_table.docx"))

```

### Same table but for all other parameters

```{r}

base_cations_table <- tidy_chem %>%
  filter(site_code %in% c("LNGR", "JOER","COMR", "HORR",  "CBRR", "PTRR", "BRNR")) %>%
  select(site_code, Year, ANC, Na, Mg, Ca, F, Cl) %>%
  pivot_longer(cols = c(ANC, Na, Mg, Ca, F, Cl),
               names_to = "parameter",
               values_to = "value") %>%
  #get the median and max for each site
  group_by(site_code, parameter) %>%
  summarise(median = median(value, na.rm = TRUE),
            max = max(value, na.rm = TRUE),
            sd = sd(value,na.rm = T)) %>%
  ungroup()%>%
  pivot_longer(cols = c(median, max,sd),
               names_to = "stat",
               values_to = "stat_value")%>%
  unite("param_stat", parameter, stat, sep = " ") %>%
  pivot_wider(names_from = param_stat, values_from = stat_value)%>%
  arrange(factor(site_code, levels = c("LNGR", "JOER", "HORR", "COMR", "CBRR", "PTRR", "BRNR")))%>%
  #round all numeric values to 2 digits
  mutate(across(where(is.numeric), ~ round(.x, 2)))


param_oi <- c("ANC", "Ca", "Cl", "F", "Mg", "Na")

params <- chem_units %>%
  filter(simple %in% param_oi) %>%
  arrange(match(simple, param_oi)) %>%
  pull(Parameters)


# Create a basic flextable
ft3 <- base_cations_table %>%
  rename(Site = site_code) %>%
  flextable() %>%
  set_header_labels(
    .labels = setNames(rep(c("mean", "max", "sd"), times = length(params)),
                      nm = as.vector(t(outer("", c("median", "max","sd"), paste, sep = " "))))) %>%
  delete_part(part = "header")%>%
  add_header_row(
    values = c("", rep(c("Median", "Max", "SD"), times = length(params))),  
    colwidths = c(1, rep(1, length(params)*3))  # Adjust colwidths for Site and parameters
  ) %>%
  add_header_row(
    values = c("Site", params), 
    colwidths = c(1, rep(3, length(params)))  # Adjust colwidths for Site and parameters
  ) %>%
  
  add_header_row(
    values = c("", 
                "mg L⁻¹", "mg L⁻¹", "mg L⁻¹", 
               "mg L⁻¹", "mg L⁻¹", "mg L⁻¹"), 
    colwidths = c(1, rep(3, length(params)))  # Adjust colwidths again to match
  ) %>%
  merge_v(j = ~Site) %>%
  merge_h(part = "header") %>%
  border_remove() %>%
  border_outer() %>%
  border_inner_v() %>%
  border_inner_h() %>%
  align(align = "center", part = "all") %>%
  valign(valign = "center", part = "all") %>%
  fontsize(size = 11, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  set_table_properties(layout = "autofit", width = 1)%>%
  autofit()

sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    width = 8.3, height = 11.7
  ),
  type = "continuous",
  page_margins = page_mar()
)

save_as_docx( ft3, path = here(paper_draft_dir,"SI_3_table.docx"))

```




## SI Plot 1: Timelines of key parameters at reservoirs 

```{r}

SI_plot <- tidy_chem %>%
  filter(site_code %in% c("LNGR", "JOER","COMR", "HORR",  "CBRR", "PTRR", "BRNR"))%>%
  pivot_longer(cols = c( ChlA, DTN, pH, SC, DOC, K, NH4, NO3, SO4, K, Ca, ANC, Na, Mg, Ca, F, Cl),
               names_to = "parameter",
               values_to = "value")%>%
  left_join(chem_units, by = c("parameter" = "simple"))%>%
  mutate(Date = as.POSIXct(Date ))


#plot 1 

Si_timeline_1 <- ggplot(SI_plot%>%filter(parameter %in% c("DOC", "DTN","pH", "SC", "ANC"))%>%
         mutate(parameter = factor(parameter, levels = c("DOC", "DTN","pH", "SC", "ANC"))),
       aes(x = Date, y = value, group = interaction(site_code, Year), color = site_code)) +
  geom_line() +
  geom_point()+
  facet_grid(rows = vars(combined),cols = vars(Year), scales = "free", switch = "y") +
  labs(x = "Date",
       y = "") +
  scale_color_manual(values = colorsBS_site_code)+
  scale_x_datetime(
    date_labels = "%b",  # e.g., May\n2021
    date_breaks = "month"
  ) +
  labs(color = "Site")+
  theme_bw() +
  theme(strip.placement = "outside", 
        legend.position = "bottom",
        # make the x ticks at 45 degrees
        axis.text.x = element_text(angle = 45, hjust = 1))

save_plot_image(Si_timeline_1,"SI_timeline_1.png")


SI_timeline_2 <- 
ggplot(SI_plot%>%filter(parameter %in% c("K", "Ca","SO4", "Na", "Cl", "Mg", "F")),
       aes(x = Date, y = value, group = interaction(site_code, Year), color = site_code)) +
  geom_line() +
  geom_point()+
  facet_grid(rows = vars(combined),cols = vars(Year), scales = "free", switch = "y") +
  labs(x = "Date",
       y = "") +
  scale_color_manual(values = colorsBS_site_code)+
  scale_x_datetime(
    date_labels = "%b",  # e.g., May\n2021
    date_breaks = "month"
  ) +
  labs(color = "Site")+
  theme_bw() +
  theme(strip.placement = "outside", 
        legend.position = "bottom",
        # make the x ticks at 45 degrees
        axis.text.x = element_text(angle = 45, hjust = 1))

save_plot_image(Si_timeline_2,"SI_timeline_2.png")

```


# Main Figures

There will be four main figures in this paper:
- Key nutrient dotplot
- Chla dot plot
- NO3 chla JWC relationship
- SF Chla (on vs off reservoir)


## Key Nutrient Dot Plot

### Function
```{r}
build_res_dotplot<- function(param = "ChlA", binwdth = 0.3, sz = 1, title_min = -.7){
  
    boxplot_res <- filter(tidy_chem, site_code %in% c("LNGR", "PTRR", "BRNR", "JOER", "CBRR") )%>%
      filter(!is.na(.data[[param]]))%>%
      mutate(site_year = paste(site_code, Year, sep = " "),
             dayofyear = yday(Date))
    param_units <- filter(chem_units, simple == param)%>%
      pull(combined)
    
    #Create plot
    ggplot(boxplot_res, aes(x=reorder(site_year,BS_order,max),y=.data[[param]], middle=mean(.data[[param]]), fill=site_code)) + 
      geom_dotplot(stackgroups = FALSE, binaxis = "y", method = "histodot",stackdir = "center", position = "dodge", binwidth = binwdth, dotsize = sz)+
      scale_color_grey(start = 0.4, end = 0)+
      #make x axis titles vertical
      theme(axis.text.x = element_text(angle = 90, hjust = 1))+
      geom_rect( inherit.aes=FALSE, aes(xmin="LNGR 2021", xmax="LNGR 2024", ymin=title_min,
                                        ymax=0), color="transparent", fill="#0072B2", alpha=0.01)+
      geom_rect( inherit.aes=FALSE, aes(xmin="JOER 2021", xmax="JOER 2024", ymin=title_min,
                                        ymax=0), color="transparent", fill="#56B4E9", alpha=0.05)+
      
      geom_rect( inherit.aes=FALSE, aes(xmin="CBRR 2021", xmax="CBRR 2024", ymin=title_min,
                                        ymax= 0), color="transparent", fill="#F9C74F", alpha=0.05)+
      
      geom_rect( inherit.aes=FALSE, aes(xmin="PTRR 2021", xmax="PTRR 2024", ymin=title_min,
                                        ymax=0), color="transparent", fill="#F3722C", alpha=0.05)+
      geom_rect( inherit.aes=FALSE, aes(xmin= 16.5, xmax= 20.5, ymin=title_min,
                                        ymax=0), color="transparent", fill="#d71a1d", alpha=0.008)+
      geom_abline(aes(slope=0, intercept=0), color='#000000', linetype="solid") +
      geom_text(aes(x= 2.5, y = (title_min/2), label= "Long Draw",  color = "#000000"), size = 6, alpha = 1)+
      geom_text(aes(x= 6.5, y = (title_min/2), label= "Joe Wright",  color = "#000000"), size = 6, alpha = 1)+
      geom_text(aes(x= 10.5, y = (title_min/2), label= "Chambers",  color = "#000000"), size = 6, alpha = 1)+
      geom_text(aes(x= 14.5, y = (title_min/2), label= "Peterson",  color = "#000000"), size = 6)+
      geom_text(aes(x= 18.5, y = (title_min/2), label= "Barnes Meadow",  color = "#000000"), size = 5, alpha = 1)+
  
      
      scale_fill_manual(name="Watershed", values = colorsBS_site_code) +
      theme_bw(base_size=25) +
      
      #theme(axis.text.x = ifelse(study_reservoir_chla$Location == "Reservoir",element_text(face = "bold"),element_text(face = "plain"))) +
      labs( x = "Site", y = param_units)+
      scale_x_discrete(labels = c("2021", "2022", "2023","2024",
                                  "","","","",
                                  "2021", "2022", "2023","2024",
                                  "","","","",
                                  "2021", "2022", "2023","2024"))+
      theme(legend.position = "none",
            axis.title = element_text(face = "bold"),
            # bold and angle x axis titles to 45
            axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"))
}


```

```{r}
build_res_dotplot_concise <- function(param = "K", binwdth = .05, sz = .5){
  
    
    boxplot_res <- filter(tidy_chem, site_code %in% c("LNGR", "PTRR", "BRNR", "JOER", "CBRR") )%>%
      filter(!is.na(.data[[param]]))%>%
      mutate(site_year = paste(site_code, Year, sep = " "),
             dayofyear = yday(Date))
    param_units <- filter(chem_units, simple == param)%>%
      pull(combined)
    
    #Create plot
    ggplot(boxplot_res, aes(x=reorder(site_year,BS_order,max),y=.data[[param]], middle=mean(.data[[param]]), fill=site_code)) + 
      geom_dotplot(stackgroups = FALSE, binaxis = "y", method = "histodot",stackdir = "center", position = "dodge", binwidth = binwdth, dotsize = sz)+
      scale_color_grey(start = 0.4, end = 0)+
      #make x axis titles vertical
      theme(axis.text.x = element_text(angle = 90, hjust = 1))+
      # geom_rect( inherit.aes=FALSE, aes(xmin="LNGR 2021", xmax="LNGR 2023", ymin=title_min,
      #                                   ymax=0), color="transparent", fill="#0072B2", alpha=0.05)+
      # geom_rect( inherit.aes=FALSE, aes(xmin="JOER 2021", xmax="JOER 2023", ymin=title_min,
      #                                   ymax=0), color="transparent", fill="#56B4E9", alpha=0.05)+
      # 
      # geom_rect( inherit.aes=FALSE, aes(xmin="CBRR 2021", xmax="CBRR 2023", ymin=title_min,
      #                                   ymax= 0), color="transparent", fill="#F9C74F", alpha=0.05)+
      # 
      # geom_rect( inherit.aes=FALSE, aes(xmin="PTRR 2021", xmax="PTRR 2023", ymin=title_min,
      #                                   ymax=0), color="transparent", fill="#F3722C", alpha=0.05)+
      # geom_rect( inherit.aes=FALSE, aes(xmin= 12.75, xmax= 15.25, ymin=title_min,
      #                                   ymax=0), color="transparent", fill="#d71a1d", alpha=0.05)+
      # geom_abline(aes(slope=0, intercept=0), color='#000000', linetype="solid") +
      # geom_text(aes(x= 2, y = (title_min/2), label= "Long Draw",  color = "#000000"), size = 6, alpha = 1)+
      # geom_text(aes(x= 5, y = (title_min/2), label= "Joe Wright",  color = "#000000"), size = 6, alpha = 1)+
      # geom_text(aes(x= 8, y = (title_min/2), label= "Chambers",  color = "#000000"), size = 6, alpha = 1)+
      # geom_text(aes(x= 11, y = (title_min/2), label= "Peterson",  color = "#000000"), size = 6)+
      # geom_text(aes(x= 14, y = (title_min/2), label= "Barnes Meadow",  color = "#000000"), size = 5, alpha = 1)+
      # 
      # 
      scale_fill_manual(name="Watershed", values = colorsBS_site_code) +
      theme_bw(base_size=25) +
      
      #theme(axis.text.x = ifelse(study_reservoir_chla$Location == "Reservoir",element_text(face = "bold"),element_text(face = "plain"))) +
      labs( x = "Site", y = param_units)+
      # scale_x_discrete(labels = c("2021", "2022", "2023","", "", "","2021", "2022", "2023","", "", "","2021", "2022", "2023", "", "", "", "2021", "2022", "2023"))+
      theme(legend.position = "none",
            axis.title = element_text(face = "bold"),
            # bold and angle x axis titles to 45
            axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"))
}
```

### NO3 DOC K NH4 Dotplots

```{r}
no3_dotplot_res<- 
  build_res_dotplot_concise(param = "NO3", binwdth = .06, sz = 1)+
  ylim(0,1.5)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black") +
  theme(axis.text.x = element_blank(), 
        axis.title.x = element_blank(), 
        axis.ticks.x = element_blank())

K_dotplot_res<- 
  build_res_dotplot_concise(param = "K", binwdth = .065, sz = 1)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black") +
  theme(axis.text.x = element_blank(), 
        axis.title.x = element_blank(), 
        axis.ticks.x  = element_blank())

DOC_dotplot_res <- 
  build_res_dotplot_concise(param = "DOC", binwdth = .25, sz = 2)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black")+
  theme(axis.text.x = element_blank(), 
        axis.title.x = element_blank(), 
        axis.ticks.x = element_blank())
NH4_dotplot_res <-
  build_res_dotplot(param = "NH4", binwdth = .01, sz = 1.5, title_min = -.025)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black")+
  ylab("NH4+ (mg/L)")+
  theme( axis.title.x = element_blank())+
  ylim(-.025,0.2)


 DOC_k_no3_nh4_dotplot <-  DOC_dotplot_res/K_dotplot_res/no3_dotplot_res/NH4_dotplot_res+
  plot_layout(guides = 'collect')
 
 DOC_k_no3_nh4_dotplot


 ggsave(filename = here(paper_draft_dir, "DOC_K_NO3_NH4_dotplot.png"), plot = DOC_k_no3_nh4_dotplot, width = 20, height = 16, units = "in", dpi = 300)

```

## Chl-a dotplot

```{r}
chla_dotplot_res <- build_res_dotplot(param = "ChlA", binwdth = .3, sz = 1)+
  # add medians
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black") +
#adding meso and eutrophic levels for cold lakes
 geom_abline(aes(slope=0, intercept=2.6), color='black', linetype="dashed") +
    geom_abline(aes(slope=0, intercept=8), color='black', linetype="dashed") +
    #annotate(geom="text", label= "MDL Limit (μg/L)", x="CBRR 2022", y=.075, vjust= -1)+
    annotate(geom="text", label= "Mesotrophic", x=1.15, y=2.6, vjust= -1, size = 4)+
    annotate(geom="text", label= "Eutrophic", x= 1.5, y=8, vjust= -1, size = 5)+
   #fixing y axis
    coord_cartesian(ylim=c(0,13))

plot(chla_dotplot_res)

save_plot_image(chla_dotplot_res, "res_chla_dotplots_2021_2024")


```

## Longitudinal Patterns

### JWC

```{r}
#Seasonal Color values
season_clp_color_vals <- c("Winter" = "#03045e",  "Rising Limb" = '#047E82',  "Peak Flow" = '#59B851', "Falling Limb"='#FA850E', "Fall Baseflow"  = "#772f1a")


jwc_longitudinal <- tidy_chem %>%
  dplyr::filter(!is.na(ChlA)) %>%
  dplyr::filter(site_code %in% c("JOEI", "JOER", "CBRR", "CBRI", "CHD", "JWC"))%>%
  select(Date,site_code, ChlA, NO3, Year,season, month,location_type)%>%
  mutate(location_type = ifelse(location_type == "Outflow"|location_type =="Inflow", "Stream", "Reservoir"))

jwc_longitudinal$site_code <- factor(jwc_longitudinal$site_code, levels = c("JOEI", "JOER", "CBRI", "CBRR", "CHD", "JWC"))
jwc_longitudinal$season <- factor(jwc_longitudinal$season, levels = c("Winter", "Rising Limb", "Peak Flow", "Falling Limb", "Fall Baseflow"))
jwc_longitudinal$month <- factor(jwc_longitudinal$month, levels = c("4", "5", "6", "7", "8", "9", "10","11"))
# Calculate the medians for ChlA
median_data <- jwc_longitudinal %>%
  group_by(site_code, Year, location_type) %>%
  summarise(median_ChlA = median(ChlA, na.rm = TRUE), 
            median_NO3 = median(NO3, na.rm = TRUE), 
            max_ChlA = max(ChlA, na.rm = T), 
            max_NO3 = max(NO3, na.rm = T)) %>%
  ungroup()%>%
  add_row(site_code = "JWC", Year = "2021", location_type = "Stream", median_ChlA = NA, median_NO3 = NA, max_ChlA = NA, max_NO3 = NA)
median_data$site_code <- factor(median_data$site_code, levels = c("JOEI", "JOER", "CBRI", "CBRR", "CHD", "JWC"))

# Create the plot
jwc_long_chla<-  ggplot() + 
  geom_point(data = jwc_longitudinal, aes(x = site_code, y = ChlA, color = season),
             size = 3 ,position = position_jitterdodge(jitter.height = 0))+
  geom_rect(data = median_data %>% filter(site_code %in% c("JOER", "CBRR")),
            aes(xmin = as.numeric(site_code) - 0.4, xmax = as.numeric(site_code) + 0.4, ymin = - Inf, ymax = Inf, group = Year),
            fill = "#59B851", alpha =0.2)+
  geom_rect(data = median_data %>% filter(site_code %in% c("JWC")),
            aes(xmin = as.numeric(site_code) - 0.4, xmax = as.numeric(site_code) + 0.4, ymin = - Inf, ymax = Inf, group = Year),
            fill = "#002FA7", alpha =0.2)+
  scale_color_manual(values = season_clp_color_vals)+
  # Add lines connecting the medians
  geom_line(data = median_data, aes(x = site_code, y = median_ChlA, group = Year), color = "black", linewidth = 1) +
  geom_point(data = median_data, aes(x = site_code, y = median_ChlA, group = Year, shape = "Yearly Site Median"), color = "black", size = 3) +
  scale_shape_manual(name = "", values = c("Yearly Site Median" = 16)) +  # 16 = solid circle
  guides(shape = guide_legend(override.aes = list(color = "black", size = 3))) +
  theme_bw(base_size = 20)+
  facet_wrap(~Year, ncol = 1, scale = "free_y")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right")+
  scale_x_discrete(labels = c( "JOEI"= "  Above\nJoe Wright", 
                               "JOER" = "", 
                               "CBRI" = "Above\nChambers",
                               "CBRR" = "", 
                               "CHD" = "Below\nChambers", 
                               "JWC" = "Joe Wright Creek"))+
  labs(y = expression(Chlorophyll ~italic("a") ~ (ug ~ L^{-1})), 
       color = "Season")



jwc_long_no3 <- 
  ggplot() + 
  geom_point(data = jwc_longitudinal, aes(x = site_code, y = NO3, color = season),size = 3 ,position = position_jitterdodge(jitter.height = 0))+
  geom_rect(data = median_data %>% filter(site_code %in% c("JOER", "CBRR")),
            aes(xmin = as.numeric(site_code) - 0.4, xmax = as.numeric(site_code) + 0.4, ymin = - Inf, ymax = Inf, group = Year),
            fill = "#59B851", alpha =0.2)+
  geom_rect(data = median_data %>% filter(site_code %in% c("JWC")),
            aes(xmin = as.numeric(site_code) - 0.4, xmax = as.numeric(site_code) + 0.4, ymin = - Inf, ymax = Inf, group = Year),
            fill = "#002FA7", alpha =0.2)+
  scale_color_manual(values = season_clp_color_vals)+
  #add a Hline for the median
  #stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black") +
  geom_line(data = median_data, aes(x = site_code, y = median_NO3, group = Year), color = "black", linewidth = 1) +
  geom_point(data = median_data, aes(x = site_code, y = median_NO3, group = Year, shape = "Yearly Site Median"), color = "black", size = 3) +
  scale_shape_manual(name = "", values = c("Yearly Site Median" = 16)) +  # 16 = solid circle
  guides(shape = guide_legend(override.aes = list(color = "black", size = 3))) +
  theme_bw(base_size = 20)+
  facet_wrap(~Year, ncol = 1, scale = "free_y")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        #remove facet title
        #strip.text.x = element_blank(),
        legend.position = "none") +
  scale_x_discrete(labels = c( "JOEI"= "  Above\nJoe Wright", 
                               "JOER" = "", 
                               "CBRI" = "Above\nChambers",
                               "CBRR" = "", 
                               "CHD" = "Below\nChambers", 
                               "JWC" = "Joe Wright Creek")) +
  labs(y = expression(NO[3]^{"-"} ~ (mg ~ L^{-1})))


jwc_long_chla_no3<-  jwc_long_chla+jwc_long_no3+
  plot_layout(guides = 'collect')+
  #add color legend to top
  plot_annotation( tag_levels = 'a')

plot(jwc_long_chla_no3)
save_plot_image(jwc_long_chla_no3 ,"jwc_long_chla_no3" )
```

### South Fork

```{r}

SF_Res = c("COMI", "COMR", "COMO", "HORI", "HORR", "HORO", "BEAV", "SFM", "PSF")
#filter and org
sf_chem<- filter(tidy_chem, site_code %in% SF_Res & !is.na(ChlA))%>%
  select(site_code,Date,ChlA,location_type, Buffer_Level, Watershed_Level, watershed, Year )%>%
  mutate(location_type = case_when(
    location_type == "Stream" ~ "Beaver Creek", 
    site_code == "PSF" ~ "Poudre Below\nSouth Fork",
    site_code == "SFM" ~ "South Fork",
    TRUE ~ location_type), 
    site_group = case_when(
      watershed == "Comanche Reservoir" ~ watershed, 
      watershed == "Hourglass Reservoir"~ watershed, 
      TRUE ~ "Stream"
    ), 
    filled_status = case_when(
      Year %in% c(2021, 2022) ~ "Reservoirs NOT Filled",
      Year  %in% c(2023, 2024) ~ "Reservoirs Filled"
      
    ))

sf_chem$location_type <- factor(sf_chem$location_type, levels = c("Inflow","Reservoir", "Outflow","Beaver Creek","South Fork", "Poudre Below\nSouth Fork"))
sf_chem$filled_status <- factor(sf_chem$filled_status, levels = c("Reservoirs NOT Filled", "Reservoirs Filled"))
#get nice units for Y axis
lab_w_units <-  filter(chem_units, simple %in% "ChlA")%>%
  pull(combined)

#plot
south_fork <-   ggplot(sf_chem, aes(x=location_type,y=ChlA,  fill = filled_status, shape = )) + 
  geom_dotplot( binaxis = "y", 
               method = "histodot",
               stackdir = "center", 
               position = position_dodge(width = 0.8),
               binwidth = .2, dotsize = 2, 
               stackratio = 1, width = .01)+
  geom_abline(aes(slope=0, intercept=2.6), color='black', linetype="dashed") +
  geom_abline(aes(slope=0, intercept=8), color='black', linetype="dashed") +
  annotate(geom="text", label= "Mesotrophic", x=.75, y=2.6, vjust= -1, size = 6)+
  annotate(geom="text", label= "Eutrophic", x= .75, y=8, vjust= -1, size = 6)+
  theme_bw(base_size = 30)+
  scale_fill_manual(name = "Reservoir Status", values = c("#D55E00","#0072B2"))+
  labs(color = "Reservoir Status",
       y = lab_w_units)+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_text(face = c('plain', 'plain', 'plain', 'bold', 'bold', 'plain')), 
        legend.position = "bottom")


south_fork

save_plot_image(south_fork ,"south_fork_chla_impact" )
```

### Mainstem

## CLP

```{r}


clp_long <- tidy_chem %>%
  dplyr::filter(!is.na(ChlA)& Year %in% c(2022, 2023, 2024)) %>%
  dplyr::filter(site_code %in% c("JOEI", "JOER", "CBRR", "CBRI", "CHD", "JWC", "PFAL", "SLEP", "PBR", "PSF", "PNF", "PBD"))%>%
  select(site_code,Date, watershed, distance_upstream_km, ChlA, NO3, PO4, K, location_type,Year, month, season )%>%
  mutate(location_type = case_when(
    location_type %in% c("Inflow", "Outflow") ~ "Res Stream", 
    location_type %in% c("Reservoir") ~ "Reservoir",
    TRUE ~ "Mainstem"),
    Year = as.character(Year),
    dist_pbd = as.character(distance_upstream_km),
    site_code  = case_when(site_code == "PFAL" ~ "SLEP", TRUE ~ site_code),
    distance_upstream_km = case_when(site_code == "SLEP" ~ 73.161, TRUE ~ distance_upstream_km),
    type = ifelse(location_type %in% c("Inflow","Outflow"),"Stream",
                  ifelse(location_type == "Reservoir", "Reservoir", "Mainstem")))



clp_long$site_code <- factor(clp_long$site_code, levels = c("JOEI", "JOER", "CBRR", "CBRI", "CHD", "JWC", "SLEP", "PBR", "PSF", "PNF", "PBD"))
clp_long$season <- factor(clp_long$season, levels = c("Winter", "Rising Limb", "Peak Flow", "Falling Limb", "Fall Baseflow"))
clp_long$month <- factor(clp_long$month, levels = c("5", "6", "7", "8", "9", "10","11"))


# Calculate the medians for ChlA
clp_median_data <- clp_long %>%
  group_by(site_code,type,distance_upstream_km, location_type) %>%
  summarise(median_ChlA = median(ChlA, na.rm = TRUE), 
            median_NO3 = median(NO3, na.rm = TRUE)) %>%
  ungroup()

clp_long_chla<- ggplot(clp_long, aes(x = distance_upstream_km, y = ChlA)) + 
  annotate("rect", xmin = 78, xmax = 99, ymin = -0.1, ymax = 9, 
           color = "#4A86C5", fill = "NA", linewidth = 1.5) +
  
  geom_jitter(aes(color = season, shape = type), size = 3, width = 0.4) +
  scale_color_manual(values = season_clp_color_vals) +
  
  # Lines and points for medians
  geom_line(data = clp_median_data, aes(x = distance_upstream_km, y = median_ChlA), 
            color = "black", linewidth = 2.5) +
  geom_point(data = clp_median_data, aes(x = distance_upstream_km, y = median_ChlA, shape = type), 
             color = "black", size = 5) +
  #Box title
  annotate("label", x = 78, y = 8, label = "Joe Wright & Chambers\nReservoirs (Prev Figure)", 
           hjust = 0, size = 6, label.size = 0.5) +
  #JWC label + arrow
  annotate("label", x = 80, y = 5.75, label = "JWC site", size = 5) +
  #Confluence with Big South
    annotate("segment", x = 73, xend = 77, y = 4, yend = 2, 
           arrow = arrow(length = unit(0.3, "cm")), color = "#4A86C5", linewidth = 1) +  
  annotate("label", x = 73, y = 4, label = "Confluence\nwith Big South", size = 5) +
  #confluence with SF
        annotate("segment", x = 40, xend = 40, y = 4, yend = 1, 
           arrow = arrow(length = unit(0.3, "cm")), color = "#4A86C5", linewidth = 1) +  
  annotate("label", x = 40, y = 4, label = "Confluence with\nSouth Fork", size = 5) +
        annotate("segment", x = 5, xend = 5, y = 4, yend = 1, 
           arrow = arrow(length = unit(0.3, "cm")), color = "#4A86C5", linewidth = 1) + 
  annotate("label", x = 8, y = 4, label = "Confluence with\nNorth Fork", size = 5) +

  # Styling and Scales
  theme_bw(base_size = 30) +
  scale_x_reverse() +
  ylim(-0.2, 9.2) +
  labs(x = "Distance from Canyon Mouth (km)", 
       y = "Chlorophyll a (µg/L)", 
       shape = "Location", 
       color = "Season") +
 theme(
    legend.position = c(0.98, 0.98),         
    legend.justification = c(1, 1), 
    
    legend.box = "horizontal",               
    legend.box.background = element_rect(color = "black", linewidth = 0.5), 
    legend.background = element_blank(),     
    legend.margin = margin(6, 6, 6, 6)       
  )

print(clp_long_chla)


save_plot_image(clp_long_chla, "clp_long_chla")

```

# Remote Sensing

## Check out Temp

```{r}
#Read in the data from B

#Temp: From NW-CLP-RS/CLP_focus_temp_only.Rmd

temp_rs <- read_parquet(here("data/derived/remote_sensing/landsat/CLP_temp_estimate_v2024-10-10.parquet"))%>%
  mutate(doy = yday(date),
         kurt_SurfaceTemp = as.numeric(kurt_SurfaceTemp),
         sd_SurfaceTemp = as.numeric(sd_SurfaceTemp),
         med_surface_C = as.numeric(med_SurfaceTemp_corr) - 273.15, # Convert from Kelvin to Celsius
         doy_check = if_else(doy >= 120 & doy <= 300, "PASS", "FAIL"), # Filter for sampling season (roughly April to October)
  kurtosis_check = if_else(kurt_SurfaceTemp > 10, "FAIL", "PASS"), #filter out for instances with high kurtosis
  sd_check = if_else(sd_SurfaceTemp > 2, "FAIL", "PASS"), #filter out instances with high SD
pass_all = (sd_check == "PASS" & kurtosis_check == "PASS" & doy_check == "PASS"),
site_code = case_when(
    gnis_name == "Seaman Reservoir"     ~ "SEMR",    
    gnis_name =="Joe Wright Reservoir"    ~ "JOER", 
    gnis_name == "Barnes Meadow Reservoir"  ~ "BRNR",
    gnis_name == "Chambers Lake" ~ "CBRR",
    gnis_name == "Peterson Lake"     ~ "PTRR", 
    gnis_name ==     "Long Draw Reservoir"    ~ "LNGR",
gnis_name =="Hourglass Reservoir"   ~ "HORR",
gnis_name ==  "Comanche Reservoir"     ~ "COMR",
gnis_name == "Halligan Reservoir"  ~ "HALL",
    TRUE ~ NA_character_))%>%
  filter(pass_all == TRUE & med_surface_C > 0)

ggplot(temp_rs, aes(x = date, y = med_surface_C, color = doy, shape = pass_all))+
  geom_point()+
  geom_vline(xintercept = as.Date(c("2020-08-01")), linetype = "dashed", color = "red") +
  facet_wrap(~gnis_name,
             scales = "free_y",
             ncol = 1) +
  theme_bw() +
  labs(x = "Date", y = "Estimated Temperature (°C)", shape = "PASS QA")

#sd_SufraceTemp, min_SurfaceTemp, kurt_SurfaceTemp
#check out distributions for each of these
ggplot(temp_rs, aes(x = yday(date), y = as.numeric(kurt_SurfaceTemp)))+
  geom_point()+
  #geom_vline(xintercept = as.Date(c("2020-08-01")), linetype = "dashed", color = "red") +
  facet_wrap(~gnis_name) +
  theme_bw() +
  labs(x = "Date", y = "Kurtosis")

ggplot(temp_rs, aes(x = date, y = as.numeric(min_SurfaceTemp)-273.15 ))+
  geom_point()+
  geom_vline(xintercept = as.Date(c("2020-08-01")), linetype = "dashed", color = "red") +
  facet_wrap(~gnis_name) +
  theme_bw() +
  labs(x = "Date", y = "Min Surface Temperature (°C)")


ggplot(temp_rs, aes(x = date, y = as.numeric(sd_SurfaceTemp)))+
  geom_point()+
  geom_vline(xintercept = as.Date(c("2020-08-01")), linetype = "dashed", color = "red") +
  facet_wrap(~gnis_name) +
  theme_bw() +
  labs(x = "Date", y = "SD Surface Temperature (°K)")

ggplot(temp_rs, aes(x = as.numeric(sd_SurfaceTemp), color = as_factor(year(date))))+
  geom_histogram()+
  facet_wrap(~gnis_name) +
  theme_bw() +
  labs(x= "SD Surface Temperature (°K)")


ggplot(temp_rs, aes(x = as.numeric(min_SurfaceTemp) - 273.15, y = as.numeric(med_SurfaceTemp_corr)-273.15))+
  geom_point()+
  geom_abline(aes(slope=1, intercept=0), color = "black", linetype = "dashed")+
  #geom_vline(xintercept = as.Date(c("2020-08-01")), linetype = "dashed", color = "red") +
  facet_wrap(~gnis_name) +
  theme_bw() +
  labs(x = "Min Surface Temp (°C)", y = "Med Surface Temperature (°C)")

```


### Seasonal Trend Plots

```{r}
formatData = function(lake){
  lakeData <- temp_rs %>% 
    filter(site_code == lake)
  as_tibble(lakeData)
}

calcSeasonAdjTrends = function(lakeTibble){
  lakeTibble %>% 
    time_decompose(med_surface_C, merge = T) 
}

graphTrendObs = function(seasonAdjTrends) {
  lakeName = seasonAdjTrends$gnis_name[1]
  ggplot(seasonAdjTrends, aes(x = date, y = observed)) +
    geom_point(color = 'gray') +
    geom_vline(xintercept = as.Date(c("2020-08-01")), linetype = "dashed", color = "red") +
    labs(x = NULL, 
         y = NULL,
         title = paste0(lakeName)) +
    geom_point(aes(y = trend), color = '#E69F00') +
    geom_smooth(aes(y = trend), lty = 2) +
    theme_bw() +
    theme(plot.title = element_text(face = 'bold', hjust = 0.5, size = 12))
}

burned_plots <- map(c("BRNR", "PTRR", "CBRR"),~
formatData(.x)%>%
 calcSeasonAdjTrends(.)%>%
graphTrendObs(.))

wrap_plots(burned_plots, ncol = 1)

unburned_plots <- map(c("JOER", "LNGR"),~
formatData(.x)%>%
 calcSeasonAdjTrends(.)%>%
graphTrendObs(.))

wrap_plots(unburned_plots, ncol = 1)

```


## Join Temp with Grabs

```{r}

temp_rs_joined <- temp_rs %>%
  select(gnis_name,rs_date = date, med_surface_C) %>%
  mutate(site_code = case_when(
    gnis_name == "Seaman Reservoir"     ~ "SEMR",    
    gnis_name =="Joe Wright Reservoir"    ~ "JOER", 
    gnis_name == "Barnes Meadow Reservoir"  ~ "BRNR",
    gnis_name == "Chambers Lake" ~ "CBRR",
    gnis_name == "Peterson Lake"     ~ "PTRR", 
    gnis_name ==     "Long Draw Reservoir"    ~ "LNGR",
gnis_name =="Hourglass Reservoir"   ~ "HORR",
gnis_name ==  "Comanche Reservoir"     ~ "COMR",
gnis_name == "Halligan Reservoir"  ~ "HALL",
    TRUE ~ NA_character_),
date_low = rs_date - days(4),
    date_high = rs_date + days(4)) %>%
  left_join(tidy_chem, 
            by = join_by(site_code, between(y$Date, x$date_low, x$date_high)))%>%
  filter(!is.na(ChlA)& !is.na(med_surface_C))%>%
  mutate(date_diff = abs(as.numeric(difftime(Date, rs_date, units = "days"))))%>%
  filter(!(site_code %in% c("COMR", "HORR") & Year %in% c(2021,2022))) # removing unfilled comanche and hourglass matchups

ggplot(temp_rs_joined, aes(x = med_surface_C, y = ChlA,shape = Year))+
  geom_point()+
  facet_wrap(~site_code)+
  theme_bw()+
  labs(x = "Estimated Surface Temperature (°C)", y = "Chlorophyll a (µg/L)")

```

Not great associations and there is only ~100 matchups between the seven reservoirs

## Check out SDD

```{r}

sdd_rs <- read_parquet("data/derived/remote_sensing/landsat/CLP_SDD_estimate_v2024-10-10.parquet")%>%
  mutate(doy = yday(date),
         site_code = case_when(
           gnis_name == "Seaman Reservoir"     ~ "SEMR",    
           gnis_name =="Joe Wright Reservoir"    ~ "JOER", 
           gnis_name == "Barnes Meadow Reservoir"  ~ "BRNR",
           gnis_name == "Chambers Lake" ~ "CBRR",
           gnis_name == "Peterson Lake"     ~ "PTRR", 
           gnis_name ==     "Long Draw Reservoir"    ~ "LNGR",
           gnis_name =="Hourglass Reservoir"   ~ "HORR",
           gnis_name ==  "Comanche Reservoir"     ~ "COMR",
           gnis_name == "Halligan Reservoir"  ~ "HALL",
           TRUE ~ NA_character_))%>%
  filter(doy >= 130 & doy <= 310)# Filter for sampling season (roughly April to October)

ggplot(sdd_rs, aes(x = date, y = mean_sdd))+
  geom_point()+
  geom_vline(xintercept = as.Date(c("2020-08-01")), linetype = "dashed", color = "red") +
  facet_wrap(~gnis_name, ncol = 1)+
  theme_bw()+
  labs(x = "Date", y = "Estimated SDD (m)")
```

### Seasonal Trend Plots

```{r}
formatData = function(lake){
  lakeData <- sdd_rs %>% 
    filter(site_code == lake)
  as_tibble(lakeData)
}

calcSeasonAdjTrends = function(lakeTibble){
  lakeTibble %>% 
    time_decompose(mean_sdd, merge = T) 
}

graphTrendObs = function(seasonAdjTrends) {
  lakeName = seasonAdjTrends$gnis_name[1]
  ggplot(seasonAdjTrends, aes(x = date, y = observed)) +
    geom_point(color = 'gray') +
    geom_vline(xintercept = as.Date(c("2020-08-01")), linetype = "dashed", color = "red") +
    labs(x = NULL, 
         y = "Estimated Mean Secchi Disk Depth (m)",
         title = paste0(lakeName)) +
    geom_point(aes(y = trend), color = '#E69F00') +
    geom_smooth(aes(y = trend), lty = 2) +
    theme_bw() +
    theme(plot.title = element_text(face = 'bold', hjust = 0.5, size = 12))
}

map(c("BRNR", "PTRR", "CBRR","JOER", "LNGR"),~
formatData(.x)%>%
 calcSeasonAdjTrends(.)%>%
graphTrendObs(.))%>%
wrap_plots(., ncol = 1, axes = "collect")


```

## Join with chem
```{r}
sdd_rs_joined <- sdd_rs %>%
  select(gnis_name,rs_date = date, mean_sdd, pred1, pred2, pred3, pred4) %>%
  mutate(site_code = case_when(
    gnis_name == "Seaman Reservoir"     ~ "SEMR",    
    gnis_name =="Joe Wright Reservoir"    ~ "JOER", 
    gnis_name == "Barnes Meadow Reservoir"  ~ "BRNR",
    gnis_name == "Chambers Lake" ~ "CBRR",
    gnis_name == "Peterson Lake"     ~ "PTRR", 
    gnis_name ==     "Long Draw Reservoir"    ~ "LNGR",
gnis_name =="Hourglass Reservoir"   ~ "HORR",
gnis_name ==  "Comanche Reservoir"     ~ "COMR",
gnis_name == "Halligan Reservoir"  ~ "HALL",
    TRUE ~ NA_character_),
date_low = rs_date - days(4),
    date_high = rs_date + days(4)) %>%
  left_join(tidy_chem, 
            by = join_by(site_code, between(y$Date, x$date_low, x$date_high)))%>%
  filter(!is.na(ChlA) & !is.na(mean_sdd))%>%
  mutate(date_diff = abs(as.numeric(difftime(Date, rs_date, units = "days"))))

ggplot(sdd_rs_joined, aes(x = mean_sdd, y = ChlA, shape = Year))+
  geom_point()+
  facet_wrap(~site_code)+
  theme_bw()+
  labs(x = "Estimated Secchi Disk Depth (M)", y = "Chlorophyll a (µg/L)")

ggplot(sdd_rs_joined%>%filter(TSS <200), aes(x = mean_sdd, y = TSS, color = site_code, shape = Year))+
  geom_point(size = 3)+
  scale_y_log10()+
  #facet_wrap(~site_code)+
  theme_bw()+
  labs(x = "Estimated Secchi Disk Depth (M)", y = "TSS (mg/L)")
```




# Stats Section

### Sample Counts

```{r}
tidy_chem%>%
  mutate(loc_type = case_when(
    location_type %in% c("Inflow", "Reservoir", "Outflow") ~ "Reservoir Network", 
    site_code == "BEAV" & !is.na(ChlA) ~ "Reservoir Network",
    location_type == "Mainstem" ~ "Mainstem",
    TRUE ~ "Stream"))%>%
  group_by(loc_type)%>%
  summarise(sample_count = n(), 
            chla_samp = sum(!is.na(ChlA)))%>%
  #add a row of total minus stream
  add_row(loc_type = "Total minus Stream", sample_count = sum(.$sample_count) - .$sample_count[.$loc_type == "Stream"], chla_samp = sum(.$chla_samp) - .$chla_samp[.$loc_type == "Stream"])


tidy_chem%>%
  group_by(location_type)%>%
  summarise(sample_count = n(), 
            chla_samp = sum(!is.na(ChlA)))
```

# Evaluating differences

## Reservoirs only

### Check to see if they are normally distributed (Shapiro-Wilk test)

```{r}
chem_transformed_res <- tidy_chem%>%
    filter(site_code %in% mainstem_Res_only) %>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4), 
               names_to = "param", 
               values_to = "value")%>%
  select(site_code, Date, Year, param, value, burn_group, burn_year)%>%
  mutate(year_site = paste0(Year, "_", site_code))%>%
  mutate(log_value = log10(value + 1))%>%
  na.omit()

ggplot(chem_transformed_res, aes(x = value, color = year_site))+
  geom_histogram()+
  facet_wrap(~param, scales = "free")
 
chem_transformed_res %>%
  group_by(param) %>%
  shapiro_test(value)%>%
  filter(p > 0.05)

chem_transformed_res %>%
  group_by(param) %>%
  shapiro_test(log_value)%>%
  arrange(p)
#
```
Nothing without transformation is normal and only Ca is normal post log transform. We will need to use Welch's Anova test to compare between burned and unburned reservoirs and we will use the log transformed data for that test since it is closer to normal.

## Do parameters vary full period between burned and unburned reservoirs?

```{r}
#Welch ANOVA test which does not assume equal variances (which is typically not the case)
kruskal_res_results <- chem_transformed_res %>%
  mutate(burn_group = factor(burn_group, levels = c("unburned", "burned"))) %>%
  group_by(param) %>%
  group_modify(~{
      res <- .x %>% kruskal_test(value ~ burn_group)
      res$test_used <- "kruskal_test"
      return(as_tibble(res))
  })%>%
  mutate(p.sig = ifelse(p < 0.05,"*" , ""))
kruskal_res_results

# Post-Hoc: Games-Howell
# Games-Howell works for both equal and unequal variances.
post_hoc_results <- chem_transformed_res %>%
  #only do post-hoc for significant results from final results
  filter(param %in% welch_res_results$param[welch_res_results$p.sig == "*"]) %>%
  mutate(burn_group = factor(burn_group, levels = c("unburned", "burned"))) %>%
  group_by(param) %>%
  games_howell_test(log_value ~ burn_group)%>%
  mutate(
      pct_change = (10^estimate - 1) * 100,
      pct_conf_low = (10^conf.low - 1) * 100,
      pct_conf_high = (10^conf.high - 1) * 100
    ) %>%
    # Rounding for readability
    mutate(across(c(pct_change, pct_conf_low, pct_conf_high), ~round(., 2)))

post_hoc_results
```

## Do parameters vary between burned and unburned reservoirs (year by year)?

```{r}
#Welch ANOVA test which does not assume equal variances (which is typically not the case)
welch_res_results <- chem_transformed_res %>%
  group_by(param) %>%
  group_modify(~{
      res <- .x %>% welch_anova_test(log_value ~ burn_year)
      res$test_used <- "Welch ANOVA"
      return(as_tibble(res))
  })%>%
  mutate(p.sig = ifelse(p < 0.05,"*" , ""))%>%
  select(param, DFd,p, test_used, p.sig)
welch_res_results

# Post-Hoc: Games-Howell
# Games-Howell works for both equal and unequal variances.
post_hoc_results <- chem_transformed_res %>%
  #only do post-hoc for significant results from final results
  filter(param %in% welch_res_results$param[welch_res_results$p.sig == "*"]) %>%
  group_by(param) %>%
  games_howell_test(log_value ~ burn_year)%>%
  mutate(
      pct_change = (10^estimate - 1) * 100,
      pct_conf_low = (10^conf.low - 1) * 100,
      pct_conf_high = (10^conf.high - 1) * 100
    ) %>%
    # Rounding for readability
    mutate(across(c(pct_change, pct_conf_low, pct_conf_high), ~round(., 2)))

check <- post_hoc_results%>%filter(param == "NO3")
```

## Do parameters vary between reservoirs across the study period ?

```{r}
#Welch ANOVA test which does not assume equal variances (which is typically not the case)
welch_res_results <- chem_transformed_res %>%
  group_by(param) %>%
  group_modify(~{
      res <- .x %>% welch_anova_test(log_value ~ site_code)
      res$test_used <- "Welch ANOVA"
      return(as_tibble(res))
  })%>%
  mutate(p.sig = ifelse(p < 0.05,"*" , ""))%>%
  select(param, DFd,p, test_used, p.sig)
welch_res_results

# Post-Hoc: Games-Howell
# Games-Howell works for both equal and unequal variances.
post_hoc_results <- chem_transformed_res %>%
  #only do post-hoc for significant results from final results
  filter(param %in% welch_res_results$param[welch_res_results$p.sig == "*"]) %>%
  group_by(param) %>%
  games_howell_test(log_value ~ site_code)%>%
  mutate(
      pct_change = (10^estimate - 1) * 100,
      pct_conf_low = (10^conf.low - 1) * 100,
      pct_conf_high = (10^conf.high - 1) * 100
    ) %>%
    # Rounding for readability
    mutate(across(c(pct_change, pct_conf_low, pct_conf_high), ~round(., 2)))

post_hoc_results%>%filter(param %in% c( "NO3"))%>%arrange(param)%>%print(n = Inf)
```

## Do parameters vary between years across the study period ?

```{r}
#Welch ANOVA test which does not assume equal variances (which is typically not the case)
welch_res_results <- chem_transformed_res %>%
  mutate(Year = factor(Year, levels = c("2021", "2022", "2023", "2024")))%>%
  group_by(param) %>%
  group_modify(~{
      res <- .x %>% welch_anova_test(log_value ~ Year)
      res$test_used <- "Welch ANOVA"
      return(as_tibble(res))
  })%>%
  mutate(p.sig = ifelse(p < 0.05,"*" , ""))%>%
  select(param, DFd,p, test_used, p.sig)
welch_res_results

# Post-Hoc: Games-Howell
# Games-Howell works for both equal and unequal variances.
post_hoc_results <- chem_transformed_res %>%
    mutate(Year = factor(Year, levels = c("2021", "2022", "2023", "2024")))%>%
  #only do post-hoc for significant results from final results
  filter(param %in% welch_res_results$param[welch_res_results$p.sig == "*"]) %>%
  group_by(param) %>%
  games_howell_test(log_value ~ Year)%>%
  mutate(
      pct_change = (10^estimate - 1) * 100,
      pct_conf_low = (10^conf.low - 1) * 100,
      pct_conf_high = (10^conf.high - 1) * 100
    ) %>%
    # Rounding for readability
    mutate(across(c(pct_change, pct_conf_low, pct_conf_high), ~round(., 2)))

post_hoc_results%>%filter(param %in% c("ChlA", "K", "DOC", "NO3", "NH4", "SO4", "TSS"))%>%arrange(param)%>%print(n = Inf)%>%
filter(param == "DOC")
```


```{r}
# do anova for each parameter (doc, k, no3, nh4, so4 and TSS for burn group)
# This compares the mean of each parameter between burned and unburned sites

sitewise_anova <- chem_to_check %>%
  #filter(param %in% c("DOC","DTN",  "K", "NO3", "NH4", "SO4", "TSS", "SC")) %>% # Filter for NO3 parameter
  na.omit() %>% # Remove missing values
  group_by(param) %>% # Group by site_code to test differences within each site
  anova_test(
    value ~ site_code)
sitewise_anova


# perform TUkeyHSD on group anova
tukey_results <- chem_to_check %>%
  #filter(param %in% c("DOC", "DTN", "K", "NH4", "SO4", "SC")) %>% # Pull significant parameters
  group_by(param) %>%
  tukey_hsd(value ~ site_code)
tukey_results


# repeat measures anova (wilcox since data is not normally distributed)
# This compares within site parameters between years to see if there is a meaningful difference between years

pairwise_results <- chem_to_check %>%
  #filter(param == parameter) %>% # Filter for NO3 parameter
  na.omit() %>% # Remove missing values
  group_by(site_code, param) %>% # Group by site_code to test differences within each site
  pairwise_wilcox_test(
    value ~ Year, 
    p.adjust.method = "bonferroni"
  ) %>%
  mutate(comparison = paste(group1, group2, sep = "-"), 
         p.adj = ifelse(p.adj < 0.05, p.adj,NA))%>% # Create a comparison column
  select(site_code,param,  comparison, p.adj)%>%# Keep only relevant columns
  pivot_wider( id_cols = c(site_code, param),
    names_from = comparison, 
    values_from = p.adj,
    names_sort = TRUE )%>%
  select(site_code, param, `2021-2022`,`2022-2023`,`2021-2023`)%>%
  filter(param %in% c("ChlA","Turbidity", "TSS", "DOC", "NO3", "NH4", "SO4"))

pairwise_results

pairwise_results_nosite <- chem_to_check %>%
  na.omit() %>% # Remove missing values
  group_by(param) %>% # Group by site_code to test differences within each site
  pairwise_wilcox_test(
    value ~ Year, 
    p.adjust.method = "bonferroni"
  ) %>%
  mutate(comparison = paste(group1, group2, sep = "-"), 
         p.adj = ifelse(p.adj < 0.05, p.adj,NA))%>% # Create a comparison column
  select(param,  comparison, p.adj)%>%# Keep only relevant columns
  pivot_wider( id_cols = c( param),
    names_from = comparison, 
    values_from = p.adj,
    names_sort = TRUE )%>%
  select(param, `2021-2022`,`2022-2023`,`2021-2023`)%>%
  filter(param %in% c("ChlA","Turbidity", "TSS", "DOC", "NO3", "NH4", "SO4"))

pairwise_results_nosite

```


## Reporting Calculations
These data and calculations will be included in the funder report to give an overview of how the system is changing year to year based on our study

### Chlorophyll a eutrophic status

In cold water lakes (our study sites), mesotrophic levels are defined as between 2.6 and 8 µg/L, and eutrophic levels are greater than 8 µg/L. Under 2.6 µg/L is considered Oligotrophic. 

```{r}

chla_status_site_year <- most_recent_chem%>%
  filter(!is.na(ChlA))%>%
  filter(site_code %in% mainstem_Res_only)%>%
  mutate(eu_status  = case_when(ChlA >= 2.6 &ChlA <= 8 ~ "Mesotrophic", 
                                ChlA > 8 ~ "Eutrophic",
                                ChlA < 2.6 ~ "Oligotrophic"))%>%
  group_by(site_code, Year, eu_status, burn_group)%>%
  summarise(eu_stats = n())%>%
  ungroup()%>%
  pivot_wider(names_from = eu_status, values_from = eu_stats)%>%
  mutate_at(vars(-Year), ~replace(., is.na(.), 0))%>%
  rowwise()%>%
  mutate(total_samples = sum(c(Eutrophic, Mesotrophic, Oligotrophic)),
    perc_eu = round((Eutrophic/total_samples)*100, digits= 1), 
         perc_meso = round((Mesotrophic/total_samples)*100, digits= 1),
         perc_meso_eu = round(((Eutrophic+Mesotrophic)/total_samples)*100, digits= 1), 
        perc_oligo = round((Oligotrophic/total_samples) *100, digits= 1),
    Year = as.character(Year))

all_burn <- chla_status_site_year%>%
  group_by(burn_group, Year)%>%
  summarise(total_samples = sum(total_samples), 
            perc_eu = sum(Eutrophic)/sum(total_samples)*100, 
            perc_meso = sum(Mesotrophic)/sum(total_samples)*100, 
            perc_meso_eu = sum(Eutrophic + Mesotrophic)/sum(total_samples)*100, 
            perc_oligo = sum(Oligotrophic)/sum(total_samples)*100)

all_years <- chla_status_site_year%>%
  group_by(burn_group)%>%
  summarise(total_samples = sum(total_samples), 
            perc_eu = sum(Eutrophic)/sum(total_samples)*100, 
            perc_meso = sum(Mesotrophic)/sum(total_samples)*100, 
            perc_meso_eu = sum(Eutrophic + Mesotrophic)/sum(total_samples)*100, 
            perc_oligo = sum(Oligotrophic)/sum(total_samples)*100)

long_chla_status <- chla_status_site_year%>%
  select(site_code, Year, total_samples, perc_eu, perc_meso, perc_meso_eu, perc_oligo)%>%
  pivot_longer(cols = c(perc_eu, perc_meso, perc_meso_eu, perc_oligo), names_to = "category", values_to = "value" )

long_chla_status$site_code = factor(long_chla_status$site_code, levels = c("LNGR","JOER" , "CBRR", "PTRR", "BRNR" ))

chla_status <- c(
  perc_eu = "% eutrophic",
  perc_meso = "% mesotrophic ",
  perc_meso_eu = "% mesotrophic & eutrophic",
  perc_oligo = "% oligotrophic"
)

#PLOT
 long_chla_status%>%
  filter(site_code %in% mainstem_Res_only )%>%
   filter(category != "perc_meso_eu")%>%
  ggplot(aes(x = Year, y = value,group = site_code, color =  site_code))+
  scale_color_manual(values = colorsBS_site_code)+
 geom_line()+
   geom_point()+
  theme_bw(base_size = 15)+
facet_wrap(~category, ncol = 1, labeller = labeller(category = chla_status))+
  labs(
    x = "Year",
    y = "Percent of chlorophyll-*a* samples", 
    color = "Reservoir")+
  theme(axis.title.y = ggtext::element_markdown())


```

## Chl-a Stats

```{r}
# grouped by site code
chla_status_site_year <- most_recent_chem%>%
  filter(!is.na(ChlA))%>%
  filter(site_code %in% mainstem_Res_only)%>%
  group_by(site_code, Year)%>%
  summarise(mean_chla = mean(ChlA), 
            median_chla = median(ChlA), 
            min_chla = min(ChlA), 
            max_chla = max(ChlA), 
            sd_chla = sd(ChlA))


# Grouped by burn status

chla_status_site_year <- most_recent_chem%>%
  filter(!is.na(ChlA))%>%
  filter(site_code %in% mainstem_Res_only)%>%
  group_by(site_code, Year, burn_group)%>%
  summarise(mean_chla = mean(ChlA), 
            median_chla = median(ChlA), 
            min_chla = min(ChlA), 
            max_chla = max(ChlA), 
            sd_chla = sd(ChlA))

#create flextable for these stats
chla_status_site_year <- chla_status_site_year%>%
  ungroup()%>%
  mutate(burn_group = factor(burn_group, levels = c("Pre", "Post")))%>%
  mutate(Year = as.character(Year))%>%
  select(-site_code)%>%
  pivot_longer(cols = c(mean_chla, median_chla, min_chla, max_chla, sd_chla), names_to = "statistic", values_to = "value")%>%
  mutate(statistic = factor(statistic, levels = c("mean_chla", "median_chla", "min_chla", "max_chla", "sd_chla")))
```

### Comparing betwen burn groups and between primary reservoir sites


```{r}
# Prep data
chla_anova_df <- most_recent_chem%>%
  filter(!is.na(ChlA) )%>% #remove NAs
  filter(site_code %in% mainstem_Res_only) # only grab primary reservoirs

#perform anova test between groups
group_anova <- aov(ChlA ~ burn_group, data = chla_anova_df)
summary(group_anova)
# perform Tukey test to get magnitude of differences for chl-a between groups
TukeyHSD(group_anova)%>%tidy()%>%
  select(-term)

#perform anova test between sites
site_anova <- aov(ChlA ~ site_code, data = chla_anova_df)

summary(site_anova)

# perform Tukey test to get magnitude of differences for chl-a between groups
TukeyHSD(site_anova)%>%tidy()%>%
  #filter(adj.p.value < 0.05)%>%
  select(-term, -null.value)%>%
  mutate(sig = ifelse(adj.p.value < 0.05, "yes", "no"))%>%
  arrange(adj.p.value)
```


### Comparing years or between sites and years

Might be too many comparisons to reasonably test

```{r}
#perform anova test between groups and year
group_year_anova <- aov(ChlA ~ burn_year, data = chla_anova_df)

summary(group_year_anova)
tukey_group_year <- TukeyHSD(group_year_anova)%>%tidy()%>%
  filter(adj.p.value < 0.05)%>%
  select(-term)
tukey_group_year

#perform anova test between sites and year
site_year_anova <- aov(ChlA ~ site_year, data = chla_anova_df)

summary(site_year_anova)
TukeyHSD(site_year_anova)%>%tidy()%>%
  filter(adj.p.value < 0.05)%>%
  select(-term)
```


```{r}
compare_site_years <- function(target_site) {
  # Filter data for the selected site
  site_data <- chla_anova_df %>%
    filter(site_code == target_site)
  
  # Perform ANOVA
  site_year_anova <- aov(ChlA ~ site_year, data = site_data)
  
  #Perform Tukey HSD and tidy the results
  tukey_results <- TukeyHSD(site_year_anova) %>%
    tidy() %>%
    #filter(adj.p.value < 0.05) %>%
    select(-term, -null.value)  # Drop the 'term' column
  
  return(tukey_results)
}

#summary(compare_site_years("CBRR"))
results <- map_df(mainstem_Res_only, compare_site_years)%>%
  bind_rows()%>%
  filter(adj.p.value < 0.05)
results

```

-	Stats (significant):
-		Between unburned- burned all years: 
	mean =  - 1.22,  conf:( -0.77 -  -1.67)
o	Between unburned- burned : 
	No difference in 2021
	2022: mean =  - 1.07,  conf:( -0.18 -  -1.95)
	2023: mean = -2.29, conf: (-0.68  - -3.9)
o	Within site between years:
	JOER 2022- 2021: mean = 1.17, conf: (0.32 – 2.02)
	BRNR 2023 – 2022: mean = 2.81, conf: (1.32,4.3)
	Otherwise no significant increases at any sites
o	Between sites:
	Full anova: pvalue = 1.52e-08
	JOER – BRNR: mean = -1.53
	LNGR – BRNR: mean = -2.19
	PTRR – BRNR: mean = -1.21
	LNGR – CBRR: mean = -1.44





## Levene's test


```{r}
library(car)


#perform levene's test between groups
levene_group <- leveneTest(ChlA ~ burn_group, data = chla_anova_df)
levene_group

#perform levene's test between groups and year
levene_group_year <- leveneTest(ChlA ~ burn_year, data = chla_anova_df)
levene_group_year

#perform levene's test between groups and year


levene_group_year2223 <- leveneTest(ChlA ~ burn_year, data = chla_anova_df%>%filter(Year >= 2022))
levene_group_year2223




# perform levene's test between sites
levene_site <- leveneTest(ChlA ~ site_code, data = chla_anova_df)
levene_site



```



# When does nitrate peak?

Nearly all sites seem to peak with snowmelt, PTRR had peaks in August of 2021/2022 (PTRR's storage timing changed in 2023 due to Comanche + Hourglass being returned to service). LNGR had a peak in August of 2023.

```{r}
no3_stats <- chem_to_check%>%
  filter(param == "NO3")%>%
  group_by(site_code, Year, burn_group)%>%
  summarise(mean_no3 = mean(value, na.rm = T), 
            max_no3 = max(value, na.rm = T),
            min_no3 = min(value, na.rm = T),
            sd_no3 = sd(value, na.rm = T),
            max_no3_season = season[which.max(value)], 
            max_no3_date = Date[which.max(value)])

```

# PO4 detections

```{r}
# How many samples per study reservoirs

tidy_chem%>%
  filter( site_code %in% mainstem_Res_only)%>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4, PO4), 
               names_to = "param", 
               values_to = "value")%>%
  filter(param == "PO4")%>%
  group_by(site_code, burn_group)%>%
  summarise(samples = n(), 
           all_samples = sum(value),
            abv_nd  = sum(value > 0))%>%
  mutate(perc_abv_nd = abv_nd/samples*100)

# How many samples per year at study reservoirs

tidy_chem%>%
  filter( site_code %in% mainstem_Res_only)%>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4, PO4), 
               names_to = "param", 
               values_to = "value")%>%
  filter(param == "PO4")%>%
  group_by(Year)%>%
  summarise(samples = n(), 
           all_samples = sum(value),
            abv_nd  = sum(value > 0))%>%
  mutate(perc_abv_nd = abv_nd/samples*100)


# How many samples at mainstem/stream sites

tidy_chem%>%
  filter( site_code %nin% all_res_system)%>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4, PO4), 
               names_to = "param", 
               values_to = "value")%>%
  filter(param == "PO4")%>%
  group_by(Year)%>%
  summarise(samples = n(), 
           all_samples = sum(value),
            abv_nd  = sum(value > 0))%>%
  mutate(perc_abv_nd = abv_nd/samples*100)

```



# Anova of LNGR NO3 vs all other sites

```{r}
lngr_df <- most_recent_chem %>%
  filter(site_code %in% mainstem_Res_only)%>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4, PO4), 
               names_to = "param", 
               values_to = "value")%>%
  filter(param == "NO3")%>%
  mutate(site = case_when(site_code == "LNGR" ~ "LNGR", 
                          TRUE ~ "Other"))
  #group_by(site)%>%
  lngr_anova <- aov(value ~ site, data = lngr_anova)
  lngr_anova
TukeyHSD(lngr_anova)%>%tidy()%>%
  filter(adj.p.value < 0.05)%>%
  select(-term)
  
  
```

# Longitudinal chla/no3

## Joe Wright Creek

```{r}
joe_wright <- most_recent_chem%>%
  filter(site_code %in% c("JOEI", "JOER", "CBRI", "CBRR", "CHD", "JWC"))%>%
  pivot_longer(cols = c(ChlA:NO3), 
               names_to = "param", 
               values_to = "value")

network_data <- joe_wright %>%
  filter(param == "ChlA")

  network_anova <- aov(value ~ site_code, data = network_data)
 summary(network_anova)
 
TukeyHSD(network_anova)%>%tidy()%>%
  #filter(adj.p.value < 0.05)%>%
  select(-term)

```

## JOER

```{r}
joer_data <- joe_wright %>%
  filter(param == "ChlA" & site_code %in% c("JOEI", "JOER", "CBRI"))
  #group_by(site)%>%
  joer_anova <- aov(value ~ site_code, data = joer_data)
 summary(joer_anova)
 
TukeyHSD(joer_anova)%>%tidy()%>%
  #filter(adj.p.value < 0.05)%>%
  select(-term)
  


```

## CBRR

```{r}
cbrr_data <- joe_wright %>%
  filter(param == "ChlA" & site_code %in% c("CBRI", "CBRR", "CHD", "JWC"))
  #group_by(site)%>%
  cbrr_anova <- aov(value ~ site_code, data = cbrr_data)
 summary(cbrr_anova)
 
TukeyHSD(cbrr_anova)%>%tidy()%>%
  #filter(adj.p.value < 0.05)%>%
  select(-term)
  


```

## Test variability and means between Inflows and JWC

```{r}

inflow_jwc <- joe_wright %>%
  filter(param == "ChlA" & site_code %in% c("JOEI", "JWC", "CBRI"))

levene_jwc_inflows <- leveneTest(value ~ location_type, data = inflow_jwc)
levene_jwc_inflows

 jwc_inflow_anova <- aov(value ~ location_type, data = inflow_jwc)
  
TukeyHSD(jwc_inflow_anova)%>%tidy()%>%
  #filter(adj.p.value < 0.05)%>%
  select(-term)
  

```






# South Fork section

```{r}
chem_transformed_res <- tidy_chem%>%
    filter(site_code %in% c("COMI", )) %>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4), 
               names_to = "param", 
               values_to = "value")%>%
  select(site_code, Date, Year, param, value, burn_group, burn_year)%>%
  mutate(year_site = paste0(Year, "_", site_code))%>%
  mutate(log_value = log10(value + 1))%>%
  na.omit()

ggplot(chem_transformed_res, aes(x = value, color = year_site))+
  geom_histogram()+
  facet_wrap(~param, scales = "free")
```



```{r}
sf_chem <- tidy_chem%>%
  filter(site_code %in% c("COMI", "HORI", "COMR", "HORR", "COMO", "HORO", "BEAV", "SFM"))%>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4, PO4), 
               names_to = "param", 
               values_to = "value")%>%
  mutate(year_site = paste0(Year, "_", site_code))%>%
  mutate(log_value = log10(value + 1), 
         fill_status = if_else(Year <= 2022, "Unfilled", "Filled"))


ggplot(sf_chem, aes(x = value, color = year_site))+
  geom_histogram()+
  facet_wrap(~param, scales = "free")
```

## Inflow vs outflow unfilled

```{r}
inflow_outflow <- sf_chem %>%
  filter(site_code %in% c("COMI", "HORI", "COMR", "HORR", "COMO", "HORO"))%>%
  filter(param == "ChlA"  & !is.na(value)& location_type %in% c("Inflow", "Outflow"))%>%
  filter(Year <= 2022)%>%
  select(site_code, location_type, Date, value)
  

 in_out_anova <- aov(value ~ location_type, data = inflow_outflow)
  summary(in_out_anova)

TukeyHSD(in_out_anova)%>%tidy()%>%
  filter(adj.p.value < 0.05)%>%
  select(-term)


```

## Outflow unfilled to downstream

```{r}
outflow_ds <- sf_chem %>%
  filter(param == "ChlA"  & !is.na(value)& location_type %in% c( "Outflow", "Stream", "Mainstem"))%>%
  filter(Year <= 2022)%>%
  select(site_code, location_type, Date, value)
  

 outflow_ds_anova <- aov(value ~ location_type, data = outflow_ds)
  summary(in_out_anova)

TukeyHSD(outflow_ds_anova)%>%tidy()%>%
  #filter(adj.p.value < 0.05)%>%
  select(-term)


```


## Test between unfilled vs filled

```{r}

sf_fill_unfill <- sf_chem 
  filter(param == "ChlA"  & !is.na(value))%>%
  select(site_code, location_type,Year, value)%>%
  mutate(fill_status = ifelse(Year <= 2022, "unfilled", "filled"))%>%
  group_by(site_code) %>% # Group by site_code to test differences within each site
  pairwise_wilcox_test(
    value ~ fill_status, 
    p.adjust.method = "bonferroni")

sf_fill_unfill

sf_chem %>%
  filter(site_code %in% c("SFM", "BEAV"))%>%
   filter(param == "ChlA"  & !is.na(value))%>%
  mutate(fill_status = ifelse(Year <= 2022, "unfilled", "filled"))%>%
  group_by(site_code, fill_status)%>%
  summarise(mean_chla = mean(value))%>%
  pivot_wider(names_from = fill_status, values_from = mean_chla)%>%
  mutate(change = (filled/unfilled)*100)

```




# Mean all params

Code to make SI tables for minimum and maximum values 
```{r}

yearly_means <- tidy_chem%>%
  filter(Campaign %nin% c("CLP  Mainstem-Fort Collins","CLP Mainstem-Canyon" )& site_code %nin% c("PJW","JWC"  ,"PBR","PBD", "PNF","PSF","SFM","SAWM", "LDRT1", "legacy","timberline","ELC","archery","PFAL", "BEAV") & Year >= 2021)%>%
  select(-c("data_source","Lat","Long","location_type","Campaign","watershed","burn_status","Buffer_Level","Watershed_Level","distance_upstream_km", "FCW_number", "Field_DO_mgL","Field_Cond_µS_cm","Field_Temp_C"))%>%
  group_by(Year, site_code, Site)%>%
  summarise(across(starts_with(all_params), mean, na.rm = TRUE),
            .groups = 'drop')
 


  
  subset_table <- yearly_means%>%
  mutate(Site = paste0(Site, " (", site_code, ")"))%>%
    #select(Site, Year, `Chlorophyll a (ug/L)` = ChlA,     `SC (uS/cm)` = SC,    pH,    `DOC (mg/L)` = DOC)%>%
  #select(Site, Year,`TDN (mg/L)` = DTN,     `NH4 (mg/L)` = NH4,   	`NO3 (mg/L)` = NO3,   `PO4 (mg/L)` = PO4 )%>%
  #select(Site, Year,`SO4 (mg/L)` = SO4,     `Na (mg/L)` = Na,   	`Ca (mg/L)` = Ca,   `Cl (mg/L)` = Cl )%>%
  select(Site, Year,`K (mg/L)` = K,     `Mg (mg/L)` = Mg,   	`F (mg/L)` = `F`, TSS)%>%
    
  pivot_longer(cols = 
                   #c(`Chlorophyll a (ug/L)`:`DOC (mg/L)`),
                  #c(`TDN (mg/L)`:`PO4 (mg/L)`),
               #c(`SO4 (mg/L)`:`Cl (mg/L)`),
               c(`K (mg/L)`:TSS),
               names_to = "Parameter", values_to = "Value")%>%
  pivot_wider(names_from = c(Parameter, Year), values_from = Value, values_fill = 0)%>%
  select(Site, order(colnames(.)))

  subset_table$Site <- factor(subset_table$Site, levels = c("Barnes Meadow Outflow (BMD)", "Barnes Meadow Reservoir (BRNR)", "Chambers Inlet (CBRI)", "Chambers Reservoir (CBRR)", "Chambers Outflow (CHD)", "Joe Wright Inlet (JOEI)", "Joe Wright Reservoir (JOER)", "Comanche Inflow (COMI)",  "Comanche Reservoir (COMR)", "Comanche Outflow (COMO)", "Hourglass Inflow (HORI)", "Hourglass Reservoir (HORR)", "Hourglass Outflow (HORO)","Long Draw Outlet (LNGO)","Long Draw Reservoir (LNGR)","Peterson Outlet (PTRO)", "Peterson Reservoir (PTRR)"     ))

  #create standard border color and width
  std_border <- fp_border_default(width = 1, color = "black")
  
ft <- flextable(subset_table)%>%
  separate_header() %>%
  align(align = "center", part = "all")%>%
  colformat_double( big.mark = "'", decimal.mark = ".", digits = 2)%>%
  set_table_properties(layout = "fixed")%>%
  border_remove()%>%
  border_outer(part="all", border = std_border )%>%
  border_inner_h(border = std_border, part="all")%>%
   border_inner_v(border = std_border, part="all")%>%
  autofit()%>%
  italic( j = "Site")%>%
  fontsize( size = 8, part = "body")

#Highlighting columns other than Site
# Max is red
# min is blue
columns_to_highlight <- names(subset_table) %>%
  keep(~ !grepl("Site", .))

  for (i in 1:length(columns_to_highlight)) {
    max_color <- "#CC3311"
    min_color <- "#0077BB"
    
    ft <- color(ft, j = i+1, 
                    i = ~ subset_table[i+1] == max(subset_table[i+1], na.rm = TRUE), 
                    color = max_color)
    
    ft <- color(ft, j = i+1, 
                    i = ~ subset_table[i+1] == min(subset_table[i+1], na.rm = TRUE), 
                    color = min_color)
  }
  

ft

  
sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    width = 8.3, height = 11.7
  ),
  type = "continuous",
  page_margins = page_mar()
)

ft

# Save the table to a Word document
save_as_docx(ft, path = "data/table_export3.docx", pr_section = sect_properties)



```

# Maps / Spatial

## Prepping Landcover, Elevation/Aspect, Snow Persistance and Soil Burn Severity Data

### NLCD

```{r}

#pull in watersheds
watersheds <- st_read(dsn = here(spatial_dir,"cpf_watersheds.shp"))%>%
  left_join(most_recent_meta, by = c("site_code"))

nlcd_tiffs <- list.files(here(external_spatial, "landcover_NLCD"), pattern = "Annual_NLCD_LndCov",  full.names = T)%>%
  #find the files that contain ".tff" and nothing after it
  .[grepl(".tiff$", .)]
# read in rasters and name them based on the year contained in the file name
nlcd_rasters <- map(nlcd_tiffs, raster::raster)
#name them based on the year contained in the file name
names(nlcd_rasters) <- nlcd_tiffs%>%
  stringr::str_extract("\\d{4}")%>%
  as.numeric()


compute_top3_landcovers <- function(year){
  
  NLCD <- nlcd_rasters[[year]]
  
  nlcd_raster <- raster::brick(NLCD)


  ws <- watersheds %>%
    select(site_code)%>%
    rownames_to_column('index')%>%
    mutate(index = as.numeric(index))
  
  # using the exactextractr package, get what is essentially a "zonal histogram" in
  # ArcGIS speak
  ws_NLCD <- exactextractr::exact_extract(nlcd_raster, ws,
                                         append_cols = "index",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           group_by(lc_code = as.character(value)) %>%
                                           summarise(count = sum(coverage_fraction))) %>%
    group_by(index) %>% 
    mutate(total = sum(count)) %>%
    ungroup() %>%
    pivot_wider(names_from = 'lc_code', values_from = 'count',
                names_glue = "{lc_code}" ) %>%
    inner_join(ws, .) -> ws_NLCD
  
  ws_percents <- ws_NLCD %>% 
    #if there are NAs, change it to 0
    mutate(across(`11`:`82`, ~replace_na(., 0))) %>%
    mutate(water=100*(`11`/total),
           ice_snow=100*(`12`/total),
           open_space=100*(`21`/total),
           developed = 100*((`22`+ `23`+ `24`)/total),
           Barren = 100*(`31`/total),
           `Decidious Forest` = 100*((`41`)/total),
           `Evergreen Forest` = 100*((`42`)/total),
           `Mixed Forest` = 100*((`43`)/total),
           `Shrubland` = 100*((`52`)/total),
           `Grassland` = 100*((`71`)/total),
           pasture = 100*((`81`)/total),
           crops = 100*((`82`)/total),
           `Wetlands` = 100*((`90`+ `95`)/total),
           total_check = rowSums(across(water:Wetlands)))%>%
    st_drop_geometry()%>%
    select(total, site_code, water:total_check)
  
  
  top_land_cover <- ws_percents %>%
    select(-total, -total_check) %>%
  # Pivot longer to get land cover types and percentages in rows
  pivot_longer(cols = water:Wetlands, 
               names_to = "land_cover", 
               values_to = "percent")%>%
  # Group by site_code and find the top 4 land covers by percent
  group_by(site_code) %>%
  arrange(desc(percent), .by_group = TRUE) %>%
  slice_head(n = 3) %>%
  mutate(rank = row_number()) %>% # Rank the top 4 land covers
  ungroup() %>%
  # Pivot wider with unique column names
  pivot_wider(
    id_cols = site_code, 
    names_from = rank, 
    values_from = c(land_cover, percent), 
    names_glue = "{.value}_{rank}"
  )%>%
    select(site_code, land_cover_1, percent_1, land_cover_2, percent_2, land_cover_3, percent_3)%>%
    mutate(nlcd_year = year)%>%
    #round all percent columns to 1 decimal places
    mutate(across(contains("percent"), ~round(., 1)))


return(top_land_cover)
  
  
}

compute_all_landcovers <- function(year){
  
  NLCD <- nlcd_rasters[[year]]
  
  nlcd_raster <- raster::brick(NLCD)


  ws <- watersheds %>%
    select(site_code)%>%
    rownames_to_column('index')%>%
    mutate(index = as.numeric(index))
  
  # using the exactextractr package, get what is essentially a "zonal histogram" in
  # ArcGIS speak
  ws_NLCD <- exactextractr::exact_extract(nlcd_raster, ws,
                                         append_cols = "index",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           group_by(lc_code = as.character(value)) %>%
                                           summarise(count = sum(coverage_fraction))) %>%
    group_by(index) %>% 
    mutate(total = sum(count)) %>%
    ungroup() %>%
    pivot_wider(names_from = 'lc_code', values_from = 'count',
                names_glue = "{lc_code}" ) %>%
    inner_join(ws, .) -> ws_NLCD
  
  all_land_cover <- ws_NLCD %>% 
    #if there are NAs, change it to 0
    mutate(across(`11`:`82`, ~replace_na(., 0))) %>%
    mutate(water=100*(`11`/total),
           ice_snow=100*(`12`/total),
           open_space=100*(`21`/total),
           developed = 100*((`22`+ `23`+ `24`)/total),
           Barren = 100*(`31`/total),
           `Decidious Forest` = 100*((`41`)/total),
           `Evergreen Forest` = 100*((`42`)/total),
           `Mixed Forest` = 100*((`43`)/total),
           `Shrubland` = 100*((`52`)/total),
           `Grassland` = 100*((`71`)/total),
           pasture = 100*((`81`)/total),
           crops = 100*((`82`)/total),
           `Wetlands` = 100*((`90`+ `95`)/total),
           total_check = rowSums(across(water:Wetlands)))%>%
    st_drop_geometry()%>%
    select( site_code, total, water:`Wetlands`)%>%
    mutate(nlcd_year = year)

return(all_land_cover)
  
  
}

years <- c("2019", "2023")

ws_lc <- map_df(years, compute_top3_landcovers)%>%
  bind_rows() %>%
  #swapping water for land_cover_3
  mutate(land_cover_2 = case_when(land_cover_2 == "water" ~ land_cover_3, 
                                  TRUE ~ land_cover_2), 
         percent_2 = case_when(land_cover_2 == "water" ~ percent_3, 
                               TRUE ~ percent_2), 
         # creating a new column that combines the top 2 land covers for table
         land_cover_top2 = paste0(land_cover_1, " (", percent_1, " %) \n",land_cover_2, " (", percent_2," %) " ))%>%
  select(site_code, nlcd_year,land_cover_top2) %>%
  pivot_wider(names_from = nlcd_year, values_from = land_cover_top2, names_prefix = "lc_top2_")

#run to get all landcover %s for watersheds
ws_all_lc <- map_df(years, compute_all_landcovers)%>%
  bind_rows()%>%
  pivot_longer(
    cols = water:Wetlands, 
    names_to = "landcover", 
    values_to = "value"
  ) %>%
  select(-total)%>%
  pivot_wider(,
    names_from = nlcd_year, 
    values_from = value,
    names_prefix = "value_" # This creates value_2019 and value_2023
  ) %>%
  select(site_code, landcover, value_2019, value_2023)%>%
  mutate(pp_change = value_2023 - value_2019)


rm(nlcd_rasters, nlcd_tiffs, years)
```

## Getting DEM data

```{r}

ws_chars<- map_dfr(unique(watersheds$site_code), function(ws_code){
  
  ws <- watersheds %>%
    filter(site_code == ws_code)
  
   ws <- st_make_valid(ws)
  
  if(ws_code %in% c("JOER", "CBRR", "BRNR", "LNGR", "PTRR", "COMR","HORR")){
    z_num = 10
  }else{
    z_num = 9
  }
  
  elev <- get_elev_raster(ws, z = z_num, verbose = F)

  
  ws_elev <- exactextractr::exact_extract(elev, ws,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           summarise(mean_elev = mean(value, na.rm = T), 
                                                     max_elev = max(value, na.rm = T), 
                                                     min_elev = min(value, na.rm = T))) %>%
    select(site_code, mean_elev, max_elev, min_elev)
  
  
  #get the slope
  slope <- raster::terrain(elev, opt = "slope", unit = "degrees", verbose = F)
  ws_slope <- exactextractr::exact_extract(slope, ws,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           summarise(mean_slope = mean(value, na.rm = T), 
                                                     max_slope = max(value, na.rm = T), 
                                                     min_slope = min(value, na.rm = T))) %>%
    select( mean_slope, max_slope)
  
  #calculate aspect
  
  aspect <- raster::terrain(elev, opt = "aspect", unit = "degrees")
  ws_aspect <- exactextractr::exact_extract(aspect, ws,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                          summarise(mean_aspect = mean(value, na.rm = T))) %>%
    select(mean_aspect)%>%
    round(2)
  # calculate watershed area
  
 
  ws_area <- st_area(ws)%>%
    #convert to km2
    units::set_units("km^2")
  
  
  bind_cols(ws_elev, ws_slope, ws_aspect)%>%
    mutate(ws_area_km2 = ws_area)
  
})

```

### Get elevational gradient for CPF fire

```{r}
ws <-  terra::rast(here(spatial_dir, "cpf_sbs_BAER.tif"))
#remove 1 values
#ws[ws == 1] <- NA
ws[ws == 0] <- NA
mapview::mapview(ws)

#convert to shapefile
# Convert raster to polygon (dissolve all cells into one)
outline <- as.polygons(ws, dissolve = TRUE, na.rm = TRUE) %>%
  st_as_sf()

outline$site_code <- "cpf"

outline <- outline %>%
  group_by(site_code) %>%
  summarise(geometry = st_union(geometry), .groups = "drop") %>%
  st_cast("POLYGON")

elev <- get_elev_raster(outline, z = 9, verbose = F)



  ws_elev <- exactextractr::exact_extract(elev, outline,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           summarise(mean_elev = mean(value, na.rm = T), 
                                                     max_elev = max(value, na.rm = T), 
                                                     min_elev = min(value, na.rm = T)))%>%
    select(site_code, mean_elev, max_elev, min_elev)

  
  ws_area <- st_area(outline)%>%
    #convert to km2
    units::set_units("km^2")
  
  
  cpf<- bind_cols(ws_elev)%>%
    mutate(ws_area_km2 = ws_area)%>%
    group_by(site_code)%>%
    summarise(mean_elev = mean_elev,
           max_elev = max(max_elev),
           min_elev = min(min_elev),
           ws_area_km2 = sum(ws_area_km2))%>%
    slice(1)
  
```

## Calculating snow persistance for watersheds

```{r}

  #get the snow persistence raster
  snow <- raster::raster(here(external_spatial, "snow_persistance/MOD10A2_SCI_2020.tif"))

ws_snow <- map_dfr(unique(watersheds$site_code),  function(ws_code){
  
  ws <- watersheds %>%
    filter(site_code == ws_code)
  
  ws_snow <- exactextractr::exact_extract(snow, ws,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           summarise(mean_snow = mean(value, na.rm = T), 
                                                     max_snow = max(value, na.rm = T), 
                                                     min_snow = min(value, na.rm = T))) %>%
    select(site_code, mean_snow, max_snow, min_snow)%>%
    mutate(across(where(is.numeric), ~round(., 0)))%>%
    mutate(days = 184 * (mean_snow/100), 
           mean_snow_free_date= format(as.Date(days, origin = "2020-01-01"), "%m-%d"))%>%
    select(site_code, mean_snow, mean_snow_free_date)
  
  ws_snow
  
})

rm(snow)
```

## Calculate watershed soil burn severity

```{r}


sbs <- terra::rast(here(spatial_dir, "cpf_sbs_BAER.tif"))

# Load the shapefile and raster
ws <- st_read(here(spatial_dir, "cpf_watersheds.shp")) %>%
  st_transform(sf::st_crs(sbs)) 

ws_raster <- ws %>%
  terra::vect()

# Ensure all NA values in the raster are already set to 5
sbs[is.na(sbs) | is.null(sbs) | sbs == 0] <- 5

# Get the extent of the ws vector
ws_extent <- terra::ext(ws_raster)

# Extend the raster to match the ws extent, assigning new areas the value of 5
sbs_extended <- terra::extend(sbs, ws_extent, fill = 5)

# Get the cell size in map units (meters)
cell_size <- terra::res(sbs_extended)

cnv_to_km2 <- (cell_size[1] * cell_size[2]) / 1e6

  # using the exactextractr package, get what is essentially a "zonal histogram" in
  # ArcGIS speak
  extracted_sbs <- exactextractr::exact_extract(sbs_extended %>% raster::raster(), ws,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           group_by(sbs_code = as.character(value)) %>%
                                           summarise(count = sum(coverage_fraction))) %>%
    group_by(site_code)%>%
    mutate(total = sum(count)) %>%
    ungroup() %>%
    pivot_wider(names_from = 'sbs_code', values_from = 'count',
                names_glue = "sbs{sbs_code}" ) %>%
    inner_join(ws, .) -> ws_sbs
  
  
#   library(tmap)
# 
# # Switch to interactive mode
# tmap_mode("view")
# 
# # Create the map
# tm_shape(sbs_extended) + 
#   tm_raster(palette = "YlOrRd", opacity = 0.7, title = "Burn Severity") +
# tm_shape(ws) + 
#   tm_borders(col = "black", lwd = 2) + 
#   tm_layout(title = "Watershed vs. SBS Overlap")
  
  
  ws_percents <- extracted_sbs %>% 
    #if a column is NA change it to 0
    mutate(across(starts_with("sbs"), ~replace_na(., 0))) %>%
    mutate(Unburned=100*((sbs5 + sbs6) /total),
           `Very Low` =100*(sbs1/total),
           Low=100*(sbs2/total),
           Moderate=100*(sbs3/total),
           High=100*(sbs4/total), 
           #total_check = rowSums(across(c(Unburned:High))), 
           Unburned_km2= (sbs5 + sbs6)*cnv_to_km2,
           `Very Low_km2` = sbs1*cnv_to_km2,
           Low_km2= sbs2*cnv_to_km2,
           Moderate_km2= sbs3*cnv_to_km2,
           High_km2=sbs4*cnv_to_km2 , 
           total_burn_km2 = rowSums(across(c(`Very Low_km2`:High_km2))))%>%
    select(site_code,total, Unburned:total_burn_km2)%>%
    ungroup()%>%
    st_drop_geometry()%>%
    mutate(across(Unburned:total_burn_km2, ~round(., 1)))%>%
    mutate(lows = `Very Low`+Low, 
           lows_km2 = `Very Low_km2`+Low_km2) %>% 
    mutate(watershed_level=((Unburned*0)+(`Very Low`*0.1)+(Low*0.4)+(Moderate*0.7)+(High*1))/(Unburned+`Very Low`+Low+Moderate+High))



ws_sbs <- ws_percents

#just burned study reservoirs (excl LNGR, JOER)
ws_sbs_res <- ws_sbs%>%filter(site_code %in% c("CBRR", "BRNR", "PTRR","COMR", "HORR", "JOER", "LNGR"))%>%
  group_by(site_code)%>%
  summarize(total_area = sum(total*cnv_to_km2), 
            total_burn_km2 = sum(total_burn_km2), 
            lows_km2 = sum(lows_km2), 
            Moderate_km2 = sum(Moderate_km2),
            High_km2 = sum(High_km2))%>%
  mutate(perc_burn = (total_burn_km2/total_area)*100, 
         perc_low = (lows_km2/total_area)*100, 
         perc_mod = (Moderate_km2/total_area)*100, 
         perc_high = (High_km2/total_area)*100)


rm(sbs, sbs_extended, ws_raster, ws_extent)

```


## Get data together for Reservoir characteristics table

```{r}
res_meta <- read_csv(here(spatial_dir, "reservoir_metadata.csv"))

all_ws_data <- most_recent_meta %>%
                    select(site_code, Site_Name)%>%
  left_join(ws_chars, by = c("site_code"))%>%
  left_join(ws_sbs%>% select(-total), by = c("site_code"))%>%
  left_join(ws_all_lc, by = c("site_code"))%>%
  left_join(ws_lc, by = c("site_code"))%>%
  #mutate all numeric columns to round to 2 decimal places
  mutate(across(where(is.numeric), ~round(., 2))) %>%
   mutate(Site = paste0(Site_Name, " (", site_code, ")"),
         total_burn_perc = round(lows+Moderate+High, 1),
         total_burn_sum = paste0(total_burn_km2, " (",total_burn_perc, ") " ))%>%
  na.omit()

concise_ws_data <- all_ws_data %>%
  select(site_code, total_burn_perc, mean_elev, max_elev, min_elev, ws_area_km2)%>%
  st_drop_geometry()%>%
  distinct()

write_csv(concise_ws_data, here(spatial_dir,"concise_ws_chars.csv"))

res_ws_data <- all_ws_data%>%
    left_join(res_meta, by = c("site_code"))%>%
  #cleaning up the names
  mutate(Site = paste0(Site_Name, " (", site_code, ")"),
         total_burn_perc = round(lows+Moderate+High, 1),
         total_burn_sum = paste0(total_burn_km2, " (",total_burn_perc, ") " ))%>%
  select( site_code, Site,
         `Watershed Area (km^2)` = ws_area_km2, 
          `Reservoir Elevation (m)` = min_elev,
         `Max Watershed Elevation (m)` = max_elev,
         `Mean Watershed Elevation (m)` = mean_elev, 
         `Normal Storage Capacity (acre ft)`, 
          `Reservoir Depth (m)`, 
         `Reservoir Operator` = Manager,
         `Total Burn Area (km^2, % watershed)` = total_burn_sum,
         `Burn Index (0-1)` = watershed_level, 
         `Percent High Severity` = High,
         `Percent Moderate Severity` = Moderate,
         `Percent Very Low & Low Severity` = lows,
         `Percent Unburned` = Unburned,
         `Mean Slope (degrees)` = mean_slope,
         `2019 Top 2 Land Covers` = lc_top2_2019,
         `2023 Top 2 Land Covers` = lc_top2_2023)%>%
  arrange((`Burn Index (0-1)`))%>%
  distinct()%>%
  filter(site_code %in% c("CBRR", "BRNR", "PTRR","COMR", "HORR", "JOER", "LNGR"))

write_csv(res_ws_data, here(spatial_dir, "reservoir_characteristics.csv"))
```

## Make a table with the land cover and watershed characteristics

```{r}

t_res_ws_data<- res_ws_data%>%
  t()%>%
  as.data.frame()%>%
  #make column names the first row
  setNames(.[1,])%>%
  # remove first row
  .[-1,]%>%
  rownames_to_column(" ")

#create standard border color and width
std_border <- fp_border_default(width = 1, color = "black")
  
ft <- flextable(t_res_ws_data)%>%
  separate_header() %>%
  align(align = "center", part = "all")%>%
  colformat_double( big.mark = ",", decimal.mark = ".", digits = 3)%>%
  set_table_properties(layout = "fixed")%>%
  border_remove()%>%
  border_outer(part="all", border = std_border )%>%
  border_inner_h(border = std_border, part="all")%>%
   border_inner_v(border = std_border, part="all")%>%
  autofit()%>%
  fontsize( size = 6, part = "body")%>%
  fontsize(size = 6, part = "header") %>% # Smaller font size for the header
  bold(part = "header") 

sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    width = 8.3, height = 11.7
  ),
  type = "continuous",
  page_margins = page_mar()
)
# 
 #ft
# 
# # Save the table to a Word document
#save_as_docx(ft, path = "data/figs_output/res_ws_data_table.docx", pr_section = sect_properties)
```

## Maps

## Pull in watersheds and sampling points and join them include metadata

```{r}

## Pull in sampling points

sampling_points <- most_recent_meta %>%
  #filter for all upper sites (or reservoirs)
  filter(grepl("Upper|Reservoir", watershed))%>%
  st_as_sf(coords = c("Long", "Lat"), crs = 4326)

sampling_points_study_res <- sampling_points%>%
  filter(grepl("Reservoir", watershed))%>%
  mutate(ws_site_code = case_when(
    watershed == "Barnes Meadow Reservoir" ~ "BRNR",
    watershed == "Joe Wright Reservoir" ~ "JOER",
    watershed == "Long Draw Reservoir" ~ "LNGR",
    watershed == "Chambers Reservoir" ~ "CBRR",
    watershed == "Comanche Reservoir" ~ "COMR",
    watershed == "Hourglass Reservoir" ~ "HORR",
    watershed == "Peterson Reservoir" ~ "PTRR"))%>%
  select(site_code, Site_Name, watershed, ws_site_code)
    
st_write(sampling_points_study_res, "data/spatial/maps_data/sampling_points.shp")

#pull in watersheds
watersheds <- st_read(dsn = "data/spatial/cpf_watersheds/cpf_watersheds.shp")%>%
  left_join(most_recent_meta, by = c("site_code"))

watersheds_study_res <- watersheds%>%
  filter(site_code %in% c("JOER", "CBRR", "BRNR", "LNGR", "PTRR","COMR", "HORR"))%>%
  select(site_code, Site_Name, watershed)

st_write(watersheds_study_res, "data/spatial/maps_data/watersheds.shp")

```

## Pull in waterbodies

```{r}

study_waterbodies <- st_read("data/spatial/study_water_bodies/study_water_bodies.shp")%>%
  st_transform(4326)%>%
  left_join(most_recent_meta, by = c("site_code"))%>%
  select(site_code, Site_Name, watershed)

st_write(study_waterbodies, "data/spatial/maps_data/study_waterbodies.shp")

#mapview::mapview(study_waterbodies)

```

## Pull in SBS tiff

```{r}

# Load the raster
sbs <- raster::raster("data/spatial/BAER_sbs_raster/cpf_sbs_BAER.tif")
#project into 4326
sbs <- projectRaster(sbs, crs = 4326)

# Remove water bodies  and blank spaces (0 & 6)
sbs[sbs == 6] <- NA
sbs[sbs == 0] <- NA
# Define burn severity colors and labels

legend <- tibble(
  value = c(1, 2, 3, 4, 5),
  legend = c("Very Low/Unburned", 
             "Low", 
             "Moderate", 
             "High",
             "Unburned"),
        colors = c("blue", "green2", "yellow", "red", "green4"))

# Map the values to colors
burn_colors <- legend$colors
names(burn_colors) <- legend$value

# Create an interactive map
mapview::mapview(sbs,
        col.regions = burn_colors,
        legend = TRUE,
        na.color = "transparent")

mapview::mapview(sbs)+
  mapview::mapview(watersheds%>%filter(site_code == "PBD"))
 


```

## Make a simple map of reservoir points, watersheds over the SBS

```{r}
mapview::mapview(sbs, col.regions = burn_colors, legend = TRUE, na.color = "transparent")+
  mapview::mapview(study_waterbodies)+
  mapview::mapview(watersheds_study_res)+
mapview::mapview(sampling_points_study_res)
  

mapview::mapview(sbs, col.regions = burn_colors, legend = TRUE, na.color = "transparent") +
  
  mapview::mapview(watersheds_study_res, zcol = "site_code" , color = colorsBS_site_code_alpha, col.regions = colorsBS_site_code_alpha,
                   alpha.regions = 0.1,   burst = T, lwd = 2)+  
  #mapview::mapview(sampling_points_study_res, zcol = "watershed_site_code" , col.regions = colorsBS_site_code_alpha, burst = T)+
  mapview::mapview(study_waterbodies ,col.regions = "blue2", burst = T)

  
```

## attempt at ggplot map

```{r}

library(ggplot2)
library(sf)
library(raster)
library(tidyverse)
library(ggspatial)
library(ggmap)
library(basemaps)

ext <- draw_ext()


ext_final <- st_transform(sbs,crs = st_crs(3857))
# set defaults for the basemap
set_defaults(map_service = "esri", map_type = "natgeo_world_map")

# load and return basemap map as class of choice, e.g. as image using magick:
basemap_magick(ext)



# Convert the raster (sbs) to a data frame for ggplot
sbs_df <- as.data.frame(raster::rasterToPoints(sbs), stringsAsFactors = FALSE)
colnames(sbs_df) <- c("x", "y", "value")

# Define the ggplot
ggplot() + 
  basemap_gglayer(ext_final)+
  
  # Add the watersheds layer
  geom_sf(data = watersheds_study_res, aes(color = site_code), fill = NA, linewidth = 1.5) +
  scale_color_manual(
    values = colorsBS_site_code_alpha,
    name = "Watersheds"
  ) +
  
  # Add the waterbodies layer
  geom_sf(data = study_waterbodies, color = "blue2", fill = "blue2", alpha = 0.3) +
  
  # Remove axis titles and text
  theme_minimal() +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) +
  
  # Add a legend title and adjust its position
  guides(
    fill = guide_legend(title = "Soil Burn Severity"),
    color = guide_legend(title = "Watersheds")
  )
```

# Merging watersheds

```{r}
library(sf)
library(mapview)



grand_ditch <- st_read("data/spatial/working/final_watersheds/grand_ditch.shp" )%>%
  st_transform(4326)%>%
  #mutate(site_name = "Michigan Ditch")%>%
  select(site_name, geometry)

michigan_ditch <- st_read("data/spatial/working/final_watersheds/michigan_ditch.shp" )%>%
  st_transform(4326)%>%
  #mutate(site_name = "Michigan Ditch")%>%
  select(site_name, geometry)

LRT <- st_read("data/spatial/working/final_watersheds/LRT_TB_final.shp" )%>%
  st_transform(4326)%>%
  mutate(site_name = "Laramie River Tunnel")%>%
  select(site_name, geometry)

st_write(LRT, "data/spatial/working/final_watersheds/Laramie_River_Tunnel_final.shp")


mapview::mapview(grand_ditch)+
  mapview::mapview(michigan_ditch)+
  mapview::mapview(LRT)

TB <- bind_rows(
  grand_ditch,
  michigan_ditch,
  LRT
)

mapview::mapview(TB, zcol = "site_name", burst = T)
st_write(TB, "data/spatial/working/final_watersheds/trans_basin_ws.shp")

# Merge polygons

#st_write(grand_corrected, "data/spatial/working/final_watersheds/grand_ditch.shp")


#st_write(michigan_final, "data/spatial/working/final_watersheds/michigan_ditch.shp")


# 
# polygon3 <- st_read("data/spatial/working/SLEP_ss/layers/globalwatershed.shp" )%>%
#   st_transform(4326)%>%
#   mutate(site_name = "SLEP_ss")%>%
#   select(site_name, geometry)
# 
# st_is_valid(polygon3)
#polygon3 <- st_make_valid(polygon3)
# 
# mapview::mapview(polygon3)

# Merge polygons
merged_geometry <- st_difference(polygon2$geometry, polygon1$geometry)
#merged_geometry <- st_union(polygon1$geometry, polygon2$geometry)

merged_polygon <- st_sf(
  site_name = "HORR",
  geometry = st_sfc(merged_geometry))%>%
  st_make_valid()

st_is_valid(merged_polygon)

mapview::mapview(merged_polygon)
# 
# merged_geometry2 <- st_union(merged_polygon$geometry, polygon3$geometry)
# 
# 
# merged_polygon2 <- st_sf(
#   site_name = "SLEP",
#   geometry = st_sfc(merged_geometry2))%>%
#   st_make_valid()
# 
# st_is_valid(merged_polygon2)
# 
# mapview::mapview(merged_polygon2)


# Visualize with different colors
mapview::mapview(merged_polygon, zcol = "site_name", col.regions = "blue")+
  #mapview::mapview(merged_polygon2, zcol = "site_name", col.regions = "green")+
  mapview::mapview(polygon1, zcol = "site_name", col.regions = "cyan")+
  mapview::mapview(polygon2, zcol = "site_name", col.regions = "red")
  #mapview::mapview(polygon3, zcol = "site_name", col.regions = "yellow")
  

# save the merged polygon in the spatial/working/final_ws folder
st_write(merged_polygon, "data/spatial/working/final_watersheds/HORR_final.shp")

st_read("data/spatial/working/final_watersheds/HORR_final.shp" )%>%
  st_is_valid()


```

Stream Stats sites

```{r}

polygon <- st_read("data/spatial/working/HORR_tweaked/HORR_tweaked.shp" )%>%
  st_transform(4326)

polygon <- polygon[1,]%>%
  mutate(site_name = "HORO")%>%
  select(site_name, geometry)

st_is_valid(polygon)
mapview::mapview(polygon)

st_write(polygon, "data/spatial/working/final_watersheds/HORO_final.shp")

```

# Read in all files

```{r}

meta <- read.csv("data/ross_clp_chem/data/metadata/location_metadata.csv")%>%
  mutate(point_lat_long = paste0(Lat, " , ", Long), 
         site_desc = Site_Name)%>%
  select(-Lat, -Long, -burn_status, -Site_Name)

# loop over the final_watersheds folder to read in all the files
files <- list.files("data/spatial/working/final_watersheds", pattern = ".shp", full.names = T)

# read in all the files
final_watersheds <- map(files, st_read)%>%
  bind_rows()

ws_detailed <- final_watersheds%>%
  mutate(tb_ws = case_when(
    site_name %in% c("Grand Ditch", "Michigan Ditch", "Laramie River Tunnel") ~ T,
    TRUE ~ FALSE), 
    # South Fk sites include no TB, PTRR is all native flow, BRNR is all native flow
  incl_tb = case_when(
    site_name %in% c("HORR", "HORO", "HORI", "COMI", "COMO", "COMR", "LBEA", "PENN","SFM",
                     "PTRO", "PTRR", "BRNR", "BMD" ) ~ F,
    TRUE ~ T))

%>%
  left_join(meta, by = c("site_name" = "site_code"))

st_write(ws_detailed, "data/spatial/cpf_watersheds_final/cpf_watersheds_final.shp")



mapview::mapview(ws_detailed %>% filter(location_type == "Reservoir"), zcol = "site_name", burst = T)
#+
 # mapview::mapview(ws_detailed %>% filter(trans_basin), zcol = "site_name", col.regions = "red")

```

## Test

```{r}
# Read in the final watersheds

check <- st_read(dsn =  "data/spatial/cpf_watersheds_final/all_watersheds_w_meta.shp")

mapview::mapview(check %>% filter(location_type == "Reservoir"), zcol = "site_name", burst = T)

```


# Flow data

```{r}
library(cdssr)
 # pull dwr claftcco data from 1881-2021
q_data <- cdssr::get_sw_ts(abbrev = "CLAFTCCO", timescale = "day", start_date = "1980-10-01", end_date = "2020-09-30")
  
  
q_data_clean <- q_data%>%
  mutate(q_cfs = value, 
           site_code = abbrev, 
         acre_ft_day = q_cfs *1.983 )%>%
    distinct()%>%
  # remove rows where flag_b is "Ice"
  filter(!grepl("Ice", flag_b))%>%
  select(abbrev, Date = datetime, q_cfs, acre_ft_day, flag_a, flag_b)%>%
  addWaterYear()

q_annual <- q_data_clean%>%
  group_by(waterYear)%>%
  summarise(t_flow_acft = round(sum(acre_ft_day), 0), 
            max_cfs = max(q_cfs), 
            #find the Date with the highest q_cfs per year
            max_cfs_date = Date[which.max(q_cfs)],
            )%>%
  ungroup()%>%
  mutate(peak_dy = yday(max_cfs_date), 
         month = month(max_cfs_date))

mean(q_annual$t_flow_acft)

#make a histogram of peak date
q_annual%>%
  ggplot(aes(x = month))+
  geom_histogram(binwidth = 1)+
  theme_bw(base_size = 15)+
  labs(title = "Peak Flow Date Distribution", x = "Day of Year", y = "Count")



check <- q_data_clean%>%
  mutate(y_day = yday(Date), 
         waterYear = as.character(waterYear))

ggplot(check, aes(y_day, q_cfs, color = waterYear))+
  geom_line()+
  theme_bw(base_size = 15)+
  labs(title = "Flow Data for 1999", x = "Date", y = "Flow (cfs)")+
  gghighlight(waterYear %in% c("2019", "2012", "2000", "1990") , use_direct_label = TRUE)

  
```
