---
title: "Figures"
author: "Sam Struthers"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---
Setup
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

### ----- Load packages ----- ###
package_load <- function(package_names){
  for(i in 1:length(package_names)){
    if(!package_names[i] %in% installed.packages()){
      install.packages(package_names[i])
    }
    library(package_names[i],character.only = TRUE)
  }
}

#vector of packages
pack_req <- c( 
  # data wrangling packages
  "janitor", "broom","readxl","here","arrow",
  #spatial packages
  "sf","terra","raster","tmap","exactextractr",
  # plotting
  "ggpubr","ggthemes","patchwork", "gghighlight","grid", 
  "geomtextpath","plotly", "ggpmisc","officer", "flextable","kableExtra",
  #external datasets
  "cdssr","nhdplusTools","elevatr","prism",
  #stats
  "trend", "anomalize","rstatix", "car","performance","lme4","glmmTMB","emmeans",
  #always load tidyverse last to avoid conflicts with other packages
  "tidyverse")
package_load(pack_req)

remove(pack_req, package_load)

#basic function to save a plot in standard size, dpi and width into figures folder
save_plot_image<- function(plot,name){
  
  ggsave(paste0("docs/cpf_reservoir/paper_draft/",name,".jpg"), dpi = 300, width = 18, height = 10)
  
}

paper_draft_dir <- here("docs/cpf_reservoir/paper_draft/")
spatial_dir <- "data/raw/spatial/res_study"
external_spatial <- "data/raw/external_datasets"


#Simple function to negate %in%

`%nin%` = Negate(`%in%`)

# # Distance order and xdistance
order_table <- tibble(
  site_code = c('JOEI', 'JOER', 'CBRI', 'CBRR', 'CHD', 'BRNR', 'BMD', 'LNGR', 'LNGO', 'PTRR', 'PTRO', 'COMI', 'COMR', 'COMO', 'HORI', 'HORR', 'HORO', 'BEAV', 'JWC', 'PJW', 'PFAL', 'SLEP', 'PBR', 'SFM', 'PSF', 'PNF', 'PBD'),
  Distance = c('1 - JOEI', '2 - JOER', '3 - CBRI', '4 - CBRR', '5 - CHD', '6', '7', '8', '9', '99', '999', '11 - COMI', '12 - COMR', '13 - COMO', '14 - HORI', '15 - HORR', '16 - HORO', '17 - BEAV', '18 - JWC', '19 - PJW','20- SLEP', '20 - SLEP', '21 - PBR', '22 - SFM', '23 - PSF', '24 - PNF', '25 - PBD'),
  Xdistance = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 99, 999, 12, 13, 14, 15, 16, 17, 18, 19,20, 20, 21, 22, 23, 24, 25, 26),
  order = c('03', '04', '05', '06', '07', '10', '11', '01', '02', '08', '09', '12', '13', '14', '15', '16', '17', '18', '19','20', '20', '21', '22', '23', '24', '25', '26'), 
  BS_order =  c(NA, '2', NA, '4', NA, '7', NA, '1', NA, '6', NA, NA, '5', NA, NA, '3', NA, NA, NA, NA, NA, NA, NA, NA, NA, NA,NA),
)


# Create the lookup table with expression syntax
# Define as a named list
chem_expressions <- list(
  "Turbidity" = expression(Turbidity ~ (NTU)),
  "TSS"       = expression(TSS ~ (mg ~ L^{-1})),
  "ChlA"      = expression(Chlorophyll ~ italic("a") ~ (mu*g ~ L^{-1})),
  "TC"        = expression(TC ~ (mg ~ L^{-1})),
  "TN"        = expression(TN ~ (mg ~ L^{-1})),
  "DOC"       = expression(DOC ~ (mg ~ L^{-1})),
  "DTN"       = expression(TDN ~ (mg ~ L^{-1})),
  "pH"        = expression(pH),
  "ANC"       = expression(ANC ~ (mu*eq ~ L^{-1})),
  "SC"        = expression(SC ~ (mu*S ~ cm^{-1})),
  "Na"        = expression(Na^{"+"} ~ (mg ~ L^{-1})),
  "NH4"       = expression(NH[4]^{"+"} ~ (mg ~ L^{-1})),
  "K"         = expression(K^{"+"} ~ (mg ~ L^{-1})),
  "Mg"        = expression(Mg^{"2+"} ~ (mg ~ L^{-1})),
  "Ca"        = expression(Ca^{"2+"} ~ (mg ~ L^{-1})),
  "F"         = expression(F^{"-"} ~ (mg ~ L^{-1})),
  "Cl"        = expression(Cl^{"-"} ~ (mg ~ L^{-1})),
  "NO3"       = expression(NO[3]^{"-"} ~ (mg ~ L^{-1})),
  "PO4"       = expression(PO[4]^{"3-"} ~ (mg ~ L^{-1})),
  "SO4"       = expression(SO[4]^{"2-"} ~ (mg ~ L^{-1}))
)

 primary_study_res <-c("CBRR", "JOER", "PTRR",  "BRNR", "LNGR")
 all_reservoirs <- c("COMR", "HORR", "CBRR", "JOER", "PTRR", "BRNR", "LNGR")
 all_res_system <- c("COMI", "COMR", "COMO", "HORI","HORR" , "HORO","CBRR", "CHD", "JOEI", "JOER", "CBRI", "PTRR", "PTRO", "BMD", "BRNR", "LNGO", "LNGR")

### ----- Color Sets ----- ###

cbbPalette <- c( "#999999","#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#Color for reservoirs, matches burn severity by site code

colorsBS_site_code <- c("BRNR" = "#d71a1d", "PTRR" ="#F3722C", "COMR" = "#F8961E", "CBRR" = "#F9C74F", "HORR" = "#F0E442", "JOER" = "#56B4E9", "LNGR" = "#0072B2")
#Seasonal Color values
season_clp_color_vals <- c("Winter" = "#03045e",  "Rising Limb" = '#047E82',  "Peak Flow" = '#59B851', "Falling Limb"='#FA850E', "Fall Baseflow" = "#772f1a")

if ({
  require(systemfonts)
  ("Roboto" %in% system_fonts()$family)
}) {
  ROSS_theme <- theme_bw() + #or theme_few()
    theme(plot.title = element_text(hjust = 0.5, face = 'bold', family = "Roboto"),
          plot.subtitle = element_text(hjust = 0.5, family = "Roboto"))
} else {
  message("You do not have the Roboto font family installed on your computer, currenly using ggplot default text family.
          See ROSS_themes.R for directions to install the font family on your computer.")
  ROSS_theme <- theme_bw() + #or theme_few()
    theme(plot.title = element_text(hjust = 0.5, face = 'bold'),
          plot.subtitle = element_text(hjust = 0.5))
}

ROSS_lt_pal <- c("#002EA3", "#E70870", "#256BF5",
                 "#745CFB", "#1E4D2B", "#56104E")

ROSS_dk_pal <- c("#1E4D2B", "#E70870", "#256BF5",
                 "#FFCA3A", "#745CFB", "#C3F73A")

ROSS_acc_pal <- c("#56104E", "#745CFB", "#FFCA3A", "#1E4D2B")
```


# Grab data

```{r}
#discrete sample data and location metadata from most recent pub
data_folder <- here("data/raw/chem/ross_clp_chem/v2025.11.14")

most_recent_chem <- read_rds(here(data_folder, "data/cleaned/CLP_chemistry_up_to_20251114.rds"))%>%
  mutate(flow_gauge_id = ifelse(flow_gauge_source == "USGS", paste0("0", flow_gauge_id), flow_gauge_id))
# most_recent_meta <- read_csv_arrow(here(data_folder,"data/metadata/location_metadata.csv"))
# units_meta_file <- here(data_folder, "data/metadata/water_chemistry_units.xlsx" )
# chem_units <- readxl::read_xlsx(units_meta_file)
# all_params <- readxl::read_xlsx(units_meta_file)%>%
#   pull(simple)

rm(units_meta_file)


tidy_chem <- most_recent_chem%>%
  #remove random pre fire data 
  filter(Year > 2020 & Year < 2025)%>%
  filter(watershed %nin% c("CLP Mainstem - Lower","CLP Tributary - Lower"))%>%
  # pfal + slep represent the same site but were moved to match future sensor location
  mutate(distance_upstream_km = case_when(site_code == "PFAL" ~ 73.9, TRUE ~ distance_upstream_km))%>%
  mutate(location_type = if_else(site_code == "PFAL", "Mainstem", location_type))%>%
  # add order table which gives the sites an order to be plotted in
  left_join(order_table, by = "site_code")%>%
  #filter site code BRNR with chla value of 0.02, seems to be sample error
  filter(!(site_code == "BRNR" & ChlA == 0.02))%>%
  #filter site code CBRR with TSS value > 1000, seems to be sample error
  mutate( TSS = case_when((site_code == "CBRR" & TSS > 250) ~ NA_real_,
                          TRUE ~ TSS),
           Turbidity = case_when((site_code == "CBRR" & Turbidity > 250) ~ NA_real_,
                          TRUE ~ Turbidity), 
          Year = as.character(Year), 
          month = as.character(month(Date)),
          season = case_when(
            month %in% c("3", "4") ~ "Rising Limb", 
            month %in% c( "5","6") ~ "Peak Flow",
            month %in% c("7", "8") ~ "Falling Limb",
            month %in% c("9", "10", "11") ~ "Fall Baseflow", 
            TRUE ~ "Winter"),
          burn_group = case_when(site_code %in% c("CBRR", "BRNR", "PTRR") ~ "burned", 
                                 TRUE ~ "unburned"), 
          burn_year = paste(burn_group, Year, sep = "_"), 
          site_year = paste(site_code, Year, sep = "_")
  )


```

## Run above to get data loaded

```{r}

```


There will be five main figures in this paper: - Key nutrient dotplot
(NO3, NH4, K, DOC) No COMR/HORR - Chla dot plot (All sites) - NO3 chla
JWC relationship - SF Chla (on vs off reservoir) - CLP Mainstem
Longitudinal (Chl-a), NO3?

# Nutrient Comparison Between reservoirs

### Functions


```{r}
#Original function with no COMR/HORR and just dotplot with title always
build_res_dotplot<- function(param = "ChlA", binwdth = 0.3, sz = 1, title_min = -.7){
  
    boxplot_res <- filter(tidy_chem, site_code %in% c("LNGR", "PTRR", "BRNR", "JOER", "CBRR") )%>%
      filter(!is.na(.data[[param]]))%>%
      mutate(site_year = paste(site_code, Year, sep = " "),
             dayofyear = yday(Date))
    param_units <- filter(chem_units, simple == param)%>%
      pull(combined)
    
    #Create plot
    ggplot(boxplot_res, aes(x=reorder(site_year,BS_order,max),y=.data[[param]], 
                            #middle=mean(.data[[param]]), 
                            fill=site_code)) + 
      geom_dotplot(stackgroups = FALSE, binaxis = "y", method = "histodot",stackdir = "center", position = "dodge", binwidth = binwdth, dotsize = sz)+
      scale_color_grey(start = 0.4, end = 0)+
      #make x axis titles vertical
      theme(axis.text.x = element_text(angle = 90, hjust = 1))+
      geom_rect( inherit.aes=FALSE, aes(xmin="LNGR 2021", xmax="LNGR 2024", ymin=title_min,
                                        ymax=0), color="transparent", fill="#0072B2", alpha=0.01)+
      geom_rect( inherit.aes=FALSE, aes(xmin="JOER 2021", xmax="JOER 2024", ymin=title_min,
                                        ymax=0), color="transparent", fill="#56B4E9", alpha=0.05)+
      
      geom_rect( inherit.aes=FALSE, aes(xmin="CBRR 2021", xmax="CBRR 2024", ymin=title_min,
                                        ymax= 0), color="transparent", fill="#F9C74F", alpha=0.05)+
      
      geom_rect( inherit.aes=FALSE, aes(xmin="PTRR 2021", xmax="PTRR 2024", ymin=title_min,
                                        ymax=0), color="transparent", fill="#F3722C", alpha=0.05)+
      geom_rect( inherit.aes=FALSE, aes(xmin= 16.5, xmax= 20.5, ymin=title_min,
                                        ymax=0), color="transparent", fill="#d71a1d", alpha=0.008)+
      geom_abline(aes(slope=0, intercept=0), color='#000000', linetype="solid") +
      geom_text(aes(x= 2.5, y = (title_min/2), label= "Long Draw",  color = "#000000"), size = 6, alpha = 1)+
      geom_text(aes(x= 6.5, y = (title_min/2), label= "Joe Wright",  color = "#000000"), size = 6, alpha = 1)+
      geom_text(aes(x= 10.5, y = (title_min/2), label= "Chambers",  color = "#000000"), size = 6, alpha = 1)+
      geom_text(aes(x= 14.5, y = (title_min/2), label= "Peterson",  color = "#000000"), size = 6)+
      geom_text(aes(x= 18.5, y = (title_min/2), label= "Barnes Meadow",  color = "#000000"), size = 5, alpha = 1)+
  
      
      scale_fill_manual(name="Watershed", values = colorsBS_site_code) +
      theme_bw(base_size=25) +
      
      #theme(axis.text.x = ifelse(study_reservoir_chla$Location == "Reservoir",element_text(face = "bold"),element_text(face = "plain"))) +
      labs( x = "Site", y = param_units)+
      scale_x_discrete(labels = c("2021", "2022", "2023","2024",
                                  "","","","",
                                  "2021", "2022", "2023","2024",
                                  "","","","",
                                  "2021", "2022", "2023","2024"))+
      theme(legend.position = "none",
            axis.title = element_text(face = "bold"),
            # bold and angle x axis titles to 45
            axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"))
}
```

```{r}
build_res_plot <- function(param = "ChlA", binwdth = 0.3, sz = 1, title_min = NULL, b_size = 20) {
  # Order: Long Draw, Joe Wright, Hourglass, Chambers, Comanche, Peterson, Barnes Meadow
  sites <- c("LNGR", "JOER", "HORR", "CBRR", "COMR", "PTRR", "BRNR")

  y_label <- chem_expressions[[param]]  


  boxplot_res <- tidy_chem %>%
    filter(site_code %in% sites) %>%
    filter(!is.na(.data[[param]])) %>%
    mutate(site_year = paste(site_code, Year, sep = " "), 
           BS_order = as.numeric(BS_order))
  
  # Inject 6 gaps between the 7 reservoirs
  gaps <- data.frame(
    site_year = paste0("Gap", 1:8),
    site_code = NA, 
    BS_order = c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5) 
  )
  
  plot_data <- bind_rows(boxplot_res, gaps)
  
  # 2. Build Base Plot
  p <- ggplot(plot_data, aes(x = reorder(site_year, BS_order, max), y = .data[[param]], fill = site_code))
  
  # 3. Conditional Annotations
  if (!is.null(title_min)) {
    p <-  p +
      # Site Grouping Backgrounds
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01),xmin = 1.5, xmax = 5.5, fill="#0072B2", alpha=0.01) + #LNGR
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01),xmin = 6.5, xmax = 10.5, fill="#56B4E9", alpha=0.05) + #JOER
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01),xmin = 11.5, xmax = 15.5, fill="#F0E442", alpha=0.05) + #HORR
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01),xmin = 16.5, xmax = 20.5, fill="#F9C74F", alpha=0.05) + #CBRR
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01),xmin = 21.5, xmax = 25.5, fill="#F8961E", alpha=0.05) + #COMR
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01),xmin = 26.5, xmax = 30.5, fill="#F3722C", alpha=0.05) + #PTRR
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01), xmin = 31.5, xmax = 35.5, fill="#d71a1d", alpha=0.008) + #BRNR
      
      # Labels positioned at the center of each 4-year block (allowing for gaps)
      annotate("text", x = 3.5,  y = title_min/2, label = "Long Draw", size = 6) +
      annotate("text", x = 8.5,  y = title_min/2, label = "Joe Wright", size = 6) +
      annotate("text", x = 13.5, y = title_min/2, label = "Hourglass", size = 6) +
      annotate("text", x = 18.5, y = title_min/2, label = "Chambers", size = 6) +
      annotate("text", x = 23.5, y = title_min/2, label = "Comanche", size = 6) +
      annotate("text", x = 28.5, y = title_min/2, label = "Peterson", size = 6) +
      annotate("text", x = 33.5, y = title_min/2, label = "Barnes Meadow", size = 5) +
      geom_dotplot(stackgroups = FALSE, binaxis = "y", method = "histodot",stackdir = "center", position = "dodge", binwidth = binwdth, dotsize = sz)+
      geom_abline(intercept = 0, slope = 0, color = "black") +
      scale_fill_manual(values = colorsBS_site_code) +
      theme_bw(base_size = b_size) +
      labs(y = y_label, x = NULL)+
      theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),
            legend.position = "none") +
      # This function cleans the X-axis: removes site names and hides "Gap" labels
      scale_x_discrete(labels = function(x) ifelse(grepl("Gap", x), "", gsub(".* ", "", x)))+
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)))
    
  } else {
    p <- p+
      geom_dotplot(stackgroups = FALSE, binaxis = "y", method = "histodot",stackdir = "center", position = "dodge", binwidth = binwdth, dotsize = sz)+
      geom_abline(intercept = 0, slope = 0, color = "black") +
      scale_fill_manual(values = colorsBS_site_code) +
      theme_bw(base_size = b_size) +
      labs(y = y_label, x = NULL)+
      theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none"
      )
  }
  
  return(p)
}
```



### Plots with COMR/HORR

```{r}
TSS_dotplot_res <-  build_res_plot(param = "TSS", binwdth = .4, sz = 5.5)+
  ylim(-0.5,50)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black")

K_dotplot_res <-  build_res_plot(param = "K", binwdth = .045, sz = 1.25)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black")

no3_dotplot_res <-   build_res_plot(param = "NO3", binwdth = 0.1, sz = 1.15)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black")+
  ylim(-0.1,2.5)

 DTN_dotplot_res <- build_res_plot(param = "DTN", binwdth = .04, sz = 1.25)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, alpha = 0.7)

DOC_dotplot_res <-  build_res_plot(param = "DOC", binwdth = .35, sz = 1.4, title_min = -2.5)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, alpha = 0.7)

TSS_dotplot_res/K_dotplot_res/no3_dotplot_res/DTN_dotplot_res/DOC_dotplot_res

ggsave("docs/cpf_reservoir/paper_draft/test_res_chem_dotplots.jpg", dpi = 300, width = 18, height = 18)

```

### SI Fig: Extra Dotplots
```{r}

 Ca_dotplot_res <- build_res_plot(param = "Ca", binwdth = .1, sz = 5.5)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black")

SC_dotplot_res <-   build_res_plot(param = "SC", binwdth = 1.25, sz = 1.5)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black")

Mg_dotplot_res <-   build_res_plot(param = "Mg", binwdth = 0.05, sz = 1.15)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black")


pH_dotplot_res <-    build_res_plot(param = "pH", binwdth = .03, sz = 1.4)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, alpha = 0.7)
  
   ANC_dotplot_res <- build_res_plot(param = "ANC", binwdth = 15, sz = 1.25, title_min = -50)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, alpha = 0.7)

SI_plot <- Ca_dotplot_res/Mg_dotplot_res/SC_dotplot_res/pH_dotplot_res/ANC_dotplot_res

ggsave(filename = "docs/cpf_reservoir/paper_draft/SI_res_chem_dotplots.jpg", 
      plot = SI_plot,
       dpi = 300, width = 18, height = 18)

```


### Stats

Reservoirs Only

#### Counts near MDL

```{r}
mdl <- tidy_chem%>%
  filter(site_code %in% primary_study_res) %>%
  pivot_longer(cols = c(Turbidity, TSS, DTN,  DOC:NO3, SO4), 
               names_to = "param", 
               values_to = "value")%>%
  select(site_code, Date, Year, param, value, burn_group,season)%>%
  mutate(burn_group = factor(burn_group, levels = c("unburned", "burned")),
         Year = factor(Year, levels = c("2021", "2022", "2023", "2024")), 
         season = factor(season),
         site_code = factor(site_code, levels = c("LNGR", "JOER", "CBRR", "PTRR", "BRNR")))
mdl%>%
  group_by(param)%>%
  summarise(count_below_mdl = sum(value <= 0.01, na.rm = TRUE), 
            count_below_quant = sum(value <= 0.05, na.rm = TRUE),
            total_count = n(), 
            pct_below_mdl = (count_below_mdl/total_count)*100)
mdl%>%
  filter(param == "NH4")%>%
  group_by(site_code)%>%
  summarise(count_below_mdl = sum(value <= 0.01, na.rm = TRUE), 
            count_above_0.05 = sum(value > 0.05, na.rm = TRUE),
            count_below_quant = sum(value <= 0.05, na.rm = TRUE),
            total_count = n(), 
            pct_below_mdl = (count_below_mdl/total_count)*100, 
            pct_above_quant = (count_above_0.05/total_count)*100)


```


##### Normality

```{r}
chem_transformed_res <- tidy_chem%>%
  filter(site_code %in% primary_study_res) %>%
  pivot_longer(cols = c(Turbidity, TSS, DTN,  DOC:NO3, SO4), 
               names_to = "param", 
               values_to = "value")%>%
  select(site_code, Date, Year, param, value, burn_group,season)%>%
  mutate(burn_group = factor(burn_group, levels = c("unburned", "burned")),
         Year = factor(Year, levels = c("2021", "2022", "2023", "2024")), 
         season = factor(season),
         site_code = factor(site_code, levels = c("LNGR", "JOER", "CBRR", "PTRR", "BRNR")))%>%
  mutate(log_value = log10(value +1 ))%>%
  na.omit()

param_oi <- c( "Turbidity", "TSS", "DTN","NH4", "K", "DOC", "NO3", "SO4")

ggplot(chem_transformed_res%>%filter(param %in% param_oi), aes(x = value, fill = burn_group))+
  geom_histogram(position = "stack")+
  geom_vline(aes(xintercept = 0.05), color = "red", linetype = "dashed")+
  facet_wrap(~param, scales = "free")

ggplot(chem_transformed_res%>%filter(param %in% param_oi), aes(x = log_value, fill = burn_group))+
  geom_histogram(position = "stack")+
  facet_wrap(~param, scales = "free")

```

Nothing without transformation is normal, we should double check model
assumptions with `check_model()` after creating a model then only
evaluate those that do. We will want to use a linear mixed effects model
to compare out burned vs unburned reservoirs

##### Burned vs Unburned

Fixed effects: burn type random effect: site

```{r}
burn_group_check <- map(c( "Turbidity", "TSS", "DTN","NH4", "K", "DOC", "NO3", "SO4"), ~{
  
  #Since NO3 does not fit a normal distribution, we will use a generalized linear mixed model with a gamma distribution and log link for NO3, and linear mixed models for the other parameters.
  if(.x == "NO3"){
        model_gamma <- glmmTMB(value ~ burn_group + (1 | site_code), 
                       data = chem_transformed_res%>%filter(param == .x & value > 0), 
                       family = Gamma(link = "log"))
        
        check_model(model_gamma)
  }
  
   df <- chem_transformed_res%>%filter(param == .x)
   
   
   model_full <- lmer(log_value ~ burn_group + (1|site_code), data = df, REML = FALSE)

    check_model(model_full)

})%>%
  set_names(param_oi)

# If they pass the model diagnostics, extract the estimate for burn_group and calculate the percent difference between burned and unburned reservoirs. Then compare the full model to a null model without burn_group using an ANOVA to get the p-value for the effect of burn_group.


map(c( "Turbidity", "TSS", "DTN", "DOC", "SO4", "K", "NO3"), ~{
  
  if(.x == "NO3"){
    model_gamma <- glmmTMB(value ~ burn_group + (1 | site_code), 
                           data = chem_transformed_res%>%filter(param == .x & value > 0), 
                           family = Gamma(link = "log"))
    
    conf <- confint(model_gamma, parm = "beta_")
    low_pct <- ((exp(conf[2,1])) - 1)*100
    high_pct <- ((exp(conf[2,2])) - 1)*100
    
    estimate <- ((summary(model_gamma)$coefficients$cond["burn_groupburned", "Estimate"]))
    pct_diff <- ((exp(estimate)) - 1)*100
    model_null <- glmmTMB(value ~ 1 + (1 | site_code), 
                          data = chem_transformed_res%>%filter(param == .x & value > 0), 
                          family = Gamma(link = "log"))
    anova_res <- anova(model_null, model_gamma)
    
    tibble(param = "NO3", estimate = estimate, low_pct = low_pct, high_pct = high_pct,
           pct_diff = pct_diff,p_value_anova = anova_res$`Pr(>Chisq)`[2])%>%
      mutate(p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
             significant_effect = ifelse(p_value_anova < 0.05, ifelse(pct_diff > 0, "Higher in Burned", "Lower in Burned"), "No Significant Difference"))
  }
  df <- chem_transformed_res%>%filter(param == .x)
  model_full <- lmer(log_value ~ burn_group + (1|site_code), data = df, REML = FALSE)
  summary(model_full) 
  
  conf <- confint(model_full, parm = "beta_")
  low_pct <- ((exp(conf[2,1])) - 1)*100
  high_pct <- ((exp(conf[2,2])) - 1)*100
  
  estimate <- model_full@beta[2]
  pct_diff <- ((exp(estimate)) - 1)*100
  model_null <- lmer(log_value ~ 1 + (1|site_code), data = df, REML = FALSE)
  anova_res <- anova(model_null, model_full)
  tibble(param = .x, estimate = estimate, low_pct = low_pct, high_pct = high_pct,
         pct_diff = pct_diff,p_value_anova = anova_res$`Pr(>Chisq)`[2])%>%
    mutate(p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
           significant_effect = ifelse(p_value_anova < 0.05, ifelse(pct_diff > 0, "Higher in Burned", "Lower in Burned"), "No Significant Difference"))
})%>%
  bind_rows()%>%
  arrange(low_pct)%>%
  filter(param %in% c("TSS", "K"))


#Levene's test for NO3
lev_model <- leveneTest(log_value ~ burn_group, data = chem_transformed_res%>%filter(param == "NO3" & value > 0))
#check_model(lev_model)
```

##### Year wise (All Sites)

Fixed effects: Year random effect: site

```{r}
year_wise_check <- map(c( "Turbidity", "TSS", "DTN","NH4", "K", "DOC", "NO3", "SO4"), ~{
  if(.x == "NO3"){
    model_gamma <- glmmTMB(value ~ Year + (1 | site_code), 
                           data = chem_transformed_res%>%filter(param == .x & value > 0), 
                           family = Gamma(link = "log"))
    
    check_model(model_gamma)
  }else{
    df <- chem_transformed_res%>%filter(param == .x)
    
    model_full <- lmer(log_value ~ Year + (1|site_code), data = df, REML = FALSE)
    
    check_model(model_full)
  }
  
})%>% set_names(c( "Turbidity", "TSS", "DTN","NH4", "K", "DOC", "NO3", "SO4"))

# If they pass the model diagnostics, extract the estimate for burn_group and calculate the percent difference between burned and unburned reservoirs. Then compare the full model to a null model without burn_group using an ANOVA to get the p-value for the effect of Year


map(c("Turbidity", "TSS", "DTN", "NO3", "DOC", "SO4", "K"), ~{
  
  if(.x == "NO3"){
    # Gamma model for NO3
    model_gamma <- glmmTMB(value ~ Year + (1 | site_code), 
                           data = chem_transformed_res %>% filter(param == .x & value > 0), 
                           family = Gamma(link = "log"))
    
    year_means <- emmeans(model_gamma, "Year")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value, 
             lower.CL = asymp.LCL, 
             upper.CL = asymp.UCL) %>%
      mutate(parameter = .x)
    
    model_null <- glmmTMB(value ~ 1 + (1 | site_code), 
                          data = chem_transformed_res %>% filter(param == .x & value > 0), 
                          family = Gamma(link = "log"))
    
    anova_res <- anova(model_null, model_gamma)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
    
  } else {
    # LMM for other parameters
    df <- chem_transformed_res %>% filter(param == .x)
    model_full <- lmer(log_value ~ Year + (1|site_code), data = df, REML = FALSE)
    
    year_means <- emmeans(model_full, "Year")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value) %>%
      mutate(parameter = .x)
    
    model_null <- lmer(log_value ~ 1 + (1|site_code), data = df, REML = FALSE)
    anova_res <- anova(model_null, model_full)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
  }
  
  # Attach ANOVA p-value to the contrast dataframe
  contrasts %>%
    mutate(p_value_anova = p_anova)
  
}) %>%
  bind_rows() %>%
  # Final calculation for percent changes and significance stars
  mutate(
    pct_est  = (exp(estimate) - 1) * 100,
    pct_low  = (exp(lower.CL) - 1) * 100,
    pct_high = (exp(upper.CL) - 1) * 100,
    p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
    p_sig_contrast = ifelse(p_value_contrast < 0.05, "*", "")
  )%>%
  select(parameter, contrast, pct_est, pct_low, pct_high, p_value_contrast, p_sig_contrast, p_value_anova, p_sig_anova)%>%
  filter(parameter == "SO4")%>%
  arrange(desc(pct_low))
```

##### Site wise (interacton with year)

Fixed effects: Year + Site random effect: season

```{r}

check_model_results <- map(c(param_oi), ~{
  df <- chem_transformed_res%>%filter(param == .x)
  
  model_full <- lmer(log_value ~ site_code + Year + (1|season), data = df, REML = FALSE)
  #summary(model_full) 
  check_model(model_full)
  
})%>%
  set_names(param_oi)

  map(c("NO3", "TSS", "DTN", "K", "DOC","SO4" ), ~{
  
  if(.x == "NO3"){
    # Gamma model for NO3
    model_gamma <- glmmTMB(value ~ Year + site_code + (1 | season), 
                           data = chem_transformed_res %>% filter(param == .x & value > 0), 
                           family = Gamma(link = "log"))
    
    year_means <- emmeans(model_gamma, "site_code")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value, 
             lower.CL = asymp.LCL, 
             upper.CL = asymp.UCL) %>%
      mutate(parameter = .x)
    
    model_null <- glmmTMB(value ~ Year + site_code  + (1 | season), 
                          data = chem_transformed_res %>% filter(param == .x & value > 0), 
                          family = Gamma(link = "log"))
    
    anova_res <- anova(model_null, model_gamma)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
    
  } else {
    # LMM for other parameters
    df <- chem_transformed_res %>% filter(param == .x)
    model_full <- lmer(log_value  ~ Year + site_code  + (1 | season), data = df, REML = FALSE)
    
    year_means <- emmeans(model_full, "site_code")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value) %>%
      mutate(parameter = .x)
    
    model_null <- lmer(log_value ~ Year  + (1 | season), data = df, REML = FALSE)
    anova_res <- anova(model_null, model_full)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
  }
  contrasts%>%
    mutate(p_value_anova = anova_res$`Pr(>Chisq)`[2])
})%>%
  bind_rows() %>%
  # Final calculation for percent changes and significance stars
  mutate(
    pct_est  = (exp(estimate) - 1) * 100,
    pct_low  = (exp(lower.CL) - 1) * 100,
    pct_high = (exp(upper.CL) - 1) * 100,
    p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
    p_sig_contrast = ifelse(p_value_contrast < 0.05, "*", "")
  )%>%
  select(parameter, contrast, pct_est, pct_low, pct_high, p_value_contrast, p_sig_contrast, p_value_anova, p_sig_anova)%>%
    filter(parameter == "DTN")
  

```
# Reservoir Chlorophyll a


## Plot
```{r}

max_comr <- tidy_chem %>%
  filter(site_code == "COMR", !is.na(ChlA)) %>%
  summarise(max_chla = max(ChlA, na.rm = TRUE)) %>%
  pull(max_chla)
max_horr <- tidy_chem %>%
  filter(site_code == "HORR", !is.na(ChlA)) %>%
  summarise(max_chla = max(ChlA, na.rm = TRUE)) %>%
  pull(max_chla)

#chla_dotplot_res <- 
  build_res_plot(param = "ChlA", binwdth = .2, sz = 1, title_min = -0.6)+
  # add medians
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black") +
#adding meso and eutrophic levels for cold lakes
 geom_abline(aes(slope=0, intercept=2.6), color='black', linetype="dashed") +
    geom_abline(aes(slope=0, intercept=8), color='black', linetype="dashed") +
    #annotate(geom="text", label= "MDL Limit (μg/L)", x="CBRR 2022", y=.075, vjust= -1)+
    annotate(geom="text", label= "Mesotrophic", x=1.15, y=2.6, vjust= -1, size = 4)+
    annotate(geom="text", label= "Eutrophic", x= 1.5, y=8, vjust= -1, size = 5)+
  # Add a gray box or label over the "empty" years
 annotate("rect", xmin = 11.5, xmax = 15.5, ymin = 0, ymax = max_horr + 0.25, alpha = 0.2, fill = "gray") +
 annotate("rect", xmin = 13.5, xmax = 15.5, ymin = 0, ymax = max_horr + 0.25, alpha = 0.2, fill = "gray") +
 annotate("text", x = 12.5, y = max_horr + 0.5, label = "Unfilled", size = 3, fontface = "italic")+
  annotate("text", x = 14.5, y = max_horr + 0.5, label = "Filled", size = 3, fontface = "italic")+
 annotate("rect", xmin = 21.5, xmax = 25.5, ymin = 0, ymax = max_comr + 0.25, alpha = 0.2, fill = "gray") +
 annotate("rect", xmin = 23.5, xmax = 25.5, ymin = 0, ymax = max_comr + 0.25, alpha = 0.2, fill = "gray") +
 annotate("text", x = 22.5, y = max_comr + 0.5, label = "Unfilled", size = 3, fontface = "italic")+
    annotate("text", x = 24.5, y = max_comr + 0.5, label = "Filled", size = 3, fontface = "italic")


ggsave("docs/cpf_reservoir/paper_draft/res_chla_dotplots_2021_2024.jpg", dpi = 300, width = 20, height = 12)

```

### NO3 + TDN + Chl-a

```{r}

winter_gaps <- all_reservoirs %>%
  map_df(~{
    seq(as.Date("2020-12-31"), as.Date("2024-12-31"), by = "year") %>%
      as_tibble() %>%
      rename(Date = value) %>%
      mutate(site_code = .x, ChlA = NA, NO3 = NA, DTN = NA)
  })


plot_data <- tidy_chem %>%
  filter(site_code %in% all_reservoirs & !is.na(NO3) & !is.na(ChlA) & !is.na(DTN)) %>%
  bind_rows(winter_gaps) %>%
  pivot_longer(cols = c(NO3, ChlA, DTN), names_to = "parameter", values_to = "value")

make_grid_plot <- function(site, param, y_limit, pos_index, total_sites) {
  
  is_left_edge <- (pos_index == 1)
  is_right_edge <- (pos_index == total_sites)
  show_title <- (param == "DTN") # Titles only on top row
  
  p <- plot_data %>%
    filter(site_code == site, parameter == param) %>%
    ggplot(aes(x = Date, y = value)) +
    geom_line(linewidth = 0.8) +
    geom_point(aes(color = season)) +
    scale_color_manual(values = season_clp_color_vals, na.translate = FALSE) +
    ylim(0, y_limit) +
    theme_bw(base_size = 12) +
    labs(x = NULL, y = chem_expressions[[param]], color = "Season")
  
  # Logic: If NOT on the left or right edge, strip the y-axis text and titles
  if (!is_left_edge && !is_right_edge) {
    p <- p + theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.title.y = element_blank()
    )
  }
  
  # Add Reservoir Titles to the top row only
  if (show_title) {
    p <- p + ggtitle(site)
  }
  
  return(p)
}

n_res <- length(all_reservoirs)

dtn_plots <- imap(all_reservoirs, ~make_grid_plot(.x, "DTN", 1, .y, n_res))
no3_plots <- imap(all_reservoirs, ~make_grid_plot(.x, "NO3", 2.5, .y, n_res))
chla_plots <- imap(all_reservoirs, ~make_grid_plot(.x, "ChlA", 10, .y, n_res))

# 5. Assemble and Collect
# We combine the lists in order: Row 1 (DTN), Row 2 (NO3), Row 3 (ChlA)
final_plot <- wrap_plots(c(dtn_plots, no3_plots, chla_plots), ncol = n_res) +
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

# Display the plot
final_plot
```

### Reservoir N and Chla Hysteresis 

```{r}

#start with BRNR, make a plot of ChlA vs NO3 and ChlA vs DTN for each reservoir separately, facet by year and include a line connecting the points in chronological order to show hysteresis patterns. Use different colors for each year to distinguish them.

 tidy_chem %>%
  filter( !is.na(NO3) & !is.na(ChlA) & !is.na(DTN)) %>%
  filter(site_code == "CBRR") %>%
    mutate(doy = yday(Date), 
           N_NO3 = NO3/ 4.443)%>% #convert from mg/L to mg/L N
  pivot_longer(cols = c(N_NO3, DTN), names_to = "N_param", values_to = "N_value") %>%
  ggplot(aes(x = N_value, y = ChlA, color = doy)) +
  geom_point() +
  #add an arrow connecting the points in chronological order
  geom_path(arrow = arrow(type = "closed", length = unit(0.1 , "inches"))) +
  facet_grid(Year ~ N_param) +
  #color from blue (early season) to red (late season) with a mid point as orange (mid season)
  scale_color_gradient2(low = "#03045e", mid = "#59B851", high = "#FA850E", midpoint = 225) +
  theme_bw(base_size = 12) +
  labs(x = "Nitrogen Parameter Value", y = "Chlorophyll a (μg/L)", color = "DOY")+
  xlim(0,1.25)

```



## Reservoir Monthly Levels

```{r}

#what months had chla > 2.6 and > 8 (count and percent of samples) per year

check <- tidy_chem %>%
  filter(site_code %in% all_reservoirs & !is.na(ChlA)) %>%
   mutate(month = month(Date)) #%>%
  # group_by(site_code, Year, month) %>%
  # summarise(
  #   count_mesotrophic = sum(ChlA > 2.6, na.rm = TRUE),
  #   count_eutrophic = sum(ChlA > 8, na.rm = TRUE),
  #   total_samples = n(),
  #   pct_mesotrophic = (count_mesotrophic / total_samples) * 100,
  #   pct_eutrophic = (count_eutrophic / total_samples) * 100
  # ) %>%
  # arrange(desc(pct_eutrophic))

check %>%
  ggplot(aes(x = ChlA, fill = site_code))+
  geom_histogram()+
  geom_vline(xintercept = 2.6, linetype = "dashed", color = "red")+
  geom_vline(xintercept = 8, linetype = "dashed", color = "blue")+
  facet_wrap(~month)


```


### C:N ratios in reservoirs

```{r}

# Determine how many reservoirs you have to create enough vertical spacing
tidy_chem %>%
  filter(site_code %in% all_res_system & !is.na(DOC) & !is.na(DTN) & !is.na(ChlA)) %>%
  mutate(C_N_ratio = DOC/DTN) %>%
  ggplot(aes(x = DOC, y = ChlA)) + 
  geom_point(aes(color = site_code)) + 
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  # Use aes(label = ..rr.label..) to switch to R-squared
  stat_cor(aes(label = ..rr.label.. , color = site_code), 
           method = "pearson", 
           label.x = .1, 
           label.y = 12, 
           size = 5) + 
  facet_wrap(~location_type) +
  theme_bw(base_size = 20) +
  labs(x = "DOC", y = "Chla")


#Calculate the correlations between DOC and ChlA, DTN and ChlA, and C:N ratio and ChlA for each reservoir separately. Use Pearson correlation and report the correlation coefficient and p-value for each reservoir.

tidy_chem %>%
  filter(site_code %in% all_reservoirs & !is.na(DOC) & !is.na(DTN) & !is.na(ChlA)) %>%
  mutate(C_N_ratio = DOC/DTN) %>%
  group_by(site_code) %>%
  summarise(
    r2_DOC_ChlA = cor(DOC, ChlA, method = "pearson")^2,
    p_DOC_ChlA = cor.test(DOC, ChlA, method = "pearson")$p.value,
    r2_DTN_ChlA = cor(DTN, ChlA, method = "pearson")^2,
    p_DTN_ChlA = cor.test(DTN, ChlA, method = "pearson")$p.value,
    r2_CN_ChlA = cor(C_N_ratio, ChlA, method = "pearson")^2,
    p_CN_ChlA = cor.test(C_N_ratio, ChlA, method = "pearson")$p.value
  )%>%
  mutate(p_significant_DOC_ChlA = ifelse(p_DOC_ChlA < 0.05, "*", ""),
         p_significant_DTN_ChlA = ifelse(p_DTN_ChlA < 0.05, "*", ""),
         p_significant_CN_ChlA = ifelse(p_CN_ChlA < 0.05, "*", ""))


# Make a plot of C:N ratios for each reservoir over time, facet by year and include ChlA for that reservoir on the secondary axis

# 1. Filter for the specific site first to determine relevant ranges
library(scales)
library(tidyverse)

library(scales)
library(tidyverse)
library(lubridate)

# 1. Filter and ensure a "complete" sequence of dates
# We use 'complete' to create rows for every month in the range 
# so that R sees the winter gaps as NAs and breaks the line.
library(scales)
library(tidyverse)
library(lubridate)

# 1. Prepare data for all reservoirs
all_sites_data <- tidy_chem %>%
  filter(site_code %in% all_reservoirs & ChlA < 15) %>%
  # Grouping by site_code is essential so complete() works per-reservoir
  group_by(site_code) %>%
  complete(Date = seq.Date(min(Date), max(Date), by = "month")) %>%
  # Data processing steps
  mutate(
    NO3_N = NO3 * 0.226, # Explicitly naming this to avoid confusion
    C_N_ratio = DOC / DTN,
    month = month(Date),
    # QA Check for nitrogen species
    qa_check = if_else(DTN < NO3_N, "Check DTN", "OK")
  ) %>%
  # Keep winter NAs to break lines, otherwise keep sampled rows
  filter(month %in% c(11, 12, 1, 2) | (!is.na(NO3) & month %in% c(3:10))) %>%
  ungroup()

# 2. Determine global ranges (using all data to keep secondary axis consistent)
prim_range <- range(c(all_sites_data$DTN, all_sites_data$NO3_N), na.rm = TRUE)
sec_range <- range(all_sites_data$ChlA, na.rm = TRUE)

# 3. Plot with Faceting
ggplot(all_sites_data, aes(x = Date)) +
  # --- Primary Axis (N Species) ---
  geom_line(aes(y = DTN, color = "DTN"), linewidth = 0.8) +
  geom_line(aes(y = NO3_N, color = "NO3"), linewidth = 0.8) +
  
  # --- Secondary Axis (ChlA) ---
  geom_line(aes(y = rescale(ChlA, to = prim_range, from = sec_range), color = "ChlA"), 
            linetype = "solid", linewidth = 0.8) +
  
  # Individual data points
  geom_point(aes(y = DTN, color = "DTN"), size = 1) +
  geom_point(aes(y = NO3_N, color = "NO3"), size = 1) +
  
  # Facet per reservoir
  facet_wrap(~site_code, ncol = 2, scales = "free_x") +
  
  scale_color_manual(
    name = "Parameter",
    values = c("DTN" = "#ED553B", "NO3" = "#20639B", "ChlA" = "#3CAEA3")
  ) +
  
  scale_y_continuous(
    name = "Nitrogen Concentration (mg/L)",
    sec.axis = sec_axis(
      ~ rescale(., to = sec_range, from = prim_range), 
      name = "Chlorophyll a (μg/L)"
    )
  ) +
  
  theme_bw(base_size = 14) +
  theme(
    legend.position = "bottom",
    axis.title.y.right = element_text(color = "#3CAEA3"),
    strip.background = element_rect(fill = "white"), # Clean facet headers
    strip.text = element_text(face = "bold")
  ) +
  labs(
    title = "Reservoir Nutrient & Algal Dynamics",
    subtitle = "Nitrogen as N (mg/L) | Lines break during winter sampling gaps",
    x = "Date"
  )

```


## Stats

#### ChlA Counts

##### Max min

```{r}
 tidy_chem%>%
  filter(!is.na(ChlA))%>%
  filter(site_code %in% primary_study_res)%>%
  summarise(mean_chla = mean(ChlA), 
            median_chla = median(ChlA), 
            min_chla = min(ChlA), 
            max_chla = max(ChlA), 
            sd_chla = sd(ChlA))
```

##### Counts Oligo/Meso/Eut

```{r}
tidy_chem%>%
  filter(!is.na(ChlA))%>%
  filter(site_code %in% primary_study_res)%>%
  mutate(eu_status  = case_when(ChlA >= 2.6 &ChlA <= 8 ~ "Mesotrophic", 
                                ChlA > 8 ~ "Eutrophic",
                                ChlA < 2.6 ~ "Oligotrophic"))%>%
  group_by(site_code,eu_status)%>%
  summarise(eu_stats = n())%>%
  ungroup()%>%
  pivot_wider(names_from = eu_status, values_from = eu_stats)%>%
  mutate_at(vars(-site_code), ~replace(., is.na(.), 0))%>%
  rowwise()%>%
  mutate(total_samples = sum(c(Eutrophic, Mesotrophic, Oligotrophic)),
    perc_eu = round((Eutrophic/total_samples)*100, digits= 1), 
         perc_meso = round((Mesotrophic/total_samples)*100, digits= 1),
         perc_meso_eu = round(((Eutrophic+Mesotrophic)/total_samples)*100, digits= 1), 
        perc_oligo = round((Oligotrophic/total_samples) *100, digits= 1))

tidy_chem%>%
  filter(!is.na(ChlA))%>%
  filter(site_code %in% primary_study_res)%>%
  mutate(eu_status  = case_when(ChlA >= 2.6 &ChlA <= 8 ~ "Mesotrophic", 
                                ChlA > 8 ~ "Eutrophic",
                                ChlA < 2.6 ~ "Oligotrophic"))%>%
  group_by(burn_group, eu_status)%>%
  summarise(eu_stats = n())%>%
  ungroup()%>%
  pivot_wider(names_from = eu_status, values_from = eu_stats)%>%
  mutate_at(vars(-burn_group), ~replace(., is.na(.), 0))%>%
  rowwise()%>%
  mutate(total_samples = sum(c(Eutrophic, Mesotrophic, Oligotrophic)),
    perc_eu = round((Eutrophic/total_samples)*100, digits= 1), 
         perc_meso = round((Mesotrophic/total_samples)*100, digits= 1),
         perc_meso_eu = round(((Eutrophic+Mesotrophic)/total_samples)*100, digits= 1), 
        perc_oligo = round((Oligotrophic/total_samples) *100, digits= 1))



tidy_chem%>%
  filter(!is.na(ChlA))%>%
  filter(site_code %in% primary_study_res)%>%
  mutate(eu_status  = case_when(ChlA >= 2.6 &ChlA <= 8 ~ "Mesotrophic", 
                                ChlA > 8 ~ "Eutrophic",
                                ChlA < 2.6 ~ "Oligotrophic"))%>%
  group_by( Year, eu_status, burn_group)%>%
  summarise(eu_stats = n())%>%
  ungroup()%>%
  pivot_wider(names_from = eu_status, values_from = eu_stats)%>%
  mutate_at(vars(-Year), ~replace(., is.na(.), 0))%>%
  rowwise()%>%
  mutate(total_samples = sum(c(Eutrophic, Mesotrophic, Oligotrophic)),
    perc_eu = round((Eutrophic/total_samples)*100, digits= 1), 
         perc_meso = round((Mesotrophic/total_samples)*100, digits= 1),
         perc_meso_eu = round(((Eutrophic+Mesotrophic)/total_samples)*100, digits= 1), 
        perc_oligo = round((Oligotrophic/total_samples) *100, digits= 1),
    Year = as.character(Year))


tidy_chem%>%
  filter(!is.na(ChlA))%>%
  filter(site_code %in% primary_study_res)%>%
  mutate(eu_status  = case_when(ChlA >= 2.6 &ChlA <= 8 ~ "Mesotrophic", 
                                ChlA > 8 ~ "Eutrophic",
                                ChlA < 2.6 ~ "Oligotrophic"))%>%
  group_by(burn_group, eu_status)%>%
  summarise(eu_stats = n())%>%
  ungroup()%>%
  pivot_wider(names_from = eu_status, values_from = eu_stats)%>%
  mutate_at(vars(-burn_group), ~replace(., is.na(.), 0))%>%
  rowwise()%>%
  mutate(total_samples = sum(c(Eutrophic, Mesotrophic, Oligotrophic)),
    perc_eu = round((Eutrophic/total_samples)*100, digits= 1), 
         perc_meso = round((Mesotrophic/total_samples)*100, digits= 1),
         perc_meso_eu = round(((Eutrophic+Mesotrophic)/total_samples)*100, digits= 1), 
        perc_oligo = round((Oligotrophic/total_samples) *100, digits= 1))

```

### Prep Data

```{r}

chem_transformed_res <- tidy_chem%>%
    filter(site_code %in% primary_study_res) %>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4), 
               names_to = "param", 
               values_to = "value")%>%
  select(site_code, Date, Year, param, value, burn_group,season)%>%
    mutate(burn_group = factor(burn_group, levels = c("unburned", "burned")),
           Year = factor(Year, levels = c("2021", "2022", "2023", "2024")), 
           season = factor(season),
           site_code = factor(site_code, levels = c("LNGR", "JOER", "CBRR", "PTRR", "BRNR")))%>%
  mutate(log_value = log10(value +1 ))%>%
  na.omit()

param_oi <- c("ChlA")

ggplot(chem_transformed_res%>%filter(param %in% param_oi), aes(x = value, fill = burn_group))+
  geom_histogram()+
  facet_wrap(~param, scales = "free")

ggplot(chem_transformed_res%>%filter(param %in% param_oi), aes(x = log_value, fill = burn_group))+
  geom_histogram()+
  facet_wrap(~param, scales = "free")

```

Nothing without transformation is normal, we should double check model
assumptions with `check_model()` after creating a model then only
evaluate those that do. We will want to use a linear mixed effects model
to compare out burned vs unburned reservoirs

##### Burned vs Unburned

Fixed effects: burn type random effect: site

```{r}
# Check for model assumptions for ChlA
df <- chem_transformed_res%>%filter(param == "ChlA")
model_full <- lmer(log_value ~ burn_group + (1|site_code), data = df, REML = FALSE)
check_model(model_full)


# If they pass the model diagnostics, extract the estimate for burn_group and calculate the percent difference between burned and unburned reservoirs. Then compare the full model to a null model without burn_group using an ANOVA to get the p-value for the effect of burn_group.

model_full <- lmer(log_value ~ burn_group + (1|site_code), data = df, REML = FALSE)

conf <- confint(model_full, parm = "beta_")
low_pct <- ((exp(conf[2,1])) - 1)*100
high_pct <- ((exp(conf[2,2])) - 1)*100

estimate <- model_full@beta[2]
pct_diff <- ((exp(estimate)) - 1)*100
model_null <- lmer(log_value ~ 1 + (1|site_code), data = df, REML = FALSE)
anova_res <- anova(model_null, model_full)

tibble(param = "ChlA", estimate = estimate, low_pct = low_pct, high_pct = high_pct,
       pct_diff = pct_diff, p_value = anova_res$`Pr(>Chisq)`[2])%>%
  mutate(p_sig = ifelse(p_value < 0.05, "*", ""))



```

##### Year wise (All Sites)

Fixed effects: Year random effect: site

```{r}
map(c("ChlA"), ~{
  df <- chem_transformed_res%>%filter(param == .x)
  
  model_full <- lmer(log_value ~ Year + (1 | site_code), data = df, REML = FALSE)
  
  check_model(model_full)
  
})

# If they pass the model diagnostics, extract the estimate for burn_group and calculate the percent difference between burned and unburned reservoirs. Then compare the full model to a null model without burn_group using an ANOVA to get the p-value for the effect of Year


map(c("ChlA"), ~{
  df <- chem_transformed_res%>%filter(param == .x)
  model_full <- lmer(log_value ~ Year + (1 | site_code), data = df, REML = FALSE)
  
  # Use type = "response" to get back-transformed estimates
  year_means <- emmeans(model_full, "Year", type = "response")
  
  contrasts <- summary(pairs(year_means), infer = TRUE) %>%
    as.data.frame() %>%
    rename(p_value_contrast = p.value)%>%
    mutate(
      pct_est  = (exp(estimate) - 1) * 100,
      pct_low  = (exp(lower.CL) - 1) * 100,
      pct_high = (exp(upper.CL) - 1) * 100)%>%
    select(contrast, pct_est, pct_low, pct_high, p_value_contrast)
  
  model_null <- lmer(log_value ~ 1  + (1 | season), data = df, REML = FALSE)
  anova_res <- anova(model_null, model_full)
  
  return(contrasts %>%
           mutate(p_value_anova = anova_res$`Pr(>Chisq)`[2],
                  parameter = .x))
})%>%
  bind_rows()%>%
  mutate(
    # Since we used type="response", pct_ratio is already exp(estimate)
    p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
    p_sig_contrast = ifelse(p_value_contrast < 0.05, "*", "")
  )

```

##### Site wise (interacton with year)

Fixed effects: Year + Site random effect: season

```{r}

map(c("ChlA"), ~{
  df <- chem_transformed_res%>%filter(param == .x)
  
  model_full <- lmer(log_value ~ site_code + Year + (1|season), data = df, REML = FALSE)
  
  check_model(model_full)
  
})

map(c("ChlA" ), ~{
  df <- chem_transformed_res%>%filter(param == .x)
  model_full <- lmer(log_value ~ site_code + Year + (1|season), data = df, REML = FALSE)
  summary(model_full) 
  
  site_means <- emmeans(model_full, "site_code")
  
  contrasts <- summary(pairs(site_means), infer = TRUE) %>%
    as.data.frame() %>%
    rename(p_value_contrast = p.value)%>%
    mutate(
      pct_est  = (exp(estimate) - 1) * 100,
      pct_low  = (exp(lower.CL) - 1) * 100,
      pct_high = (exp(upper.CL) - 1) * 100)%>%
    select(contrast, pct_est, pct_low, pct_high, p_value_contrast)
  
  model_null <- lmer(log_value ~ 1 + Year + (1|season), data = df, REML = FALSE)
  anova_res <- anova(model_null, model_full)
  
  return(contrasts %>%
           mutate(p_value_anova = anova_res$`Pr(>Chisq)`[2],
                  parameter = .x))
})%>%
  bind_rows()%>%
  mutate(
    # Since we used type="response", pct_ratio is already exp(estimate)
    p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
    p_sig_contrast = ifelse(p_value_contrast < 0.05, "*", "")
  )
```


# Nutrient/ChlA Interaction 

## Plots

```{r}
#Seasonal Color values
season_clp_color_vals <- c("Winter" = "#03045e",  "Rising Limb" = '#047E82',  "Peak Flow" = '#59B851', "Falling Limb"='#FA850E', "Fall Baseflow"  = "#772f1a")

jwc_longitudinal <- tidy_chem %>%
  dplyr::filter(!is.na(ChlA)) %>%
  dplyr::filter(site_code %in% c("JOEI", "JOER", "CBRR", "CBRI", "CHD", "JWC"))%>%
  select(Date,site_code, ChlA, NO3, DTN,DOC, Year,season, month,location_type)%>%
  mutate(location_type = ifelse(location_type == "Outflow"|location_type =="Inflow", "Stream", "Reservoir"))

jwc_longitudinal$site_code <- factor(jwc_longitudinal$site_code, levels = c("JOEI", "JOER", "CBRI", "CBRR", "CHD", "JWC"))
jwc_longitudinal$season <- factor(jwc_longitudinal$season, levels = c("Winter", "Rising Limb", "Peak Flow", "Falling Limb", "Fall Baseflow"))
jwc_longitudinal$month <- factor(jwc_longitudinal$month, levels = c("4", "5", "6", "7", "8", "9", "10","11"))
# Calculate the medians for ChlA
median_data <- jwc_longitudinal %>%
  pivot_longer(cols = c(ChlA, NO3, DTN, DOC), names_to = "parameter", values_to = "value") %>%
  group_by(site_code, Year, location_type, parameter) %>%
  summarise(median = median(value, na.rm = TRUE))%>%
  ungroup()%>%
  pivot_wider(names_from = parameter, values_from = median, names_prefix = "median_") %>%
  add_row(site_code = "JWC", Year = "2021", location_type = "Stream", median_ChlA = NA, median_NO3 = NA)

median_data$site_code <- factor(median_data$site_code, levels = c("JOEI", "JOER", "CBRI", "CBRR", "CHD", "JWC"))


jwc_long_plot <- function(param = "ChlA", dot_size = 3, b_size = 20, facet_scales = "fixed"){
  
  med_col_name <- paste0("median_", param)
# Create the plot
ggplot() + 
  geom_point(data = jwc_longitudinal, aes(x = site_code, y = .data[[param]], color = season),
             size = dot_size, position = position_jitterdodge(jitter.height = 0))+
  geom_rect(data = median_data %>% filter(site_code %in% c("JOER", "CBRR")),
            aes(xmin = as.numeric(site_code) - 0.4, xmax = as.numeric(site_code) + 0.4, ymin = - Inf, ymax = Inf, group = Year),
            fill = "#59B851", alpha =0.2)+
  geom_rect(data = median_data %>% filter(site_code %in% c("JWC")),
            aes(xmin = as.numeric(site_code) - 0.4, xmax = as.numeric(site_code) + 0.4, ymin = - Inf, ymax = Inf, group = Year),
            fill = "#002FA7", alpha =0.2)+
  scale_color_manual(values = season_clp_color_vals)+
  # Add lines connecting the medians
  geom_line(data = median_data, aes(x = site_code, y = .data[[med_col_name]], group = Year), color = "black", linewidth = 1) +
  geom_point(data = median_data, aes(x = site_code, y = .data[[med_col_name]], group = Year, shape = "Yearly Site\n Median"), color = "black", size = dot_size) +
  scale_shape_manual(name = "", values = c("Yearly Site\n Median" = 16)) +  # 16 = solid circle
  guides(shape = guide_legend(override.aes = list(color = "black", size = 3))) +
  theme_bw(base_size = b_size)+
  facet_wrap(~Year, ncol = 1, scales = facet_scales)+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right")+
  scale_x_discrete(labels = c( "JOEI"= "  Above\nJoe Wright", 
                               "JOER" = "Joe Wright\nReservoir", 
                               "CBRI" = "Above\nChambers",
                               "CBRR" = "Chambers\nReservoir", 
                               "CHD" = "Below\nChambers", 
                               "JWC" = "Joe Wright Creek"))+
  labs(y = chem_expressions[[param]], 
       color = "Season")
}

jwc_long_chla <- jwc_long_plot(param = "ChlA", dot_size = 3, b_size = 24)
jwc_long_no3 <-  jwc_long_plot(param = "NO3", dot_size = 3, b_size = 24)
jwc_long_dtn <-  jwc_long_plot(param = "DTN", dot_size = 3, b_size = 20)
jwc_long_doc <-  jwc_long_plot(param = "DOC", dot_size = 3, b_size = 20)                               

jwc_long_chla+jwc_long_no3+
  plot_layout(guides = 'collect', axes = "collect")+
  #add color legend to top
  plot_annotation( tag_levels = 'a')& 
  theme(legend.position = "right")

ggsave("docs/cpf_reservoir/paper_draft/jwc_longitudinal_plots.jpg", dpi = 300, width = 18, height = 10)
```



-   Do reservoirs transform the nutrient signal?

    -   Test of variability in ChlA/NO3 between sites

        -   Levene’s test between inflow and reservoir, and reservoir
            and outflo

        -   Test of means/medians between inflow and reservoir, and reservoir and
            outflow (ChlA and NO3)

        -   Test of means/medians between streams/reservoirs (ChlA and
            NO3)

            -   Linear mixed effect model (log_value \~ site_code +
                season + (1\|year)

-   Longitudinal Plot of ChlA + NO3

## Stats

```{r}
jwc_longitudinal_stats  <- tidy_chem %>%
  dplyr::filter(!is.na(ChlA)) %>%
  dplyr::filter(site_code %in% c("JOEI", "JOER", "CBRR", "CBRI", "CHD", "JWC"))%>%
  select(Date,site_code, ChlA, NO3, DTN,DOC,K,Year,season, month,location_type)%>%
  mutate(location_type = ifelse(location_type == "Outflow"|location_type =="Inflow", "Stream", "Reservoir"))%>%
  pivot_longer(cols = c(ChlA, NO3, DOC, DTN, K), 
               names_to = "param", 
               values_to = "value")%>%
  mutate(log_value = log10(value + 1), 
         loc_type = factor(case_when(site_code %in% c("JOEI", "CBRI") ~ "Inflow",
                              site_code %in% c("JOER", "CBRR") ~ "Reservoir",
                              site_code %in% c( "CHD") ~ "Outflow", 
                              site_code == "JWC" ~ "Downstream"), levels = c("Inflow", "Reservoir", "Outflow", "Downstream")), 
         watershed = factor(case_when(site_code %in% c("JOEI", "JOER", "JOEO") ~ "Joe Wright",
                              site_code %in% c("CBRI", "CBRR", "CHD") ~ "Chambers",
                              site_code %in% c( "JWC") ~ "Joe Wright Creek"), levels = c("Joe Wright", "Chambers", "Joe Wright Creek")), 
         Year = factor(Year, levels = c("2021", "2022", "2023", "2024")),
         season = factor(season, levels = c("Winter", "Rising Limb", "Peak Flow", "Falling Limb", "Fall Baseflow")))


ggplot(jwc_longitudinal_stats, aes(x = log_value, fill = loc_type))+
  geom_histogram()+
  facet_grid(~param, scales = "free")

```

### JOER
Test inflow vs outflow: variation and means

Linear relationship between no3 & Chl-a 

```{r}

joer_longitudinal_stats<- jwc_longitudinal_stats %>%
   filter( site_code %in% c("JOEI","JOER", "CBRI"))

site_check <- map(c("DTN", "K", "DOC", "NO3", "ChlA"), ~{
  
  #Since NO3 does not fit a normal distribution, we will use a generalized linear mixed model with a gamma distribution and log link for NO3, and linear mixed models for the other parameters.
  if(.x == "NO3"){
        model_gamma <- glmmTMB(value ~ site_code + season + (1 | Year), 
                       data = joer_longitudinal_stats%>%filter(param == .x & value > 0), 
                       family = Gamma(link = "log"))
        
        return(check_model(model_gamma))
  }
  
   df <- joer_longitudinal_stats%>%filter(param == .x)
   
   
   model_full <- lmer(log_value ~ site_code + season + (1|Year), data = df, REML = FALSE)

    check_model(model_full)
  })%>%
  set_names(c("DTN", "K", "DOC", "NO3","ChlA"))

# If they pass the model diagnostics, extract the estimate for burn_group and calculate the percent difference between burned and unburned reservoirs. Then compare the full model to a null model without burn_group using an ANOVA to get the p-value for the effect of burn_group.


map(c("DTN", "K", "DOC", "NO3", "ChlA"), ~{
  df <- joer_longitudinal_stats %>% filter(param == .x & value > 0) 
  
  if(.x == "NO3"){
    # Gamma model for NO3
    model_full <- glmmTMB(value ~ site_code + season + (1 | Year), 
                           data = df, 
                           family = Gamma(link = "log"))
    
    # Use type = "response" to get ratios directly
    year_means <- emmeans(model_full, "site_code", type = "response")
    
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value)%>%
      #gamma returns ratios, so we can directly calculate percent change from 1
      mutate(pct_est  = (ratio - 1) * 100,
             pct_reduction = (1 - (1 / ratio)) * 100,
             pct_low  = (asymp.LCL - 1) * 100,
             pct_high = (asymp.UCL - 1) * 100)%>%
      select(contrast, pct_est, pct_reduction, pct_low, pct_high, p_value_contrast)
    
    model_null <- glmmTMB(value ~ season + (1 | Year), 
                          data = df, 
                          family = Gamma(link = "log"))
    
    anova_res <- anova(model_null, model_full)
    
    return(contrasts %>%mutate(p_value_anova = anova_res$`Pr(>Chisq)`[2],
           parameter = .x))
    
  } else {
    # LMM for other parameters
    model_full <- lmer(log_value ~ site_code + season + (1|Year), data = df, REML = FALSE)
    
    # Use type = "response" to get back-transformed estimates
    year_means <- emmeans(model_full, "site_code", type = "response")
    
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value)%>%
      mutate(
        pct_est  = (exp(estimate) - 1) * 100,
        pct_low  = (exp(lower.CL) - 1) * 100,
        pct_high = (exp(upper.CL) - 1) * 100)%>%
      select(contrast, pct_est, pct_low, pct_high, p_value_contrast)
    
    model_null <- lmer(log_value ~ season + (1|Year), data = df, REML = FALSE)
    anova_res <- anova(model_null, model_full)
    
    return(contrasts %>%
    mutate(p_value_anova = anova_res$`Pr(>Chisq)`[2],
           parameter = .x))
  }
}) %>%
  bind_rows() %>%
  mutate(
    # Since we used type="response", pct_ratio is already exp(estimate)
    p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
    p_sig_contrast = ifelse(p_value_contrast < 0.05, "*", "")
  ) %>%
  select(parameter, contrast, pct_est,pct_reduction, pct_low, pct_high, p_value_contrast, p_sig_contrast, p_value_anova, p_sig_anova)



#perform levene's test between groups
leveneTest(log_value ~ site_code, data = joer_longitudinal_stats%>%filter(param == "ChlA"))

leveneTest(log_value ~ site_code, data = joer_longitudinal_stats%>%filter(param == "NO3"))

leveneTest(log_value ~ site_code, data = joer_longitudinal_stats%>%filter(param == "DTN"))
```

### CBRR
Test inflow vs outflow: variation and means

Linear relationship between no3 & Chl-a 

```{r}

cbrr_longitudinal_stats<- jwc_longitudinal_stats %>%
   filter( site_code %in% c("CBRI","CBRR", "CHD"))

site_check <- map(c("DTN", "K", "DOC", "NO3", "ChlA"), ~{
  df <- cbrr_longitudinal_stats%>%filter(param == .x & value > 0)
  #Since NO3 does not fit a normal distribution, we will use a generalized linear mixed model with a gamma distribution and log link for NO3, and linear mixed models for the other parameters.
  if(.x == "NO3"){
    model_gamma <- glmmTMB(value ~ site_code + season + (1 | Year), 
                           data = df, 
                           family = Gamma(link = "log"))
    
    return(check_model(model_gamma))
  }
  model_full <- lmer(log_value ~ site_code + season + (1|Year), data = df, REML = FALSE)
  
  check_model(model_full)
})%>%
  set_names(c("DTN", "K", "DOC", "NO3","ChlA"))

# If they pass the model diagnostics, extract the estimate for burn_group and calculate the percent difference between burned and unburned reservoirs. Then compare the full model to a null model without burn_group using an ANOVA to get the p-value for the effect of burn_group.

map(c("DTN", "K", "DOC", "NO3", "ChlA"), ~{
  df <- cbrr_longitudinal_stats %>% filter(param == .x & value > 0) 
  
  if(.x == "NO3"){
    # Gamma model for NO3
    model_full <- glmmTMB(value ~ site_code + season + (1 | Year), 
                           data = df, 
                           family = Gamma(link = "log"))
    
    # Use type = "response" to get ratios directly
    year_means <- emmeans(model_full, "site_code", type = "response")
    
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value)%>%
      #gamma returns ratios, so we can directly calculate percent change from 1
      mutate(pct_est  = (ratio - 1) * 100,
             pct_reduction = (1 - (1 / ratio)) * 100,
             pct_low  = (asymp.LCL - 1) * 100,
             pct_high = (asymp.UCL - 1) * 100)%>%
      select(contrast, pct_est, pct_reduction, pct_low, pct_high, p_value_contrast)
    
    model_null <- glmmTMB(value ~ season + (1 | Year), 
                          data = df, 
                          family = Gamma(link = "log"))
    
    anova_res <- anova(model_null, model_full)
    
    return(contrasts %>%mutate(p_value_anova = anova_res$`Pr(>Chisq)`[2],
           parameter = .x))
    
  } else {
    # LMM for other parameters
    model_full <- lmer(log_value ~ site_code + season + (1|Year), data = df, REML = FALSE)
    
    # Use type = "response" to get back-transformed estimates
    year_means <- emmeans(model_full, "site_code", type = "response")
    
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value)%>%
      mutate(
        pct_est  = (exp(estimate) - 1) * 100,
        pct_low  = (exp(lower.CL) - 1) * 100,
        pct_high = (exp(upper.CL) - 1) * 100)%>%
      select(contrast, pct_est, pct_low, pct_high, p_value_contrast)
    
    model_null <- lmer(log_value ~ season + (1|Year), data = df, REML = FALSE)
    anova_res <- anova(model_null, model_full)
    
    return(contrasts %>%
    mutate(p_value_anova = anova_res$`Pr(>Chisq)`[2],
           parameter = .x))
  }
}) %>%
  bind_rows() %>%
  mutate(
    # Since we used type="response", pct_ratio is already exp(estimate)
    p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
    p_sig_contrast = ifelse(p_value_contrast < 0.05, "*", "")
  ) %>%
  select(parameter, contrast, pct_est,pct_reduction, pct_low, pct_high, p_value_contrast, p_sig_contrast, p_value_anova, p_sig_anova)
  
filter(parameter %in% c( "ChlA", "NO3", "DTN"))


#perform levene's test between groups
leveneTest(log_value ~ site_code, data = cbrr_longitudinal_stats%>%filter(param == "ChlA"))

leveneTest(log_value ~ site_code, data = cbrr_longitudinal_stats%>%filter(param == "NO3"))

leveneTest(log_value ~ site_code, data = cbrr_longitudinal_stats%>%filter(param == "DTN"))


```

### CHD vs JWC 

```{r}
chd_jwc_stats<- jwc_longitudinal_stats %>%
   filter( site_code %in% c( "CHD", "JWC"))

site_check <- map(c("DTN", "K", "DOC", "NO3", "ChlA"), ~{
  df <- chd_jwc_stats%>%filter(param == .x & value > 0)
  #Since NO3 does not fit a normal distribution, we will use a generalized linear mixed model with a gamma distribution and log link for NO3, and linear mixed models for the other parameters.
  if(.x == "NO3"){
    model_gamma <- glmmTMB(value ~ site_code + season + (1 | Year), 
                           data = df, 
                           family = Gamma(link = "log"))
    
    return(check_model(model_gamma))
  }
  model_full <- lmer(log_value ~ site_code + season + (1|Year), data = df, REML = FALSE)
  
  check_model(model_full)
})%>%
  set_names(c("DTN", "K", "DOC", "NO3","ChlA"))

# If they pass the model diagnostics, extract the estimate for burn_group and calculate the percent difference between burned and unburned reservoirs. Then compare the full model to a null model without burn_group using an ANOVA to get the p-value for the effect of burn_group.

map(c("DTN", "K", "DOC", "NO3","ChlA"), ~{
  df <- chd_jwc_stats%>%filter(param == .x & value > 0) 
  
  if(.x == "NO3"){
    # Gamma model for NO3
    model_gamma <- glmmTMB(value ~ site_code + season + (1 | Year), 
                           data = df, 
                           family = Gamma(link = "log"))
    
    year_means <- emmeans(model_gamma, "site_code")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value, 
             lower.CL = asymp.LCL, 
             upper.CL = asymp.UCL)
    
    model_null <- glmmTMB(value ~  season + (1 | Year), 
                          data = df, 
                          family = Gamma(link = "log"))
    
    anova_res <- anova(model_null, model_gamma)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
    
  } else {
    # LMM for other parameters
    model_full <- lmer(log_value ~ site_code + season + (1|Year), data = df, REML = FALSE)
    
    year_means <- emmeans(model_full, "site_code")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value)
    
    model_null <- lmer(log_value ~ season + (1|Year), data = df, REML = FALSE)
    anova_res <- anova(model_null, model_full)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
  }
  contrasts%>%
    mutate(p_value_anova = anova_res$`Pr(>Chisq)`[2],
           parameter = .x)
})%>%
  bind_rows()%>%
  # Final calculation for percent changes and significance stars
  mutate(
    pct_est  = (exp(estimate) - 1) * 100,
    pct_low  = (exp(lower.CL) - 1) * 100,
    pct_high = (exp(upper.CL) - 1) * 100,
    p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
    p_sig_contrast = ifelse(p_value_contrast < 0.05, "*", "")
  )%>%
  select(parameter, contrast, pct_est, pct_low, pct_high, p_value_contrast, p_sig_contrast, p_value_anova, p_sig_anova)#%>%
  filter(parameter %in% c( "ChlA", "NO3", "DTN"))


#perform levene's test between groups
leveneTest(log_value ~ site_code, data = chd_jwc_stats%>%filter(param == "ChlA"))

leveneTest(log_value ~ site_code, data = chd_jwc_stats%>%filter(param == "NO3"))

leveneTest(log_value ~ site_code, data = chd_jwc_stats%>%filter(param == "DTN"))
```

# CLP Longitudinal
Show how reservoir impact dies as additional tributaries dilute the signal. We can also show how the reservoirs impact changes across seasons (minimal). 

## Plot

```{r}


clp_long <- tidy_chem %>%
  filter(!is.na(ChlA)& Year %in% c(2022, 2023, 2024)) %>%
  filter(site_code %in% c("JOEI", "JOER", "CBRR", "CBRI", "CHD", "JWC", "PFAL", "SLEP", "PBR", "PSF","PMAN", "PNF", "PBD", "SFM", "PJW"))%>%
  select(site_code,Date, watershed, distance_upstream_km, ChlA, NO3, PO4, K, location_type,Year, month, season )%>%
  mutate(location_type = case_when(
    location_type %in% c("Inflow", "Outflow") ~ "Res Stream", 
    location_type %in% c("Reservoir") ~ "Reservoir",
    TRUE ~ "Mainstem"),
    Year = as.character(Year),
    dist_pbd = as.character(distance_upstream_km),
    site_code  = case_when(site_code == "SLEP" ~ "PFAL",
                           site_code == "PMAN" ~ "PNF",
                           TRUE ~ site_code),
    distance_upstream_km = case_when(site_code == "PFAL" ~ 73.161, 
                                     site_code == "PNF" ~ 7.47,
                                     site_code == "PJW" ~ 77, 
                                      site_code == "SFM" ~ 40,
                                     TRUE ~ distance_upstream_km),
    type = ifelse(location_type %in% c("Inflow","Outflow"),"Stream",
                  ifelse(location_type == "Reservoir", "Reservoir", "Mainstem")))


# Calculate the medians for ChlA
clp_median_data <- clp_long %>%
  group_by(site_code,type,distance_upstream_km, location_type) %>%
  summarise(median_ChlA = median(ChlA, na.rm = TRUE), 
            median_NO3 = median(NO3, na.rm = TRUE)) %>%
  ungroup()

trib_sites <- clp_long %>% 
  filter(site_code %in% c("SFM", "PJW"))


main_plot_data <- clp_long %>% 
  filter(!site_code %in% c("SFM", "PJW"))

main_plot_data$site_code <- factor(main_plot_data$site_code, levels = c("JOEI", "JOER", "CBRR", "CBRI", "CHD", "JWC", "PFAL", "PBR", "PSF", "PNF", "PBD"))
main_plot_data$season <- factor(main_plot_data$season, levels = c("Winter", "Rising Limb", "Peak Flow", "Falling Limb", "Fall Baseflow"))
main_plot_data$month <- factor(main_plot_data$month, levels = c("5", "6", "7", "8", "9", "10","11"))



# Updated Plot
clp_long_chla_extra <- ggplot(main_plot_data, aes(x = distance_upstream_km, y = ChlA)) + 
  annotate("rect", xmin = 78, xmax = 99, ymin = -0.1, ymax = 9, 
           color = "#4A86C5", fill = "NA", linewidth = 1.5) +

  # Original Jitter for main sites
  geom_jitter(aes(color = season, shape = type), size = 3, width = 0.4) +
    
  # Grey Tributary points (SFM and PJW)
  geom_jitter(data = trib_sites, aes(x = distance_upstream_km, y = ChlA), color = "grey60", alpha = 0.4, size = 3, width = 0.4) +
  
  scale_color_manual(values = season_clp_color_vals) +
  
  # Lines and points for medians
  geom_line(data = clp_median_data %>% filter(!site_code %in% c("SFM", "PJW")), 
            aes(x = distance_upstream_km, y = median_ChlA), 
            color = "black", linewidth = 2.5) +
  geom_point(data = clp_median_data %>% filter(!site_code %in% c("SFM", "PJW")), 
             aes(x = distance_upstream_km, y = median_ChlA, shape = type), 
             color = "black", size = 5) +
  
  # Annotations (Your existing arrows and labels)
  annotate("label", x = 78, y = 8, label = "Joe Wright & Chambers\nReservoirs (Prev Figure)", 
           hjust = 0, size = 6, label.size = 0.5) +
  annotate("label", x = 80, y = 5.75, label = "JWC site", size = 5) +
  
  # Confluence with Big South (PJW is at ~77km)
  annotate("segment", x = 73, xend = 77, y = 5, yend = 2, 
           arrow = arrow(length = unit(0.3, "cm")), color = "#4A86C5", linewidth = 1) +  
  annotate("label", x = 73, y = 5, label = "Confluence with\n Big South (PJW)", size = 5) +
  
  # Confluence with South Fork (SFM is at ~40km)
  annotate("segment", x = 45, xend = 40, y = 5, yend = 1, 
           arrow = arrow(length = unit(0.3, "cm")), color = "#4A86C5", linewidth = 1) +  
  annotate("label", x = 45, y = 5, label = "Confluence with\nSouth Fork (SFM)", size = 5) +
  
  annotate("segment", x = 8, xend = 5, y = 5, yend = 1, 
           arrow = arrow(length = unit(0.3, "cm")), color = "#4A86C5", linewidth = 1) + 
  annotate("label", x = 8, y = 5, label = "Confluence with\nNorth Fork", size = 5) +

  # Styling and Scales
  theme_bw(base_size = 30) +
  scale_x_reverse() +
  ylim(-0.2, 9.2) +
  labs(x = "Distance from Canyon Mouth (km)", 
       y = "Chlorophyll a (µg/L)", 
       shape = "Location", 
       color = "Season") +
  theme(
    legend.position = c(0.98, 0.98),          
    legend.justification = c(1, 1), 
    legend.box = "horizontal",                
    legend.box.background = element_rect(color = "black", linewidth = 0.5), 
    legend.background = element_blank(),      
    legend.margin = margin(6, 6, 6, 6)        
  )

save_plot_image(clp_long_chla_extra, "clp_long_chla_extra")



clp_long_chla<- ggplot(clp_long, aes(x = distance_upstream_km, y = ChlA)) + 
  annotate("rect", xmin = 78, xmax = 99, ymin = -0.1, ymax = 9, 
           color = "#4A86C5", fill = "NA", linewidth = 1.5) +
  
  geom_jitter(aes(color = season, shape = type), size = 3, width = 0.4) +
  scale_color_manual(values = season_clp_color_vals) +
  
  # Lines and points for medians
  geom_line(data = clp_median_data, aes(x = distance_upstream_km, y = median_ChlA), 
            color = "black", linewidth = 2.5) +
  geom_point(data = clp_median_data, aes(x = distance_upstream_km, y = median_ChlA, shape = type), 
             color = "black", size = 5) +
  #Box title
  annotate("label", x = 78, y = 8, label = "Joe Wright & Chambers\nReservoirs (Prev Figure)", 
           hjust = 0, size = 6, label.size = 0.5) +
  #JWC label + arrow
  annotate("label", x = 80, y = 5.75, label = "JWC site", size = 5) +
  #Confluence with Big South
    annotate("segment", x = 73, xend = 77, y = 4, yend = 2, 
           arrow = arrow(length = unit(0.3, "cm")), color = "#4A86C5", linewidth = 1) +  
  annotate("label", x = 73, y = 4, label = "Confluence\nwith Big South", size = 5) +
  #confluence with SF
        annotate("segment", x = 40, xend = 40, y = 4, yend = 1, 
           arrow = arrow(length = unit(0.3, "cm")), color = "#4A86C5", linewidth = 1) +  
  annotate("label", x = 40, y = 4, label = "Confluence with\nSouth Fork", size = 5) +
        annotate("segment", x = 5, xend = 5, y = 4, yend = 1, 
           arrow = arrow(length = unit(0.3, "cm")), color = "#4A86C5", linewidth = 1) + 
  annotate("label", x = 8, y = 4, label = "Confluence with\nNorth Fork", size = 5) +

  # Styling and Scales
  theme_bw(base_size = 30) +
  scale_x_reverse() +
  ylim(-0.2, 9.2) +
  labs(x = "Distance from Canyon Mouth (km)", 
       y = "Chlorophyll a (µg/L)", 
       shape = "Location", 
       color = "Season") +
 theme(
    legend.position = c(0.98, 0.98),         
    legend.justification = c(1, 1), 
    
    legend.box = "horizontal",               
    legend.box.background = element_rect(color = "black", linewidth = 0.5), 
    legend.background = element_blank(),     
    legend.margin = margin(6, 6, 6, 6)       
  )

print(clp_long_chla)


save_plot_image(clp_long_chla, "clp_long_chla")

```

## Stats
Compare site chla variability with levene test and means using mixed model 
Fixed: Site
Random: Season + Year

Calculate the metric of reservoir impact: acre ft impounded water storage/ watershed area 
Does this correlate with Chl-a across the basin?

```{r}

clp_long_stats <-  tidy_chem  %>%
  filter(!is.na(ChlA)& Year %in% c(2022, 2023, 2024)) %>%
  mutate(site_code  = case_when(site_code == "SLEP" ~ "PFAL",
                           site_code == "PMAN" ~ "PNF",
                           TRUE ~ site_code))%>%
  dplyr::filter(site_code %in% c("JWC", "PJW", "PFAL", "PBR", "PSF", "PNF", "PBD"))%>%
  select(Date,site_code, ChlA, NO3, DTN,DOC,K,Year,season, month,location_type)%>%
  mutate(location_type = ifelse(location_type == "Outflow"|location_type =="Inflow", "Stream", "Reservoir"))%>%
  pivot_longer(cols = c(ChlA, NO3, DOC, DTN, K), 
               names_to = "param", 
               values_to = "value")%>%
  mutate(log_value = log10(value + 1),
         Year = factor(Year, levels = c("2021", "2022", "2023", "2024")),
         season = factor(season, levels = c("Winter", "Rising Limb", "Peak Flow", "Falling Limb", "Fall Baseflow")))

site_check <- map(c("DTN", "K", "DOC", "NO3", "ChlA"), ~{
  df <- clp_long_stats%>%filter(param == .x & value > 0)
  #Since NO3 does not fit a normal distribution, we will use a generalized linear mixed model with a gamma distribution and log link for NO3, and linear mixed models for the other parameters.
  if(.x == "NO3"){
    model_gamma <- glmmTMB(value ~ site_code + (1|Year) + (1|season), 
                           data = df, 
                           family = Gamma(link = "log"))
    
    return(check_model(model_gamma))
  }
  model_full <- lmer(log_value ~ site_code + season + (1|Year) + (1|season), data = df, REML = FALSE)
  
  check_model(model_full)
})%>%
  set_names(c("DTN", "K", "DOC", "NO3","ChlA"))

# If they pass the model diagnostics, extract the estimate for burn_group and calculate the percent difference between burned and unburned reservoirs. Then compare the full model to a null model without burn_group using an ANOVA to get the p-value for the effect of burn_group.

map(c("DTN", "K", "DOC", "NO3","ChlA"), ~{
  df <- clp_long_stats%>%filter(param == .x & value > 0) 
  
  if(.x == "NO3"){
    # Gamma model for NO3
    model_gamma <- glmmTMB(value ~ site_code + (1|Year) + (1|season), 
                           data = df, 
                           family = Gamma(link = "log"))
    
    year_means <- emmeans(model_gamma, "site_code")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value, 
             lower.CL = asymp.LCL, 
             upper.CL = asymp.UCL)
    
    model_null <- glmmTMB(value ~  season + (1 | Year), 
                          data = df, 
                          family = Gamma(link = "log"))
    
    anova_res <- anova(model_null, model_gamma)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
    
  } else {
    # LMM for other parameters
    model_full <- lmer(log_value ~ site_code  + (1|Year) + (1|season), data = df, REML = FALSE)
    
    year_means <- emmeans(model_full, "site_code")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value)
    
    model_null <- lmer(log_value ~ season + (1|Year), data = df, REML = FALSE)
    anova_res <- anova(model_null, model_full)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
  }
  contrasts%>%
    mutate(p_value_anova = anova_res$`Pr(>Chisq)`[2],
           parameter = .x)
})%>%
  bind_rows()%>%
  # Final calculation for percent changes and significance stars
  mutate(
    pct_est  = (exp(estimate) - 1) * 100,
    pct_low  = (exp(lower.CL) - 1) * 100,
    pct_high = (exp(upper.CL) - 1) * 100,
    p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
    p_sig_contrast = ifelse(p_value_contrast < 0.05, "*", "")
  )%>%
  select(parameter, contrast, pct_est, pct_low, pct_high, p_value_contrast, p_sig_contrast, p_value_anova, p_sig_anova)%>%
  filter(parameter %in% c( "ChlA"))

```

### Paired up-downstream regressions

Pairs: 
- CHD- JWC
- CHD - PFAL
- HORO/COMO - SFM
- HORO/COMO - BEAV

```{r}
#find paired samples (ie collected within the same week)

up_down_corr <- tidy_chem %>%
  filter(site_code %in% c("CBRR", "CHD", "JWC", "PFAL", "PBR"))%>%
  select(site_code, Date, ChlA)%>%
  mutate(year_week = paste(year(Date), week(Date), sep = "-"))%>%
  summarise(mean_chla = mean(ChlA, na.rm = TRUE), .by = c("site_code", "year_week"))%>%
  pivot_wider(id_cols = year_week, names_from = site_code, values_from = mean_chla)

  chd_jwc <- ggplot(data = up_down_corr, aes(x = CHD, y = JWC)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  #add R2 and p-value
    stat_cor(aes(label = paste(after_stat(rr.label), after_stat(p.label), sep = "~`,`~")),
           label.x = 0.5, label.y = 5.5)+
  labs(x = "CHD Chlorophyll a (µg/L)", 
         y = "JWC Chlorophyll a (µg/L)") +
  theme_bw(base_size = 20)+
    coord_cartesian(xlim = c(0, 6), ylim = c(0, 6))
  
  chd_pfal <- ggplot(data = up_down_corr, aes(x = CHD, y = PFAL)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  #add R2 and p-value
    stat_cor(aes(label = paste(after_stat(rr.label), after_stat(if_else(p.label < 0.001, "0.001", p.label)), sep
 = "~`,`~")),
           label.x = 0.5, label.y = 5.5)+
  labs(x = "CHD Chlorophyll a (µg/L)",
       y = "PFAL Chlorophyll a (µg/L)") +
  theme_bw(base_size = 20)+
    coord_cartesian(xlim = c(0, 6), ylim = c(0, 6))
  
  
  chd_pbr <- ggplot(data = up_down_corr, aes(x = CHD, y = PBR)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  #add R2 and p-value
    stat_cor(aes(label = paste(after_stat(rr.label), after_stat(if_else(p.label
 < 0.001, "0.001", p.label)), sep = "~`,`~")),
           label.x = 0.5, label.y = 5.5)+
  labs(x = "CHD Chlorophyll a (µg/L)",
       y = "PBR Chlorophyll a (µg/L)") +
  theme_bw(base_size = 20)+
    coord_cartesian(xlim = c(0, 6), ylim = c(0, 6))
  
  
  chd_jwc + chd_pfal + chd_pbr
  

up_down_corr_sf <- tidy_chem %>%
  filter(site_code %in% c("HORO", "COMO", "BEAV", "SFM", "PSF"))%>%
  select(site_code, Date, ChlA)%>%
  filter(year(Date) > 2022)%>%
  mutate(year_week = paste(year(Date), week(Date), sep = "-"))%>%#, 
         #site_code = ifelse(site_code %in% c("HORO", "COMO"), "Combined Outflows", site_code))%>%
  summarise(mean_chla = mean(ChlA, na.rm = TRUE), .by = c("site_code", "year_week"))%>%
  pivot_wider(id_cols = year_week, names_from = site_code, values_from = mean_chla)

  outflows_beav <- ggplot(data = up_down_corr_sf, aes(x = COMO, y = BEAV)) +
  geom_point() +
  geom_point(aes(x = HORO, y = BEAV), color = "red") +
  geom_smooth(method = "lm", se = F) +
  #add R2 and p-value
    stat_cor(aes(label = paste(after_stat(rr.label), after_stat(p.label), sep = "~`,`~")),
           label.x = 0.5, label.y = 8.5)+
  labs(x = "Outflow Chlorophyll a (µg/L)", 
         y = "BEAV Chlorophyll a (µg/L)") +
  theme_bw(base_size = 20)+
    coord_cartesian(xlim = c(0, 10), ylim = c(0, 10))
  
  outflows_sfm <- ggplot(data = up_down_corr_sf, aes(x = COMO, y = SFM)) +
  geom_point() +
  geom_point(aes(x = HORO, y = BEAV), color = "red") +
  geom_smooth(method = "lm", se = F) +
  #add R2 and p-value
    stat_cor(aes(label = paste(after_stat(rr.label), after_stat(p.label), sep = "~`,`~")),
           label.x = 0.5, label.y = 8.5)+
  labs(x = "Outflow Chlorophyll a (µg/L)", 
         y = "SFM Chlorophyll a (µg/L)") +
  theme_bw(base_size = 20)+
    coord_cartesian(xlim = c(0, 10), ylim = c(0, 10))
  
  
  outflows_psf <- ggplot(data = up_down_corr_sf, aes(x = COMO, y = PSF)) +
  geom_point() +
  geom_point(aes(x = HORO, y = BEAV), color = "red") +
  geom_smooth(method = "lm", se = F) +
  #add R2 and p-value
    stat_cor(aes(label = paste(after_stat(rr.label), after_stat(p.label), sep = "~`,`~")),
           label.x = 0.5, label.y = 8.5)+
  labs(x = "Outflow Chlorophyll a (µg/L)", 
         y = "PSF Chlorophyll a (µg/L)") +
  theme_bw(base_size = 20)+
    coord_cartesian(xlim = c(0, 10), ylim = c(0, 10))
  
  
  outflows_beav + outflows_sfm + outflows_psf
  
```


### Res Prop/Impact Metric


```{r}
# NID = https://nid.sec.usace.army.mil/nid/#/downloads
dams <- st_read('data/raw/external_datasets/inventory_of_dams/nation_20260313.gpkg') %>%
  filter(state %in% "Colorado",
         grepl("Water Supply|Irrigation", purposeIds, ignore.case = TRUE)) %>%
  st_transform(4326)%>%
  select(name, nidId, ownerNames, latitude, longitude, riverName,nidStorage,purposeIds,normalStorage,maxStorage )


#read in watersheds 
watersheds <- st_read(dsn = here(spatial_dir,"cpf_watersheds.shp"))%>%
  left_join(most_recent_meta, by = c("site_code"))%>%
  filter(location_type != c("Reservoir"))

dams_watersheds <- map(1:nrow(watersheds), ~{
  
  watershed_sel <- watersheds[.x,]
  
  dams_in_watershed <- st_join(dams, watershed_sel, join = st_intersects) %>%
    filter(!is.na(site_code)) 
  
  if(nrow(dams_in_watershed) == 0){
    return(tibble(site_code = watershed_sel$site_code, 
                  location_type = watershed_sel$location_type,
                  total_storage_acre_ft = 0, 
                  watershed_area_km2 = as.numeric(st_area(watershed_sel) / 1e6)))
  }
  
  
  
  return(tibble(watershed_area_km2 = as.numeric(st_area(watershed_sel) / 1e6), 
                site_code = unique(watershed_sel$site_code), 
                  location_type = watershed_sel$location_type,
                total_storage_acre_ft = sum(as.numeric(dams_in_watershed$normalStorage), na.rm = TRUE)))
  
}) %>%
  bind_rows()%>%
  mutate(location_type = case_when(site_code == "CBRI" ~ "Outflow", 
                                   site_code %in% c("SFM", "JWC") ~ "Stream",
                                   site_code == "PFAL" ~ "Mainstem",
                                   TRUE ~location_type))%>%
  filter(location_type != "Outflow" & site_code %nin% c("LBEA", "PENN"))

highlight_sites <- c("pbd", "pnf", "beav", "jwc", "pfal", "pbr")%>%toupper()

tidy_chem %>%
  filter(!is.na(ChlA)) %>%
  select(-location_type)%>%
  left_join(dams_watersheds, by = "site_code") %>%
  mutate(impounded_ws_area = total_storage_acre_ft / watershed_area_km2) %>%
  filter(!is.na(impounded_ws_area)) %>%
  ggplot(aes(x = total_storage_acre_ft, y = ChlA, color = location_type)) +
  geom_boxplot(aes(group = site_code), width = 1000) +
  geom_smooth(method = "lm", se = FALSE)+
  #scale_y_log10()+
  scale_color_manual(values = ROSS_dk_pal) + 
  labs(x = "Normal Impounded Water Storage (acre-ft)", 
       y = "Chlorophyll a (µg/L)", 
       color = "Site Code")+
  ylim(0,10)

dams_watersheds%>%
  left_join(tidy_chem %>% select(site_code, ChlA) %>% group_by(site_code) %>% summarise(median_ChlA = median(ChlA, na.rm = TRUE)), by = "site_code")%>%
  mutate(metric = total_storage_acre_ft / watershed_area_km2)%>%
  ggplot(aes(x = watershed_area_km2, y = metric, color = site_code))+
  geom_point()


tmap_mode("view")
tm_basemap("Esri.WorldImagery")+
tm_shape(watersheds%>%filter(site_code == "PBD")) +
  tm_fill(col = "lightblue", alpha = 0.5) +
  tm_borders(col = "black", lwd = 2) +
tm_shape(filtered_dams) +
  tm_dots(size = 1)

#for each watershed calculate the total storage of dams contained in that watershed and the total watershed area


```


### Prop of samples below 2.6

```{r}

clp_long_stats%>%
  filter(param == "ChlA")%>%
  mutate(group = if_else(site_code == "JWC", "JWC", "Other Sites"))%>%
  group_by(site_code)%>%
  summarize(prop_below_2.6 = mean(value <= 2.6, na.rm = TRUE) *100, 
            #how many times was a sample above 2.6
            prop_above_2.6 = sum(value > 2.6, na.rm = TRUE))

```

## Seasonality of high chl-a

```{r}

clp_long_stats%>%
  filter(param == "ChlA")%>%
  mutate(group = case_when(site_code == "JWC" ~ "JWC",
                           site_code == "PFAL" ~ "PFAL",
                           TRUE ~"Other Sites"))%>%
  ggplot(aes(x = Date, y = value))+
  geom_point()+
  geom_line()+
  geom_hline(yintercept = 2.6, linetype = "dashed")+
  #scale_color_manual(values = c("black", "red"), labels = c("Below 2.6", "Above 2.6"))+
  facet_wrap(~site_code)
  labs(x = "Date", y = chem_expressions$ChlA, fill = "Site Code")+
  theme_bw(base_size = 30)

clp_long_stats%>%
  filter(param == "ChlA")%>%
  mutate(group = case_when(site_code == "JWC" ~ "JWC",
                           site_code == "PFAL" ~ "PFAL",
                           TRUE ~"Other Sites"))%>%
  summarize(prop_below_2.6 = mean(value <= 2.6, na.rm = TRUE) *100, 
            prop_above_2.6 = mean(value > 2.6, na.rm = TRUE), .by = c( "season"))%>%
  arrange(prop_above_2.6)

 tidy_chem%>%
  filter(location_type %nin% c("Outflow", "Reservoir", "Inflow", "Stream") & !is.na(ChlA))%>%
  summarize(prop_below_2.6 = mean(ChlA <= 2.6, na.rm = TRUE) *100,
            prop_above_2.6 = mean(ChlA > 2.6, na.rm = TRUE) *100, 
            count = n(),
            .by = "season")%>%
  arrange(prop_above_2.6)

```



# South Fork

## Plots

```{r}

SF_Res = c("COMI", "COMR", "COMO", "HORI", "HORR", "HORO", "BEAV", "SFM", "PSF")
#filter and org
sf_chem<- filter(tidy_chem, site_code %in% SF_Res & !is.na(ChlA))%>%
  select(site_code,Date,ChlA,location_type, Buffer_Level, Watershed_Level, watershed, Year )%>%
  mutate(location_type = case_when(
    location_type == "Stream" ~ "Beaver Creek", 
    site_code == "PSF" ~ "Poudre Below\nSouth Fork",
    site_code == "SFM" ~ "South Fork",
    TRUE ~ location_type), 
    site_group = case_when(
      watershed == "Comanche Reservoir" ~ watershed, 
      watershed == "Hourglass Reservoir"~ watershed, 
      TRUE ~ "Stream"
    ), 
    filled_status = case_when(
      Year %in% c(2021, 2022) ~ "Reservoirs NOT Filled",
      Year  %in% c(2023, 2024) ~ "Reservoirs Filled"
      
    )) 

sf_chem$location_type <- factor(sf_chem$location_type, levels = c("Inflow","Reservoir", "Outflow","Beaver Creek","South Fork", "Poudre Below\nSouth Fork"))
sf_chem$filled_status <- factor(sf_chem$filled_status, levels = c("Reservoirs NOT Filled", "Reservoirs Filled"))
#get nice units for Y axis

#plot
south_fork <-   ggplot(sf_chem, aes(x=location_type,y=ChlA,  fill = filled_status, shape = )) + 
  geom_dotplot( binaxis = "y", 
                method = "histodot",
                stackdir = "center", 
                position = position_dodge(width = 0.8),
                binwidth = .2, dotsize = 2, 
                stackratio = 1, width = .01)+
  geom_abline(aes(slope=0, intercept=2.6), color='black', linetype="dashed") +
  geom_abline(aes(slope=0, intercept=8), color='black', linetype="dashed") +
  annotate(geom="text", label= "Mesotrophic", x=.75, y=2.6, vjust= -1, size = 6)+
  annotate(geom="text", label= "Eutrophic", x= .75, y=8, vjust= -1, size = 6)+
  theme_bw(base_size = 30)+
  scale_fill_manual(name = "Reservoir Status", values = c("#D55E00","#0072B2"))+
  labs(color = "Reservoir Status",
       y = chem_expressions$ChlA)+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_text(face = c('plain', 'plain', 'plain', 'plain', 'plain', 'plain')), 
        legend.position = "bottom")


south_fork

save_plot_image(south_fork ,"south_fork_chla_impact" )
```

## Stats

```{r}
sf_chem <- tidy_chem %>%
  filter(site_code %in% c("COMI", "HORI", "COMR", "HORR", "COMO", "HORO", "BEAV", "SFM", "PSF"))%>%
  select(site_code,Date,ChlA, Turbidity, TSS, DTN, DOC, NO3, SO4, PO4, K, location_type, watershed, Year, season )%>%
  mutate(location = case_when(
    site_code == "BEAV" ~ "Beaver Creek", 
    site_code %in% c("HORO", "COMO") ~ "Outflows",
    site_code %in% c("HORR", "COMR") ~ "Reservoirs",
    site_code %in% c("HORI", "COMI") ~ "Inflows",
    site_code == "SFM" ~ "South Fork",
    site_code == "PSF" ~ "Poudre Below South Fork",
    TRUE ~ location_type))%>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4, PO4, K), 
               names_to = "param", 
               values_to = "value")%>%
  mutate(year_site = paste0(Year, "_", site_code))%>%
  mutate(log_value = log10(value + 1), 
         fill_status = factor(if_else(Year <= 2022, "Unfilled", "Filled")),
         Year = factor(Year, levels = c("2021", "2022", "2023", "2024")),
         season = factor(season, levels = c("Winter", "Rising Limb", "Peak Flow", "Falling Limb", "Fall Baseflow")))%>%
  na.omit()

ggplot(sf_chem, aes(x = value, fill = site_code))+
  geom_histogram()+
  facet_wrap(~param, scales = "free")
```

### Compare site-wise (Unfilled)

```{r}
sf_chem_unfilled <- sf_chem %>% filter(fill_status == "Unfilled")

site_check <- map(c("DTN", "K", "DOC", "NO3", "ChlA"), ~{
  df <- sf_chem_unfilled%>%filter(param == .x & value > 0)
  #Since NO3 does not fit a normal distribution, we will use a generalized linear mixed model with a gamma distribution and log link for NO3, and linear mixed models for the other parameters.
  if(.x == "NO3"){
    model_gamma <- glmmTMB(log_value ~ location + (1|season), 
                           data = df, 
                           family = Gamma(link = "log"))
    
    return(check_model(model_gamma))
  }
  model_full <- lmer(log_value ~ location + (1|season), data = df, REML = FALSE)
  
  check_model(model_full)
})%>%
  set_names(c("DTN", "K", "DOC", "NO3","ChlA"))


map(c("DTN", "K", "DOC", "NO3","ChlA"), ~{
  df <- sf_chem_unfilled%>%filter(param == .x & value > 0) 
  
  if(.x == "NO3"){
    # Gamma model for NO3
    model_gamma <- glmmTMB(value ~ location + (1|season), 
                           data = df, 
                           family = Gamma(link = "log"))
    
    year_means <- emmeans(model_gamma, "location")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value, 
             lower.CL = asymp.LCL, 
             upper.CL = asymp.UCL)
    
    model_null <- glmmTMB(value ~  (1 | season), 
                          data = df, 
                          family = Gamma(link = "log"))
    
    anova_res <- anova(model_null, model_gamma)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
    
  } else {
    # LMM for other parameters
    model_full <- lmer(log_value ~ location + (1|season), data = df, REML = FALSE)
    
    year_means <- emmeans(model_full, "location")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value)
    
    model_null <- lmer(log_value ~ (1|season), data = df, REML = FALSE)
    anova_res <- anova(model_null, model_full)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
  }
  contrasts%>%
    mutate(p_value_anova = anova_res$`Pr(>Chisq)`[2],
           parameter = .x)
})%>%
  bind_rows()%>%
  # Final calculation for percent changes and significance stars
  mutate(
    pct_est  = (exp(estimate) - 1) * 100,
    pct_low  = (exp(lower.CL) - 1) * 100,
    pct_high = (exp(upper.CL) - 1) * 100,
    p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
    p_sig_contrast = ifelse(p_value_contrast < 0.05, "*", "")
  )%>%
  select(parameter, contrast, pct_est, pct_low, pct_high, p_value_contrast, p_sig_contrast, p_value_anova, p_sig_anova)%>%
  filter(parameter %in% c( "ChlA"))

```

### Compare site-wise (Filled)

```{r}
sf_chem_filled <- sf_chem %>% filter(fill_status == "Filled")

site_check <- map(c("DTN", "K", "DOC", "NO3", "ChlA"), ~{
  df <- sf_chem_filled%>%filter(param == .x & value > 0)
  #Since NO3 does not fit a normal distribution, we will use a generalized linear mixed model with a gamma distribution and log link for NO3, and linear mixed models for the other parameters.
  if(.x == "NO3"){
    model_gamma <- glmmTMB(log_value ~ location + (1|season), 
                           data = df, 
                           family = Gamma(link = "log"))
    
    return(check_model(model_gamma))
  }
  model_full <- lmer(log_value ~ location + (1|season), data = df, REML = FALSE)
  
  check_model(model_full)
})%>%
  set_names(c("DTN", "K", "DOC", "NO3","ChlA"))


map(c("DTN", "K", "DOC", "NO3","ChlA"), ~{
  df <- sf_chem_filled%>%filter(param == .x & value > 0) 
  
  if(.x == "NO3"){
    # Gamma model for NO3
    model_gamma <- glmmTMB(value ~ location + (1|season), 
                           data = df, 
                           family = Gamma(link = "log"))
    
    year_means <- emmeans(model_gamma, "location")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value, 
             lower.CL = asymp.LCL, 
             upper.CL = asymp.UCL)
    
    model_null <- glmmTMB(value ~  (1 | season), 
                          data = df, 
                          family = Gamma(link = "log"))
    
    anova_res <- anova(model_null, model_gamma)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
    
  } else {
    # LMM for other parameters
    model_full <- lmer(log_value ~ location + (1|season), data = df, REML = FALSE)
    
    year_means <- emmeans(model_full, "location")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value)
    
    model_null <- lmer(log_value ~ (1|season), data = df, REML = FALSE)
    anova_res <- anova(model_null, model_full)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
  }
  contrasts%>%
    mutate(p_value_anova = anova_res$`Pr(>Chisq)`[2],
           parameter = .x)
})%>%
  bind_rows()%>%
  # Final calculation for percent changes and significance stars
  mutate(
    pct_est  = (exp(estimate) - 1) * 100,
    pct_low  = (exp(lower.CL) - 1) * 100,
    pct_high = (exp(upper.CL) - 1) * 100,
    p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
    p_sig_contrast = ifelse(p_value_contrast < 0.05, "*", "")
  )%>%
  select(parameter, contrast, pct_est, pct_low, pct_high, p_value_contrast, p_sig_contrast, p_value_anova, p_sig_anova)%>%
  filter(parameter %in% c( "ChlA"))

```


### Downstream sites (filled vs unfilled)

When a reservoir is active, do downstream sites have increased chl-a and if so, how far is that impact "felt"?

```{r}

site_check <- map(c("DTN", "K", "DOC", "NO3", "ChlA"), ~{
  df <- sf_chem%>%filter(param == .x & value > 0)
  #Since NO3 does not fit a normal distribution, we will use a generalized linear mixed model with a gamma distribution and log link for NO3, and linear mixed models for the other parameters.
  if(.x == "NO3"){
    model_gamma <- glmmTMB(log_value ~ location * fill_status + (1|season) + (1|Year), 
                           data = df, 
                           family = Gamma(link = "log"))
    
    return(check_model(model_gamma))
  }
  model_full <- lmer(log_value ~ location * fill_status + (1|season) + (1|Year), data = df, REML = FALSE)
  
  check_model(model_full)
})%>%
  set_names(c("DTN", "K", "DOC", "NO3","ChlA"))


map(c("DTN", "K", "DOC", "NO3","ChlA"), ~{
  df <- sf_chem%>%filter(param == .x & value > 0) 
  
  if(.x == "NO3"){
    # Gamma model for NO3
    model_gamma <- glmmTMB(value ~ location * fill_status, 
                           data = df, 
                           family = Gamma(link = "log"))
    
   loc_fill_means <- emmeans(model_gamma, ~ fill_status | location, type = "response")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value)
    
    model_null <- glmmTMB(value ~  (1 | season), 
                          data = df, 
                          family = Gamma(link = "log"))
    
    anova_res <- anova(model_null, model_gamma)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
    
  } else {
    # LMM for other parameters
   model_full <- lmer(log_value ~ location * fill_status + (1|season), data = df, REML = FALSE)
    
   loc_fill_means <- emmeans(model_full, ~ fill_status | location, type = "response")

    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(loc_fill_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value)
    
    model_null <- lmer(log_value ~ (1|season), data = df, REML = FALSE)
    anova_res <- anova(model_null, model_full)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
  }
  contrasts%>%
    mutate(p_value_anova = anova_res$`Pr(>Chisq)`[2],
           parameter = .x)
})%>%
  bind_rows()%>%
  # Final calculation for percent changes and significance stars
  mutate(
    pct_est  = (exp(estimate) - 1) * 100,
    pct_low  = (exp(lower.CL) - 1) * 100,
    pct_high = (exp(upper.CL) - 1) * 100,
    p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
    p_sig_contrast = ifelse(p_value_contrast < 0.05, "*", "")
  )%>%
  select(parameter, location, contrast, pct_est, pct_low, pct_high, p_value_contrast, p_sig_contrast, p_value_anova, p_sig_anova)%>%
  filter(parameter %in% c( "ChlA"))
```



# Remote Sensing

## Import datasets and join with chem data

```{r}
#Read in the data from B

#Temp: From NW-CLP-RS/CLP_focus_temp_only.Rmd

temp_rs <- read_parquet(here("data/derived/remote_sensing/landsat/CLP_temp_estimate_v2024-10-10.parquet"))%>%
  mutate(doy = yday(date),
         kurt_SurfaceTemp = as.numeric(kurt_SurfaceTemp),
         sd_SurfaceTemp = as.numeric(sd_SurfaceTemp),
         med_surface_C = as.numeric(med_SurfaceTemp_corr) - 273.15, # Convert from Kelvin to Celsius
         doy_check = if_else(doy >= 120 & doy <= 300, "PASS", "FAIL"), # Filter for sampling season (roughly April to October)
  kurtosis_check = if_else(kurt_SurfaceTemp > 10, "FAIL", "PASS"), #filter out for instances with high kurtosis
  sd_check = if_else(sd_SurfaceTemp > 2, "FAIL", "PASS"), #filter out instances with high SD
pass_all = (sd_check == "PASS" & kurtosis_check == "PASS" & doy_check == "PASS"))

all_rs <- temp_rs %>%
  left_join(read_parquet("data/derived/remote_sensing/landsat/CLP_SDD_estimate_v2024-10-10.parquet"), by = c("gnis_name", "date"))%>%
  mutate(
    site_code = case_when(
    gnis_name == "Seaman Reservoir"     ~ "SEMR",    
    gnis_name =="Joe Wright Reservoir"    ~ "JOER", 
    gnis_name == "Barnes Meadow Reservoir"  ~ "BRNR",
    gnis_name == "Chambers Lake" ~ "CBRR",
    gnis_name == "Peterson Lake"     ~ "PTRR", 
    gnis_name ==     "Long Draw Reservoir"    ~ "LNGR",
gnis_name =="Hourglass Reservoir"   ~ "HORR",
gnis_name ==  "Comanche Reservoir"     ~ "COMR",
gnis_name == "Halligan Reservoir"  ~ "HALL",
    TRUE ~ NA_character_))%>%
  filter(year(date) >= 1990)

# How many images did each reservoir get over the study period
all_rs %>% 
  mutate(year = year(date))%>% 
  filter( year >= 2021 & !is.na(med_surface_C) & (doy >= 120 & doy <= 300))%>%
  summarise(count = n()/22.5, .by = c("site_code","year"))%>%
  pivot_wider(names_from = year, values_from = count)%>%
  filter(site_code %nin% c("HORR", "COMR", "SEMR", "HALL"))
# Total expected images (no clouds)  (DOY filter: 300-120)/8 day pass time = 22.5


```

## Seasonal Trends in Temperatures

```{r}

graphTrendObs_temp = function(lake) {
  seasonAdjTrends <- all_rs %>% 
    mutate(doy = yday(date))%>%
    filter(site_code == lake & !is.na(med_surface_C) & (doy >= 120 & doy <= 300))%>%
    filter(!c(site_code %in% c("HORR", "COMR") & year(date) %in% c(2021,2022)))%>% 
    select(gnis_name, date, med_surface_C)%>%
    as_tibble()%>% 
    arrange(date)%>%
    time_decompose(med_surface_C, merge = T)
  
  lakeName = seasonAdjTrends$gnis_name[1]
  
  ggplot(seasonAdjTrends, aes(x = date, y = observed)) +
    geom_point(color = 'gray') +
    geom_vline(xintercept = as.Date(c("2020-08-01")), linetype = "dashed", color = "red", lwd = 1) +
    # Removed y-axis title from here
    labs(x = NULL, y = NULL, title = paste0(lakeName)) + 
    geom_point(aes(y = trend), color = '#E69F00') +
    geom_smooth(aes(y = trend), lty = 2) +
    theme_bw(base_size = 14) + # Dropped base_size slightly so labels fit better side-by-side
    theme(plot.title = element_text(face = 'bold', hjust = 0.5, size = 12)) +
    ylim(0,30)+
    ROSS_theme
}

temp_plots <- map(c("BRNR", "PTRR","CBRR","JOER", "LNGR"), ~graphTrendObs_temp(.x))
#ggsave("docs/cpf_reservoir/paper_draft/temp_trends.jpg", dpi = 300, width = 10, height = 8)

```
## Seasonal Trends in SDD

```{r}
graphTrendObs_SDD = function(lake) {
  seasonAdjTrends <- all_rs %>% 
    mutate(doy = yday(date))%>%
    filter(site_code == lake & !is.na(mean_sdd) & (doy >= 120 & doy <= 300))%>%
    filter(!c(site_code %in% c("HORR", "COMR") & year(date) %in% c(2021,2022)))%>% 
    select(gnis_name, date, mean_sdd)%>%
    as_tibble() %>% 
    arrange(date) %>%
    time_decompose(mean_sdd, merge = T) 
    
  ggplot(seasonAdjTrends, aes(x = date, y = observed)) +
    geom_point(color = 'gray') +
    geom_vline(xintercept = as.Date(c("2020-08-01")), linetype = "dashed", color = "red", lwd = 1) +
    # Removed title and y-axis title from here so they don't double up
    labs(x = NULL, y = NULL, title = NULL) + 
    geom_point(aes(y = trend), color = '#E69F00') +
    geom_smooth(aes(y = trend), lty = 2) +
    theme_bw(base_size = 14) +
    ylim(0,6)+
        scale_y_continuous(position = "right", 
                           #add labels every 2 from 0-6
                           breaks = seq(0, 6, by = 2), 
                           limits = c(0,6))+
    ROSS_theme

}

sdd_plots <- map(c("BRNR", "PTRR", "CBRR","JOER", "LNGR"), ~graphTrendObs_SDD(.x))
ggsave("docs/cpf_reservoir/paper_draft/sdd_trends.jpg", plot = sdd_plots, dpi = 300, width = 10, height = 8)
```

```{r}

# 1. Interleave the plots: Temp 1, SDD 1, Temp 2, SDD 2...
all_plots <- c(rbind(temp_plots, sdd_plots))

# 2. Wrap them into a 2-column layout and span the site titles across the pairs
combined_grid <- wrap_plots(all_plots, ncol = 2, axes = "collect_x") & 
  theme(plot.title = element_text(hjust = 0, face = "bold")) # Spans the title toward the center

# 3. Add single global y-axis titles to the left and right sides
final_plot<- wrap_elements(panel = textGrob("Water Temperature (°C)", rot = 90, gp = gpar(fontsize = 14))) + 
  combined_grid + 
  wrap_elements(panel = textGrob("Secchi Depth (m)", rot = -90, gp = gpar(fontsize = 14))) +
  plot_layout(widths = c(1, 40, 1)) # Gives the side-labels just enough breathing room

# Save the master composite image
ggsave("docs/cpf_reservoir/paper_draft/combined_trends.jpg", final_plot, dpi = 300, width = 8, height = 6)
```

### Pre vs Post Fire Tests

```{r}

rs_tests <- all_rs %>%
  mutate(doy = yday(date), 
         year = year(date))%>%
  filter(site_code %nin% c("COMR", "HORR", "SEMR", "HALL"))%>%
  filter(doy >= 120 & doy <= 300) %>%
  filter(year >= 2016 & year != 2020)%>%
  select(site_code, gnis_name, date, doy, year, med_surface_C, mean_sdd)%>%
  pivot_longer(cols = c(med_surface_C, mean_sdd), names_to = "param", values_to = "value")%>%
  mutate(log_value = log(value), 
         fire_status = factor(if_else(year < 2020, "Pre-Fire", "Post-Fire")))

ggplot(rs_tests, aes(x = value, fill = site_code))+
  geom_histogram()+
  facet_wrap(param~fire_status, scales = "free")+
  theme_bw() 

check_model(lmer(
    value ~ site_code * fire_status + (1|year),
    data = rs_tests%>%filter(param == "med_surface_C" & value > 0),
    REML = FALSE
  ))
check_model(lmer(
    value ~ site_code * fire_status + (1|year),
    data = rs_tests%>%filter(param == "mean_sdd" & value > 0),
    REML = FALSE
  ))

map(c("med_surface_C", "mean_sdd"), ~{
  
  df <- rs_tests %>%
    filter(param == .x & value > 0)
  
  # Full model - site × fire interaction
  model_full <- lmer(
    value ~ site_code * fire_status + (1|year),
    data = df,
    REML = FALSE
  )
  
  # Null model - no interaction
  model_null <- lmer(
    value ~ site_code + fire_status + (1|year),
    data = df,
    REML = FALSE
  )
  
  anova_res <- anova(model_null, model_full)
  
  # Estimated marginal means for each lake × fire period
  emm <- emmeans(model_full, ~ fire_status | site_code)
  
  contrasts <- summary(pairs(emm, reverse = TRUE),  # Post - Pre
                       infer = TRUE) %>%
    as.data.frame() %>%
    rename(p_value_contrast = p.value)
  
  # Pre-fire baseline per lake for % change
  pre_means <- as.data.frame(emm) %>%
    filter(fire_status == "Pre-Fire") %>%
    select(site_code, pre_mean = emmean)
  
  contrasts %>%
    left_join(pre_means, by = "site_code") %>%
    mutate(
      p_value_anova = anova_res$`Pr(>Chisq)`[2],
      parameter     = .x
    )
  
}) %>%
  bind_rows() %>%
  mutate(
    diff      = estimate,
    diff_low  = lower.CL,
    diff_high = upper.CL,
    sig_stars = case_when(
      p_value_contrast < 0.001 ~ "***",
      p_value_contrast < 0.01  ~ "**",
      p_value_contrast < 0.05  ~ "*",
      TRUE                     ~ "ns"
    )
  ) %>%
  select(
    site_code, contrast, parameter, pre_mean,
    diff, diff_low, diff_high,
    p_value_contrast, sig_stars, p_value_anova
  ) %>%
  mutate(across(where(is.numeric), ~round(.x, 3)))%>%
  mutate(interp = case_when(p_value_contrast < 0.05 & parameter == "med_surface_C"& diff > 0 ~ "Temp Increase",
                            p_value_contrast < 0.05 & parameter == "med_surface_C"& diff < 0 ~ "Temp Decrease",
                            p_value_contrast < 0.05 & parameter == "mean_sdd"& diff > 0 ~ "Clarity Decrease",
                            p_value_contrast < 0.05 & parameter == "mean_sdd"& diff < 0 ~ "Clarity Increase",
                            TRUE ~ "No Significant Change"))

```


## RS vs grabs

```{r}
chem_rs <- all_rs%>%
  rename(rs_date = date)%>%
  mutate(date_low = rs_date - days(3),
    date_high = rs_date + days(3)) %>%
  left_join(tidy_chem, 
            by = join_by(site_code, between(y$Date, x$date_low, x$date_high)))%>%
  filter(!is.na(med_surface_C))%>%
  mutate(date_diff = abs(as.numeric(difftime(Date, rs_date, units = "days"))))%>%
  filter(!(site_code %in% c("COMR", "HORR") & Year %in% c(2021,2022)))%>% # removing unfilled comanche and hourglass matchups
  select(rs_date,sample_date = Date, date_diff, site_code, gnis_name, med_surface_C, mean_sdd, ChlA, Turbidity, TSS, NO3, DTN, DOC, Year)
```



Not great associations and there is only \~120 matchups between the seven reservoirs

```{r}
ggplot(chem_rs%>%filter(!is.na(ChlA)), aes(x = med_surface_C, y = ChlA,shape = Year))+
  geom_point()+
  facet_wrap(~site_code)+
  theme_bw()+
  labs(x = "Estimated Surface Temperature (°C)", y = "Sample Chlorophyll a (µg/L)")


ggplot(chem_rs%>%filter(!is.na(ChlA)), aes(x = mean_sdd, y = TSS, shape = Year))+
  geom_point()+
  facet_wrap(~site_code)+
  theme_bw()+
  labs(x = "Estimated Mean Secchi Disk Depth (m)", y = "Sample TSS (mg/L)")

ggplot(chem_rs%>%filter(!is.na(ChlA)), aes(x = mean_sdd, y = Turbidity, shape = Year))+
  geom_point()+
  facet_wrap(~site_code)+
  theme_bw()+
  labs(x = "Estimated Mean Secchi Disk Depth (m)", y = "Sample Turbidity (NTU)")

ggplot(chem_rs%>%filter(!is.na(ChlA)), aes(x = mean_sdd, y = ChlA,shape = Year))+
  geom_point()+
  facet_wrap(~site_code)+
  theme_bw()+
  labs(x = "Estimated Mean Secchi Disk Depth (m)", y = "Sample Chlorophyll a (µg/L)")

ggplot(chem_rs%>%filter(!is.na(ChlA)), aes(x = mean_sdd, y = DOC, shape = Year))+
  geom_point()+
  facet_wrap(~site_code)+
  theme_bw()+
  labs(x = "Estimated Mean Secchi Disk Depth (m)", y = "Sample DOC (mg/L)")


```


# Scratch

## k/ca trends

```{r}
k_ca_plot <- tidy_chem %>%
  filter(!is.na(K) & !is.na(Ca) & site_code %in% all_reservoirs) %>%
  mutate(K_Ca = K/Ca) %>%
  mutate(site_code = factor(site_code, levels = rev(names(colorsBS_site_code)))) %>%
  ggplot(aes(x = site_code, y = K_Ca, fill = site_code)) +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.5) +
  scale_fill_manual(values = colorsBS_site_code) +
  #facet_wrap(~Year, scales = "free_y") +
  stat_summary(fun = "median", geom = "crossbar", size =1, color = "black") +
  labs(
    x = "Reservoir", 
    y = bquote("Ratio: " ~ K^"+" ~ "/" ~ Ca^"2+"), 
    fill = "Site Code"
  ) +
  theme_bw(base_size = 20)



tidy_chem%>%
  filter(!is.na(K) & !is.na(Ca) & site_code %in% all_reservoirs)%>%
  mutate(K_Ca = K/Ca)%>%
  #arrange site code by colorBS_site_code
  mutate(site_code = factor(site_code, levels = rev(names(colorsBS_site_code))))%>%
  ggplot(aes(x = K_Ca, fill = site_code))+
  geom_histogram()+
  scale_fill_manual(values = colorsBS_site_code)+
  facet_wrap(~Year)+
  labs(x = paste0("Ratio of K/Ca"), fill = "Site Code")



   df <- tidy_chem%>%
  filter(!is.na(K) & !is.na(Ca) & site_code %in% c("LNGR", "JOER", "CBRR", "PTRR", "BRNR"))%>%
  mutate(K_Ca = K/Ca)%>%
  #arrange site code by colorBS_site_code
  mutate(site_code = factor(site_code, levels = rev(names(colorsBS_site_code))))%>%
  mutate(log_K_Ca = log10(K_Ca + 1))%>%
       mutate(burn_group = factor(burn_group, levels = c("unburned", "burned")),
         Year = factor(Year, levels = c("2021", "2022", "2023", "2024")), 
         season = factor(season),
         site_code = factor(site_code, levels = c("LNGR", "JOER", "CBRR", "PTRR", "BRNR")))
   
anova_res <- aov(log_K_Ca ~ burn_group, data = df)
summary(anova_res)

```


# SI

## SI Table 1: PO4 detections over time


Looking at where PO4 is present (above 0.01 detection limit) and if that
is changing over time
```{r}
# How many samples per study reservoirs

tidy_chem%>%
  filter( site_code %in% primary_study_res)%>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4, PO4), 
               names_to = "param", 
               values_to = "value")%>%
  filter(param == "PO4")%>%
  group_by(site_code, burn_group)%>%
  summarise(samples = n(), 
           all_samples = sum(value),
            abv_nd  = sum(value > 0))%>%
  mutate(perc_abv_nd = abv_nd/samples*100)

# How many samples per year at study reservoirs

tidy_chem%>%
  filter( site_code %in% primary_study_res)%>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4, PO4), 
               names_to = "param", 
               values_to = "value")%>%
  filter(param == "PO4")%>%
  group_by(Year)%>%
  summarise(samples = n(), 
           all_samples = sum(value),
            abv_nd  = sum(value > 0))%>%
  mutate(perc_abv_nd = abv_nd/samples*100)


# How many samples at mainstem/stream sites

tidy_chem%>%
  filter( site_code %nin% all_reservoirs)%>%
  filter(site_code %nin% c("BEAV", "PENN", "SAWM", "LBEA" ))%>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4, PO4), 
               names_to = "param", 
               values_to = "value")%>%
  filter(param == "PO4")%>%
  group_by(location_type)%>%
  summarise(samples = n(), 
           all_samples = sum(value),
            abv_nd  = sum(value > 0))%>%
  mutate(perc_abv_nd = abv_nd/samples*100)%>%
  arrange(desc(perc_abv_nd))

```

```{r}

upper_PO4 <- filter(tidy_chem, watershed %nin% c("CLP Mainstem - Lower","CLP Tributary - Lower"))%>%
  select(site_code,location_type, Date, Turbidity, PO4, Year)%>%
  mutate(location_type = case_when(
    site_code %in% c("JWC", "BEAV", "LBEA", "PENN", "SFM", "SAWM", "JOEI", "CBRI", "COMI", "HORI") ~ "Tributary",
    TRUE ~ location_type))%>%
  group_by(location_type, Year)%>%
  # fina what percent of each year had PO4 levels above 0.01 and get the mean level for each location type and year
  summarise(total_samples = n(),
            PO4_perc_dect = sum(PO4 >= 0.01, na.rm = TRUE)/n()* 100,
            #calculate mean of PO4 for only samples above 0.01, excluding samples lower than 0.01
            mean_PO4 = mean(PO4[PO4 >= 0.01], na.rm = TRUE),
            samples = paste0(sum(PO4 >= 0.01, na.rm = TRUE),"/",n() ))%>%
  # pivot wider by year
  pivot_wider(names_from = Year, values_from = c(total_samples, PO4_perc_dect, mean_PO4, samples)) %>%
  mutate(across(total_samples_2021:mean_PO4_2024, ~replace_na(.x, 0)))%>%
  mutate(across(samples_2021:samples_2024, ~replace_na(.x, "0/0")))%>%
  # if a column is numeric, round it to 2 decimal places
  mutate(across(where(is.numeric), ~round(.x, 2)))%>%
  arrange(factor(location_type, levels = c("Inflow", "Reservoir", "Outflow", "Mainstem", "Tributary")))%>%
  select(location_type, total_samples_2021:mean_PO4_2024)

#### Creating Table ####
# Add spacer columns
upper_PO4_with_spacers <- upper_PO4 %>%
  mutate(sp1 = NA, sp2 = NA, sp3 = NA) %>%
  select(1, 2:5, sp1, 6:9, sp2, 10:13, sp3)

# Rename columns with valid names
colnames(upper_PO4_with_spacers) <- c(
  "Site_Type",           # 1
  "n_2021", "n_2022", "n_2023", "n_2024", "sp1",  # 2-6
  "pct_2021", "pct_2022", "pct_2023", "pct_2024", "sp2",  # 7-11
  "mean_2021", "mean_2022", "mean_2023", "mean_2024", "sp3" # 12-16
)

# Create the flextable
ft1 <- flextable(upper_PO4_with_spacers) %>%
  set_header_labels(
    Site_Type = "",
    n_2021 = "2021", n_2022 = "2022", n_2023 = "2023", n_2024 = "2024",
    pct_2021 = "2021", pct_2022 = "2022", pct_2023 = "2023", pct_2024 = "2024",
    mean_2021 = "2021", mean_2022 = "2022", mean_2023 = "2023", mean_2024 = "2024",
    sp1 = "", sp2 = "", sp3 = ""
  ) %>%
  add_header_row(
    values = c("Site Type", "Samples per Year", "", "Percent above Detection Limit", "", 
               "Mean PO4 Above Detection Limit", ""),
    colwidths = c(1, 4, 1, 4, 1, 4, 1)
  ) %>%
  theme_vanilla() %>%
  align(align = "center", part = "all") %>%
  valign(valign = "center", part = "all") %>%
  fontsize(size = 11, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  set_table_properties(layout = "autofit", width = 1)%>%
  autofit()

ft1
save_as_docx(ft1,path = here(paper_draft_dir,"SI_1_Table.docx"))


```

## SI Table 2: Mean, SD, Max for nutrients for reservoirs

Calculate mean, sd and max for key constituents for the 7 main study
reservoirs

```{r}

key_table <- tidy_chem %>%
  filter(site_code %in% c("LNGR", "JOER","COMR", "HORR",  "CBRR", "PTRR", "BRNR")) %>%
  select(site_code, Year, pH,SC,TSS, DOC, DTN, K,NH4,NO3, SO4) %>%
  pivot_longer(cols = c(pH,SC,TSS, DOC, DTN, K,NH4,NO3, SO4),
               names_to = "parameter",
               values_to = "value") %>%
  #get the median and max for each site
  group_by(site_code, parameter) %>%
  summarise(median = median(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE),
            max = max(value, na.rm = TRUE)) %>%
  ungroup()%>%
  pivot_longer(cols = c(median, max, sd),
               names_to = "stat",
               values_to = "stat_value")%>%
  unite("param_stat", parameter, stat, sep = " ") %>%
  pivot_wider(names_from = param_stat, values_from = stat_value)%>%
  arrange(factor(site_code, levels = c("LNGR", "JOER", "HORR", "COMR", "CBRR", "PTRR", "BRNR")))%>%
  select(site_code:`NO3 sd`,`SO4 median`:`pH sd`, `SC median`,`SC max`,`SC sd`)%>%
  #round all numeric values to 2 digits
  mutate(across(where(is.numeric), ~ round(.x, 2)))


param_oi <- c( "DOC", "DTN", "K", "NH4", "NO3", "SO4", "TSS", "pH", "SC")

params <- chem_units %>%
  filter(simple %in% param_oi) %>%
  arrange(match(simple, param_oi)) %>%
  pull(Parameters)


# Create a basic flextable
ft2 <- key_table %>%
  rename(Site = site_code) %>%
  flextable() %>%
  set_header_labels(
    .labels = setNames(rep(c("mean", "max", "sd"), times = length(params)),
                      nm = as.vector(t(outer("", c("median", "max","sd"), paste, sep = " "))))) %>%
  delete_part(part = "header")%>%
  add_header_row(
    values = c("", rep(c("Median", "Max", "SD"), times = length(params))),  
    colwidths = c(1, rep(1, length(params)*3))  # Adjust colwidths for Site and parameters
  ) %>%
  add_header_row(
    values = c("Site", params), 
    colwidths = c(1, rep(3, length(params)))  # Adjust colwidths for Site and parameters
  ) %>%
  
  add_header_row(
    values = c("", 
                "mg L⁻¹", "mg L⁻¹", "mg L⁻¹", 
               "mg L⁻¹", "mg L⁻¹", "mg L⁻¹", "mg L⁻¹",  " ", "µS cm⁻¹"), 
    colwidths = c(1, rep(3, length(params)))  # Adjust colwidths again to match
  ) %>%
  merge_v(j = ~Site) %>%
  merge_h(part = "header") %>%
  border_remove() %>%
  border_outer() %>%
  border_inner_v() %>%
  border_inner_h() %>%
  align(align = "center", part = "all") %>%
  valign(valign = "center", part = "all") %>%
  fontsize(size = 11, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  set_table_properties(layout = "autofit", width = 1)%>%
  autofit()

sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    width = 8.3, height = 11.7
  ),
  type = "continuous",
  page_margins = page_mar()
)

save_as_docx( ft2, path = here(paper_draft_dir,"SI_2_table.docx"))

```

## SI Table 3: Mean, SD, Max for remainders for reservoirs

```{r}

base_cations_table <- tidy_chem %>%
  filter(site_code %in% c("LNGR", "JOER","COMR", "HORR",  "CBRR", "PTRR", "BRNR")) %>%
  select(site_code, Year, ANC, Na, Mg, Ca, F, Cl) %>%
  pivot_longer(cols = c(ANC, Na, Mg, Ca, F, Cl),
               names_to = "parameter",
               values_to = "value") %>%
  #get the median and max for each site
  group_by(site_code, parameter) %>%
  summarise(median = median(value, na.rm = TRUE),
            max = max(value, na.rm = TRUE),
            sd = sd(value,na.rm = T)) %>%
  ungroup()%>%
  pivot_longer(cols = c(median, max,sd),
               names_to = "stat",
               values_to = "stat_value")%>%
  unite("param_stat", parameter, stat, sep = " ") %>%
  pivot_wider(names_from = param_stat, values_from = stat_value)%>%
  arrange(factor(site_code, levels = c("LNGR", "JOER", "HORR", "COMR", "CBRR", "PTRR", "BRNR")))%>%
  #round all numeric values to 2 digits
  mutate(across(where(is.numeric), ~ round(.x, 2)))


param_oi <- c("ANC", "Ca", "Cl", "F", "Mg", "Na")

params <- chem_units %>%
  filter(simple %in% param_oi) %>%
  arrange(match(simple, param_oi)) %>%
  pull(Parameters)


# Create a basic flextable
ft3 <- base_cations_table %>%
  rename(Site = site_code) %>%
  flextable() %>%
  set_header_labels(
    .labels = setNames(rep(c("mean", "max", "sd"), times = length(params)),
                      nm = as.vector(t(outer("", c("median", "max","sd"), paste, sep = " "))))) %>%
  delete_part(part = "header")%>%
  add_header_row(
    values = c("", rep(c("Median", "Max", "SD"), times = length(params))),  
    colwidths = c(1, rep(1, length(params)*3))  # Adjust colwidths for Site and parameters
  ) %>%
  add_header_row(
    values = c("Site", params), 
    colwidths = c(1, rep(3, length(params)))  # Adjust colwidths for Site and parameters
  ) %>%
  
  add_header_row(
    values = c("", 
                "mg L⁻¹", "mg L⁻¹", "mg L⁻¹", 
               "mg L⁻¹", "mg L⁻¹", "mg L⁻¹"), 
    colwidths = c(1, rep(3, length(params)))  # Adjust colwidths again to match
  ) %>%
  merge_v(j = ~Site) %>%
  merge_h(part = "header") %>%
  border_remove() %>%
  border_outer() %>%
  border_inner_v() %>%
  border_inner_h() %>%
  align(align = "center", part = "all") %>%
  valign(valign = "center", part = "all") %>%
  fontsize(size = 11, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  set_table_properties(layout = "autofit", width = 1)%>%
  autofit()

sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    width = 8.3, height = 11.7
  ),
  type = "continuous",
  page_margins = page_mar()
)

save_as_docx( ft3, path = here(paper_draft_dir,"SI_3_table.docx"))

```

## RS Results

```{r}

```

## SI Plot 1: Timelines of key parameters at reservoirs

```{r}

SI_plot <- tidy_chem %>%
  filter(site_code %in% c("LNGR", "JOER","COMR", "HORR",  "CBRR", "PTRR", "BRNR"))%>%
  pivot_longer(cols = c( ChlA, DTN, pH, SC, DOC, K, NH4, NO3, SO4, K, Ca, ANC, Na, Mg, Ca, F, Cl),
               names_to = "parameter",
               values_to = "value")%>%
  left_join(chem_units, by = c("parameter" = "simple"))%>%
  mutate(Date = as.POSIXct(Date ))


#plot 1 

Si_timeline_1 <- ggplot(SI_plot%>%filter(parameter %in% c("DOC", "DTN","pH", "SC", "ANC"))%>%
         mutate(parameter = factor(parameter, levels = c("DOC", "DTN","pH", "SC", "ANC"))),
       aes(x = Date, y = value, group = interaction(site_code, Year), color = site_code)) +
  geom_line() +
  geom_point()+
  facet_grid(rows = vars(combined),cols = vars(Year), scales = "free", switch = "y") +
  labs(x = "Date",
       y = "") +
  scale_color_manual(values = colorsBS_site_code)+
  scale_x_datetime(
    date_labels = "%b",  # e.g., May\n2021
    date_breaks = "month"
  ) +
  labs(color = "Site")+
  theme_bw() +
  theme(strip.placement = "outside", 
        legend.position = "bottom",
        # make the x ticks at 45 degrees
        axis.text.x = element_text(angle = 45, hjust = 1))

save_plot_image(Si_timeline_1,"SI_timeline_1.png")


SI_timeline_2 <- 
ggplot(SI_plot%>%filter(parameter %in% c("K", "Ca","SO4", "Na", "Cl", "Mg", "F")),
       aes(x = Date, y = value, group = interaction(site_code, Year), color = site_code)) +
  geom_line() +
  geom_point()+
  facet_grid(rows = vars(combined),cols = vars(Year), scales = "free", switch = "y") +
  labs(x = "Date",
       y = "") +
  scale_color_manual(values = colorsBS_site_code)+
  scale_x_datetime(
    date_labels = "%b",  # e.g., May\n2021
    date_breaks = "month"
  ) +
  labs(color = "Site")+
  theme_bw() +
  theme(strip.placement = "outside", 
        legend.position = "bottom",
        # make the x ticks at 45 degrees
        axis.text.x = element_text(angle = 45, hjust = 1))

save_plot_image(Si_timeline_2,"SI_timeline_2.png")

```

## SI Plot 2: Percent of reservoir samples eutrophic status

In cold water lakes (our study sites), mesotrophic levels are defined as
between 2.6 and 8 µg/L, and eutrophic levels are greater than 8 µg/L.
Under 2.6 µg/L is considered Oligotrophic.

```{r}

chla_status_site_year <- tidy_chem%>%
  filter(!is.na(ChlA))%>%
  filter(site_code %in% primary_study_res)%>%
  mutate(eu_status  = case_when(ChlA >= 2.6 &ChlA <= 8 ~ "Mesotrophic", 
                                ChlA > 8 ~ "Eutrophic",
                                ChlA < 2.6 ~ "Oligotrophic"))%>%
  group_by(site_code, Year, eu_status, burn_group)%>%
  summarise(eu_stats = n())%>%
  ungroup()%>%
  pivot_wider(names_from = eu_status, values_from = eu_stats)%>%
  mutate_at(vars(-Year), ~replace(., is.na(.), 0))%>%
  rowwise()%>%
  mutate(total_samples = sum(c(Eutrophic, Mesotrophic, Oligotrophic)),
    perc_eu = round((Eutrophic/total_samples)*100, digits= 1), 
         perc_meso = round((Mesotrophic/total_samples)*100, digits= 1),
         perc_meso_eu = round(((Eutrophic+Mesotrophic)/total_samples)*100, digits= 1), 
        perc_oligo = round((Oligotrophic/total_samples) *100, digits= 1),
    Year = as.character(Year))

all_years <- chla_status_site_year%>%
  group_by(burn_group)%>%
  summarise(total_samples = sum(total_samples), 
            perc_eu = sum(Eutrophic)/sum(total_samples)*100, 
            perc_meso = sum(Mesotrophic)/sum(total_samples)*100, 
            perc_meso_eu = sum(Eutrophic + Mesotrophic)/sum(total_samples)*100, 
            perc_oligo = sum(Oligotrophic)/sum(total_samples)*100)

long_chla_status <- chla_status_site_year%>%
  select(site_code, Year, total_samples, perc_eu, perc_meso, perc_meso_eu, perc_oligo)%>%
  pivot_longer(cols = c(perc_eu, perc_meso, perc_meso_eu, perc_oligo), names_to = "category", values_to = "value" )

long_chla_status$site_code = factor(long_chla_status$site_code, levels = c("LNGR","JOER" , "CBRR", "PTRR", "BRNR" ))

chla_status <- c(
  perc_eu = "% eutrophic",
  perc_meso = "% mesotrophic ",
  perc_meso_eu = "% mesotrophic & eutrophic",
  perc_oligo = "% oligotrophic"
)

#PLOT
 long_chla_status%>%
  filter(site_code %in% primary_study_res )%>%
   filter(category != "perc_meso_eu")%>%
  ggplot(aes(x = Year, y = value,group = site_code, color =  site_code))+
  scale_color_manual(values = colorsBS_site_code)+
 geom_line()+
   geom_point()+
  theme_bw(base_size = 15)+
facet_wrap(~category, ncol = 1, labeller = labeller(category = chla_status))+
  labs(
    x = "Year",
    y = "Percent of chlorophyll-*a* samples", 
    color = "Reservoir")+
  theme(axis.title.y = ggtext::element_markdown())

```


# Watershed Characteristics (Spatial)

## Merging transbasin diversion watersheds

```{r}

grand_ditch <- st_read("data/raw/spatial/res_study/working_tb/grand_ditch.shp" )%>%
  st_transform(4326)

michigan_ditch <- st_read("data/raw/spatial/res_study/working_tb/michigan_ditch.shp" )%>%
  st_transform(4326)

LRT <- st_read("data/raw/spatial/res_study/working_tb/Laramie_River_Tunnel.shp" )%>%
  st_transform(4326)


mapview::mapview(grand_ditch)+
  mapview::mapview(michigan_ditch)+
  mapview::mapview(LRT)

TB <- bind_rows(
  grand_ditch,
  michigan_ditch,
  LRT
)

#mapview::mapview(TB, zcol = "site_name", burst = T)
#st_write(TB, "data/raw/spatial/res_study/trans_basin_watersheds.shp")

#Old code to merge stream stats generated watershed with relevant transbasin
# 
# polygon3 <- st_read("data/spatial/working/SLEP_ss/layers/globalwatershed.shp" )%>%
#   st_transform(4326)%>%
#   mutate(site_name = "site_streamstats")%>%
#   select(site_name, geometry)
# 
# st_is_valid(polygon3)
#polygon3 <- st_make_valid(polygon3)
# 
# mapview::mapview(polygon3)

# Merge polygons
# merged_geometry <- st_difference(polygon2$geometry, polygon1$geometry)
# #merged_geometry <- st_union(polygon1$geometry, polygon2$geometry)
# 
# merged_polygon <- st_sf(
#   site_name = "HORR",
#   geometry = st_sfc(merged_geometry))%>%
#   st_make_valid()
# 
# st_is_valid(merged_polygon)
# 
# mapview::mapview(merged_polygon)
```


## Landcover

```{r}

#pull in watersheds
watersheds <- st_read(dsn = here(spatial_dir,"cpf_watersheds.shp"))%>%
  left_join(most_recent_meta, by = c("site_code"))

nlcd_tiffs <- list.files(here(external_spatial, "landcover_NLCD"), pattern = "Annual_NLCD_LndCov",  full.names = T)%>%
  #find the files that contain ".tff" and nothing after it
  .[grepl(".tiff$", .)]
# read in rasters and name them based on the year contained in the file name
nlcd_rasters <- map(nlcd_tiffs, raster::raster)
#name them based on the year contained in the file name
names(nlcd_rasters) <- nlcd_tiffs%>%
  stringr::str_extract("\\d{4}")%>%
  as.numeric()


compute_top3_landcovers <- function(year){
  
  NLCD <- nlcd_rasters[[year]]
  
  nlcd_raster <- raster::brick(NLCD)


  ws <- watersheds %>%
    select(site_code)%>%
    rownames_to_column('index')%>%
    mutate(index = as.numeric(index))
  
  # using the exactextractr package, get what is essentially a "zonal histogram" in
  # ArcGIS speak
  ws_NLCD <- exactextractr::exact_extract(nlcd_raster, ws,
                                         append_cols = "index",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           group_by(lc_code = as.character(value)) %>%
                                           summarise(count = sum(coverage_fraction))) %>%
    group_by(index) %>% 
    mutate(total = sum(count)) %>%
    ungroup() %>%
    pivot_wider(names_from = 'lc_code', values_from = 'count',
                names_glue = "{lc_code}" ) %>%
    inner_join(ws, .) -> ws_NLCD
  
  ws_percents <- ws_NLCD %>% 
    #if there are NAs, change it to 0
    mutate(across(`11`:`82`, ~replace_na(., 0))) %>%
    mutate(water=100*(`11`/total),
           ice_snow=100*(`12`/total),
           open_space=100*(`21`/total),
           developed = 100*((`22`+ `23`+ `24`)/total),
           Barren = 100*(`31`/total),
           `Decidious Forest` = 100*((`41`)/total),
           `Evergreen Forest` = 100*((`42`)/total),
           `Mixed Forest` = 100*((`43`)/total),
           `Shrubland` = 100*((`52`)/total),
           `Grassland` = 100*((`71`)/total),
           pasture = 100*((`81`)/total),
           crops = 100*((`82`)/total),
           `Wetlands` = 100*((`90`+ `95`)/total),
           total_check = rowSums(across(water:Wetlands)))%>%
    st_drop_geometry()%>%
    select(total, site_code, water:total_check)
  
  
  top_land_cover <- ws_percents %>%
    select(-total, -total_check) %>%
  # Pivot longer to get land cover types and percentages in rows
  pivot_longer(cols = water:Wetlands, 
               names_to = "land_cover", 
               values_to = "percent")%>%
  # Group by site_code and find the top 4 land covers by percent
  group_by(site_code) %>%
  arrange(desc(percent), .by_group = TRUE) %>%
  slice_head(n = 3) %>%
  mutate(rank = row_number()) %>% # Rank the top 4 land covers
  ungroup() %>%
  # Pivot wider with unique column names
  pivot_wider(
    id_cols = site_code, 
    names_from = rank, 
    values_from = c(land_cover, percent), 
    names_glue = "{.value}_{rank}"
  )%>%
    select(site_code, land_cover_1, percent_1, land_cover_2, percent_2, land_cover_3, percent_3)%>%
    mutate(nlcd_year = year)%>%
    #round all percent columns to 1 decimal places
    mutate(across(contains("percent"), ~round(., 1)))


return(top_land_cover)
  
  
}

compute_all_landcovers <- function(year){
  
  NLCD <- nlcd_rasters[[year]]
  
  nlcd_raster <- raster::brick(NLCD)


  ws <- watersheds %>%
    select(site_code)%>%
    rownames_to_column('index')%>%
    mutate(index = as.numeric(index))
  
  # using the exactextractr package, get what is essentially a "zonal histogram" in
  # ArcGIS speak
  ws_NLCD <- exactextractr::exact_extract(nlcd_raster, ws,
                                         append_cols = "index",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           group_by(lc_code = as.character(value)) %>%
                                           summarise(count = sum(coverage_fraction))) %>%
    group_by(index) %>% 
    mutate(total = sum(count)) %>%
    ungroup() %>%
    pivot_wider(names_from = 'lc_code', values_from = 'count',
                names_glue = "{lc_code}" ) %>%
    inner_join(ws, .) -> ws_NLCD
  
  all_land_cover <- ws_NLCD %>% 
    #if there are NAs, change it to 0
    mutate(across(`11`:`82`, ~replace_na(., 0))) %>%
    mutate(water=100*(`11`/total),
           ice_snow=100*(`12`/total),
           open_space=100*(`21`/total),
           developed = 100*((`22`+ `23`+ `24`)/total),
           Barren = 100*(`31`/total),
           `Decidious Forest` = 100*((`41`)/total),
           `Evergreen Forest` = 100*((`42`)/total),
           `Mixed Forest` = 100*((`43`)/total),
           `Shrubland` = 100*((`52`)/total),
           `Grassland` = 100*((`71`)/total),
           pasture = 100*((`81`)/total),
           crops = 100*((`82`)/total),
           `Wetlands` = 100*((`90`+ `95`)/total),
           total_check = rowSums(across(water:Wetlands)))%>%
    st_drop_geometry()%>%
    select( site_code, total, water:`Wetlands`)%>%
    mutate(nlcd_year = year)

return(all_land_cover)
  
  
}

years <- c("2019", "2023")

ws_lc <- map_df(years, compute_top3_landcovers)%>%
  bind_rows() %>%
  #swapping water for land_cover_3
  mutate(land_cover_2 = case_when(land_cover_2 == "water" ~ land_cover_3, 
                                  TRUE ~ land_cover_2), 
         percent_2 = case_when(land_cover_2 == "water" ~ percent_3, 
                               TRUE ~ percent_2), 
         # creating a new column that combines the top 2 land covers for table
         land_cover_top2 = paste0(land_cover_1, " (", percent_1, " %) \n",land_cover_2, " (", percent_2," %) " ))%>%
  select(site_code, nlcd_year,land_cover_top2) %>%
  pivot_wider(names_from = nlcd_year, values_from = land_cover_top2, names_prefix = "lc_top2_")

#run to get all landcover %s for watersheds
ws_all_lc <- map_df(years, compute_all_landcovers)%>%
  bind_rows()%>%
  pivot_longer(
    cols = water:Wetlands, 
    names_to = "landcover", 
    values_to = "value"
  ) %>%
  select(-total)%>%
  pivot_wider(,
    names_from = nlcd_year, 
    values_from = value,
    names_prefix = "value_" # This creates value_2019 and value_2023
  ) %>%
  select(site_code, landcover, value_2019, value_2023)%>%
  mutate(pp_change = value_2023 - value_2019)


rm(years)
```

## Slope, elevation and aspect

```{r}

ws_chars<- map_dfr(unique(watersheds$site_code), function(ws_code){
  
  ws <- watersheds %>%
    filter(site_code == ws_code)
  
   ws <- st_make_valid(ws)
  
  if(ws_code %in% c("JOER", "CBRR", "BRNR", "LNGR", "PTRR", "COMR","HORR")){
    z_num = 10
  }else{
    z_num = 9
  }
  
  elev <- get_elev_raster(ws, z = z_num, verbose = F)

  ws_elev <- exactextractr::exact_extract(elev, ws,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           summarise(mean_elev = mean(value, na.rm = T), 
                                                     max_elev = max(value, na.rm = T), 
                                                     min_elev = min(value, na.rm = T))) %>%
    select(site_code, mean_elev, max_elev, min_elev)
  
  
  #calculate slope
  slope <- raster::terrain(elev, opt = "slope", unit = "degrees", verbose = F)
  ws_slope <- exactextractr::exact_extract(slope, ws,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           summarise(mean_slope = mean(value, na.rm = T), 
                                                     max_slope = max(value, na.rm = T), 
                                                     min_slope = min(value, na.rm = T))) %>%
    select( mean_slope, max_slope)
  
  #calculate aspect
  aspect <- raster::terrain(elev, opt = "aspect", unit = "degrees")
  ws_aspect <- exactextractr::exact_extract(aspect, ws,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                          summarise(mean_aspect = mean(value, na.rm = T))) %>%
    select(mean_aspect)%>%
    round(2)
  # calculate watershed area
  ws_area <- st_area(ws)%>%
    #convert to km2
    units::set_units("km^2")
  
  bind_cols(ws_elev, ws_slope, ws_aspect)%>%
    mutate(ws_area_km2 = ws_area)
  
})

```
## Calculating snow persistance for watersheds

```{r}

#get the snow persistence raster
snow <- raster::raster(here(external_spatial, "snow_persistance/MOD10A2_SCI_2020.tif"))

ws_snow <- map_dfr(unique(watersheds$site_code),  function(ws_code){
  
  ws <- watersheds %>%
    filter(site_code == ws_code)
  exactextractr::exact_extract(snow, ws,
                                          append_cols = "site_code",
                                          summarize_df = T,
                                          fun=function(x) x %>%
                                            summarise(mean_snow = mean(value, na.rm = T), 
                                                      max_snow = max(value, na.rm = T), 
                                                      min_snow = min(value, na.rm = T))) %>%
    select(site_code, mean_snow, max_snow, min_snow)%>%
    mutate(across(where(is.numeric), ~round(., 0)))%>%
    mutate(days = 184 * (mean_snow/100), 
           mean_snow_free_date= format(as.Date(days, origin = "2020-01-01"), "%m-%d"))%>%
    select(site_code, mean_snow, mean_snow_free_date)
})

```

## CPF Elevational Gradient

```{r}
ws <-  terra::rast(here(spatial_dir, "cpf_sbs_BAER.tif"))
#remove 0 values
#ws[ws == 1] <- NA
ws[ws == 0] <- NA

#convert to shapefile
# Convert raster to polygon (dissolve all cells into one)
outline <- as.polygons(ws, dissolve = TRUE, na.rm = TRUE) %>%
  st_as_sf()

outline$site_code <- "cpf"

outline <- outline %>%
  group_by(site_code) %>%
  summarise(geometry = st_union(geometry), .groups = "drop") %>%
  st_cast("POLYGON")

elev <- get_elev_raster(outline, z = 9, verbose = F)



ws_elev <- exactextractr::exact_extract(elev, outline,
                                        append_cols = "site_code",
                                        summarize_df = T,
                                        fun=function(x) x %>%
                                          summarise(mean_elev = mean(value, na.rm = T), 
                                                    max_elev = max(value, na.rm = T), 
                                                    min_elev = min(value, na.rm = T)))%>%
  select(site_code, mean_elev, max_elev, min_elev)


ws_area <- st_area(outline)%>%
  #convert to km2
  units::set_units("km^2")


cpf<- bind_cols(ws_elev)%>%
  mutate(ws_area_km2 = ws_area)%>%
  group_by(site_code)%>%
  summarise(mean_elev = max(mean_elev),
            max_elev = max(max_elev),
            min_elev = min(min_elev),
            ws_area_km2 = sum(ws_area_km2))
```



## Soil Burn Severity

From Rhoades paper : "We partitioned the catchments into groups based on the proportional area burned at high + moderate severity (Extensive:>50%, High: 25–50%, Mid:<25%). Unburned catchments and those with <25% total burn extent, no high severity and little moderate severity (0–13%) were included in a No/Low class."

```{r}

sbs <- terra::rast(here(spatial_dir, "cpf_sbs_BAER.tif"))

# Load the shapefile and raster
ws <- st_read(here(spatial_dir, "cpf_watersheds.shp")) %>%
  st_transform(sf::st_crs(sbs)) 

ws_raster <- ws %>%
  terra::vect()

# Ensure all NA values in the raster are already set to 5
sbs[is.na(sbs) | is.null(sbs) | sbs == 0] <- 5

# Get the extent of the ws vector
ws_extent <- terra::ext(ws_raster)

# Extend the raster to match the ws extent, assigning new areas the value of 5 (Unburned)
sbs_extended <- terra::extend(sbs, ws_extent, fill = 5)

# Get the cell size in map units (meters)
cell_size <- terra::res(sbs_extended)

cnv_to_km2 <- (cell_size[1] * cell_size[2]) / 1e6

# using the exactextractr package, get what is essentially a "zonal histogram" in
# ArcGIS speak
ws_extracted_sbs <- exactextractr::exact_extract(sbs_extended %>% raster::raster(), ws,
                                              append_cols = "site_code",
                                              summarize_df = T,
                                              fun=function(x) x %>%
                                                group_by(sbs_code = as.character(value)) %>%
                                                summarise(count = sum(coverage_fraction))) %>%
  group_by(site_code)%>%
  mutate(total = sum(count)) %>%
  ungroup() %>%
  pivot_wider(names_from = 'sbs_code', values_from = 'count',
              names_glue = "sbs{sbs_code}" ) %>%
  inner_join(ws, .) -> ws_sbs

ws_sbs_full <- ws_extracted_sbs %>% 
  ungroup()%>%
  st_drop_geometry()%>%
  #if a column is NA change it to 0
  mutate(across(starts_with("sbs"), ~replace_na(., 0))) %>%
  mutate(
    # 1. Define raw pixel counts for burn classes
    Unburned = sbs5 + sbs6,
    `Very Low` = sbs1,
    Low = sbs2,
    Moderate = sbs3,
    High = sbs4,
    
    # 2. Convert to Square Kilometers (km2)
    Unburned_km2 = Unburned * cnv_to_km2,
    VeryLow_km2 = `Very Low` * cnv_to_km2,
    Low_km2 = Low * cnv_to_km2,
    Moderate_km2 = Moderate * cnv_to_km2,
    High_km2 = High * cnv_to_km2,
    total_km2 = total * cnv_to_km2,
    
    # 3. Calculate Percentages (Assumes 'total' column exists in raw pixels)
    Unburned_pct = (Unburned / total) * 100,
    VeryLow_pct = (`Very Low` / total) * 100,
    Low_pct = (Low / total) * 100,
    Moderate_pct = (Moderate / total) * 100,
    High_pct = (High / total) * 100
  )%>%
  mutate(across(Unburned:High_pct, ~round(., 1)))%>%
  mutate(extent_burned = ((`Very Low` + Low + Moderate + High)),
         extent_burned_km2 = extent_burned * cnv_to_km2,
         extent_burned_pct = (extent_burned / total) * 100,
         high_burn = (Moderate + High)/total*100,
         burn_class = case_when(high_burn > 50 ~ "Extensive", 
                                high_burn > 25 & high_burn <= 50 ~ "High", 
                                high_burn < 25 & Moderate_pct <= 13 & high_burn <= 5 ~ "No/Low",
                                TRUE ~ "Mid"))

ws_sbs_concise <- ws_sbs_full%>%
  #filter(site_code %in% c("CBRR", "BRNR", "PTRR","COMR", "HORR", "JOER", "LNGR"))%>%
  select(site_code, 
         `Total Area (km2)` = total_km2,
         `Extent Burned (km2)` = extent_burned_km2,
         `Extent Burned (%)` = extent_burned_pct,
         `High Burn (%)` = high_burn, 
         `Watershed Burn Class` = burn_class)
  


```

### Buffer SBS

```{r}
waterbody <- st_read("data/raw/spatial/res_study/study_waterbodies.shp")%>%
  st_transform(sf::st_crs(sbs)) #transform to coordinates in sbs

waterbody_buffer <- waterbody%>%
  #add a buffer of 200 meters to the waterbodies
  st_buffer(200) 

# pick out watersheds of reservoirs
res_ws <- ws%>%
  filter(site_code %in% c("JOER","LNGR","COMR","HORR","CBRR","PTRR", "BRNR"))%>%
  st_transform(sf::st_crs(sbs))

#clip buffer to it's watershed
waterbody_buffer <- st_intersection(waterbody_buffer,res_ws )%>%
  #only keep the waterbodies that are in the same watershed as the reservoir
  filter(site_code == site_code.1)

#remove water body from buffer
final_buffer <- st_difference(waterbody_buffer, waterbody)%>%
  filter(site_code == site_code.1 & site_code.1 == site_code.2)

#map to view
# tm_basemap("OpenStreetMap") +
# tm_shape(final_buffer) +
#   tm_polygons(col = "site_code", palette = "Set2", alpha = 0.5, border.col = "black", border.lwd = 1, title = "Buffered Waterbodies")+
#   #put the legend inside the plot
#   tm_layout(legend.position = c("right", "bottom"), title = "Buffered Waterbodies within Watersheds", title.position = c("right", "top"))


# Get the extent of the ws vector
waterbody_buffer_extent <- terra::ext(final_buffer)

# Extend the raster to match the ws extent, assigning new areas the value of 5
sbs_extended <- terra::extend(sbs, waterbody_buffer_extent, fill = 5)

# Get the cell size in map units (meters)
cell_size <- terra::res(sbs_extended)

  extracted_sbs <- exactextractr::exact_extract(sbs_extended %>% raster::raster(), final_buffer,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           group_by(sbs_code = as.character(value)) %>%
                                           summarise(count = sum(coverage_fraction))) %>%
    group_by(site_code)%>%
    mutate(total = sum(count)) %>%
    ungroup() %>%
    pivot_wider(names_from = 'sbs_code', values_from = 'count',
                names_glue = "sbs{sbs_code}" ) %>%
    inner_join(ws, .) -> ws_sbs
  
  

  
  waterbody_buffer_percents <- extracted_sbs %>% 
    ungroup()%>%
    st_drop_geometry()%>%
    #if a column is NA change it to 0
    mutate(across(starts_with("sbs"), ~replace_na(., 0))) %>%
    mutate(
      # 1. Define raw pixel counts for burn classes
      Unburned = sbs5 + sbs6,
      `Very Low` = sbs1,
      Low = sbs2,
      Moderate = sbs3,
      High = sbs4,
      
      # 2. Convert to Square Kilometers (km2)
      Unburned_km2 = Unburned * cnv_to_km2,
      VeryLow_km2 = `Very Low` * cnv_to_km2,
      Low_km2 = Low * cnv_to_km2,
      Moderate_km2 = Moderate * cnv_to_km2,
      High_km2 = High * cnv_to_km2,
      total_km2 = total * cnv_to_km2,
      
      # 3. Calculate Percentages (Assumes 'total' column exists in raw pixels)
      Unburned_pct = (Unburned / total) * 100,
      VeryLow_pct = (`Very Low` / total) * 100,
      Low_pct = (Low / total) * 100,
      Moderate_pct = (Moderate / total) * 100,
      High_pct = (High / total) * 100
    )%>%
    mutate(across(Unburned:High_pct, ~round(., 1)))%>%
    mutate(extent_burned = ((`Very Low` + Low + Moderate + High)),
           extent_burned_km2 = extent_burned * cnv_to_km2,
           extent_burned_pct = (extent_burned / total) * 100,
           high_burn = (Moderate + High)/total*100,
           burn_class = case_when(high_burn > 50 ~ "Extensive", 
                                  high_burn > 25 & high_burn <= 50 ~ "High", 
                                  high_burn < 25 & Moderate_pct <= 13 & high_burn <= 5 ~ "No/Low",
                                  TRUE ~ "Mid"))
  
  ws_burn_info <- ws_sbs_concise %>%
    left_join(waterbody_buffer_percents %>% 
                select(site_code,
                       `200m Buffer Total Area (km2)` = total_km2,
                       `200m Buffer Burn (km2)` = extent_burned_km2,
                       `200m Buffer Burn (%)` = extent_burned_pct,
                       `200m Buffer High Burn (%)` = high_burn,
                       `200m Buffer Burn Class` = burn_class), by = "site_code")
  
```

## Get data together for Reservoir characteristics table

```{r}
# Reservoir area, normal storage capacity and reservoir depth from NID database
res_meta <- read_csv(here(spatial_dir, "reservoir_metadata.csv"))%>% rename(`Reservoir Operator` = Manager)

all_ws_data <- most_recent_meta %>%
                    select(site_code, Site_Name)%>%
  left_join(ws_chars, by = c("site_code"))%>%
  left_join(ws_burn_info, by = c("site_code"))%>%
  #left_join(ws_all_lc, by = c("site_code"))%>%
  left_join(ws_lc, by = c("site_code"))%>%
  left_join(res_meta, by = c("site_code"))%>%
  #mutate all numeric columns to round to 2 decimal places
  mutate(across(where(is.numeric), ~round(., 2))) %>%
   mutate(Site = paste0(Site_Name, " (", site_code, ")"))%>%
  filter(!is.na(min_elev))%>%
  #cleaning up the names
  select( `Site (site code)` = Site,
          site_code, 
          `Watershed Area (km^2)` = ws_area_km2,
          `Reservoir Elevation (m)` = min_elev,
          `Max Watershed Elevation (m)` = max_elev,
          `Mean Watershed Elevation (m)` = mean_elev, 
          `Normal Storage Capacity (acre ft)`, 
          `Reservoir Depth (m)`, 
          `Reservoir Operator`,
          `Extent Burned (km2)`,
          `Extent Burned (%)`,
          `High Burn (%)`,
          `Watershed Burn Class`,
          `200m Buffer Total Area (km2)` ,
          #`200m Buffer Burn (km2)`,
          `200m Buffer Burn (%)`,
          #`200m Buffer High Burn (%)`,
          `200m Buffer Burn Class`,
          `Mean Slope (degrees)` = mean_slope,
          `2019 Top 2 Land Covers` = lc_top2_2019,
          `2023 Top 2 Land Covers` = lc_top2_2023)

res_ws_data <- all_ws_data%>%
    filter(site_code %in% c("CBRR", "BRNR", "PTRR","COMR", "HORR", "JOER", "LNGR"))%>%
  select(-site_code)

write_csv(res_ws_data, here(spatial_dir, "reservoir_characteristics.csv"))
write_csv(all_ws_data, here(spatial_dir, "watershed_characteristics.csv"))
```

## Make a table with the land cover and watershed characteristics

```{r}
res_ws_data <- read_csv(here(spatial_dir, "reservoir_characteristics.csv"))
watershed_characteristics <- read_csv(here(spatial_dir, "watershed_characteristics.csv"))

t_res_ws_data<- res_ws_data%>%
  t()%>%
  as.data.frame()%>%
  #make column names the first row
  setNames(.[1,])%>%
  # remove first row
  .[-1,]%>%
  rownames_to_column(" ")

#create standard border color and width
std_border <- fp_border_default(width = 1, color = "black")

ft <- flextable(t_res_ws_data)%>%
  separate_header() %>%
  align(align = "center", part = "all")%>%
  colformat_double( big.mark = ",", decimal.mark = ".", digits = 3)%>%
  set_table_properties(layout = "fixed")%>%
  border_remove()%>%
  border_outer(part="all", border = std_border )%>%
  border_inner_h(border = std_border, part="all")%>%
  border_inner_v(border = std_border, part="all")%>%
  autofit()%>%
  fontsize( size = 6, part = "body")%>%
  fontsize(size = 6, part = "header") %>% # Smaller font size for the header
  bold(part = "header") 

sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    width = 8.3, height = 11.7
  ),
  type = "continuous",
  page_margins = page_mar()
)
# 
#ft
# 
# # Save the table to a Word document
#save_as_docx(ft, path = "data/figs_output/res_ws_data_table.docx", pr_section = sect_properties)
```


# Maps

### Read in shapefiles for watersheds, sampling locs and waterbodies

```{r}

## Pull in sampling points

sampling_points <- most_recent_meta %>%
  #filter for all upper sites (or reservoirs)
  filter(grepl("Upper|Reservoir", watershed))%>%
  st_as_sf(coords = c("Long", "Lat"), crs = 4326)

sampling_points_study_res <- sampling_points%>%
  filter(grepl("Reservoir", watershed))%>%
  mutate(ws_site_code = case_when(
    watershed == "Barnes Meadow Reservoir" ~ "BRNR",
    watershed == "Joe Wright Reservoir" ~ "JOER",
    watershed == "Long Draw Reservoir" ~ "LNGR",
    watershed == "Chambers Reservoir" ~ "CBRR",
    watershed == "Comanche Reservoir" ~ "COMR",
    watershed == "Hourglass Reservoir" ~ "HORR",
    watershed == "Peterson Reservoir" ~ "PTRR"))%>%
  select(site_code, Site_Name, watershed, ws_site_code)
    
#st_write(sampling_points_study_res, "data/raw/spatial/sampling_points.shp")


study_waterbodies <- st_read("data/raw/spatial/res_study/study_waterbodies.shp")%>%
  st_transform(4326)%>%
  select(site_code, Site_Name, watershed)



```

### Reservoirs with SBS

```{r}

# Load the raster
sbs <- raster::raster("data/raw/spatial/res_study/cpf_sbs_BAER.tif")
#project into 4326
sbs <- projectRaster(sbs, crs = 4326)

# Remove water bodies  and blank spaces (0 & 6)
sbs[sbs == 6] <- NA
sbs[sbs == 0] <- NA
# Define burn severity colors and labels

#clip sbs_extended to watershed extent
ws_extent <- ws%>%filter(site_code %in% c("JOER","COMR","HORR","CBRR","PTRR", "BRNR"))%>%
  st_transform(sf::st_crs(sbs_extended))%>%
  st_bbox() %>%
  #add a buffer of 1000 meters to the extent
  st_as_sfc() %>%
  st_buffer(1000)

sbs_clip <- terra::crop(sbs_extended, ws_extent)
#   library(tmap)
# 
# # Switch to interactive mode
tmap_mode(mode = "plot")
severity_cols <- c(
  "#31A354", # 1: Very Low (Pale Yellow)
  "#FFFFB2", # 2: Low (Yellow-Orange)
  "#FD8D3C", # 3: Moderate (Orange)
  "#E31A1C", # 4: High (Red)
  "#31A354", # 5: Unburned (Green)
  "steelblue"  # 6: Water (Blue)
)
  
# Create the map
tm_shape(sbs_clip) +
tm_raster(
  palette = severity_cols,
  breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5),
  labels = c("Very Low", "Low", "Moderate", "High", "Unburned", "Water"),
  opacity = 0.7,
  title = "Burn Severity"
) +
  tm_shape(ws%>%filter(site_code %in% c("JOER","LNGR","COMR","HORR","CBRR","PTRR", "BRNR"))) +
  tm_polygons(
    col = "site_code", 
    palette = "Set2",       # Qualitative palette for distinct site colors
    title = "Monitoring Site",
    alpha = 0.6             # Adds slight transparency to see terrain underneath
  ) +
  tm_borders(col = "black", lwd = 2) +
  tm_shape(study_waterbodies)+
  tm_polygons(col = "blue", alpha = 0.5, border.col = "black", border.lwd = 1, title = "Waterbodies") 
  tm_layout(title = "Watershed vs. SBS Overlap")
```
# Extra Analysis

## Flow data

```{r}

api_creds <- yaml::read_yaml(file = "creds/CDWRCreds.yml")$api_key

# pull dwr claftcco data from 1980-2021
q_data <- get_sw_ts(abbrev = "CLAFTCCO", timescale = "day", start_date = "1980-10-01", end_date = "2020-09-30", api_key = api_creds)
  
q_data_clean <- q_data%>%
  mutate(q_cfs = value, 
           site_code = abbrev, 
         acre_ft_day = q_cfs *1.983 )%>%
    distinct()%>%
  # remove rows where flag_b is "Ice"
  filter(!grepl("Ice", flag_b))%>%
  select(abbrev, Date = datetime, q_cfs, acre_ft_day, flag_a, flag_b)%>%
  dataRetrieval::addWaterYear()

q_annual <- q_data_clean%>%
  group_by(waterYear)%>%
  summarise(t_flow_acft = round(sum(acre_ft_day), 0), 
            max_cfs = max(q_cfs), 
            #find the Date with the highest q_cfs per year
            max_cfs_date = Date[which.max(q_cfs)],
            )%>%
  ungroup()%>%
  mutate(peak_dy = yday(max_cfs_date), 
         month = month(max_cfs_date), 
         t_flow_cm = round(t_flow_acft * 1233.48, 0)) # convert acre-feet to cubic meters

mean(q_annual$t_flow_acft)
mean(q_annual$t_flow_cm)/1000

```

## Prism Precip

```{r}

# 1. Setup
# Set a directory for PRISM to store the large .bil files
prism_set_dl_dir("data/raw/external_datasets/prism/precip", create = T)
aoi <- st_read("data/raw/spatial/res_study/cpf_watersheds.shp")%>%
  filter(site_code == "PFAL")

# 2. Download Data
# Download 800m Normals (PPT)
get_prism_normals(type = "ppt", resolution = "4km", annual = TRUE, keepZip = FALSE)


# Download Annual data for 2021-2024
get_prism_annual(type = "ppt", years = 2021:2026,resolution = "4km", keepZip = FALSE)


# 1. Load the Annual Data (2021-2024)
annual_files <- prism_archive_subset(
  type = "ppt", 
  years = 2021:2026, 
  resolution = "4km", 
  temp_period = "annual"
)
r_annuals <- terra::rast(pd_to_file(annual_files))

# 2. Load the 30-Year Normals (manual due to file naming issues)
r_normal <- terra::rast("data/raw/external_datasets/prism/precip/prism_ppt_us_25m_2020_avg_30y/prism_ppt_us_25m_2020_avg_30y.tif")


# 5. Spatial Extraction
# Ensure AOI and Rasters are in the same CRS (PRISM is usually NAD83)
aoi_transformed <- st_transform(aoi, st_crs(r_normal))

# Calculate mean precipitation across the AOI for each layer
precip_stats <- tibble(
  period = c("Normal", names(r_annuals)),
  precip_mm = unlist(c(
    exact_extract(r_normal, aoi_transformed, 'mean'),
    exact_extract(r_annuals, aoi_transformed, 'mean')
  ))
)

# 6. Clean and Compare
comparison_df <- precip_stats %>%
  mutate(
    # Extract year from PRISM filename or label
    year = case_when(
      period == "Normal" ~ "1991-2020 Normal",
      TRUE ~ str_extract(period, "\\d{4}")
    )
  ) %>%
  select(year, precip_mm) %>%
  # Calculate deviation from normal
  mutate(
    normal_val = precip_mm[year == "1991-2020 Normal"],
    diff_mm = precip_mm - normal_val,
    pct_of_normal = (precip_mm / normal_val) * 100
  )

print(comparison_df)

other_years <- comparison_df%>%filter(!(year %in% c("1991-2020 Normal", "2025")))
other_years$diff_mm
mean(other_years$diff_mm)

# Quick visualization
ggplot(comparison_df %>% filter(year != "1991-2020 Normal"), 
       aes(x = year, y = precip_mm)) +
  geom_col(fill = "steelblue") +
  geom_hline(aes(yintercept = normal_val), linetype = "dashed", color = "red") +
  labs(title = "Annual Precipitation vs. 30-Year Normal",
       subtitle = "Red dashed line indicates 1991-2020 normal",
       y = "Precipitation (mm)", x = "Year") +
  theme_minimal()

```

## PRISM Summer Temp

```{r}

# 2. Download Data
# Download Monthly Normals (TMEAN) for June-Sept
get_prism_normals(type = "tmean", resolution = "4km", mon = 4:10, keepZip = FALSE)

# Download Monthly data for June-Sept for years 2021-2025 
# (Note: 2026 data is not yet available for all summer months)
get_prism_monthlys(type = "tmean", years = 2021:2025, mon = 4:10, keepZip = FALSE)

# 3. Load and Average the Summer Normals (June-Sept)
folders <- list.files("data/raw/external_datasets/prism/precip",pattern = "avg",full.names = T)%>%grep("tmean",., value = T)

#files in each folder with nothing after ".tif"
files <- map(folders, list.files, full.names = T, recursive = T, pattern = "\\.tif$")%>%
  flatten_chr()
#read in files
r_normals <- rast(files)


# 2. Load all Individual Monthly Rasters (2021-2025)
# We grab all months at once instead of averaging them in a loop
annual_monthly_folders <- prism_archive_subset(type = "tmean", temp_period = "monthly", years = 2021:2025, mon = 4:10, resolution = "4km")
r_monthlies <- rast(pd_to_file(annual_monthly_folders))

# 3. Spatial Extraction
aoi_transformed <- st_transform(aoi, st_crs(r_normals))

# Extract Normals
df_normals <- tibble(
  layer_name = names(r_normals),
  normal_temp = unlist(exact_extract(r_normals, aoi_transformed, 'mean'))
) %>%
  mutate(month = str_extract(layer_name, "(?<=2020)\\d{2}")) # Extract month from "...._06"

# Extract Actuals
df_actuals <- tibble(
  layer_name = names(r_monthlies),
  actual_temp = unlist(exact_extract(r_monthlies, aoi_transformed, 'mean'))
) %>%
  mutate(
    year = str_extract(layer_name, "\\d{4}"),
    month = str_extract(layer_name, "\\d{4}(\\d{2})", group = 1) # Extract month
  )

# 4. Join and Compare
comparison_df <- df_actuals %>%
  left_join(df_normals %>% select(month, normal_temp), by = "month") %>%
  mutate(
    diff_c = as.numeric(actual_temp - normal_temp),
    month_name = month.abb[as.numeric(month)]
  ) %>%
  # Order months correctly for plotting
  mutate(month_name = factor(month_name, levels = c("Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct")))

print(comparison_df)

# 5. Visualization: Faceted by Month
ggplot(comparison_df, aes(x = year, y = actual_temp)) +
  geom_col(aes(fill = diff_c > 0)) +
  facet_wrap(~month_name, scales = "free_y") +
  geom_errorbar(aes(yintercept = normal_temp, ymin = normal_temp, ymax = normal_temp, color = "Normal"), 
                linewidth = 1, linetype = "dashed", width = 0.8) +
  scale_fill_manual(values = c("steelblue", "firebrick"), labels = c("Cooler", "Warmer"), name = "Status") +
  scale_color_manual(values = c("Normal" = "black"), name = "Reference") +
  labs(
    title = "Monthly Temperature Comparison (June-Sept)",
    subtitle = "Comparing individual months to their respective 1991-2020 Normals",
    y = "Mean Temperature (°C)",
    x = "Year"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

