---
title: "Figures"
author: "Sam Struthers"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---
Setup
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

### ----- Load packages ----- ###
package_load <- function(package_names){
  for(i in 1:length(package_names)){
    if(!package_names[i] %in% installed.packages()){
      install.packages(package_names[i])
    }
    library(package_names[i],character.only = TRUE)
  }
}

#vector of packages
pack_req <- c( 
  # data wrangling packages
  "janitor", "broom","readxl","here","arrow",
  #spatial packages
  "sf","terra","raster","tmap",
  # plotting
  "ggpubr","ggthemes","patchwork", "geomtextpath","plotly", "ggpmisc","officer", "flextable","kableExtra",
  #external datasets
  "cdssr","nhdplusTools","elevatr",
  #stats
  "trend", "anomalize","rstatix", "car","performance","lme4","glmmTMB","emmeans",
  #always load tidyverse last to avoid conflicts with other packages
  "tidyverse")
package_load(pack_req)

remove(pack_req, package_load)

#basic function to save a plot in standard size, dpi and width into figures folder
save_plot_image<- function(plot,name){
  
  ggsave(paste0("docs/cpf_reservoir/paper_draft/",name,".jpg"), dpi = 300, width = 18, height = 10)
  
}

paper_draft_dir <- here("docs/cpf_reservoir/paper_draft/")
spatial_dir <- "data/raw/spatial/res_study"
external_spatial <- "data/raw/external_datasets"


#Simple function to negate %in%

`%nin%` = Negate(`%in%`)

# # Distance order and xdistance
order_table <- tibble(
  site_code = c('JOEI', 'JOER', 'CBRI', 'CBRR', 'CHD', 'BRNR', 'BMD', 'LNGR', 'LNGO', 'PTRR', 'PTRO', 'COMI', 'COMR', 'COMO', 'HORI', 'HORR', 'HORO', 'BEAV', 'JWC', 'PJW', 'PFAL', 'SLEP', 'PBR', 'SFM', 'PSF', 'PNF', 'PBD'),
  Distance = c('1 - JOEI', '2 - JOER', '3 - CBRI', '4 - CBRR', '5 - CHD', '6', '7', '8', '9', '99', '999', '11 - COMI', '12 - COMR', '13 - COMO', '14 - HORI', '15 - HORR', '16 - HORO', '17 - BEAV', '18 - JWC', '19 - PJW','20- SLEP', '20 - SLEP', '21 - PBR', '22 - SFM', '23 - PSF', '24 - PNF', '25 - PBD'),
  Xdistance = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 99, 999, 12, 13, 14, 15, 16, 17, 18, 19,20, 20, 21, 22, 23, 24, 25, 26),
  order = c('03', '04', '05', '06', '07', '10', '11', '01', '02', '08', '09', '12', '13', '14', '15', '16', '17', '18', '19','20', '20', '21', '22', '23', '24', '25', '26'), 
  BS_order =  c(NA, '2', NA, '4', NA, '7', NA, '1', NA, '6', NA, NA, '5', NA, NA, '3', NA, NA, NA, NA, NA, NA, NA, NA, NA, NA,NA),
)


# Create the lookup table with expression syntax
# Define as a named list
chem_expressions <- list(
  "Turbidity" = expression(Turbidity ~ (NTU)),
  "TSS"       = expression(TSS ~ (mg ~ L^{-1})),
  "ChlA"      = expression(Chlorophyll ~ italic("a") ~ (mu*g ~ L^{-1})),
  "TC"        = expression(TC ~ (mg ~ L^{-1})),
  "TN"        = expression(TN ~ (mg ~ L^{-1})),
  "DOC"       = expression(DOC ~ (mg ~ L^{-1})),
  "DTN"       = expression(DTN ~ (mg ~ L^{-1})),
  "pH"        = expression(pH),
  "ANC"       = expression(ANC ~ (mu*eq ~ L^{-1})),
  "SC"        = expression(SC ~ (mu*S ~ cm^{-1})),
  "Na"        = expression(Na^{"+"} ~ (mg ~ L^{-1})),
  "NH4"       = expression(NH[4]^{"+"} ~ (mg ~ L^{-1})),
  "K"         = expression(K^{"+"} ~ (mg ~ L^{-1})),
  "Mg"        = expression(Mg^{"2+"} ~ (mg ~ L^{-1})),
  "Ca"        = expression(Ca^{"2+"} ~ (mg ~ L^{-1})),
  "F"         = expression(F^{"-"} ~ (mg ~ L^{-1})),
  "Cl"        = expression(Cl^{"-"} ~ (mg ~ L^{-1})),
  "NO3"       = expression(NO[3]^{"-"} ~ (mg ~ L^{-1})),
  "PO4"       = expression(PO[4]^{"3-"} ~ (mg ~ L^{-1})),
  "SO4"       = expression(SO[4]^{"2-"} ~ (mg ~ L^{-1}))
)

 primary_study_res <-c("CBRR", "JOER", "PTRR",  "BRNR", "LNGR")
 all_reservoirs <- c("COMR", "HORR", "CBRR", "JOER", "PTRR", "BRNR", "LNGR")
 all_res_system <- c("COMI", "COMR", "COMO", "HORI","HORR" , "HORO","CBRR", "CHD", "JOEI", "JOER", "CBRI", "PTRR", "PTRO", "BMD", "BRNR", "LNGO", "LNGR")

### ----- Color Sets ----- ###

cbbPalette <- c( "#999999","#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#Color for reservoirs, matches burn severity by site code

colorsBS_site_code <- c("BRNR" = "#d71a1d", "PTRR" ="#F3722C", "COMR" = "#F8961E", "CBRR" = "#F9C74F", "HORR" = "#F0E442", "JOER" = "#56B4E9", "LNGR" = "#0072B2")
#Seasonal Color values
season_clp_color_vals <- c("Winter" = "#03045e",  "Rising Limb" = '#047E82',  "Peak Flow" = '#59B851', "Falling Limb"='#FA850E', "Fall" = "#772f1a")

```


# Grab data

```{r}
#discrete sample data and location metadata from most recent pub
data_folder <- here("data/raw/chem/ross_clp_chem/v2025.11.14")

most_recent_chem <- read_rds(here(data_folder, "data/cleaned/CLP_chemistry_up_to_20251114.rds"))%>%
  mutate(flow_gauge_id = ifelse(flow_gauge_source == "USGS", paste0("0", flow_gauge_id), flow_gauge_id))
most_recent_meta <- read_csv_arrow(here(data_folder,"data/metadata/location_metadata.csv"))
units_meta_file <- here(data_folder, "data/metadata/water_chemistry_units.xlsx" )
chem_units <- readxl::read_xlsx(units_meta_file)
all_params <- readxl::read_xlsx(units_meta_file)%>%
  pull(simple)

rm(units_meta_file)


tidy_chem <- most_recent_chem%>%
  #remove random pre fire data 
  filter(Year > 2020 & Year < 2025)%>%
  filter(watershed %nin% c("CLP Mainstem - Lower","CLP Tributary - Lower"))%>%
  # pfal + slep represent the same site but were moved to match future sensor location
  mutate(distance_upstream_km = case_when(site_code == "PFAL" ~ 73.9, TRUE ~ distance_upstream_km))%>%
  mutate(location_type = if_else(site_code == "PFAL", "Mainstem", location_type))%>%
  # add order table which gives the sites an order to be plotted in
  left_join(order_table, by = "site_code")%>%
  #filter site code BRNR with chla value of 0.02, seems to be sample error
  filter(!(site_code == "BRNR" & ChlA == 0.02))%>%
  #filter site code CBRR with TSS value > 1000, seems to be sample error
  mutate( TSS = case_when((site_code == "CBRR" & TSS > 250) ~ NA_real_,
                          TRUE ~ TSS),
           Turbidity = case_when((site_code == "CBRR" & Turbidity > 250) ~ NA_real_,
                          TRUE ~ Turbidity), 
          Year = as.character(Year), 
          month = as.character(month(Date)),
          season = case_when(
            month %in% c("3", "4") ~ "Rising Limb", 
            month %in% c( "5","6") ~ "Peak Flow",
            month %in% c("7", "8") ~ "Falling Limb",
            month %in% c("9", "10", "11") ~ "Fall Baseflow", 
            TRUE ~ "Winter"),
          burn_group = case_when(site_code %in% c("CBRR", "BRNR", "PTRR") ~ "burned", 
                                 TRUE ~ "unburned"), 
          burn_year = paste(burn_group, Year, sep = "_"), 
          site_year = paste(site_code, Year, sep = "_")
  )

```

## Run above to get data loaded

```{r}

```


There will be five main figures in this paper: - Key nutrient dotplot
(NO3, NH4, K, DOC) No COMR/HORR - Chla dot plot (All sites) - NO3 chla
JWC relationship - SF Chla (on vs off reservoir) - CLP Mainstem
Longitudinal (Chl-a), NO3?

# Nutrient Comparison Between reservoirs

### Functions


```{r}

#Code for violin plot plus jittered points with median line

# ggplot(boxplot_res, aes(x = reorder(site_year, BS_order, max), 
#                        y = .data[[param]], 
#                        fill = site_code)) + 
#   # Switched to violin plot with width proportional to sample size
# geom_violin(scale = "count", width = 4, alpha = 0.5, color = NA) + 
#   geom_jitter(color = "black", size = 0.8, alpha = 0.3, 
#               position = position_jitter(width = 0.15)) +  
#   stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black") +
#   # Custom rectangles for site grouping
#   geom_rect(inherit.aes = FALSE, aes(xmin = "LNGR 2021", xmax = "LNGR 2024", ymin = title_min, ymax = 0), 
#             color = "transparent", fill = "#0072B2", alpha = 0.01) + 
#   geom_rect(inherit.aes = FALSE, aes(xmin = "JOER 2021", xmax = "JOER 2024", ymin = title_min, ymax = 0), 
#             color = "transparent", fill = "#56B4E9", alpha = 0.05) + 
#   geom_rect(inherit.aes = FALSE, aes(xmin = "CBRR 2021", xmax = "CBRR 2024", ymin = title_min, ymax = 0), 
#             color = "transparent", fill = "#F9C74F", alpha = 0.05) + 
#   geom_rect(inherit.aes = FALSE, aes(xmin = "PTRR 2021", xmax = "PTRR 2024", ymin = title_min, ymax = 0), 
#             color = "transparent", fill = "#F3722C", alpha = 0.05) + 
#   geom_rect(inherit.aes = FALSE, aes(xmin = 16.5, xmax = 20.5, ymin = title_min, ymax = 0), 
#             color = "transparent", fill = "#d71a1d", alpha = 0.008) + 
#   
#   # Reference line and site labels
#   geom_abline(aes(slope = 0, intercept = 0), color = '#000000', linetype = "solid") + 
#   geom_text(aes(x = 2.5, y = (title_min/2), label = "Long Draw"), color = "#000000", size = 6) + 
#   geom_text(aes(x = 6.5, y = (title_min/2), label = "Joe Wright"), color = "#000000", size = 6) + 
#   geom_text(aes(x = 10.5, y = (title_min/2), label = "Chambers"), color = "#000000", size = 6) + 
#   geom_text(aes(x = 14.5, y = (title_min/2), label = "Peterson"), color = "#000000", size = 6) + 
#   geom_text(aes(x = 18.5, y = (title_min/2), label = "Barnes Meadow"), color = "#000000", size = 5) + 
#   
#   # Scales and Themes
#   scale_fill_manual(name = "Watershed", values = colorsBS_site_code) + 
#   scale_color_grey(start = 0.4, end = 0) + 
#   theme_bw(base_size = 25) + 
#   labs(x = "Site", y = param_units) + 
#   scale_x_discrete(labels = c("2021", "2022", "2023", "2024", "", "", "", "", 
#                               "2021", "2022", "2023", "2024", "", "", "", "", 
#                               "2021", "2022", "2023", "2024")) + 
#   theme(legend.position = "none",
#         axis.title = element_text(face = "bold"),
#         axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"))
```


```{r}
#Original function with no COMR/HORR and just dotplot with title always
build_res_dotplot<- function(param = "ChlA", binwdth = 0.3, sz = 1, title_min = -.7){
  
    boxplot_res <- filter(tidy_chem, site_code %in% c("LNGR", "PTRR", "BRNR", "JOER", "CBRR") )%>%
      filter(!is.na(.data[[param]]))%>%
      mutate(site_year = paste(site_code, Year, sep = " "),
             dayofyear = yday(Date))
    param_units <- filter(chem_units, simple == param)%>%
      pull(combined)
    
    #Create plot
    ggplot(boxplot_res, aes(x=reorder(site_year,BS_order,max),y=.data[[param]], 
                            #middle=mean(.data[[param]]), 
                            fill=site_code)) + 
      geom_dotplot(stackgroups = FALSE, binaxis = "y", method = "histodot",stackdir = "center", position = "dodge", binwidth = binwdth, dotsize = sz)+
      scale_color_grey(start = 0.4, end = 0)+
      #make x axis titles vertical
      theme(axis.text.x = element_text(angle = 90, hjust = 1))+
      geom_rect( inherit.aes=FALSE, aes(xmin="LNGR 2021", xmax="LNGR 2024", ymin=title_min,
                                        ymax=0), color="transparent", fill="#0072B2", alpha=0.01)+
      geom_rect( inherit.aes=FALSE, aes(xmin="JOER 2021", xmax="JOER 2024", ymin=title_min,
                                        ymax=0), color="transparent", fill="#56B4E9", alpha=0.05)+
      
      geom_rect( inherit.aes=FALSE, aes(xmin="CBRR 2021", xmax="CBRR 2024", ymin=title_min,
                                        ymax= 0), color="transparent", fill="#F9C74F", alpha=0.05)+
      
      geom_rect( inherit.aes=FALSE, aes(xmin="PTRR 2021", xmax="PTRR 2024", ymin=title_min,
                                        ymax=0), color="transparent", fill="#F3722C", alpha=0.05)+
      geom_rect( inherit.aes=FALSE, aes(xmin= 16.5, xmax= 20.5, ymin=title_min,
                                        ymax=0), color="transparent", fill="#d71a1d", alpha=0.008)+
      geom_abline(aes(slope=0, intercept=0), color='#000000', linetype="solid") +
      geom_text(aes(x= 2.5, y = (title_min/2), label= "Long Draw",  color = "#000000"), size = 6, alpha = 1)+
      geom_text(aes(x= 6.5, y = (title_min/2), label= "Joe Wright",  color = "#000000"), size = 6, alpha = 1)+
      geom_text(aes(x= 10.5, y = (title_min/2), label= "Chambers",  color = "#000000"), size = 6, alpha = 1)+
      geom_text(aes(x= 14.5, y = (title_min/2), label= "Peterson",  color = "#000000"), size = 6)+
      geom_text(aes(x= 18.5, y = (title_min/2), label= "Barnes Meadow",  color = "#000000"), size = 5, alpha = 1)+
  
      
      scale_fill_manual(name="Watershed", values = colorsBS_site_code) +
      theme_bw(base_size=25) +
      
      #theme(axis.text.x = ifelse(study_reservoir_chla$Location == "Reservoir",element_text(face = "bold"),element_text(face = "plain"))) +
      labs( x = "Site", y = param_units)+
      scale_x_discrete(labels = c("2021", "2022", "2023","2024",
                                  "","","","",
                                  "2021", "2022", "2023","2024",
                                  "","","","",
                                  "2021", "2022", "2023","2024"))+
      theme(legend.position = "none",
            axis.title = element_text(face = "bold"),
            # bold and angle x axis titles to 45
            axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"))
}
```

```{r}
build_res_plot <- function(param = "ChlA", binwdth = 0.3, sz = 1, title_min = NULL, b_size = 20) {
  # Order: Long Draw, Joe Wright, Hourglass, Chambers, Comanche, Peterson, Barnes Meadow
  sites <- c("LNGR", "JOER", "HORR", "CBRR", "COMR", "PTRR", "BRNR")

  y_label <- chem_expressions[[param]]  


  boxplot_res <- tidy_chem %>%
    filter(site_code %in% sites) %>%
    filter(!is.na(.data[[param]])) %>%
    mutate(site_year = paste(site_code, Year, sep = " "), 
           BS_order = as.numeric(BS_order))
  
  # Inject 6 gaps between the 7 reservoirs
  gaps <- data.frame(
    site_year = paste0("Gap", 1:8),
    site_code = NA, 
    BS_order = c(0.5,1.5,2.5,3.5,4.5,5.5,6.5,7.5) 
  )
  
  plot_data <- bind_rows(boxplot_res, gaps)
  
  # 2. Build Base Plot
  p <- ggplot(plot_data, aes(x = reorder(site_year, BS_order, max), y = .data[[param]], fill = site_code))
  
  # 3. Conditional Annotations
  if (!is.null(title_min)) {
    p <-  p +
      # Site Grouping Backgrounds
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01),xmin = 1.5, xmax = 5.5, fill="#0072B2", alpha=0.01) + #LNGR
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01),xmin = 6.5, xmax = 10.5, fill="#56B4E9", alpha=0.05) + #JOER
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01),xmin = 11.5, xmax = 15.5, fill="#F0E442", alpha=0.05) + #HORR
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01),xmin = 16.5, xmax = 20.5, fill="#F9C74F", alpha=0.05) + #CBRR
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01),xmin = 21.5, xmax = 25.5, fill="#F8961E", alpha=0.05) + #COMR
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01),xmin = 26.5, xmax = 30.5, fill="#F3722C", alpha=0.05) + #PTRR
      geom_rect(inherit.aes=FALSE, aes( ymin=title_min, ymax=-0.01), xmin = 31.5, xmax = 35.5, fill="#d71a1d", alpha=0.008) + #BRNR
      
      # Labels positioned at the center of each 4-year block (allowing for gaps)
      annotate("text", x = 3.5,  y = title_min/2, label = "Long Draw", size = 6) +
      annotate("text", x = 8.5,  y = title_min/2, label = "Joe Wright", size = 6) +
      annotate("text", x = 13.5, y = title_min/2, label = "Hourglass", size = 6) +
      annotate("text", x = 18.5, y = title_min/2, label = "Chambers", size = 6) +
      annotate("text", x = 23.5, y = title_min/2, label = "Comanche", size = 6) +
      annotate("text", x = 28.5, y = title_min/2, label = "Peterson", size = 6) +
      annotate("text", x = 33.5, y = title_min/2, label = "Barnes Meadow", size = 5) +
      geom_dotplot(stackgroups = FALSE, binaxis = "y", method = "histodot",stackdir = "center", position = "dodge", binwidth = binwdth, dotsize = sz)+
      geom_abline(intercept = 0, slope = 0, color = "black") +
      scale_fill_manual(values = colorsBS_site_code) +
      theme_bw(base_size = b_size) +
      labs(y = y_label, x = NULL)+
      theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),
            legend.position = "none") +
      # This function cleans the X-axis: removes site names and hides "Gap" labels
      scale_x_discrete(labels = function(x) ifelse(grepl("Gap", x), "", gsub(".* ", "", x)))+
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)))
    
  } else {
    p <- p+
      geom_dotplot(stackgroups = FALSE, binaxis = "y", method = "histodot",stackdir = "center", position = "dodge", binwidth = binwdth, dotsize = sz)+
      geom_abline(intercept = 0, slope = 0, color = "black") +
      scale_fill_manual(values = colorsBS_site_code) +
      theme_bw(base_size = b_size) +
      labs(y = y_label, x = NULL)+
      theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none"
      )
  }
  
  return(p)
}
```



### Plots with COMR/HORR

```{r}
TSS_dotplot_res <-  build_res_plot(param = "TSS", binwdth = .4, sz = 5)+
  ylim(-0.5,50)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black")

K_dotplot_res <-  build_res_plot(param = "K", binwdth = .045, sz = 1)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black")

no3_dotplot_res <-   build_res_plot(param = "NO3", binwdth = 0.09, sz = 1)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black")+
  ylim(-0.1,2.5)

DOC_dotplot_res <-  build_res_plot(param = "DOC", binwdth = .35, sz = 1.25, title_min = -2.5)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, alpha = 0.7)

TSS_dotplot_res/K_dotplot_res/no3_dotplot_res/DOC_dotplot_res

ggsave("docs/cpf_reservoir/paper_draft/res_chem_dotplots.jpg", dpi = 300, width = 18, height = 14)

```

### SI Fig: Extra Dotplots
```{r}
 build_res_plot(param = "DTN", binwdth = .02, sz = 1, title_min = -0.2)+
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, alpha = 0.7)

```


### Stats

Reservoirs Only

##### Normality

```{r}
chem_transformed_res <- tidy_chem%>%
  filter(site_code %in% primary_study_res) %>%
  pivot_longer(cols = c(Turbidity, TSS, DTN,  DOC:NO3, SO4), 
               names_to = "param", 
               values_to = "value")%>%
  select(site_code, Date, Year, param, value, burn_group,season)%>%
  mutate(burn_group = factor(burn_group, levels = c("unburned", "burned")),
         Year = factor(Year, levels = c("2021", "2022", "2023", "2024")), 
         season = factor(season),
         site_code = factor(site_code, levels = c("LNGR", "JOER", "CBRR", "PTRR", "BRNR")))%>%
  mutate(log_value = log10(value +1 ))%>%
  na.omit()

param_oi <- c( "Turbidity", "TSS", "DTN","NH4", "K", "DOC", "NO3", "SO4")

ggplot(chem_transformed_res%>%filter(param %in% param_oi), aes(x = value, fill = burn_group))+
  geom_histogram(position = "stack")+
  geom_vline(aes(xintercept = 0.05), color = "red", linetype = "dashed")+
  facet_wrap(~param, scales = "free")

ggplot(chem_transformed_res%>%filter(param %in% param_oi), aes(x = log_value, fill = burn_group))+
  geom_histogram(position = "stack")+
  facet_wrap(~param, scales = "free")

```

Nothing without transformation is normal, we should double check model
assumptions with `check_model()` after creating a model then only
evaluate those that do. We will want to use a linear mixed effects model
to compare out burned vs unburned reservoirs

##### Burned vs Unburned

Fixed effects: burn type random effect: site

```{r}
burn_group_check <- map(c( "Turbidity", "TSS", "DTN","NH4", "K", "DOC", "NO3", "SO4"), ~{
  
  #Since NO3 does not fit a normal distribution, we will use a generalized linear mixed model with a gamma distribution and log link for NO3, and linear mixed models for the other parameters.
  if(.x == "NO3"){
        model_gamma <- glmmTMB(value ~ burn_group + (1 | site_code), 
                       data = chem_transformed_res%>%filter(param == .x & value > 0), 
                       family = Gamma(link = "log"))
        
        check_model(model_gamma)
  }
  
   df <- chem_transformed_res%>%filter(param == .x)
   
   
   model_full <- lmer(log_value ~ burn_group + (1|site_code), data = df, REML = FALSE)

    check_model(model_full)

})%>%
  set_names(param_oi)

# If they pass the model diagnostics, extract the estimate for burn_group and calculate the percent difference between burned and unburned reservoirs. Then compare the full model to a null model without burn_group using an ANOVA to get the p-value for the effect of burn_group.


map(c( "Turbidity", "TSS", "DTN", "DOC", "SO4", "K", "NO3"), ~{
  
  if(.x == "NO3"){
    model_gamma <- glmmTMB(value ~ burn_group + (1 | site_code), 
                           data = chem_transformed_res%>%filter(param == .x & value > 0), 
                           family = Gamma(link = "log"))
    
    conf <- confint(model_gamma, parm = "beta_")
    low_pct <- ((exp(conf[2,1])) - 1)*100
    high_pct <- ((exp(conf[2,2])) - 1)*100
    
    estimate <- ((summary(model_gamma)$coefficients$cond["burn_groupburned", "Estimate"]))
    pct_diff <- ((exp(estimate)) - 1)*100
    model_null <- glmmTMB(value ~ 1 + (1 | site_code), 
                          data = chem_transformed_res%>%filter(param == .x & value > 0), 
                          family = Gamma(link = "log"))
    anova_res <- anova(model_null, model_gamma)
    
    tibble(param = "NO3", estimate = estimate, low_pct = low_pct, high_pct = high_pct,
           pct_diff = pct_diff,p_value_anova = anova_res$`Pr(>Chisq)`[2])%>%
      mutate(p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
             significant_effect = ifelse(p_value_anova < 0.05, ifelse(pct_diff > 0, "Higher in Burned", "Lower in Burned"), "No Significant Difference"))
  }
  df <- chem_transformed_res%>%filter(param == .x)
  model_full <- lmer(log_value ~ burn_group + (1|site_code), data = df, REML = FALSE)
  summary(model_full) 
  
  conf <- confint(model_full, parm = "beta_")
  low_pct <- ((exp(conf[2,1])) - 1)*100
  high_pct <- ((exp(conf[2,2])) - 1)*100
  
  estimate <- model_full@beta[2]
  pct_diff <- ((exp(estimate)) - 1)*100
  model_null <- lmer(log_value ~ 1 + (1|site_code), data = df, REML = FALSE)
  anova_res <- anova(model_null, model_full)
  tibble(param = .x, estimate = estimate, low_pct = low_pct, high_pct = high_pct,
         pct_diff = pct_diff,p_value_anova = anova_res$`Pr(>Chisq)`[2])%>%
    mutate(p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
           significant_effect = ifelse(p_value_anova < 0.05, ifelse(pct_diff > 0, "Higher in Burned", "Lower in Burned"), "No Significant Difference"))
})%>%
  bind_rows()%>%
  arrange(low_pct)%>%
  filter(param == "DTN")


#Levene's test for NO3
lev_model <- leveneTest(log_value ~ burn_group, data = chem_transformed_res%>%filter(param == "NO3" & value > 0))
check_model(lev_model)
```

##### Year wise (All Sites)

Fixed effects: Year random effect: site

```{r}
year_wise_check <- map(c( "Turbidity", "TSS", "DTN","NH4", "K", "DOC", "NO3", "SO4"), ~{
  if(.x == "NO3"){
    model_gamma <- glmmTMB(value ~ Year + (1 | site_code), 
                           data = chem_transformed_res%>%filter(param == .x & value > 0), 
                           family = Gamma(link = "log"))
    
    check_model(model_gamma)
  }else{
    df <- chem_transformed_res%>%filter(param == .x)
    
    model_full <- lmer(log_value ~ Year + (1|site_code), data = df, REML = FALSE)
    
    check_model(model_full)
  }
  
})%>% set_names(c( "Turbidity", "TSS", "DTN","NH4", "K", "DOC", "NO3", "SO4"))

# If they pass the model diagnostics, extract the estimate for burn_group and calculate the percent difference between burned and unburned reservoirs. Then compare the full model to a null model without burn_group using an ANOVA to get the p-value for the effect of Year


map(c("Turbidity", "TSS", "DTN", "NO3", "DOC", "SO4", "K"), ~{
  
  if(.x == "NO3"){
    # Gamma model for NO3
    model_gamma <- glmmTMB(value ~ Year + (1 | site_code), 
                           data = chem_transformed_res %>% filter(param == .x & value > 0), 
                           family = Gamma(link = "log"))
    
    year_means <- emmeans(model_gamma, "Year")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value, 
             lower.CL = asymp.LCL, 
             upper.CL = asymp.UCL) %>%
      mutate(parameter = .x)
    
    model_null <- glmmTMB(value ~ 1 + (1 | site_code), 
                          data = chem_transformed_res %>% filter(param == .x & value > 0), 
                          family = Gamma(link = "log"))
    
    anova_res <- anova(model_null, model_gamma)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
    
  } else {
    # LMM for other parameters
    df <- chem_transformed_res %>% filter(param == .x)
    model_full <- lmer(log_value ~ Year + (1|site_code), data = df, REML = FALSE)
    
    year_means <- emmeans(model_full, "Year")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value) %>%
      mutate(parameter = .x)
    
    model_null <- lmer(log_value ~ 1 + (1|site_code), data = df, REML = FALSE)
    anova_res <- anova(model_null, model_full)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
  }
  
  # Attach ANOVA p-value to the contrast dataframe
  contrasts %>%
    mutate(p_value_anova = p_anova)
  
}) %>%
  bind_rows() %>%
  # Final calculation for percent changes and significance stars
  mutate(
    pct_est  = (exp(estimate) - 1) * 100,
    pct_low  = (exp(lower.CL) - 1) * 100,
    pct_high = (exp(upper.CL) - 1) * 100,
    p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
    p_sig_contrast = ifelse(p_value_contrast < 0.05, "*", "")
  )%>%
  select(parameter, contrast, pct_est, pct_low, pct_high, p_value_contrast, p_sig_contrast, p_value_anova, p_sig_anova)%>%
  filter(parameter == "DTN")
```

##### Site wise (interacton with year)

Fixed effects: Year + Site random effect: season

```{r}

check_model_results <- map(c(param_oi), ~{
  df <- chem_transformed_res%>%filter(param == .x)
  
  model_full <- lmer(log_value ~ site_code + Year + (1|season), data = df, REML = FALSE)
  #summary(model_full) 
  check_model(model_full)
  
})%>%
  set_names(param_oi)

  map(c("NO3", "TSS", "DTN", "K", "DOC","SO4" ), ~{
  
  if(.x == "NO3"){
    # Gamma model for NO3
    model_gamma <- glmmTMB(value ~ Year + site_code + (1 | season), 
                           data = chem_transformed_res %>% filter(param == .x & value > 0), 
                           family = Gamma(link = "log"))
    
    year_means <- emmeans(model_gamma, "site_code")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value, 
             lower.CL = asymp.LCL, 
             upper.CL = asymp.UCL) %>%
      mutate(parameter = .x)
    
    model_null <- glmmTMB(value ~ Year + site_code  + (1 | season), 
                          data = chem_transformed_res %>% filter(param == .x & value > 0), 
                          family = Gamma(link = "log"))
    
    anova_res <- anova(model_null, model_gamma)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
    
  } else {
    # LMM for other parameters
    df <- chem_transformed_res %>% filter(param == .x)
    model_full <- lmer(log_value  ~ Year + site_code  + (1 | season), data = df, REML = FALSE)
    
    year_means <- emmeans(model_full, "site_code")
    
    # Extract estimates and confidence intervals for the contrasts
    contrasts <- summary(pairs(year_means), infer = TRUE) %>%
      as.data.frame() %>%
      rename(p_value_contrast = p.value) %>%
      mutate(parameter = .x)
    
    model_null <- lmer(log_value ~ Year  + (1 | season), data = df, REML = FALSE)
    anova_res <- anova(model_null, model_full)
    p_anova <- anova_res$`Pr(>Chisq)`[2]
  }
  contrasts%>%
    mutate(p_value_anova = anova_res$`Pr(>Chisq)`[2])
})%>%
  bind_rows() %>%
  # Final calculation for percent changes and significance stars
  mutate(
    pct_est  = (exp(estimate) - 1) * 100,
    pct_low  = (exp(lower.CL) - 1) * 100,
    pct_high = (exp(upper.CL) - 1) * 100,
    p_sig_anova = ifelse(p_value_anova < 0.05, "*", ""), 
    p_sig_contrast = ifelse(p_value_contrast < 0.05, "*", "")
  )%>%
  select(parameter, contrast, pct_est, pct_low, pct_high, p_value_contrast, p_sig_contrast, p_value_anova, p_sig_anova)%>%
    filter(parameter == "DTN")
  

```

# Nutrient/ChlA response 

## Plots

```{r}
#Seasonal Color values
season_clp_color_vals <- c("Winter" = "#03045e",  "Rising Limb" = '#047E82',  "Peak Flow" = '#59B851', "Falling Limb"='#FA850E', "Fall Baseflow"  = "#772f1a")


jwc_longitudinal <- tidy_chem %>%
  dplyr::filter(!is.na(ChlA)) %>%
  dplyr::filter(site_code %in% c("JOEI", "JOER", "CBRR", "CBRI", "CHD", "JWC"))%>%
  select(Date,site_code, ChlA, NO3, Year,season, month,location_type)%>%
  mutate(location_type = ifelse(location_type == "Outflow"|location_type =="Inflow", "Stream", "Reservoir"))

jwc_longitudinal$site_code <- factor(jwc_longitudinal$site_code, levels = c("JOEI", "JOER", "CBRI", "CBRR", "CHD", "JWC"))
jwc_longitudinal$season <- factor(jwc_longitudinal$season, levels = c("Winter", "Rising Limb", "Peak Flow", "Falling Limb", "Fall Baseflow"))
jwc_longitudinal$month <- factor(jwc_longitudinal$month, levels = c("4", "5", "6", "7", "8", "9", "10","11"))
# Calculate the medians for ChlA
median_data <- jwc_longitudinal %>%
  group_by(site_code, Year, location_type) %>%
  summarise(median_ChlA = median(ChlA, na.rm = TRUE), 
            median_NO3 = median(NO3, na.rm = TRUE), 
            max_ChlA = max(ChlA, na.rm = T), 
            max_NO3 = max(NO3, na.rm = T)) %>%
  ungroup()%>%
  add_row(site_code = "JWC", Year = "2021", location_type = "Stream", median_ChlA = NA, median_NO3 = NA, max_ChlA = NA, max_NO3 = NA)
median_data$site_code <- factor(median_data$site_code, levels = c("JOEI", "JOER", "CBRI", "CBRR", "CHD", "JWC"))

# Create the plot
jwc_long_chla<-  ggplot() + 
  geom_point(data = jwc_longitudinal, aes(x = site_code, y = ChlA, color = season),
             size = 3 ,position = position_jitterdodge(jitter.height = 0))+
  geom_rect(data = median_data %>% filter(site_code %in% c("JOER", "CBRR")),
            aes(xmin = as.numeric(site_code) - 0.4, xmax = as.numeric(site_code) + 0.4, ymin = - Inf, ymax = Inf, group = Year),
            fill = "#59B851", alpha =0.2)+
  geom_rect(data = median_data %>% filter(site_code %in% c("JWC")),
            aes(xmin = as.numeric(site_code) - 0.4, xmax = as.numeric(site_code) + 0.4, ymin = - Inf, ymax = Inf, group = Year),
            fill = "#002FA7", alpha =0.2)+
  scale_color_manual(values = season_clp_color_vals)+
  # Add lines connecting the medians
  geom_line(data = median_data, aes(x = site_code, y = median_ChlA, group = Year), color = "black", linewidth = 1) +
  geom_point(data = median_data, aes(x = site_code, y = median_ChlA, group = Year, shape = "Yearly Site Median"), color = "black", size = 3) +
  scale_shape_manual(name = "", values = c("Yearly Site Median" = 16)) +  # 16 = solid circle
  guides(shape = guide_legend(override.aes = list(color = "black", size = 3))) +
  theme_bw(base_size = 20)+
  facet_wrap(~Year, ncol = 1, scale = "free_y")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right")+
  scale_x_discrete(labels = c( "JOEI"= "  Above\nJoe Wright", 
                               "JOER" = "", 
                               "CBRI" = "Above\nChambers",
                               "CBRR" = "", 
                               "CHD" = "Below\nChambers", 
                               "JWC" = "Joe Wright Creek"))+
  labs(y = expression(Chlorophyll ~italic("a") ~ (ug ~ L^{-1})), 
       color = "Season")



jwc_long_no3 <- 
  ggplot() + 
  geom_point(data = jwc_longitudinal, aes(x = site_code, y = NO3, color = season),size = 3 ,position = position_jitterdodge(jitter.height = 0))+
  geom_rect(data = median_data %>% filter(site_code %in% c("JOER", "CBRR")),
            aes(xmin = as.numeric(site_code) - 0.4, xmax = as.numeric(site_code) + 0.4, ymin = - Inf, ymax = Inf, group = Year),
            fill = "#59B851", alpha =0.2)+
  geom_rect(data = median_data %>% filter(site_code %in% c("JWC")),
            aes(xmin = as.numeric(site_code) - 0.4, xmax = as.numeric(site_code) + 0.4, ymin = - Inf, ymax = Inf, group = Year),
            fill = "#002FA7", alpha =0.2)+
  scale_color_manual(values = season_clp_color_vals)+
  #add a Hline for the median
  #stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black") +
  geom_line(data = median_data, aes(x = site_code, y = median_NO3, group = Year), color = "black", linewidth = 1) +
  geom_point(data = median_data, aes(x = site_code, y = median_NO3, group = Year, shape = "Yearly Site Median"), color = "black", size = 3) +
  scale_shape_manual(name = "", values = c("Yearly Site Median" = 16)) +  # 16 = solid circle
  guides(shape = guide_legend(override.aes = list(color = "black", size = 3))) +
  theme_bw(base_size = 20)+
  facet_wrap(~Year, ncol = 1, scale = "free_y")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        #remove facet title
        #strip.text.x = element_blank(),
        legend.position = "none") +
  scale_x_discrete(labels = c( "JOEI"= "  Above\nJoe Wright", 
                               "JOER" = "", 
                               "CBRI" = "Above\nChambers",
                               "CBRR" = "", 
                               "CHD" = "Below\nChambers", 
                               "JWC" = "Joe Wright Creek")) +
  labs(y = expression(NO[3]^{"-"} ~ (mg ~ L^{-1})))


jwc_long_chla_no3<-  jwc_long_chla+jwc_long_no3+
  plot_layout(guides = 'collect')+
  #add color legend to top
  plot_annotation( tag_levels = 'a')

plot(jwc_long_chla_no3)
save_plot_image(jwc_long_chla_no3 ,"jwc_long_chla_no3" )
```



-   Do reservoirs transform the nutrient signal?

    -   Test of variability in ChlA/NO3 between sites

        -   Levene’s test between inflow and reservoir, and reservoir
            and outflo

        -   

        -   Test of means/medians between streams/reservoirs (ChlA and
            NO3)

            -   Linear mixed effect model (log_value \~ site_code +
                season + (1\|year)

-   Longitudinal Plot of ChlA + NO3

## Stats

```{r}
jwc_longitudinal_stats <- jwc_longitudinal%>%
  pivot_longer(cols = c(ChlA:NO3), 
               names_to = "param", 
               values_to = "value")%>%
  mutate(log_value = log10(value + 1))

```

### JOER
Test inflow vs outflow: variation and means

Linear relationship between no3 & Chl-a 
```{r}

joer_longitudinal_stats<- jwc_longitudinal_stats %>%
   filter( site_code %in% c("JOEI", "JOER", "CBRI"))
#perform levene's test between groups
levene_site_jwc_chla <- leveneTest(log_value ~ site_code, data = joer_longitudinal_stats%>%filter(param == "ChlA"))
levene_site_jwc_chla

levene_site_jwc_no3 <-leveneTest(log_value ~ site_code, data = joer_longitudinal_stats%>%filter(param == "NO3"))
levene_site_jwc_no3
```

### CBRR
Test inflow vs outflow: variation and means

Linear relationship between no3 & Chl-a 

```{r}

```


# Reservoir Chlorophyll a


## Plot
```{r}

chla_dotplot_res <- build_res_dotplot(param = "ChlA", binwdth = .3, sz = 1)+
  # add medians
  stat_summary(fun = median, geom = "crossbar", position = position_dodge(width = 0.8), width = 0.8, fill = "black") +
#adding meso and eutrophic levels for cold lakes
 geom_abline(aes(slope=0, intercept=2.6), color='black', linetype="dashed") +
    geom_abline(aes(slope=0, intercept=8), color='black', linetype="dashed") +
    #annotate(geom="text", label= "MDL Limit (μg/L)", x="CBRR 2022", y=.075, vjust= -1)+
    annotate(geom="text", label= "Mesotrophic", x=1.15, y=2.6, vjust= -1, size = 4)+
    annotate(geom="text", label= "Eutrophic", x= 1.5, y=8, vjust= -1, size = 5)+
   #fixing y axis
    coord_cartesian(ylim=c(0,13))

plot(chla_dotplot_res)

save_plot_image(chla_dotplot_res, "res_chla_dotplots_2021_2024")


```

## Stats

#### Eutrophic status

```{r}
# grouped by site code
chla_status_site_year <- tidy_chem%>%
  filter(!is.na(ChlA))%>%
  filter(site_code %in% primary_study_res)%>%
  group_by(site_code, Year)%>%
  summarise(mean_chla = mean(ChlA), 
            median_chla = median(ChlA), 
            min_chla = min(ChlA), 
            max_chla = max(ChlA), 
            sd_chla = sd(ChlA))


# Grouped by burn status

chla_status_site_year <- tidy_chem%>%
  filter(!is.na(ChlA))%>%
  filter(site_code %in% primary_study_res)%>%
  group_by(site_code, Year, burn_group)%>%
  summarise(mean_chla = mean(ChlA), 
            median_chla = median(ChlA), 
            min_chla = min(ChlA), 
            max_chla = max(ChlA), 
            sd_chla = sd(ChlA))

#create flextable for these stats
chla_status_site_year <- chla_status_site_year%>%
  ungroup()%>%
  mutate(burn_group = factor(burn_group, levels = c("Pre", "Post")))%>%
  mutate(Year = as.character(Year))%>%
  select(-site_code)%>%
  pivot_longer(cols = c(mean_chla, median_chla, min_chla, max_chla, sd_chla), names_to = "statistic", values_to = "value")%>%
  mutate(statistic = factor(statistic, levels = c("mean_chla", "median_chla", "min_chla", "max_chla", "sd_chla")))
```


### Prep Data

```{r}

chem_transformed_res <- tidy_chem%>%
    filter(site_code %in% primary_study_res) %>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4), 
               names_to = "param", 
               values_to = "value")%>%
  select(site_code, Date, Year, param, value, burn_group,season)%>%
    mutate(burn_group = factor(burn_group, levels = c("unburned", "burned")),
           Year = factor(Year, levels = c("2021", "2022", "2023", "2024")), 
           season = factor(season),
           site_code = factor(site_code, levels = c("LNGR", "JOER", "CBRR", "PTRR", "BRNR")))%>%
  mutate(log_value = log10(value +1 ))%>%
  na.omit()

param_oi <- c("ChlA")

ggplot(chem_transformed_res%>%filter(param %in% param_oi), aes(x = value, fill = burn_group))+
  geom_histogram()+
  facet_wrap(~param, scales = "free")

ggplot(chem_transformed_res%>%filter(param %in% param_oi), aes(x = log_value, fill = burn_group))+
  geom_histogram()+
  facet_wrap(~param, scales = "free")

```

Nothing without transformation is normal, we should double check model
assumptions with `check_model()` after creating a model then only
evaluate those that do. We will want to use a linear mixed effects model
to compare out burned vs unburned reservoirs

##### Burned vs Unburned

Fixed effects: burn type random effect: site

```{r}
map(c("ChlA"), ~{
   df <- chem_transformed_res%>%filter(param == .x)
   
#    outliers_list <- check_outliers(df$log_value, method = "zscore", threshold = 3)
#    
# df <- df %>% mutate(outlier = outliers_list)%>%
#   filter(outlier == FALSE)

   model_full <- lmer(log_value ~ burn_group + (1|site_code), data = df, REML = FALSE)

    plot(check_model(model_full))+
     plot_annotation(title = .x)

})

# If they pass the model diagnostics, extract the estimate for burn_group and calculate the percent difference between burned and unburned reservoirs. Then compare the full model to a null model without burn_group using an ANOVA to get the p-value for the effect of burn_group.


map(c("ChlA"), ~{
  df <- chem_transformed_res%>%filter(param == .x)
  model_full <- lmer(log_value ~ burn_group + (1|site_code), data = df, REML = FALSE)
  summary(model_full) 
  
  conf <- confint(model_full, parm = "beta_")
  low_pct <- ((exp(conf[2,1])) - 1)*100
  high_pct <- ((exp(conf[2,2])) - 1)*100
  
  estimate <- model_full@beta[2]
  pct_diff <- ((exp(estimate)) - 1)*100
  model_null <- lmer(log_value ~ 1 + (1|site_code), data = df, REML = FALSE)
  anova_res <- anova(model_null, model_full)
  tibble(param = .x, estimate = estimate, low_pct = low_pct, high_pct = high_pct,
         pct_diff = pct_diff, p_value = anova_res$`Pr(>Chisq)`[2])%>%
    mutate(p_sig = ifelse(p_value < 0.05, "*", ""))
})%>%
  bind_rows()


```

##### Year wise (All Sites)

Fixed effects: Year random effect: site

```{r}
map(c("ChlA"), ~{
   df <- chem_transformed_res%>%filter(param == .x)

   model_full <- lmer(log_value ~ Year + (1|site_code), data = df, REML = FALSE)

    plot(check_model(model_full))+
     plot_annotation(title = .x)

})

# If they pass the model diagnostics, extract the estimate for burn_group and calculate the percent difference between burned and unburned reservoirs. Then compare the full model to a null model without burn_group using an ANOVA to get the p-value for the effect of Year


map(c("ChlA"), ~{
  df <- chem_transformed_res%>%filter(param == .x)
  model_full <- lmer(log_value ~ Year + (1|site_code), data = df, REML = FALSE)
  summary(model_full) 
  year_means <- emmeans(model_full, "Year")
  
  contrasts <- pairs(year_means, combine = TRUE) %>%
  as.data.frame()%>%
  mutate(parameter = .x)
  # 
  # conf <- confint(model_full, parm = "beta_")
  # low_pct <- ((exp(conf[2,1])) - 1)*100
  # high_pct <- ((exp(conf[2,2])) - 1)*100
  # 
  # estimate <- model_full@beta[2]
  # pct_diff <- ((exp(estimate)) - 1)*100
  model_null <- lmer(log_value ~ 1 + (1|site_code), data = df, REML = FALSE)
  anova_res <- anova(model_null, model_full)
  
  contrasts%>%
    mutate(p_value = anova_res$`Pr(>Chisq)`[2], 
           p_sig = ifelse(p_value < 0.05, "*", ""))
})%>%
  bind_rows()

```

##### Site wise (interacton with year)

Fixed effects: Year + Site random effect: season

```{r}

check_model_results <- map(c(param_oi), ~{
  df <- chem_transformed_res%>%filter(param == .x)
  
  model_full <- lmer(log_value ~ site_code + Year + (1|season), data = df, REML = FALSE)
  #summary(model_full) 
  check_model(model_full)
  
})%>%
  set_names(param_oi)

  map(c("ChlA" ), ~{
  df <- chem_transformed_res%>%filter(param == .x)
  model_full <- lmer(log_value ~ site_code + Year + (1|season), data = df, REML = FALSE)
  summary(model_full) 
  
  site_means <- emmeans(model_full, "site_code")
  
  contrasts <- pairs(site_means, combine = TRUE) %>%
  as.data.frame()%>%
  mutate(parameter = .x, 
         p_sig_contrast = ifelse(p.value < 0.05, "*", ""), 
         pct_diff = ((exp(estimate)) - 1)*100)%>%
    select(parameter, contrast, estimate, pct_diff, p.value, p_sig_contrast)

  model_null <- lmer(log_value ~ 1 + (1|site_code), data = df, REML = FALSE)
  anova_res <- anova(model_null, model_full)
  
  contrasts%>%
    mutate(p_sig_anova = ifelse(anova_res$`Pr(>Chisq)`[2] < 0.05, "*", ""))
})%>%
  bind_rows()
```


# South Fork

## Plots

```{r}

SF_Res = c("COMI", "COMR", "COMO", "HORI", "HORR", "HORO", "BEAV", "SFM", "PSF")
#filter and org
sf_chem<- filter(tidy_chem, site_code %in% SF_Res & !is.na(ChlA))%>%
  select(site_code,Date,ChlA,location_type, Buffer_Level, Watershed_Level, watershed, Year )%>%
  mutate(location_type = case_when(
    location_type == "Stream" ~ "Beaver Creek", 
    site_code == "PSF" ~ "Poudre Below\nSouth Fork",
    site_code == "SFM" ~ "South Fork",
    TRUE ~ location_type), 
    site_group = case_when(
      watershed == "Comanche Reservoir" ~ watershed, 
      watershed == "Hourglass Reservoir"~ watershed, 
      TRUE ~ "Stream"
    ), 
    filled_status = case_when(
      Year %in% c(2021, 2022) ~ "Reservoirs NOT Filled",
      Year  %in% c(2023, 2024) ~ "Reservoirs Filled"
      
    )) 

sf_chem$location_type <- factor(sf_chem$location_type, levels = c("Inflow","Reservoir", "Outflow","Beaver Creek","South Fork", "Poudre Below\nSouth Fork"))
sf_chem$filled_status <- factor(sf_chem$filled_status, levels = c("Reservoirs NOT Filled", "Reservoirs Filled"))
#get nice units for Y axis
lab_w_units <-  filter(chem_units, simple %in% "ChlA")%>%
  pull(combined)

#plot
south_fork <-   ggplot(sf_chem, aes(x=location_type,y=ChlA,  fill = filled_status, shape = )) + 
  geom_dotplot( binaxis = "y", 
                method = "histodot",
                stackdir = "center", 
                position = position_dodge(width = 0.8),
                binwidth = .2, dotsize = 2, 
                stackratio = 1, width = .01)+
  geom_abline(aes(slope=0, intercept=2.6), color='black', linetype="dashed") +
  geom_abline(aes(slope=0, intercept=8), color='black', linetype="dashed") +
  annotate(geom="text", label= "Mesotrophic", x=.75, y=2.6, vjust= -1, size = 6)+
  annotate(geom="text", label= "Eutrophic", x= .75, y=8, vjust= -1, size = 6)+
  theme_bw(base_size = 30)+
  scale_fill_manual(name = "Reservoir Status", values = c("#D55E00","#0072B2"))+
  labs(color = "Reservoir Status",
       y = lab_w_units)+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_text(face = c('plain', 'plain', 'plain', 'bold', 'bold', 'plain')), 
        legend.position = "bottom")


south_fork

save_plot_image(south_fork ,"south_fork_chla_impact" )
```
## Stats

```{r}
sf_chem <- tidy_chem %>%
  filter(site_code %in% c("COMI", "HORI", "COMR", "HORR", "COMO", "HORO", "BEAV", "SFM"))%>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4, PO4), 
               names_to = "param", 
               values_to = "value")%>%
  mutate(year_site = paste0(Year, "_", site_code))%>%
  mutate(log_value = log10(value + 1), 
         fill_status = if_else(Year <= 2022, "Unfilled", "Filled"))%>%
  mutate(log_value = log10(value + 1))%>%
  na.omit()

ggplot(sf_chem, aes(x = value, color = year_site))+
  geom_histogram()+
  facet_wrap(~param, scales = "free")
```

### Inflows vs Beaver 
Do we see a similar impact to the JWC network? ie increase across the basin

```{r}

```

### Downstream sites (filled vs unfilled)

When a reservoir is active, do downstream sites have increased chl-a and if so, how far is that impact "felt"?

```{r}

```


# CLP Longitudinal
Show how reservoir impact dies as additional tributaries dilute the signal. We can also show how the reservoirs impact changes across seasons (minimal). 

## Plot

```{r}


clp_long <- tidy_chem %>%
  dplyr::filter(!is.na(ChlA)& Year %in% c(2022, 2023, 2024)) %>%
  dplyr::filter(site_code %in% c("JOEI", "JOER", "CBRR", "CBRI", "CHD", "JWC", "PFAL", "SLEP", "PBR", "PSF", "PNF", "PBD"))%>%
  select(site_code,Date, watershed, distance_upstream_km, ChlA, NO3, PO4, K, location_type,Year, month, season )%>%
  mutate(location_type = case_when(
    location_type %in% c("Inflow", "Outflow") ~ "Res Stream", 
    location_type %in% c("Reservoir") ~ "Reservoir",
    TRUE ~ "Mainstem"),
    Year = as.character(Year),
    dist_pbd = as.character(distance_upstream_km),
    site_code  = case_when(site_code == "PFAL" ~ "SLEP", TRUE ~ site_code),
    distance_upstream_km = case_when(site_code == "SLEP" ~ 73.161, TRUE ~ distance_upstream_km),
    type = ifelse(location_type %in% c("Inflow","Outflow"),"Stream",
                  ifelse(location_type == "Reservoir", "Reservoir", "Mainstem")))



clp_long$site_code <- factor(clp_long$site_code, levels = c("JOEI", "JOER", "CBRR", "CBRI", "CHD", "JWC", "SLEP", "PBR", "PSF", "PNF", "PBD"))
clp_long$season <- factor(clp_long$season, levels = c("Winter", "Rising Limb", "Peak Flow", "Falling Limb", "Fall Baseflow"))
clp_long$month <- factor(clp_long$month, levels = c("5", "6", "7", "8", "9", "10","11"))


# Calculate the medians for ChlA
clp_median_data <- clp_long %>%
  group_by(site_code,type,distance_upstream_km, location_type) %>%
  summarise(median_ChlA = median(ChlA, na.rm = TRUE), 
            median_NO3 = median(NO3, na.rm = TRUE)) %>%
  ungroup()

clp_long_chla<- ggplot(clp_long, aes(x = distance_upstream_km, y = ChlA)) + 
  annotate("rect", xmin = 78, xmax = 99, ymin = -0.1, ymax = 9, 
           color = "#4A86C5", fill = "NA", linewidth = 1.5) +
  
  geom_jitter(aes(color = season, shape = type), size = 3, width = 0.4) +
  scale_color_manual(values = season_clp_color_vals) +
  
  # Lines and points for medians
  geom_line(data = clp_median_data, aes(x = distance_upstream_km, y = median_ChlA), 
            color = "black", linewidth = 2.5) +
  geom_point(data = clp_median_data, aes(x = distance_upstream_km, y = median_ChlA, shape = type), 
             color = "black", size = 5) +
  #Box title
  annotate("label", x = 78, y = 8, label = "Joe Wright & Chambers\nReservoirs (Prev Figure)", 
           hjust = 0, size = 6, label.size = 0.5) +
  #JWC label + arrow
  annotate("label", x = 80, y = 5.75, label = "JWC site", size = 5) +
  #Confluence with Big South
    annotate("segment", x = 73, xend = 77, y = 4, yend = 2, 
           arrow = arrow(length = unit(0.3, "cm")), color = "#4A86C5", linewidth = 1) +  
  annotate("label", x = 73, y = 4, label = "Confluence\nwith Big South", size = 5) +
  #confluence with SF
        annotate("segment", x = 40, xend = 40, y = 4, yend = 1, 
           arrow = arrow(length = unit(0.3, "cm")), color = "#4A86C5", linewidth = 1) +  
  annotate("label", x = 40, y = 4, label = "Confluence with\nSouth Fork", size = 5) +
        annotate("segment", x = 5, xend = 5, y = 4, yend = 1, 
           arrow = arrow(length = unit(0.3, "cm")), color = "#4A86C5", linewidth = 1) + 
  annotate("label", x = 8, y = 4, label = "Confluence with\nNorth Fork", size = 5) +

  # Styling and Scales
  theme_bw(base_size = 30) +
  scale_x_reverse() +
  ylim(-0.2, 9.2) +
  labs(x = "Distance from Canyon Mouth (km)", 
       y = "Chlorophyll a (µg/L)", 
       shape = "Location", 
       color = "Season") +
 theme(
    legend.position = c(0.98, 0.98),         
    legend.justification = c(1, 1), 
    
    legend.box = "horizontal",               
    legend.box.background = element_rect(color = "black", linewidth = 0.5), 
    legend.background = element_blank(),     
    legend.margin = margin(6, 6, 6, 6)       
  )

print(clp_long_chla)


save_plot_image(clp_long_chla, "clp_long_chla")

```

## Stats
Compare site chla variability with levene test and means using mixed model 
Fixed: Site
Random: Season + Year

Calculate the metric of reservoir impact: acre ft impounded water storage/ watershed area 
Does this correlate with Chl-a across the basin?


# Scratch

# Remote Sensing

## Import datasets and join with chem data

```{r}
#Read in the data from B

#Temp: From NW-CLP-RS/CLP_focus_temp_only.Rmd

temp_rs <- read_parquet(here("data/derived/remote_sensing/landsat/CLP_temp_estimate_v2024-10-10.parquet"))%>%
  mutate(doy = yday(date),
         kurt_SurfaceTemp = as.numeric(kurt_SurfaceTemp),
         sd_SurfaceTemp = as.numeric(sd_SurfaceTemp),
         med_surface_C = as.numeric(med_SurfaceTemp_corr) - 273.15, # Convert from Kelvin to Celsius
         doy_check = if_else(doy >= 120 & doy <= 300, "PASS", "FAIL"), # Filter for sampling season (roughly April to October)
  kurtosis_check = if_else(kurt_SurfaceTemp > 10, "FAIL", "PASS"), #filter out for instances with high kurtosis
  sd_check = if_else(sd_SurfaceTemp > 2, "FAIL", "PASS"), #filter out instances with high SD
pass_all = (sd_check == "PASS" & kurtosis_check == "PASS" & doy_check == "PASS"))

all_rs <- temp_rs %>%
  left_join(read_parquet("data/derived/remote_sensing/landsat/CLP_SDD_estimate_v2024-10-10.parquet"), by = c("gnis_name", "date"))%>%
  mutate(
    site_code = case_when(
    gnis_name == "Seaman Reservoir"     ~ "SEMR",    
    gnis_name =="Joe Wright Reservoir"    ~ "JOER", 
    gnis_name == "Barnes Meadow Reservoir"  ~ "BRNR",
    gnis_name == "Chambers Lake" ~ "CBRR",
    gnis_name == "Peterson Lake"     ~ "PTRR", 
    gnis_name ==     "Long Draw Reservoir"    ~ "LNGR",
gnis_name =="Hourglass Reservoir"   ~ "HORR",
gnis_name ==  "Comanche Reservoir"     ~ "COMR",
gnis_name == "Halligan Reservoir"  ~ "HALL",
    TRUE ~ NA_character_))

```

## Seasonal Trends in Temperatures

```{r}

graphTrendObs_temp = function(lake) {
  seasonAdjTrends <- all_rs %>% 
    mutate(doy = yday(date))%>%
    filter(site_code == lake & !is.na(med_surface_C) & (doy >= 120 & doy <= 300))%>%
    select(gnis_name, date, med_surface_C)%>%
    as_tibble()%>% 
    time_decompose(med_surface_C, merge = T)
  
  lakeName = seasonAdjTrends$gnis_name[1]
  ggplot(seasonAdjTrends, aes(x = date, y = observed)) +
    geom_point(color = 'gray') +
    geom_vline(xintercept = as.Date(c("2020-08-01")), linetype = "dashed", color = "red") +
    labs(x = NULL, 
         y = "Estimated lake surface temperature (C)",
         title = paste0(lakeName)) +
    geom_point(aes(y = trend), color = '#E69F00') +
    geom_smooth(aes(y = trend), lty = 2) +
    theme_bw() +
    theme(plot.title = element_text(face = 'bold', hjust = 0.5, size = 12))+
    ylim(0,30)
}

temp_plots <- map(c("BRNR", "PTRR", "CBRR","JOER", "LNGR"),~graphTrendObs_temp(.x))

wrap_plots(temp_plots, ncol = 1)+ plot_layout(axes = "collect")

```
## Seasonal Trends in SDD

```{r}

graphTrendObs_SDD = function(lake) {
    seasonAdjTrends <- all_rs %>% 
    mutate(doy = yday(date))%>%
    filter(site_code == lake & !is.na(mean_sdd) & (doy >= 120 & doy <= 300))%>%
    select(gnis_name, date, mean_sdd)%>%
    as_tibble() %>% 
    time_decompose(mean_sdd, merge = T) 
    
  lakeName = seasonAdjTrends$gnis_name[1]
  ggplot(seasonAdjTrends, aes(x = date, y = observed)) +
    geom_point(color = 'gray') +
    geom_vline(xintercept = as.Date(c("2020-08-01")), linetype = "dashed", color = "red") +
    labs(x = NULL, 
         y = "Estimated Mean Secchi Disk Depth (m)",
         title = paste0(lakeName)) +
    geom_point(aes(y = trend), color = '#E69F00') +
    geom_smooth(aes(y = trend), lty = 2) +
    theme_bw() +
    theme(plot.title = element_text(face = 'bold', hjust = 0.5, size = 12))+
    ylim(0,6)
}

sdd_plots <- map(c("BRNR", "PTRR", "CBRR","JOER", "LNGR"),~graphTrendObs_SDD(.x))

wrap_plots(sdd_plots, ncol = 1, axes = "collect")


```

## RS vs grabs

```{r}
chem_rs <- all_rs%>%
  rename(rs_date = date)%>%
  mutate(date_low = rs_date - days(4),
    date_high = rs_date + days(4)) %>%
  left_join(tidy_chem, 
            by = join_by(site_code, between(y$Date, x$date_low, x$date_high)))%>%
  filter(!is.na(med_surface_C))%>%
  mutate(date_diff = abs(as.numeric(difftime(Date, rs_date, units = "days"))))%>%
  filter(!(site_code %in% c("COMR", "HORR") & Year %in% c(2021,2022)))%>% # removing unfilled comanche and hourglass matchups
  select(rs_date,sample_date = Date, date_diff, site_code, gnis_name, med_surface_C, mean_sdd, ChlA, Turbidity, TSS, NO3, DTN, DOC, Year)
```



Not great associations and there is only \~120 matchups between the seven reservoirs

```{r}
ggplot(chem_rs%>%filter(!is.na(ChlA)), aes(x = med_surface_C, y = ChlA,shape = Year))+
  geom_point()+
  facet_wrap(~site_code)+
  theme_bw()+
  labs(x = "Estimated Surface Temperature (°C)", y = "Sample Chlorophyll a (µg/L)")


ggplot(chem_rs%>%filter(!is.na(ChlA)), aes(x = mean_sdd, y = TSS, shape = Year))+
  geom_point()+
  facet_wrap(~site_code)+
  theme_bw()+
  labs(x = "Estimated Mean Secchi Disk Depth (m)", y = "Sample TSS (mg/L)")

ggplot(chem_rs%>%filter(!is.na(ChlA)), aes(x = mean_sdd, y = Turbidity, shape = Year))+
  geom_point()+
  facet_wrap(~site_code)+
  theme_bw()+
  labs(x = "Estimated Mean Secchi Disk Depth (m)", y = "Sample Turbidity (NTU)")

ggplot(chem_rs%>%filter(!is.na(ChlA)), aes(x = mean_sdd, y = ChlA,shape = Year))+
  geom_point()+
  facet_wrap(~site_code)+
  theme_bw()+
  labs(x = "Estimated Mean Secchi Disk Depth (m)", y = "Sample Chlorophyll a (µg/L)")

ggplot(chem_rs%>%filter(!is.na(ChlA)), aes(x = mean_sdd, y = DOC, shape = Year))+
  geom_point()+
  facet_wrap(~site_code)+
  theme_bw()+
  labs(x = "Estimated Mean Secchi Disk Depth (m)", y = "Sample DOC (mg/L)")


```


# SI

## SI Table 1: PO4 detections over time


Looking at where PO4 is present (above 0.01 detection limit) and if that
is changing over time
```{r}
# How many samples per study reservoirs

tidy_chem%>%
  filter( site_code %in% primary_study_res)%>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4, PO4), 
               names_to = "param", 
               values_to = "value")%>%
  filter(param == "PO4")%>%
  group_by(site_code, burn_group)%>%
  summarise(samples = n(), 
           all_samples = sum(value),
            abv_nd  = sum(value > 0))%>%
  mutate(perc_abv_nd = abv_nd/samples*100)

# How many samples per year at study reservoirs

tidy_chem%>%
  filter( site_code %in% primary_study_res)%>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4, PO4), 
               names_to = "param", 
               values_to = "value")%>%
  filter(param == "PO4")%>%
  group_by(Year)%>%
  summarise(samples = n(), 
           all_samples = sum(value),
            abv_nd  = sum(value > 0))%>%
  mutate(perc_abv_nd = abv_nd/samples*100)


# How many samples at mainstem/stream sites

tidy_chem%>%
  filter( site_code %nin% all_res_system)%>%
  pivot_longer(cols = c(ChlA,Turbidity, TSS, DTN,  DOC:NO3, SO4, PO4), 
               names_to = "param", 
               values_to = "value")%>%
  filter(param == "PO4")%>%
  group_by(Year)%>%
  summarise(samples = n(), 
           all_samples = sum(value),
            abv_nd  = sum(value > 0))%>%
  mutate(perc_abv_nd = abv_nd/samples*100)

```

```{r}

upper_PO4 <- filter(tidy_chem, watershed %nin% c("CLP Mainstem - Lower","CLP Tributary - Lower"))%>%
  select(site_code,location_type, Date, Turbidity, PO4, Year)%>%
  mutate(location_type = case_when(
    site_code %in% c("JWC", "BEAV", "LBEA", "PENN", "SFM", "SAWM", "JOEI", "CBRI", "COMI", "HORI") ~ "Tributary",
    TRUE ~ location_type))%>%
  group_by(location_type, Year)%>%
  # fina what percent of each year had PO4 levels above 0.01 and get the mean level for each location type and year
  summarise(total_samples = n(),
            PO4_perc_dect = sum(PO4 >= 0.01, na.rm = TRUE)/n()* 100,
            #calculate mean of PO4 for only samples above 0.01, excluding samples lower than 0.01
            mean_PO4 = mean(PO4[PO4 >= 0.01], na.rm = TRUE),
            samples = paste0(sum(PO4 >= 0.01, na.rm = TRUE),"/",n() ))%>%
  # pivot wider by year
  pivot_wider(names_from = Year, values_from = c(total_samples, PO4_perc_dect, mean_PO4, samples)) %>%
  mutate(across(total_samples_2021:mean_PO4_2024, ~replace_na(.x, 0)))%>%
  mutate(across(samples_2021:samples_2024, ~replace_na(.x, "0/0")))%>%
  # if a column is numeric, round it to 2 decimal places
  mutate(across(where(is.numeric), ~round(.x, 2)))%>%
  arrange(factor(location_type, levels = c("Inflow", "Reservoir", "Outflow", "Mainstem", "Tributary")))%>%
  select(location_type, total_samples_2021:mean_PO4_2024)

#### Creating Table ####
# Add spacer columns
upper_PO4_with_spacers <- upper_PO4 %>%
  mutate(sp1 = NA, sp2 = NA, sp3 = NA) %>%
  select(1, 2:5, sp1, 6:9, sp2, 10:13, sp3)

# Rename columns with valid names
colnames(upper_PO4_with_spacers) <- c(
  "Site_Type",           # 1
  "n_2021", "n_2022", "n_2023", "n_2024", "sp1",  # 2-6
  "pct_2021", "pct_2022", "pct_2023", "pct_2024", "sp2",  # 7-11
  "mean_2021", "mean_2022", "mean_2023", "mean_2024", "sp3" # 12-16
)

# Create the flextable
ft1 <- flextable(upper_PO4_with_spacers) %>%
  set_header_labels(
    Site_Type = "",
    n_2021 = "2021", n_2022 = "2022", n_2023 = "2023", n_2024 = "2024",
    pct_2021 = "2021", pct_2022 = "2022", pct_2023 = "2023", pct_2024 = "2024",
    mean_2021 = "2021", mean_2022 = "2022", mean_2023 = "2023", mean_2024 = "2024",
    sp1 = "", sp2 = "", sp3 = ""
  ) %>%
  add_header_row(
    values = c("Site Type", "Samples per Year", "", "Percent above Detection Limit", "", 
               "Mean PO4 Above Detection Limit", ""),
    colwidths = c(1, 4, 1, 4, 1, 4, 1)
  ) %>%
  theme_vanilla() %>%
  align(align = "center", part = "all") %>%
  valign(valign = "center", part = "all") %>%
  fontsize(size = 11, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  set_table_properties(layout = "autofit", width = 1)%>%
  autofit()

ft1
save_as_docx(ft1,path = here(paper_draft_dir,"SI_1_Table.docx"))


```

## SI Table 2: Mean, SD, Max for nutrients for reservoirs

Calculate mean, sd and max for key constituents for the 7 main study
reservoirs

```{r}

key_table <- tidy_chem %>%
  filter(site_code %in% c("LNGR", "JOER","COMR", "HORR",  "CBRR", "PTRR", "BRNR")) %>%
  select(site_code, Year, pH,SC,TSS, DOC, DTN, K,NH4,NO3, SO4) %>%
  pivot_longer(cols = c(pH,SC,TSS, DOC, DTN, K,NH4,NO3, SO4),
               names_to = "parameter",
               values_to = "value") %>%
  #get the median and max for each site
  group_by(site_code, parameter) %>%
  summarise(median = median(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE),
            max = max(value, na.rm = TRUE)) %>%
  ungroup()%>%
  pivot_longer(cols = c(median, max, sd),
               names_to = "stat",
               values_to = "stat_value")%>%
  unite("param_stat", parameter, stat, sep = " ") %>%
  pivot_wider(names_from = param_stat, values_from = stat_value)%>%
  arrange(factor(site_code, levels = c("LNGR", "JOER", "HORR", "COMR", "CBRR", "PTRR", "BRNR")))%>%
  select(site_code:`NO3 sd`,`SO4 median`:`pH sd`, `SC median`,`SC max`,`SC sd`)%>%
  #round all numeric values to 2 digits
  mutate(across(where(is.numeric), ~ round(.x, 2)))


param_oi <- c( "DOC", "DTN", "K", "NH4", "NO3", "SO4", "TSS", "pH", "SC")

params <- chem_units %>%
  filter(simple %in% param_oi) %>%
  arrange(match(simple, param_oi)) %>%
  pull(Parameters)


# Create a basic flextable
ft2 <- key_table %>%
  rename(Site = site_code) %>%
  flextable() %>%
  set_header_labels(
    .labels = setNames(rep(c("mean", "max", "sd"), times = length(params)),
                      nm = as.vector(t(outer("", c("median", "max","sd"), paste, sep = " "))))) %>%
  delete_part(part = "header")%>%
  add_header_row(
    values = c("", rep(c("Median", "Max", "SD"), times = length(params))),  
    colwidths = c(1, rep(1, length(params)*3))  # Adjust colwidths for Site and parameters
  ) %>%
  add_header_row(
    values = c("Site", params), 
    colwidths = c(1, rep(3, length(params)))  # Adjust colwidths for Site and parameters
  ) %>%
  
  add_header_row(
    values = c("", 
                "mg L⁻¹", "mg L⁻¹", "mg L⁻¹", 
               "mg L⁻¹", "mg L⁻¹", "mg L⁻¹", "mg L⁻¹",  " ", "µS cm⁻¹"), 
    colwidths = c(1, rep(3, length(params)))  # Adjust colwidths again to match
  ) %>%
  merge_v(j = ~Site) %>%
  merge_h(part = "header") %>%
  border_remove() %>%
  border_outer() %>%
  border_inner_v() %>%
  border_inner_h() %>%
  align(align = "center", part = "all") %>%
  valign(valign = "center", part = "all") %>%
  fontsize(size = 11, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  set_table_properties(layout = "autofit", width = 1)%>%
  autofit()

sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    width = 8.3, height = 11.7
  ),
  type = "continuous",
  page_margins = page_mar()
)

save_as_docx( ft2, path = here(paper_draft_dir,"SI_2_table.docx"))

```

## SI Table 3: Mean, SD, Max for remainders for reservoirs

```{r}

base_cations_table <- tidy_chem %>%
  filter(site_code %in% c("LNGR", "JOER","COMR", "HORR",  "CBRR", "PTRR", "BRNR")) %>%
  select(site_code, Year, ANC, Na, Mg, Ca, F, Cl) %>%
  pivot_longer(cols = c(ANC, Na, Mg, Ca, F, Cl),
               names_to = "parameter",
               values_to = "value") %>%
  #get the median and max for each site
  group_by(site_code, parameter) %>%
  summarise(median = median(value, na.rm = TRUE),
            max = max(value, na.rm = TRUE),
            sd = sd(value,na.rm = T)) %>%
  ungroup()%>%
  pivot_longer(cols = c(median, max,sd),
               names_to = "stat",
               values_to = "stat_value")%>%
  unite("param_stat", parameter, stat, sep = " ") %>%
  pivot_wider(names_from = param_stat, values_from = stat_value)%>%
  arrange(factor(site_code, levels = c("LNGR", "JOER", "HORR", "COMR", "CBRR", "PTRR", "BRNR")))%>%
  #round all numeric values to 2 digits
  mutate(across(where(is.numeric), ~ round(.x, 2)))


param_oi <- c("ANC", "Ca", "Cl", "F", "Mg", "Na")

params <- chem_units %>%
  filter(simple %in% param_oi) %>%
  arrange(match(simple, param_oi)) %>%
  pull(Parameters)


# Create a basic flextable
ft3 <- base_cations_table %>%
  rename(Site = site_code) %>%
  flextable() %>%
  set_header_labels(
    .labels = setNames(rep(c("mean", "max", "sd"), times = length(params)),
                      nm = as.vector(t(outer("", c("median", "max","sd"), paste, sep = " "))))) %>%
  delete_part(part = "header")%>%
  add_header_row(
    values = c("", rep(c("Median", "Max", "SD"), times = length(params))),  
    colwidths = c(1, rep(1, length(params)*3))  # Adjust colwidths for Site and parameters
  ) %>%
  add_header_row(
    values = c("Site", params), 
    colwidths = c(1, rep(3, length(params)))  # Adjust colwidths for Site and parameters
  ) %>%
  
  add_header_row(
    values = c("", 
                "mg L⁻¹", "mg L⁻¹", "mg L⁻¹", 
               "mg L⁻¹", "mg L⁻¹", "mg L⁻¹"), 
    colwidths = c(1, rep(3, length(params)))  # Adjust colwidths again to match
  ) %>%
  merge_v(j = ~Site) %>%
  merge_h(part = "header") %>%
  border_remove() %>%
  border_outer() %>%
  border_inner_v() %>%
  border_inner_h() %>%
  align(align = "center", part = "all") %>%
  valign(valign = "center", part = "all") %>%
  fontsize(size = 11, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%
  set_table_properties(layout = "autofit", width = 1)%>%
  autofit()

sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    width = 8.3, height = 11.7
  ),
  type = "continuous",
  page_margins = page_mar()
)

save_as_docx( ft3, path = here(paper_draft_dir,"SI_3_table.docx"))

```

## SI Plot 1: Timelines of key parameters at reservoirs

```{r}

SI_plot <- tidy_chem %>%
  filter(site_code %in% c("LNGR", "JOER","COMR", "HORR",  "CBRR", "PTRR", "BRNR"))%>%
  pivot_longer(cols = c( ChlA, DTN, pH, SC, DOC, K, NH4, NO3, SO4, K, Ca, ANC, Na, Mg, Ca, F, Cl),
               names_to = "parameter",
               values_to = "value")%>%
  left_join(chem_units, by = c("parameter" = "simple"))%>%
  mutate(Date = as.POSIXct(Date ))


#plot 1 

Si_timeline_1 <- ggplot(SI_plot%>%filter(parameter %in% c("DOC", "DTN","pH", "SC", "ANC"))%>%
         mutate(parameter = factor(parameter, levels = c("DOC", "DTN","pH", "SC", "ANC"))),
       aes(x = Date, y = value, group = interaction(site_code, Year), color = site_code)) +
  geom_line() +
  geom_point()+
  facet_grid(rows = vars(combined),cols = vars(Year), scales = "free", switch = "y") +
  labs(x = "Date",
       y = "") +
  scale_color_manual(values = colorsBS_site_code)+
  scale_x_datetime(
    date_labels = "%b",  # e.g., May\n2021
    date_breaks = "month"
  ) +
  labs(color = "Site")+
  theme_bw() +
  theme(strip.placement = "outside", 
        legend.position = "bottom",
        # make the x ticks at 45 degrees
        axis.text.x = element_text(angle = 45, hjust = 1))

save_plot_image(Si_timeline_1,"SI_timeline_1.png")


SI_timeline_2 <- 
ggplot(SI_plot%>%filter(parameter %in% c("K", "Ca","SO4", "Na", "Cl", "Mg", "F")),
       aes(x = Date, y = value, group = interaction(site_code, Year), color = site_code)) +
  geom_line() +
  geom_point()+
  facet_grid(rows = vars(combined),cols = vars(Year), scales = "free", switch = "y") +
  labs(x = "Date",
       y = "") +
  scale_color_manual(values = colorsBS_site_code)+
  scale_x_datetime(
    date_labels = "%b",  # e.g., May\n2021
    date_breaks = "month"
  ) +
  labs(color = "Site")+
  theme_bw() +
  theme(strip.placement = "outside", 
        legend.position = "bottom",
        # make the x ticks at 45 degrees
        axis.text.x = element_text(angle = 45, hjust = 1))

save_plot_image(Si_timeline_2,"SI_timeline_2.png")

```

## SI Plot 2: Percent of reservoir samples eutrophic status

In cold water lakes (our study sites), mesotrophic levels are defined as
between 2.6 and 8 µg/L, and eutrophic levels are greater than 8 µg/L.
Under 2.6 µg/L is considered Oligotrophic.

```{r}

chla_status_site_year <- tidy_chem%>%
  filter(!is.na(ChlA))%>%
  filter(site_code %in% primary_study_res)%>%
  mutate(eu_status  = case_when(ChlA >= 2.6 &ChlA <= 8 ~ "Mesotrophic", 
                                ChlA > 8 ~ "Eutrophic",
                                ChlA < 2.6 ~ "Oligotrophic"))%>%
  group_by(site_code, Year, eu_status, burn_group)%>%
  summarise(eu_stats = n())%>%
  ungroup()%>%
  pivot_wider(names_from = eu_status, values_from = eu_stats)%>%
  mutate_at(vars(-Year), ~replace(., is.na(.), 0))%>%
  rowwise()%>%
  mutate(total_samples = sum(c(Eutrophic, Mesotrophic, Oligotrophic)),
    perc_eu = round((Eutrophic/total_samples)*100, digits= 1), 
         perc_meso = round((Mesotrophic/total_samples)*100, digits= 1),
         perc_meso_eu = round(((Eutrophic+Mesotrophic)/total_samples)*100, digits= 1), 
        perc_oligo = round((Oligotrophic/total_samples) *100, digits= 1),
    Year = as.character(Year))

all_years <- chla_status_site_year%>%
  group_by(burn_group)%>%
  summarise(total_samples = sum(total_samples), 
            perc_eu = sum(Eutrophic)/sum(total_samples)*100, 
            perc_meso = sum(Mesotrophic)/sum(total_samples)*100, 
            perc_meso_eu = sum(Eutrophic + Mesotrophic)/sum(total_samples)*100, 
            perc_oligo = sum(Oligotrophic)/sum(total_samples)*100)

long_chla_status <- chla_status_site_year%>%
  select(site_code, Year, total_samples, perc_eu, perc_meso, perc_meso_eu, perc_oligo)%>%
  pivot_longer(cols = c(perc_eu, perc_meso, perc_meso_eu, perc_oligo), names_to = "category", values_to = "value" )

long_chla_status$site_code = factor(long_chla_status$site_code, levels = c("LNGR","JOER" , "CBRR", "PTRR", "BRNR" ))

chla_status <- c(
  perc_eu = "% eutrophic",
  perc_meso = "% mesotrophic ",
  perc_meso_eu = "% mesotrophic & eutrophic",
  perc_oligo = "% oligotrophic"
)

#PLOT
 long_chla_status%>%
  filter(site_code %in% primary_study_res )%>%
   filter(category != "perc_meso_eu")%>%
  ggplot(aes(x = Year, y = value,group = site_code, color =  site_code))+
  scale_color_manual(values = colorsBS_site_code)+
 geom_line()+
   geom_point()+
  theme_bw(base_size = 15)+
facet_wrap(~category, ncol = 1, labeller = labeller(category = chla_status))+
  labs(
    x = "Year",
    y = "Percent of chlorophyll-*a* samples", 
    color = "Reservoir")+
  theme(axis.title.y = ggtext::element_markdown())

```


# Watershed Characteristics (Spatial)

## Merging transbasin diversion watersheds

```{r}

grand_ditch <- st_read("data/raw/spatial/res_study/working_tb/grand_ditch.shp" )%>%
  st_transform(4326)

michigan_ditch <- st_read("data/raw/spatial/res_study/working_tb/michigan_ditch.shp" )%>%
  st_transform(4326)

LRT <- st_read("data/raw/spatial/res_study/working_tb/Laramie_River_Tunnel.shp" )%>%
  st_transform(4326)


mapview::mapview(grand_ditch)+
  mapview::mapview(michigan_ditch)+
  mapview::mapview(LRT)

TB <- bind_rows(
  grand_ditch,
  michigan_ditch,
  LRT
)

#mapview::mapview(TB, zcol = "site_name", burst = T)
#st_write(TB, "data/raw/spatial/res_study/trans_basin_watersheds.shp")

#Old code to merge stream stats generated watershed with relevant transbasin
# 
# polygon3 <- st_read("data/spatial/working/SLEP_ss/layers/globalwatershed.shp" )%>%
#   st_transform(4326)%>%
#   mutate(site_name = "site_streamstats")%>%
#   select(site_name, geometry)
# 
# st_is_valid(polygon3)
#polygon3 <- st_make_valid(polygon3)
# 
# mapview::mapview(polygon3)

# Merge polygons
# merged_geometry <- st_difference(polygon2$geometry, polygon1$geometry)
# #merged_geometry <- st_union(polygon1$geometry, polygon2$geometry)
# 
# merged_polygon <- st_sf(
#   site_name = "HORR",
#   geometry = st_sfc(merged_geometry))%>%
#   st_make_valid()
# 
# st_is_valid(merged_polygon)
# 
# mapview::mapview(merged_polygon)
```


## Landcover

```{r}

#pull in watersheds
watersheds <- st_read(dsn = here(spatial_dir,"cpf_watersheds.shp"))%>%
  left_join(most_recent_meta, by = c("site_code"))

nlcd_tiffs <- list.files(here(external_spatial, "landcover_NLCD"), pattern = "Annual_NLCD_LndCov",  full.names = T)%>%
  #find the files that contain ".tff" and nothing after it
  .[grepl(".tiff$", .)]
# read in rasters and name them based on the year contained in the file name
nlcd_rasters <- map(nlcd_tiffs, raster::raster)
#name them based on the year contained in the file name
names(nlcd_rasters) <- nlcd_tiffs%>%
  stringr::str_extract("\\d{4}")%>%
  as.numeric()


compute_top3_landcovers <- function(year){
  
  NLCD <- nlcd_rasters[[year]]
  
  nlcd_raster <- raster::brick(NLCD)


  ws <- watersheds %>%
    select(site_code)%>%
    rownames_to_column('index')%>%
    mutate(index = as.numeric(index))
  
  # using the exactextractr package, get what is essentially a "zonal histogram" in
  # ArcGIS speak
  ws_NLCD <- exactextractr::exact_extract(nlcd_raster, ws,
                                         append_cols = "index",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           group_by(lc_code = as.character(value)) %>%
                                           summarise(count = sum(coverage_fraction))) %>%
    group_by(index) %>% 
    mutate(total = sum(count)) %>%
    ungroup() %>%
    pivot_wider(names_from = 'lc_code', values_from = 'count',
                names_glue = "{lc_code}" ) %>%
    inner_join(ws, .) -> ws_NLCD
  
  ws_percents <- ws_NLCD %>% 
    #if there are NAs, change it to 0
    mutate(across(`11`:`82`, ~replace_na(., 0))) %>%
    mutate(water=100*(`11`/total),
           ice_snow=100*(`12`/total),
           open_space=100*(`21`/total),
           developed = 100*((`22`+ `23`+ `24`)/total),
           Barren = 100*(`31`/total),
           `Decidious Forest` = 100*((`41`)/total),
           `Evergreen Forest` = 100*((`42`)/total),
           `Mixed Forest` = 100*((`43`)/total),
           `Shrubland` = 100*((`52`)/total),
           `Grassland` = 100*((`71`)/total),
           pasture = 100*((`81`)/total),
           crops = 100*((`82`)/total),
           `Wetlands` = 100*((`90`+ `95`)/total),
           total_check = rowSums(across(water:Wetlands)))%>%
    st_drop_geometry()%>%
    select(total, site_code, water:total_check)
  
  
  top_land_cover <- ws_percents %>%
    select(-total, -total_check) %>%
  # Pivot longer to get land cover types and percentages in rows
  pivot_longer(cols = water:Wetlands, 
               names_to = "land_cover", 
               values_to = "percent")%>%
  # Group by site_code and find the top 4 land covers by percent
  group_by(site_code) %>%
  arrange(desc(percent), .by_group = TRUE) %>%
  slice_head(n = 3) %>%
  mutate(rank = row_number()) %>% # Rank the top 4 land covers
  ungroup() %>%
  # Pivot wider with unique column names
  pivot_wider(
    id_cols = site_code, 
    names_from = rank, 
    values_from = c(land_cover, percent), 
    names_glue = "{.value}_{rank}"
  )%>%
    select(site_code, land_cover_1, percent_1, land_cover_2, percent_2, land_cover_3, percent_3)%>%
    mutate(nlcd_year = year)%>%
    #round all percent columns to 1 decimal places
    mutate(across(contains("percent"), ~round(., 1)))


return(top_land_cover)
  
  
}

compute_all_landcovers <- function(year){
  
  NLCD <- nlcd_rasters[[year]]
  
  nlcd_raster <- raster::brick(NLCD)


  ws <- watersheds %>%
    select(site_code)%>%
    rownames_to_column('index')%>%
    mutate(index = as.numeric(index))
  
  # using the exactextractr package, get what is essentially a "zonal histogram" in
  # ArcGIS speak
  ws_NLCD <- exactextractr::exact_extract(nlcd_raster, ws,
                                         append_cols = "index",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           group_by(lc_code = as.character(value)) %>%
                                           summarise(count = sum(coverage_fraction))) %>%
    group_by(index) %>% 
    mutate(total = sum(count)) %>%
    ungroup() %>%
    pivot_wider(names_from = 'lc_code', values_from = 'count',
                names_glue = "{lc_code}" ) %>%
    inner_join(ws, .) -> ws_NLCD
  
  all_land_cover <- ws_NLCD %>% 
    #if there are NAs, change it to 0
    mutate(across(`11`:`82`, ~replace_na(., 0))) %>%
    mutate(water=100*(`11`/total),
           ice_snow=100*(`12`/total),
           open_space=100*(`21`/total),
           developed = 100*((`22`+ `23`+ `24`)/total),
           Barren = 100*(`31`/total),
           `Decidious Forest` = 100*((`41`)/total),
           `Evergreen Forest` = 100*((`42`)/total),
           `Mixed Forest` = 100*((`43`)/total),
           `Shrubland` = 100*((`52`)/total),
           `Grassland` = 100*((`71`)/total),
           pasture = 100*((`81`)/total),
           crops = 100*((`82`)/total),
           `Wetlands` = 100*((`90`+ `95`)/total),
           total_check = rowSums(across(water:Wetlands)))%>%
    st_drop_geometry()%>%
    select( site_code, total, water:`Wetlands`)%>%
    mutate(nlcd_year = year)

return(all_land_cover)
  
  
}

years <- c("2019", "2023")

ws_lc <- map_df(years, compute_top3_landcovers)%>%
  bind_rows() %>%
  #swapping water for land_cover_3
  mutate(land_cover_2 = case_when(land_cover_2 == "water" ~ land_cover_3, 
                                  TRUE ~ land_cover_2), 
         percent_2 = case_when(land_cover_2 == "water" ~ percent_3, 
                               TRUE ~ percent_2), 
         # creating a new column that combines the top 2 land covers for table
         land_cover_top2 = paste0(land_cover_1, " (", percent_1, " %) \n",land_cover_2, " (", percent_2," %) " ))%>%
  select(site_code, nlcd_year,land_cover_top2) %>%
  pivot_wider(names_from = nlcd_year, values_from = land_cover_top2, names_prefix = "lc_top2_")

#run to get all landcover %s for watersheds
ws_all_lc <- map_df(years, compute_all_landcovers)%>%
  bind_rows()%>%
  pivot_longer(
    cols = water:Wetlands, 
    names_to = "landcover", 
    values_to = "value"
  ) %>%
  select(-total)%>%
  pivot_wider(,
    names_from = nlcd_year, 
    values_from = value,
    names_prefix = "value_" # This creates value_2019 and value_2023
  ) %>%
  select(site_code, landcover, value_2019, value_2023)%>%
  mutate(pp_change = value_2023 - value_2019)


rm(years)
```

## Slope, elevation and aspect

```{r}

ws_chars<- map_dfr(unique(watersheds$site_code), function(ws_code){
  
  ws <- watersheds %>%
    filter(site_code == ws_code)
  
   ws <- st_make_valid(ws)
  
  if(ws_code %in% c("JOER", "CBRR", "BRNR", "LNGR", "PTRR", "COMR","HORR")){
    z_num = 10
  }else{
    z_num = 9
  }
  
  elev <- get_elev_raster(ws, z = z_num, verbose = F)

  ws_elev <- exactextractr::exact_extract(elev, ws,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           summarise(mean_elev = mean(value, na.rm = T), 
                                                     max_elev = max(value, na.rm = T), 
                                                     min_elev = min(value, na.rm = T))) %>%
    select(site_code, mean_elev, max_elev, min_elev)
  
  
  #calculate slope
  slope <- raster::terrain(elev, opt = "slope", unit = "degrees", verbose = F)
  ws_slope <- exactextractr::exact_extract(slope, ws,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           summarise(mean_slope = mean(value, na.rm = T), 
                                                     max_slope = max(value, na.rm = T), 
                                                     min_slope = min(value, na.rm = T))) %>%
    select( mean_slope, max_slope)
  
  #calculate aspect
  aspect <- raster::terrain(elev, opt = "aspect", unit = "degrees")
  ws_aspect <- exactextractr::exact_extract(aspect, ws,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                          summarise(mean_aspect = mean(value, na.rm = T))) %>%
    select(mean_aspect)%>%
    round(2)
  # calculate watershed area
  ws_area <- st_area(ws)%>%
    #convert to km2
    units::set_units("km^2")
  
  bind_cols(ws_elev, ws_slope, ws_aspect)%>%
    mutate(ws_area_km2 = ws_area)
  
})

```
## Calculating snow persistance for watersheds

```{r}

#get the snow persistence raster
snow <- raster::raster(here(external_spatial, "snow_persistance/MOD10A2_SCI_2020.tif"))

ws_snow <- map_dfr(unique(watersheds$site_code),  function(ws_code){
  
  ws <- watersheds %>%
    filter(site_code == ws_code)
  exactextractr::exact_extract(snow, ws,
                                          append_cols = "site_code",
                                          summarize_df = T,
                                          fun=function(x) x %>%
                                            summarise(mean_snow = mean(value, na.rm = T), 
                                                      max_snow = max(value, na.rm = T), 
                                                      min_snow = min(value, na.rm = T))) %>%
    select(site_code, mean_snow, max_snow, min_snow)%>%
    mutate(across(where(is.numeric), ~round(., 0)))%>%
    mutate(days = 184 * (mean_snow/100), 
           mean_snow_free_date= format(as.Date(days, origin = "2020-01-01"), "%m-%d"))%>%
    select(site_code, mean_snow, mean_snow_free_date)
})

```

## CPF Elevational Gradient

```{r}
ws <-  terra::rast(here(spatial_dir, "cpf_sbs_BAER.tif"))
#remove 0 values
#ws[ws == 1] <- NA
ws[ws == 0] <- NA

#convert to shapefile
# Convert raster to polygon (dissolve all cells into one)
outline <- as.polygons(ws, dissolve = TRUE, na.rm = TRUE) %>%
  st_as_sf()

outline$site_code <- "cpf"

outline <- outline %>%
  group_by(site_code) %>%
  summarise(geometry = st_union(geometry), .groups = "drop") %>%
  st_cast("POLYGON")

elev <- get_elev_raster(outline, z = 9, verbose = F)



ws_elev <- exactextractr::exact_extract(elev, outline,
                                        append_cols = "site_code",
                                        summarize_df = T,
                                        fun=function(x) x %>%
                                          summarise(mean_elev = mean(value, na.rm = T), 
                                                    max_elev = max(value, na.rm = T), 
                                                    min_elev = min(value, na.rm = T)))%>%
  select(site_code, mean_elev, max_elev, min_elev)


ws_area <- st_area(outline)%>%
  #convert to km2
  units::set_units("km^2")


cpf<- bind_cols(ws_elev)%>%
  mutate(ws_area_km2 = ws_area)%>%
  group_by(site_code)%>%
  summarise(mean_elev = max(mean_elev),
            max_elev = max(max_elev),
            min_elev = min(min_elev),
            ws_area_km2 = sum(ws_area_km2))
```



## Soil Burn Severity

From Rhoades paper : "We partitioned the catchments into groups based on the proportional area burned at high + moderate severity (Extensive:>50%, High: 25–50%, Mid:<25%). Unburned catchments and those with <25% total burn extent, no high severity and little moderate severity (0–13%) were included in a No/Low class."

```{r}

sbs <- terra::rast(here(spatial_dir, "cpf_sbs_BAER.tif"))

# Load the shapefile and raster
ws <- st_read(here(spatial_dir, "cpf_watersheds.shp")) %>%
  st_transform(sf::st_crs(sbs)) 

ws_raster <- ws %>%
  terra::vect()

# Ensure all NA values in the raster are already set to 5
sbs[is.na(sbs) | is.null(sbs) | sbs == 0] <- 5

# Get the extent of the ws vector
ws_extent <- terra::ext(ws_raster)

# Extend the raster to match the ws extent, assigning new areas the value of 5 (Unburned)
sbs_extended <- terra::extend(sbs, ws_extent, fill = 5)

# Get the cell size in map units (meters)
cell_size <- terra::res(sbs_extended)

cnv_to_km2 <- (cell_size[1] * cell_size[2]) / 1e6

# using the exactextractr package, get what is essentially a "zonal histogram" in
# ArcGIS speak
ws_extracted_sbs <- exactextractr::exact_extract(sbs_extended %>% raster::raster(), ws,
                                              append_cols = "site_code",
                                              summarize_df = T,
                                              fun=function(x) x %>%
                                                group_by(sbs_code = as.character(value)) %>%
                                                summarise(count = sum(coverage_fraction))) %>%
  group_by(site_code)%>%
  mutate(total = sum(count)) %>%
  ungroup() %>%
  pivot_wider(names_from = 'sbs_code', values_from = 'count',
              names_glue = "sbs{sbs_code}" ) %>%
  inner_join(ws, .) -> ws_sbs

ws_sbs_full <- ws_extracted_sbs %>% 
  ungroup()%>%
  st_drop_geometry()%>%
  #if a column is NA change it to 0
  mutate(across(starts_with("sbs"), ~replace_na(., 0))) %>%
  mutate(
    # 1. Define raw pixel counts for burn classes
    Unburned = sbs5 + sbs6,
    `Very Low` = sbs1,
    Low = sbs2,
    Moderate = sbs3,
    High = sbs4,
    
    # 2. Convert to Square Kilometers (km2)
    Unburned_km2 = Unburned * cnv_to_km2,
    VeryLow_km2 = `Very Low` * cnv_to_km2,
    Low_km2 = Low * cnv_to_km2,
    Moderate_km2 = Moderate * cnv_to_km2,
    High_km2 = High * cnv_to_km2,
    total_km2 = total * cnv_to_km2,
    
    # 3. Calculate Percentages (Assumes 'total' column exists in raw pixels)
    Unburned_pct = (Unburned / total) * 100,
    VeryLow_pct = (`Very Low` / total) * 100,
    Low_pct = (Low / total) * 100,
    Moderate_pct = (Moderate / total) * 100,
    High_pct = (High / total) * 100
  )%>%
  mutate(across(Unburned:High_pct, ~round(., 1)))%>%
  mutate(extent_burned = ((`Very Low` + Low + Moderate + High)),
         extent_burned_km2 = extent_burned * cnv_to_km2,
         extent_burned_pct = (extent_burned / total) * 100,
         high_burn = (Moderate + High)/total*100,
         burn_class = case_when(high_burn > 50 ~ "Extensive", 
                                high_burn > 25 & high_burn <= 50 ~ "High", 
                                high_burn < 25 & Moderate_pct <= 13 & high_burn <= 5 ~ "No/Low",
                                TRUE ~ "Mid"))

ws_sbs_concise <- ws_sbs_full%>%
  #filter(site_code %in% c("CBRR", "BRNR", "PTRR","COMR", "HORR", "JOER", "LNGR"))%>%
  select(site_code, 
         `Total Area (km2)` = total_km2,
         `Extent Burned (km2)` = extent_burned_km2,
         `Extent Burned (%)` = extent_burned_pct,
         `High Burn (%)` = high_burn, 
         `Watershed Burn Class` = burn_class)
  


```

### Buffer SBS

```{r}
waterbody <- st_read("data/raw/spatial/res_study/study_waterbodies.shp")%>%
  st_transform(sf::st_crs(sbs)) #transform to coordinates in sbs

waterbody_buffer <- waterbody%>%
  #add a buffer of 200 meters to the waterbodies
  st_buffer(200) 

# pick out watersheds of reservoirs
res_ws <- ws%>%
  filter(site_code %in% c("JOER","LNGR","COMR","HORR","CBRR","PTRR", "BRNR"))%>%
  st_transform(sf::st_crs(sbs))

#clip buffer to it's watershed
waterbody_buffer <- st_intersection(waterbody_buffer,res_ws )%>%
  #only keep the waterbodies that are in the same watershed as the reservoir
  filter(site_code == site_code.1)

#remove water body from buffer
final_buffer <- st_difference(waterbody_buffer, waterbody)%>%
  filter(site_code == site_code.1 & site_code.1 == site_code.2)

#map to view
# tm_basemap("OpenStreetMap") +
# tm_shape(final_buffer) +
#   tm_polygons(col = "site_code", palette = "Set2", alpha = 0.5, border.col = "black", border.lwd = 1, title = "Buffered Waterbodies")+
#   #put the legend inside the plot
#   tm_layout(legend.position = c("right", "bottom"), title = "Buffered Waterbodies within Watersheds", title.position = c("right", "top"))


# Get the extent of the ws vector
waterbody_buffer_extent <- terra::ext(final_buffer)

# Extend the raster to match the ws extent, assigning new areas the value of 5
sbs_extended <- terra::extend(sbs, waterbody_buffer_extent, fill = 5)

# Get the cell size in map units (meters)
cell_size <- terra::res(sbs_extended)

  extracted_sbs <- exactextractr::exact_extract(sbs_extended %>% raster::raster(), final_buffer,
                                         append_cols = "site_code",
                                         summarize_df = T,
                                         fun=function(x) x %>%
                                           group_by(sbs_code = as.character(value)) %>%
                                           summarise(count = sum(coverage_fraction))) %>%
    group_by(site_code)%>%
    mutate(total = sum(count)) %>%
    ungroup() %>%
    pivot_wider(names_from = 'sbs_code', values_from = 'count',
                names_glue = "sbs{sbs_code}" ) %>%
    inner_join(ws, .) -> ws_sbs
  
  

  
  waterbody_buffer_percents <- extracted_sbs %>% 
    ungroup()%>%
    st_drop_geometry()%>%
    #if a column is NA change it to 0
    mutate(across(starts_with("sbs"), ~replace_na(., 0))) %>%
    mutate(
      # 1. Define raw pixel counts for burn classes
      Unburned = sbs5 + sbs6,
      `Very Low` = sbs1,
      Low = sbs2,
      Moderate = sbs3,
      High = sbs4,
      
      # 2. Convert to Square Kilometers (km2)
      Unburned_km2 = Unburned * cnv_to_km2,
      VeryLow_km2 = `Very Low` * cnv_to_km2,
      Low_km2 = Low * cnv_to_km2,
      Moderate_km2 = Moderate * cnv_to_km2,
      High_km2 = High * cnv_to_km2,
      total_km2 = total * cnv_to_km2,
      
      # 3. Calculate Percentages (Assumes 'total' column exists in raw pixels)
      Unburned_pct = (Unburned / total) * 100,
      VeryLow_pct = (`Very Low` / total) * 100,
      Low_pct = (Low / total) * 100,
      Moderate_pct = (Moderate / total) * 100,
      High_pct = (High / total) * 100
    )%>%
    mutate(across(Unburned:High_pct, ~round(., 1)))%>%
    mutate(extent_burned = ((`Very Low` + Low + Moderate + High)),
           extent_burned_km2 = extent_burned * cnv_to_km2,
           extent_burned_pct = (extent_burned / total) * 100,
           high_burn = (Moderate + High)/total*100,
           burn_class = case_when(high_burn > 50 ~ "Extensive", 
                                  high_burn > 25 & high_burn <= 50 ~ "High", 
                                  high_burn < 25 & Moderate_pct <= 13 & high_burn <= 5 ~ "No/Low",
                                  TRUE ~ "Mid"))
  
  ws_burn_info <- ws_sbs_concise %>%
    left_join(waterbody_buffer_percents %>% 
                select(site_code,
                       `200m Buffer Total Area (km2)` = total_km2,
                       `200m Buffer Burn (km2)` = extent_burned_km2,
                       `200m Buffer Burn (%)` = extent_burned_pct,
                       `200m Buffer High Burn (%)` = high_burn,
                       `200m Buffer Burn Class` = burn_class), by = "site_code")
  
```

## Get data together for Reservoir characteristics table

```{r}
# Reservoir area, normal storage capacity and reservoir depth from NID database
res_meta <- read_csv(here(spatial_dir, "reservoir_metadata.csv"))%>% rename(`Reservoir Operator` = Manager)

all_ws_data <- most_recent_meta %>%
                    select(site_code, Site_Name)%>%
  left_join(ws_chars, by = c("site_code"))%>%
  left_join(ws_burn_info, by = c("site_code"))%>%
  #left_join(ws_all_lc, by = c("site_code"))%>%
  left_join(ws_lc, by = c("site_code"))%>%
  left_join(res_meta, by = c("site_code"))%>%
  #mutate all numeric columns to round to 2 decimal places
  mutate(across(where(is.numeric), ~round(., 2))) %>%
   mutate(Site = paste0(Site_Name, " (", site_code, ")"))%>%
  filter(!is.na(min_elev))%>%
  #cleaning up the names
  select( `Site (site code)` = Site,
          site_code, 
          `Watershed Area (km^2)` = ws_area_km2,
          `Reservoir Elevation (m)` = min_elev,
          `Max Watershed Elevation (m)` = max_elev,
          `Mean Watershed Elevation (m)` = mean_elev, 
          `Normal Storage Capacity (acre ft)`, 
          `Reservoir Depth (m)`, 
          `Reservoir Operator`,
          `Extent Burned (km2)`,
          `Extent Burned (%)`,
          `High Burn (%)`,
          `Watershed Burn Class`,
          `200m Buffer Total Area (km2)` ,
          #`200m Buffer Burn (km2)`,
          `200m Buffer Burn (%)`,
          #`200m Buffer High Burn (%)`,
          `200m Buffer Burn Class`,
          `Mean Slope (degrees)` = mean_slope,
          `2019 Top 2 Land Covers` = lc_top2_2019,
          `2023 Top 2 Land Covers` = lc_top2_2023)

res_ws_data <- all_ws_data%>%
    filter(site_code %in% c("CBRR", "BRNR", "PTRR","COMR", "HORR", "JOER", "LNGR"))%>%
  select(-site_code)

write_csv(res_ws_data, here(spatial_dir, "reservoir_characteristics.csv"))
write_csv(ws_all_lc, here(spatial_dir, "watershed_characteristics.csv"))
```

## Make a table with the land cover and watershed characteristics

```{r}
res_ws_data <- read_csv(here(spatial_dir, "reservoir_characteristics.csv"))
watershed_characteristics <- read_csv(here(spatial_dir, "watershed_characteristics.csv"))

t_res_ws_data<- res_ws_data%>%
  t()%>%
  as.data.frame()%>%
  #make column names the first row
  setNames(.[1,])%>%
  # remove first row
  .[-1,]%>%
  rownames_to_column(" ")

#create standard border color and width
std_border <- fp_border_default(width = 1, color = "black")

ft <- flextable(t_res_ws_data)%>%
  separate_header() %>%
  align(align = "center", part = "all")%>%
  colformat_double( big.mark = ",", decimal.mark = ".", digits = 3)%>%
  set_table_properties(layout = "fixed")%>%
  border_remove()%>%
  border_outer(part="all", border = std_border )%>%
  border_inner_h(border = std_border, part="all")%>%
  border_inner_v(border = std_border, part="all")%>%
  autofit()%>%
  fontsize( size = 6, part = "body")%>%
  fontsize(size = 6, part = "header") %>% # Smaller font size for the header
  bold(part = "header") 

sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    width = 8.3, height = 11.7
  ),
  type = "continuous",
  page_margins = page_mar()
)
# 
#ft
# 
# # Save the table to a Word document
#save_as_docx(ft, path = "data/figs_output/res_ws_data_table.docx", pr_section = sect_properties)
```


# Maps

### Read in shapefiles for watersheds, sampling locs and waterbodies

```{r}

## Pull in sampling points

sampling_points <- most_recent_meta %>%
  #filter for all upper sites (or reservoirs)
  filter(grepl("Upper|Reservoir", watershed))%>%
  st_as_sf(coords = c("Long", "Lat"), crs = 4326)

sampling_points_study_res <- sampling_points%>%
  filter(grepl("Reservoir", watershed))%>%
  mutate(ws_site_code = case_when(
    watershed == "Barnes Meadow Reservoir" ~ "BRNR",
    watershed == "Joe Wright Reservoir" ~ "JOER",
    watershed == "Long Draw Reservoir" ~ "LNGR",
    watershed == "Chambers Reservoir" ~ "CBRR",
    watershed == "Comanche Reservoir" ~ "COMR",
    watershed == "Hourglass Reservoir" ~ "HORR",
    watershed == "Peterson Reservoir" ~ "PTRR"))%>%
  select(site_code, Site_Name, watershed, ws_site_code)
    
#st_write(sampling_points_study_res, "data/raw/spatial/sampling_points.shp")


study_waterbodies <- st_read("data/raw/spatial/res_study/study_waterbodies.shp")%>%
  st_transform(4326)%>%
  select(site_code, Site_Name, watershed)



```

### Reservoirs with SBS

```{r}

# Load the raster
sbs <- raster::raster("data/raw/spatial/res_study/cpf_sbs_BAER.tif")
#project into 4326
sbs <- projectRaster(sbs, crs = 4326)

# Remove water bodies  and blank spaces (0 & 6)
sbs[sbs == 6] <- NA
sbs[sbs == 0] <- NA
# Define burn severity colors and labels

#clip sbs_extended to watershed extent
ws_extent <- ws%>%filter(site_code %in% c("JOER","COMR","HORR","CBRR","PTRR", "BRNR"))%>%
  st_transform(sf::st_crs(sbs_extended))%>%
  st_bbox() %>%
  #add a buffer of 1000 meters to the extent
  st_as_sfc() %>%
  st_buffer(1000)

sbs_clip <- terra::crop(sbs_extended, ws_extent)
#   library(tmap)
# 
# # Switch to interactive mode
tmap_mode(mode = "plot")
severity_cols <- c(
  "#31A354", # 1: Very Low (Pale Yellow)
  "#FFFFB2", # 2: Low (Yellow-Orange)
  "#FD8D3C", # 3: Moderate (Orange)
  "#E31A1C", # 4: High (Red)
  "#31A354", # 5: Unburned (Green)
  "steelblue"  # 6: Water (Blue)
)
  
# Create the map
tm_shape(sbs_clip) +
tm_raster(
  palette = severity_cols,
  breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5),
  labels = c("Very Low", "Low", "Moderate", "High", "Unburned", "Water"),
  opacity = 0.7,
  title = "Burn Severity"
) +
  tm_shape(ws%>%filter(site_code %in% c("JOER","LNGR","COMR","HORR","CBRR","PTRR", "BRNR"))) +
  tm_polygons(
    col = "site_code", 
    palette = "Set2",       # Qualitative palette for distinct site colors
    title = "Monitoring Site",
    alpha = 0.6             # Adds slight transparency to see terrain underneath
  ) +
  tm_borders(col = "black", lwd = 2) +
  tm_shape(study_waterbodies)+
  tm_polygons(col = "blue", alpha = 0.5, border.col = "black", border.lwd = 1, title = "Waterbodies") 
  tm_layout(title = "Watershed vs. SBS Overlap")
```

# Flow data

```{r}

# pull dwr claftcco data from 1980-2021
q_data <- get_sw_ts(abbrev = "CLAFTCCO", timescale = "day", start_date = "1980-10-01", end_date = "2020-09-30")
  
q_data_clean <- q_data%>%
  mutate(q_cfs = value, 
           site_code = abbrev, 
         acre_ft_day = q_cfs *1.983 )%>%
    distinct()%>%
  # remove rows where flag_b is "Ice"
  filter(!grepl("Ice", flag_b))%>%
  select(abbrev, Date = datetime, q_cfs, acre_ft_day, flag_a, flag_b)%>%
  addWaterYear()

q_annual <- q_data_clean%>%
  group_by(waterYear)%>%
  summarise(t_flow_acft = round(sum(acre_ft_day), 0), 
            max_cfs = max(q_cfs), 
            #find the Date with the highest q_cfs per year
            max_cfs_date = Date[which.max(q_cfs)],
            )%>%
  ungroup()%>%
  mutate(peak_dy = yday(max_cfs_date), 
         month = month(max_cfs_date))

mean(q_annual$t_flow_acft)

```
