---
title: "testing"
author: "Sam Struthers"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("paper_analysis/00_analysis_setup.R")

#basic function to save a plot in standard size, dpi and width into figures folder
save_plot_image<- function(plot,name){
  
  ggsave(paste0("data/figs_output/2025_figs/",name,".jpg"), dpi = 300, width = 18, height = 10)
  
}

```

# Grab data

```{r}

#discrete sample data and location metadata from most recent pub



DOI = "14397591"
data_version = "v.2024.12.11"


source("paper_analysis/functions/grab_zenodo_data.R")

 all_params <- chem_units%>%
  pull(simple)
 
 tidy_chem <- most_recent_chem%>%
  #remove random pre fire data and remove 2024 data which will be omitted from publication
  filter(Year > 2020 & Year != 2024)%>%
  # pfal + slep represent the same site but were moved to match future sensor location
  mutate(distance_upstream_km = case_when(site_code == "PFAL" ~ 73.9, TRUE ~ distance_upstream_km))%>%
  # add order table which gives the sites an order to be plotted in
  left_join(order_table, by = "site_code")%>%
  mutate(n_nh4 = NH4 * 0.7765, 
         n_no3 = NO3 * 0.2259, 
         DIN = n_nh4 + n_no3,
         DON = DTN - DIN)


#site order now added to tidy chem and thus not needed
rm(order_table, DOI, data_version)
print("Data Successfully loaded for analysis")
```

## Run above to get data loaded

```{r}

```


# Testing New Ideas:
- DOC in lakes 
  does high DOC in reservoirs result in lower chla?
  - look at chla vs DOC in lakes
  - Selecting samples for genomics
  
# Figs for presentation

## Barnes chla timeline

```{r}

monthly_res_chla <- tidy_chem %>%
  filter(Year %in% c(2021, 2022, 2023) & !is.na(ChlA) & site_code %in% mainstem_Res_only) %>%
  select(site_code, Date, NO3, DTN, ChlA, Year) %>%
  mutate(
    month = factor(month(Date, label = TRUE, abbr = TRUE), 
                   levels = c( "May", "Jun", "Jul", "Aug", "Sep", "Oct")), # Ensure correct month order
    month_year = factor(paste(month, Year), 
                        levels = unique(paste(month(Date, label = TRUE, abbr = TRUE), Year))) # Preserve order
  ) %>%
  group_by(site_code, Year, month, month_year) %>%
  summarise(
    monthly_mean_chla = mean(ChlA, na.rm = TRUE),
    monthly_mean_NO3 = mean(NO3, na.rm = TRUE),
    monthly_mean_DTN = mean(DTN, na.rm = TRUE)
  )%>%
  mutate(Year = as.character(Year))%>%
  filter(!is.na(month))%>%
  mutate(site_name = case_when(
    site_code == "LNGR" ~ "Long Draw",
    site_code == "JOER" ~ "Joe Wright",
    site_code == "CBRR" ~ "Chambers",
    site_code == "PTRR" ~ "Peterson",
    site_code == "BRNR" ~ "Barnes Meadow"))%>%
  filter(site_name == "Barnes Meadow")

monthly_res_chla$Year <- factor(monthly_res_chla$Year, levels = c("2021", "2022", "2023"))
# monthly_res_chla$site_code <- factor(monthly_res_chla$site_code, levels = c("LNGR", "JOER", "CBRR", "PTRR", "BRNR"))
monthly_res_chla$site_name <- factor(monthly_res_chla$site_name, levels = c("Long Draw", "Joe Wright", "Chambers", "Peterson", "Barnes Meadow"))

barnes_monthly_chla <-   
  ggplot(data = monthly_res_chla, aes(x=month, y=monthly_mean_chla, color=Year, group=Year)) +
  geom_line( linewidth=1.5) +
  geom_point( size=5) +
  scale_color_manual(values = c("2021" = "#009E73",
                                "2022" = "#56B4E9",
                                "2023" = "#D55E00")
                       )+
  geom_abline(aes(slope=0, intercept=2.6), color='black', linetype="dashed") +
  geom_abline(aes(slope=0, intercept = 8), color='black', linetype="dashed") +
  labs(x = "Date", y = " Chlorophyll a (μg/L)") +
  theme_few(base_size=28) +
  #coord_cartesian(ylim=c(0,10))+
  #scale_x_date(date_breaks = "months", date_labels = "%b")+
  facet_wrap(~site_name, 
             #scales = "free_y",
             nrow = 1)
# 
# "#002EA3", "#E70870", "#256BF5", 
#                  "#745CFB", "#1E4D2B",
  
  
save_plot_image(barnes_monthly_chla, "barnes_monthly_chla")

```

  
  
  
  
  ## Genomic

```{r}

# filter tidy chem for the dates that genomics were collected
dates <- ymd(c("2022-06-22", "2022-08-25", "2022-09-19" ))

genomes <- tidy_chem%>%
  filter(Date %in% dates)%>%
  #none of these samples have tn, tc and no detectable PO4
  select(site_code, Date, DT_mst, Turbidity:SO4, -TN, -TC,-PO4, watershed)

write_csv(genomes, "data/genomes.csv")

genomes_long <- genomes%>%
  pivot_longer(cols = Turbidity:SO4, names_to = "parameter", values_to = "value")

# make a plot of date and each parameter, color by site
ggplot(data = genomes_long%>%filter(parameter %in% c("Turbidity", "ChlA", "NO3", "DOC", "K")), aes(x = Date, y = value , color = site_code, shape = watershed))+
  geom_point()+
  facet_wrap(~parameter, scales = "free_y")+
  theme_bw()

```
  
# TOC dist

```{r}
#make distribution of TOC, facet at each site


toc_chem <- tidy_chem %>%
  filter(!is.na(TOC)& !is.na(WQ_sensor)) %>%
  filter(Date >= "2023-09-01")%>%
  filter(site_code %nin% c("timberline", "prospect", "boxelder", "riverbluffs", "PNF"))

sample_counts <- toc_chem%>%
  count(Site, name = "n")

unique(toc_chem$Site)

toc_chem$Site <- factor(toc_chem$Site, levels = c( "Joe Wright Inlet","Chambers Inlet", "Chambers Outflow","Poudre below Poudre Falls",   "Poudre Below Rustic", "Little Beaver", "Pennock (Signal Mountain)","Poudre South Fork","Poudre at Manner's Bridge","Poudre at Bellvue Diversion","Poudre at Lincoln" , "Poudre at Archery Range","Boxelder Creek above Sanitation District","Spring Creek at Poudre Trail"))

# Plot with sample count as label
toc_plot <- ggplot(data = toc_chem, aes(x = TOC, fill = Site)) +
  geom_histogram(bins = 20) +
  #facet_wrap(~Site)+#, scales = "free_y") +
  # geom_text(data = sample_counts,
  #           aes(x = -Inf, y = Inf, label = paste0("n = ", n)),
  #           hjust = -3, vjust = 2, inherit.aes = FALSE) +
  theme_bw() +
  scale_fill_viridis_d(option = "turbo")+
  labs(x = "TOC (mg/L)", y = "Count", fill = "Site")
save_plot_image(toc_plot, "toc_dist_all")
```

```{r}
# Filter and prepare counts

chla_chem <- tidy_chem %>%
  filter(!is.na(ChlA)& !is.na(WQ_sensor)) %>%
  filter(ChlA < 10)%>%
  filter(Date >= "2023-09-01")%>%
  filter(site_code %nin% c("timberline", "prospect", "boxelder", "riverbluffs", "PNF"))

sample_counts <- chla_chem%>%
  count(Site, name = "n")

unique(chla_chem$Site)

chla_chem$Site <- factor(chla_chem$Site, levels = c( "Joe Wright Inlet","Chambers Inlet", "Chambers Outflow","Poudre below Poudre Falls",   "Poudre Below Rustic", "Little Beaver", "Pennock (Signal Mountain)","Poudre South Fork","Poudre at Manner's Bridge","Poudre at Bellvue Diversion","Poudre at Lincoln" , "Poudre at Archery Range","Boxelder Creek above Sanitation District","Spring Creek at Poudre Trail"))

# Plot with sample count as label
chla_dist <- ggplot(data = chla_chem, aes(x = ChlA, fill = Site)) +
  geom_histogram(bins = 20) +
  #facet_wrap(~Site)+#, scales = "free_y") +
  # geom_text(data = sample_counts,
  #           aes(x = -Inf, y = Inf, label = paste0("n = ", n)),
  #           hjust = -3, vjust = 2, inherit.aes = FALSE) +
  theme_bw() +
  scale_fill_viridis_d(option = "turbo")+
  labs(x = "Chl-a (ug/L)", y = "Count", fill = "Site")


save_plot_image(chla_dist, "chla_dist_all")
```
  
#chla and TOC by season

```{r}

sensor_chem <- tidy_chem %>%
  filter(!is.na(WQ_sensor)) %>%
  filter(Date >= "2023-09-01")%>%
  filter(ChlA <10)%>%
  filter(site_code %nin% c("timberline", "prospect", "boxelder", "riverbluffs", "PNF"))%>%
  mutate(m_d = format(Date, "%m-%d"), 
         plot_date = as.Date(paste0("2023-", m_d), format = "%Y-%m-%d"))%>%
  select(Site, plot_date, TOC, ChlA, TN)%>%
  pivot_longer(values_to = "value", names_to = "parameter", cols = c(TOC, ChlA, TN))

sensor_chem$Site <- factor(sensor_chem$Site, levels = c( "Joe Wright Inlet","Chambers Inlet", "Chambers Outflow","Poudre below Poudre Falls",   "Poudre Below Rustic", "Little Beaver", "Pennock (Signal Mountain)","Poudre South Fork","Poudre at Manner's Bridge","Poudre at Bellvue Diversion","Poudre at Lincoln" , "Poudre at Archery Range","Boxelder Creek above Sanitation District","Spring Creek at Poudre Trail"))

ggplot(data = sensor_chem, aes(x = plot_date, y = value, color = Site))+
  geom_point()+
  theme_bw() +
  facet_wrap(~parameter, scales = "free_y")+
  scale_color_viridis_d(option = "turbo")+
  #geom_smooth(method = "lm")+
  theme_bw()+
  labs(x = "Date", y = "Value", color = "Site")

```
  
  
  
## Chla vs DOC and NO3:
  No huge pattern
```{r}
ggplot(data = filter(tidy_chem, site_code %in% c("CBRR", "JOER","BRNR","LNGR", "PTRR", "COMR", "HORR")), aes(x = ChlA, y = DOC, color = site_code))+ 
  geom_point()+
  scale_color_manual(values = colorsBS_site_code)+
  facet_wrap(~Year)+
  #geom_smooth(method = "lm")+
  theme_bw()

#chla no3 relationship
ggplot(data = filter(tidy_chem, site_code %in% c("CBRR", "JOER","BRNR","LNGR", "PTRR", "COMR", "HORR")), aes(x = ChlA, y = NO3))+
  geom_point()+
  scale_color_manual(values = colorsBS_site_code)+
  #facet_wrap(~Year)+
  geom_smooth(method = "lm")+
  theme_bw()

no3_rolling <- tidy_chem%>%
  filter(site_code %in% c("CBRR", "JOER","BRNR","LNGR", "PTRR"))%>%
  select(site_code, Date, NO3, ChlA, Year)

all_no3 <- tibble()

for(yr in unique(no3_rolling$Year)){
  
for(i in unique(no3_rolling$site_code)){
  site_year <- no3_rolling%>%
    filter(site_code == i & Year == yr )%>%
     arrange(site_code, Date)%>%
  mutate(back1_no3 = dplyr::lag(NO3, n = 1), 
         back1_chla = dplyr::lag(ChlA, n = 1),
         slope1_no3 = (NO3 - back1_no3)/2,
         slope1_chla = (ChlA - back1_chla)/2)
  
  all_no3 <- bind_rows(all_no3, site_year)
}
}
 
 ggplot(data = all_no3, aes(x = slope1_no3, y = slope1_chla, color = site_code))+
  geom_point()+
  #add 1:1 line
  geom_abline(intercept = 0, slope = 0, linetype = "solid", color = "red")+
  #add vertical line at x = 0
  geom_vline(xintercept = 0, linetype = "dashed", color = "blue")+
  scale_color_manual(values = colorsBS_site_code)+
  facet_wrap(~site_code)+
  theme_bw()


```


## Nitrogen stuff!!

```{r}

chem_n <- tidy_chem%>%
  filter(watershed %nin% c("CLP Mainstem - Lower","CLP Tributary - Lower"))%>%
  filter(!is.na(NO3)& !is.na(DTN) & !is.na(NH4))%>%
  mutate(n_nh4 = NH4 * 0.7765, 
         n_no3 = NO3 * 0.2259, 
         DIN = n_nh4 + n_no3,
         DON = DTN - DIN)

# DIN vs DON
ggplot(data = chem_n, aes(x = DIN, y = DON, colour = location_type))+
  geom_point()+
  facet_wrap(~watershed)+
  theme_bw()

means_n <- filter(chem_n, watershed %nin% c("CLP Mainstem - Upper", "CLP Tributary - Upper"))%>%
  group_by(site_code, Year, Watershed_Level, watershed)%>%
  summarise(mean_NO3 = mean(NO3, na.rm = TRUE),
            mean_DON = mean(DON, na.rm = TRUE),
            mean_DIN = mean(DIN, na.rm = TRUE), 
            mean_DTN = mean(DTN, na.rm = TRUE))

#TDN and NO3 on log scale with burn metric 

ggplot(data = means_n, aes(x = Watershed_Level , y = mean_DTN))+
  geom_point()+
  geom_smooth(method = "lm")+
  facet_wrap(~Year)+
  theme_bw()


```


## Flow impact on river WQ

```{r}


library(cdssr)


clp_structures <- cdssr::get_structures(water_district = 3)

our_res <- clp_structures%>%
  filter(wdid %in% c("0303679","0303683", "0303678","0303677", "0303676", "0303720", "0303686"))%>%
  select(wdid, structure_name)%>%
  mutate(site_code = case_when(structure_name == "LONG DRAW RESERVOIR" ~ "LNGR",
                               structure_name == "MOUNTAIN SUPPLY RES 20" ~ "JOER",
                               structure_name == "BIG BEAVER RES" ~ "HORR",
                               structure_name == "BARNES MEADOW RESERVOIR" ~ "BRNR",
                               structure_name == "PETERSON LAKE RES" ~ "PTRR",
                               structure_name == "COMANCHE RES" ~ "COMR",
                               structure_name == "CHAMBERS LAKE RES" ~ "CBRR", 
                               TRUE ~NA))


get_res_volume <- function(id = "0303679"){
  res_volume <- cdssr::get_structure_stage_ts(wdid = id, start_date = "2019-11-01", end_date = "2023-12-31")%>%
  arrange(datetime)%>%
    mutate(change = volume - lag(volume, n = 1),
           days = as.numeric(difftime(datetime, lag(datetime, n = 1), units = "days")),
           daily_change = change/days)
  return(res_volume)
}

res_storage_tidy <- map_dfr(our_res$wdid, get_res_volume)%>%
  left_join(our_res, by = "wdid")%>%
  mutate(day_of_year = yday(datetime))


get_res_dev_rel <- function(id = "0303679"){
  res_dv <- cdssr::get_structures_divrec_ts(wdid = id, start_date = "2019-11-01", end_date = "2023-12-31", timescale = "day")
  res_rel <- cdssr::get_structures_divrec_ts(wdid = id, wc_identifier = "release", start_date = "2019-11-01", end_date = "2023-12-31", timescale = "day")
  
  res_dv_rel <- bind_rows(res_dv, res_rel)
  return(res_dv_rel)
}

res_diversions_releases <- map_dfr(our_res$wdid, get_res_dev_rel)%>%
  left_join(our_res, by = "wdid")%>%
  mutate(day_of_year = yday(datetime), 
         #if wc_identifier contains "Release", type is release
         type = ifelse(str_detect(wc_identifier, "Release"), "release", "diversion"))

ggplot(data = res_diversions_releases%>%filter(site_code == "BRNR"& year(datetime)!=2019), aes(x = yday(datetime), y = data_value, color = type))+
  geom_point()+
  #geom_line(data = JOEO_Q, aes(x = date, y = JOEO_Q_cfs), color = "black")+
  labs( x = "Day of Year", y = "Outflow (cfs)")+
  facet_wrap(~year(datetime))+
  theme_bw()


upper_Q_gauges <- tidy_chem%>%
  filter(!is.na(flow_gauge_id)&flow_gauge_source != "Larimer County")%>%
  select(site_code, flow_gauge_id, flow_gauge_source)%>%
  distinct()%>%
  mutate(flow_gauge_id = ifelse(flow_gauge_source == "USGS", paste0("0",flow_gauge_id), flow_gauge_id))%>%
  select(site_cd = site_code, id = flow_gauge_id, source = flow_gauge_source)

grab_daily_q <- function(site_cd, id, source){
  
  if(source == "USGS"){
    q_data <- readNWISdv(siteNumbers = id,startDate = "2019-01-01",endDate = "2023-12-31" ,  parameterCd = "00060")%>%
    select(source = agency_cd,site = site_no,Date,
           Q_cfs = X_00060_00003, Q_flag = X_00060_00003_cd)%>%
    distinct()%>%
    mutate( site_code = site_cd )
    
} else if(source == "CDWR"){
  q_data <- cdssr::get_sw_ts(abbrev = id, start_date = "2019-11-01", end_date = "2023-12-31")%>%
    mutate(Q_cfs = value, 
           site_code = site_cd, 
           Q_flag = paste(flag_a, flag_b, sep = "-"))%>%
    distinct()%>%
    select(Date = datetime, Q_cfs, site_code, site = abbrev, source = data_source, Q_flag)
  }
  return(q_data)
}


CLP_daily_Q <- pmap_dfr(upper_Q_gauges, grab_daily_q)%>%
  filter(!grepl('Ice', Q_flag) & !grepl('Ssn', Q_flag)& Q_cfs > 0)%>%
  left_join(tidy_chem, by = c("site_code", "Date"))

some_CLP_daily_Q <- filter(CLP_daily_Q, site_code %nin% c("lincoln", "ELC", "boxelder")& Date > as.Date("2021-03-01"))%>%
  filter(site_code %in%c("CHD", "PBD") )


ggplot(some_CLP_daily_Q, aes(Date, Q_cfs))+
  geom_line(aes(color = site_code))+
  geom_point(data = some_CLP_daily_Q%>%filter(!is.na(Site)), aes(fill = ChlA, size = NO3, shape = site_code))+
  scale_fill_continuous(limits = c(0, 4), low = "blue", high = "green")+
  #add a vertical line for every day there site_name is not NA
  #geom_vline(data = some_CLP_daily_Q%>%filter(!is.na(Site)), aes(xintercept = as.numeric(Date)), color = "red")+
  labs(x = "Date", y = "Flow (cfs)", color = "ChlA", size = "NO3")+
  #facet_wrap(~site_code, scales = "free_y")+
  theme_bw()
  
  prop_chd_pbd <- left_join(chd_Q, pbd_Q, by = "date")%>%
    mutate(prop_chd_pbd = chd_Q_cfs / pbd_Q_cfs, 
           prop_chd_pbd =ifelse(prop_chd_pbd>=1, 1, prop_chd_pbd), 
           month = as.character(month(date)),
           season = case_when(
          month %in% c( "12", "1", "2", "3") ~ "Winter",
           month %in% c("4", "5", "6") ~ "Spring",
           month %in% c("7", "8") ~ "Summer",
           month %in% c("9", "10","11") ~ "Fall")
    )
 

  ggplot(tidy_chem%>%filter(location_type == "Reservoir"), aes(x = ChlA, TSS, color = location_type))+
    geom_point()+
    geom_smooth(method = "lm")+
    ylim(0,50)+
    theme_bw()
  
  
  
chem_to_join <- tidy_chem%>%
  filter(site_code %in% c( "JWC", "PBR", "PBD"))%>%
  mutate(date = as.POSIXct(Date))

chem_q_join <- left_join(prop_chd_pbd, chem_to_join, by = "date")%>%
  filter(season %nin% c("Winter"))


ggplot(chem_q_join, aes(x = prop_chd_pbd)) +
  geom_point(aes(y = ChlA, shape = site_code, color = season), size = 5)+
  theme_bw(base_size = 20)

```


 
 ## Make a bar plot of SC and then break up that bar by the proportion of total for each ion for each sampling date
 
```{r}
#make a bar plot of SC and then break up that bar by the proportion of total for each ion


data<- tidy_chem%>%
  filter(site_code %nin% c("PNF", "PMAN", "PBR", "boxelder", "prospect", "timberline", "riverbluffs"))%>%
  select(site_code, Date, SC, DOC, Na:SO4)%>%
  mutate(ion_sum = rowSums(across(c(DOC:SO4)), na.rm = TRUE), 
         base= rowSums(across(c(Ca, Cl,Na, Mg)), na.rm = TRUE))%>%
  mutate(across(c(base, DOC, F,K, NO3, NH4, PO4, SO4), ~ . / ion_sum * 100, .names = "prop_{.col}")) %>%
  select(site_code, Date, SC,  starts_with("prop_"))

# Prepare the data in long format
data_long <- data %>%
  pivot_longer(cols = c(prop_DOC:prop_SO4, prop_base),
               names_to = "Ion", values_to = "Proportion") %>%
  mutate(Ion = str_remove(Ion, "prop_"))  # Clean up ion names




# Plot
ggplot(data_long%>%filter(year(Date) == 2022), aes(x = Date, y = Proportion, fill = Ion)) +
  geom_col(position = "stack", width = 3) +  # Stack the bars
  #geom_line(data = data, aes(x = Date, y = SC), color = "black", size = 1, inherit.aes = FALSE) + # SC line
  labs(x = "Date", 
       y = "Proportion of Total Ions (%)", 
       fill = "Ion") +
  theme_bw()+
  facet_wrap(~site_code)+
  ylim(0,100)# Separate by site code



# make a column graph of SC for each date
ggplot(data, aes(x = Date, y = SC))+
  geom_col()+
  geom_col(aes(y = prop_Na, fill = "Na"))+
  geom_col(aes(y = prop_Cl, fill = "Cl"))+
  geom_col(aes(y = prop_SO4, fill = "SO4"))+
  geom_col(aes(y = prop_DOC, fill = "DOC"))+
  geom_col(aes(y = prop_NH4, fill = "NH4"))+
  geom_col(aes(y = prop_Ca, fill = "Ca"))+
  geom_col(aes(y = prop_F, fill = "F"))+
  geom_col(aes(y = prop_Mg, fill = "Mg"))+
  geom_col(aes(y = prop_K, fill = "K"))+
  geom_col(aes(y = prop_NO3, fill = "NO3"))+
  geom_col(aes(y = prop_PO4, fill = "PO4"))+
  labs(x = "Date", y = "SC (uS/cm)", fill = "Ion")+
  theme_bw()
```
 
 
